### 2022-01-19 Manuscript/StandardizedEvaluation.Rmd
Add Rmd and journal template

---
title: "Towards a standardized evaluation of multiple imputation routines"
author: "Hanne Oberman and Gerko Vink"
date: "19-1-2022"
abstract: "Developing new imputation methodology has become a very active field. Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods. In this paper I propose a move towards a standardized evaluation of imputation methods. To demonstrate the need for standardization, I highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines. This may lead to suboptimal use of multiple imputation in practice. Additionally, I suggest a course of action for simulating and evaluating missing data problems."
bibliography: StandardizedEvaluation.bib  
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

### 2022-01-21 Manuscript/StandardizedEvaluation.Rmd
Convert latex manuscript to markdown

@@
 # Introduction
+
+Multiple imputation @rubin1987 is a state-of-the-art technique for drawing valid conclusions from incomplete data, and is increasingly used in many substantive fields, such as medicine see e.g.
+@pulkki2015cumulative, @alkire2015global, @stinesen2015reduced, psychology e.g., @sorbi2014medium, economics e.g., @wong2015economic, @dafoe2013democratic, @allan2015determinants, sociology e.g., @crowley2014flexible, @de2014religiosity and official statistics e.g., @little2012calibrated, @klausch2015selection.
+
+The technique has earned a permanent spot in research and policy making, demonstrated e.g.
+by the detailed manual created by the National Research Council @little2012prevention.
+Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing multiple imputation techniques.
+After all, the principle of multiple imputation is very intuitive.
+
+The idea behind multiple imputation is to impute (fill in) the missing values multiple times to obtain a valid estimate of what could have been.
+The multiple datasets that are thus obtained can be analyzed by standard techniques and the analysis results can be combined into a single inference.
+In contrast to ad hoc methods for dealing with missing values (e.g.
+list-wise deletion, mean imputation, regression imputation, last observation carried forward), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
+
+The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
+In general it holds that modeling missing data becomes more challenging when the amount of missingness increases.
+However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
+In that case, multiple imputation would be substantially more efficient than the ubiquitous list-wise deletion.
+
+When evaluating the statistical properties (and thereby the practical applicability) of new imputation methodology, researchers most often make use of simulation studies.
+In such studies a complete data set is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of the imputation method.
+However, no golden standard has been established and, as a result, validity of the simulation and evaluation procedures may differ tremendously from one developer to another.
+
+The purpose of this paper is threefold: First, I raise some concerns with respect to evaluating imputation methodology.
+These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals.
+Second, I provide imputation methodologists with a suggested course of action when simulating missing data problems.
+This suggested approach should identify a common ground, but is in no way intended as an absolute solution.
+This identifies the third purpose of this paper: discussion.
+I hope to elicit critical thinking with respect to the problems at hand.
+We are all convinced that our methodology has some merit.
+But for sake of progress it would be much more advantageous if the aim of our evaluations would go beyond \emph{proving the point} and would legitimately consider the statistical properties.
+
+## Why some evaluations should not be trusted
+
+As of today, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
+Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers.
+This brings forth a chain of potential problems in the objective assessment of the performance of imputation routines that may lead to suboptimal use of multiple imputation in practice.
+To demonstrate the broad impact of these problems, I subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation.
+I further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.
+
+## Data generation problems
+
+To evaluate the ability of an imputation routine to handle the nonresponse model, a form of truth has to be established.
+Those who perform simulation studies are in the luxury position to establish the truth beforehand, either by generating data from a statistical model, or by sampling data from a sufficiently large observed set.
+However, when generating data, two easily overlooked problems may arise.
+
+First, data are often generated following the model (or distribution) that is also used for imputing the data.
+The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied.
+Although such procedures may be statistically relevant, the approach would be no good for even the simplest case.
+Such simulations pose serious threats to the applicability of the method as real-life data hardly ever follow a given theoretical distribution.
+
+Second, data are often generated such that the problem that is being studied is most pronounced.
+This results in simulated data that contains very valuable information structures, i.e.
+the correlations between groups of variables may be very pronounced.
+In other words, no matter what type of missingness is induced, the observed parts of the data may still hold much, if not all, of the information about the missing part and the performance of the imputation procedure is not surprisingly evaluated as very good.
+Also, it may be that the relations in real-life data are rarely as pronounced as the problem that is being studied.
+
+## Missingness generation problems
+
+In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng1994multiple.
+The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
+Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable see e.g.
+@rubin1976inference, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
+
+With MCAR missingness mechanisms, the probability to be missing is the same for all cases.
+This is a necessary simulation condition for evaluating the performance of imputation procedures.
+If an imputation method is not able to solve the problem (i.e.
+yield valid inference) under MCAR, the statistical properties of the procedure are not sound.
+Sadly, the straightforward case of MCAR is often neglected from simulation studies and focus is drawn to the evaluation of MAR mechanisms only.
+Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in my view may be far too simplistic.
+
+```{r mis_mech, echo=FALSE, fig.cap="Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
+Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$."}
+knitr::include_graphics("img/plot_mmech.pdf")
+```
+
+With MAR missingness mechanisms, the probability to be missing is the same within groups of cases defined by the observed data only (e.g.
+males are less likely to disclose their weight, but gender is observed).
+The essence here is that the observed relations in the data are used to induce missingness during simulation (e.g.
+weight is made incomplete based on gender), but these relations may be weak, or non-existent.
+
+When MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid.
+When MAR is induced from data without multivariate relations, the inferential implication of the missingness would mimic those of MCAR.
+This problem is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.
+Figure \@ref(fig:mis_mech) demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
+
+It is obvious that the MAR missingness generated from data with low correlation would yield similar inference to the MCAR missingness, even though the missingness is clearly random.
+The simulation results from Table 1 confirm this.
+A detailed simulation setup can be found in Appendix A.
+
+As expected, under genuine MAR missingness, i.e.
+the missingness probabilities differ conditional on the level of the incomplete variable, the performance of complete case analyses drops dramatically.
+For \emph{spurious} MAR, i.e.
+the conditional probabilities remain the same over the level of the incomplete variable, the performance is comparable to that under MCAR.
+In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation @fimd, p.
+48: the missingness does not depend on the incomplete variable.
+This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.
+
+Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit @molenberghs2008every - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
+In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis see e.g.
+@molenberghs2014handbook, part 5.
+
+<!-- table -->
+<!-- Inferences obtained by complete case analysis (CCA) and multiple imputation over 1000 simulations.
+Displayed are the bias, coverage rate (cov) of the 95% confidence interval and the confidence interval width (ciw).
+Imputations are generated by Bayesian linear regression imputation.
+-->
+
+The amount of missingness that is used to simulate the performance of imputation methods differs between studies.
+Some studies use only 10% missing data where other studies push the limits to additionally investigate the performance under at least 50% missing data.
+This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
+This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
+
+## Performance evaluation
+
+The evaluation criteria used to assess imputation performance vary from one developer to another.
+This is not surprising as people from different fields could have a different focus on the problem at hand.
+However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated.
+I highlight four aspects of evaluation that should be considered when evaluation imputation routines.
+
+First, developers often only inquire about the 'accuracy' (i.e.
+how well can the method reproduce the original data).
+The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
+We are interested in the correct answer to the research question; not in the truth itself.
+This means that, given the framework provided by @rubin1987, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
+After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neyman1934two, p.
+591.
+
+Second, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
+For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large.
+While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.
+Rather, one would prefer an imputation technique to yield both valid inference and plausible imputations.
+It should be studied if an imputation method is prone to deliver such impractical results, and if so, under what conditions.
+
+Third, sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
+However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink2014pooling.
+After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
+
+Last, when the evaluator is on the verge of drawing conclusions about the performance of the imputation routine, the performance should be carefully qualified.
+Comparing the performance of an imputation routine given a population (or true) parameter allows for quantitative evaluation.
+However, in order to pose qualitative statements about the performance on simulated conditions, comparative methodology is required.
+For example, when claiming that imputation performance is unacceptable when deviations from normality become rather stringent, such performance is highly dependent on the simulation conditions that are used.
+For a well-balanced judgement about the severity of the performance drop, comparative simulations with e.g.
+nonparametric models should be executed.
+A method may perform badly, but if it still outperforms every other approach, it may yet be of great practical relevance.
+
+# Suggested course of action
+
+## Step 1. Obtain truth
+
+The evaluator must to decide whether there is a need for sampling variance in the simulations scheme. If sampling variance cannot be omitted from the simulation scheme, two approaches are possible:
+
+- Model-based simulation: draw samples from a known probability distribution, such as the multivariate normal distribution.
+The theoretical parameters under which the samples are obtained will serve as the comparative truth during the simulations.
+- Design-based simulation: sample data from a sufficiently large set.
+The parameters of the sufficiently large set will serve as the comparative truth during the simulations.
+This approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
+Applications of design-based simulation are often found in official statistics.
+A benefit of design-based simulation is the ability to use real-life observed data structures.
+
+For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches.
+The parameters of that single complete set will serve as comparative truth.
+This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations.
+One needs to realize, however, that the conventional pooling rules cf.
+@rubin1987, pp.
+76-77, do not apply for finite population inference and that alternative pooling rules need to be used @raghunathan2003multiple, @vink2014pooling.
+
+## Step 2. Induce missingness
+
+First, one should always consider MCAR missingness, i.e.
+the scenario where the missing values are missing at random and the observed values are observed at random.
+Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
+
+Next, missing data should be induced conform a model that is dependent on the observed data.
+A straightforward technique for inducing different forms of univariate MAR missingness is described in @fimd, p.
+63; see Figure \@(fig:MAR) and a generalization to multivariate MAR missingness can be found in @van2006fully, Appendix B, and @brand1999development, par.
+5.2.3.
+If the missingness is to be induced in longitudinal data, auto-regressive MAR models e.g.
+cf.
+@shara2015randomly, model 2 and model 3, can be useful.
+
+```{r MAR, echo=FALSE, fig.cap="Examples of four random missingness mechanisms that yield approximately 50 percent missingness.
+For negatively correlated missingness covariates the resulting mechanisms are reversed."}
+knitr::include_graphics("img/plot_mar.pdf")
+```
+
+Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
+Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
+This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching @little1988missing, @rubin1986statistical may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+It would be a shame to overlook such results due to the focus on a single MAR mechanism.
+
+Fourth, the amount of missingness must be varied.
+Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
+
+The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
+If there is no missingness - or if there is no data, for that matter - the inference does not depend on the assumption.
+Alternatively, the validity of assumptions become increasingly important when the missingness increases.
+Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
+I therefor propose, for all mechanisms, to evaluate at least the following univariate missingness percentages when evaluating imputation routines:
+
+- 10% missingness: Depending on the size of the data, this percentage can be considered as a lower bound of realistic evaluation.
+Anything less than 10% may be of little influence on the true data inference.
+Performance of a missing data method should at least be acceptable for most missing data problems.
+- 25% missingness: This is a fair amount of missingness and will, depending on the observed data information, have a noticeable influence on the completed data inference.
+When compared to the condition with 10% missingness, the inference obtained under 25% missingness should be less certain (i.e.
+confidence/credibility interval width should increase), but estimates should still be properly covered and the statistical properties of the missing data method should be sound.
+In practice, at least to my experience in social sciences and official statistics, 25% univariate missingness can easily be considered as a realistic missingness percentage.
+- 50% missingness: Performance under 50% percent simulated missingness will most likely be impacted severely.
+Depending on how the missingness mechanism interacts with the simulated data, some imputation techniques may yield estimates that are under-covered such that the completed data inference should not be deemed valid anymore.
+If a method yields acceptable inference under 50% MAR missingness, we can determine that the statistical properties of the imputation methodology are sound.
+
+Although I limit the focus here to ignorable non-response, the above suggested proportions are equally applicable to simulations under non-ignorable non-response.
+
+Fifth, the used missingness mechanism should be detailed, either graphically or written as a function of the data.
+Often, when inducing missingness, authors remain vague about the actual missingness mechanism under investigation and, even worse, some authors only report something like
+\emph{We generated missing data following a MAR missingness mechanism.}
+This should be considered unacceptable as claims about the validity of the multiple imputation inference depends heavily on the simulated missingness mechanism.
+
+## Step 3. Evaluate performance
+It is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
+We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance.
+The evaluation criteria may depend on the estimand.
+When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
+Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
+In general, I would say that each multiple imputation routine should be evaluated on at least the following points:
+
+- \emph{Bias:} Results should preferable be unbiased.
+However, the way bias is considered can greatly influence the interpretation of the results.
+For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.
+The way bias is computed should therefore be carefully chosen and described.
+- \emph{Interval width:} The way the confidence interval is calculated should be described.
+Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference.
+However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.
+- \emph{Coverage:} Coverage of a 95% interval should in theory be $\geq 95$, where a coverage of 95% would be most efficient.
+Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
+- \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data.
+Under anything but the MCAR assumption, this can be expected.
+When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed see @abayomi2008diagnostics for an detailed overview of diagnostic evaluation for multivariate imputations.
+When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.
+- \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference.
+However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property.
+When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
+- \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers, PIGS, @li2012imputing.
+The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
+
+For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
+This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
+Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
+However, the real-world missingness would then be omitted.
+
+# Discussion
+
+This document is aimed at establishing a common ground for the evaluation of imputation routines.
+Such a common ground would be the basis of a standardized evaluation.
+This allows for fair and efficient comparisons between imputation techniques.
+Ultimately, it would be desirable to evaluate every imputation routine against the same standardized set in order to quantify the statistical properties across imputation routines.
+If properly executed, this would allow for careful matching of imputation methodologies to new missing data problems.
+
+# Appendix A: Simulation setup
+
+Let $Y=(Y_{obs},Y_{mis})$ be an incomplete semi-continuous variable, where $Y_{obs}$ and $Y_{mis}$ denote the observed values and the missing values in $Y$, respectively.
+Further, $X=(X_1,...,X_k)$ is a set of $k$ fully observed covariates, where $X_{obs}$ and $X_{mis}$ correspond to the observed an missing parts in $Y$ and where $X_j$ would indicate the $j$th variable in $X$, with $j=1,\dots, k$.
+Finally, let $R$ be a response indicator that is 1 if $Y$ is observed and 0 if $Y$ is missing.
+
+To prove the point, I generated a single data set from the multivariate normal distribution with means
+$$
+\mu= \bordermatrix{&	 \cr
+ Y	&5	\cr
+{ X}_1	&10	\cr
+{ X}_2	&10	},
+$$
+and covariance matrix
+$$
+\Sigma = \bordermatrix{& Y &{ X}_1 &{ X}_2	 \cr
+ Y	&1	&0.8	&0\cr
+{ X}_1	&0.8	&1	&0\cr
+{ X}_2	&0	&0	&1}.
+$$
+
+Three incomplete sets were then generated, wherein random missingness was imposed in $Y$ in three ways.
+First, I left the observed data \emph{observed at random} to impose MCAR missingness following the mechanism
+$$
+P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0)=.50.
+$$
+Next, I created random missingness in $Y$ according to the following MAR missingness mechanism
+$$
+P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0|Y_{obs}, X_j),
+$$
+by using a random draw from a binomial distribution of the same length as $Y$ and of size 1 with missingness probability equal to the inverse logit
+$$
+P(R=0)=\frac{e^{a}}{(1+e^{a})}.
+$$
+where $a=(\mu_{X_j}-X_{ij})/\sqrt{\mathbb{E}[X_{ij} - \mu_{X_j}]}$ in order to create approximately 50% right-tailed MAR missingness. I followed this procedure twice; once for each $X_j$.
+
+The three resulting incomplete data sets were then imputed with package `mice` @mice, v2.25, in `R` @R, v3.2.3, with Bayesian linear regression imputation (`mice.impute.norm`) as the imputation routine.
+The completed-data analyses were combined into a single inference following the rules described in @vink2014pooling.
+
+# References

### 2022-01-25 Manuscript/StandardizedEvaluation.Rmd
Add to do list and some references

@@
 
+# TO DO
+
+- add ADEMP paper [@morr18], diagnostic checking paper [@zhao22]
+- add PPC (refer to Mingyang, and Gelman & King)
+- expand convergence section
+- add RMSE
+- add insights from literature review for chapter on missing data in EHR
+
 # Introduction
 
-Multiple imputation @rubin1987 is a state-of-the-art technique for drawing valid conclusions from incomplete data, and is increasingly used in many substantive fields, such as medicine see e.g.
-@pulkki2015cumulative, @alkire2015global, @stinesen2015reduced, psychology e.g., @sorbi2014medium, economics e.g., @wong2015economic, @dafoe2013democratic, @allan2015determinants, sociology e.g., @crowley2014flexible, @de2014religiosity and official statistics e.g., @little2012calibrated, @klausch2015selection.
+Multiple imputation @rubin1987 is a state-of-the-art technique for drawing valid conclusions from incomplete data, and is increasingly used in many substantive fields, such as medicine [see e.g.
+@pulkki2015cumulative; @alkire2015global; @stinesen2015reduced], psychology [e.g.
+@sorbi2014medium], economics [e.g.
+@wong2015economic; @dafoe2013democratic; @allan2015determinants], sociology [e.g.
+@crowley2014flexible; @de2014religiosity], and official statistics [e.g.
+@little2012calibrated; @klausch2015selection].
 
 The technique has earned a permanent spot in research and policy making, demonstrated e.g.
-by the detailed manual created by the National Research Council @little2012prevention.
+by the detailed manual created by the National Research Council [@little2012prevention].
 Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing multiple imputation techniques.
@@
 
-To evaluate the ability of an imputation routine to handle the nonresponse model, a form of truth has to be established.
+To evaluate the ability of an imputation routine to handle the non-response model, a form of truth has to be established.
 Those who perform simulation studies are in the luxury position to establish the truth beforehand, either by generating data from a statistical model, or by sampling data from a sufficiently large observed set.
@@
 
-```{r mis_mech, echo=FALSE, fig.cap="Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
+```{r mis_mech, echo=FALSE, fig.cap="\\label{fig:mis_mech}Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
 Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$."}
@@
 This problem is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.
-Figure \@ref(fig:mis_mech) demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
+Figure \ref{fig:mis_mech} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
 
@@
 A straightforward technique for inducing different forms of univariate MAR missingness is described in @fimd, p.
-63; see Figure \@(fig:MAR) and a generalization to multivariate MAR missingness can be found in @van2006fully, Appendix B, and @brand1999development, par.
+63; see Figure \ref{fig:MAR} and a generalization to multivariate MAR missingness can be found in @van2006fully, Appendix B, and @brand1999development, par.
 5.2.3.
@@
 
-```{r MAR, echo=FALSE, fig.cap="Examples of four random missingness mechanisms that yield approximately 50 percent missingness.
+```{r MAR, echo=FALSE, fig.cap="\\label{fig:MAR}Examples of four random missingness mechanisms that yield approximately 50 percent missingness.
 For negatively correlated missingness covariates the resulting mechanisms are reversed."}
@@
 Often, when inducing missingness, authors remain vague about the actual missingness mechanism under investigation and, even worse, some authors only report something like
-\emph{We generated missing data following a MAR missingness mechanism.}
+
+> We generated missing data following a MAR missingness mechanism.
+
 This should be considered unacceptable as claims about the validity of the multiple imputation inference depends heavily on the simulated missingness mechanism.

### 2022-01-25 Manuscript/StandardizedEvaluation.Rmd
Add script for the small simulation from the manuscript

@@
 - add insights from literature review for chapter on missing data in EHR
+- add link to simulation script, and $n$
+- add simulation results
 

### 2022-01-25 R/run_sim.R

@@ -1,0 +1,57 @@
+# small simulation study for illustration
+
+# set-up 
+set.seed(123)
+library(dplyr)
+rmse <- function(obs, pred = 0){sqrt(mean((obs-pred)^2))}
+
+# parameters
+n_sim <- c(100, 1000)
+n_obs <- 500 #c(100, 500, 1000)
+n_var <- 2 #c(2, 5, 20)
+rho <- c(0, .8)
+mis_mech <- c("MCAR", "MAR", "MNAR")
+mis_type <- "right" #c("left", "right", "mid", "tail")
+mis_prop <- c(.1, .25, .5)
+n_imp <- c(1, 5)
+n_it <- c(1, 5, 10)
+imp_meth <- c("mean", "norm.predict", "norm")
+est <- c("mean(Y)", "var(Y)", "cov(X,Y)", "lm(Y~X)", "lm(X~Y)")
+# other: cca, model vs design based; with or without sampling variance
+
+# data generating mechanism
+mu <- c(5, 10, 10)
+sigma <- matrix(0, 3, 3)
+diag(sigma) <- 1
+sigma[1, 2] <- sigma[2, 1] <- 0.8 #make this rho
+
+# generate data
+dat <- data.frame(mvtnorm::rmvnorm(n_obs, mean = mu, sigma = sigma))
+
+# ampute 
+amp.mcar.8 <- mice::ampute(dat[, c("X1", "X2")], prop = .5, mech = "MCAR")$amp
+amp.mcar.0 <- mice::ampute(dat[, c("X1", "X3")], prop = .5, mech = "MCAR")$amp
+amp.mar.8 <- mice::ampute(dat[, c("X1", "X2")], prop = .5, mech = "MAR")$amp
+amp.mar.0 <- mice::ampute(dat[, c("X1", "X3")], prop = .5, mech = "MAR")$amp
+
+# impute and fit analysis model
+imp.mcar.8 <- amp.mcar.8 %>% 
+  mice::mice(method = "norm", print = FALSE) %>% 
+  with(lm(X1 ~ X2))
+
+# pool results
+res.mcar.8 <- imp.mcar.8 %>% 
+  mice::pool() %>% 
+  broom::tidy(conf.int = TRUE) %>% 
+  select("estimate", "conf.low", "conf.high") %>% 
+  .[2, ] %>% 
+  c(rmse_pred = purrr::map_dbl(
+    1:5, 
[…]

### 2022-01-25 Manuscript/StandardizedEvaluation.Rmd
Outline simulation function

@@
 ```{r mis_mech, echo=FALSE, fig.cap="\\label{fig:mis_mech}Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
-Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$."}
+Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\\rho= 0$."}
 knitr::include_graphics("img/plot_mmech.pdf")
@@
 
-> We generated missing data following a MAR missingness mechanism.
+> *We generated missing data following a MAR missingness mechanism.*
 

### 2022-01-25 R/run_sim.R

@@ -3,55 +3,88 @@
 # set-up 
 set.seed(123)
 library(dplyr)
-rmse <- function(obs, pred = 0){sqrt(mean((obs-pred)^2))}
+source("R/utils.R")
 
 # parameters
 n_sim <- c(100, 1000)
 n_obs <- 500 #c(100, 500, 1000)
 n_var <- 2 #c(2, 5, 20)
-rho <- c(0, .8)
+means <- c(10, 100)
+corr <- c(0, .8)
 mis_mech <- c("MCAR", "MAR", "MNAR")
-mis_type <- "right" #c("left", "right", "mid", "tail")
+mis_type <- "RIGHT" #c("left", "right", "mid", "tail")
 mis_prop <- c(.1, .25, .5)
 n_imp <- c(1, 5)
+imp_meth <- c("mean", "norm.predict", "norm") #cca
 n_it <- c(1, 5, 10)
-imp_meth <- c("mean", "norm.predict", "norm")
 est <- c("mean(Y)", "var(Y)", "cov(X,Y)", "lm(Y~X)", "lm(X~Y)")
-# other: cca, model vs design based; with or without sampling variance
+# model vs design based
+# with or without sampling variance
 
 # data generating mechanism
-mu <- c(5, 10, 10)
-sigma <- matrix(0, 3, 3)
-diag(sigma) <- 1
-sigma[1, 2] <- sigma[2, 1] <- 0.8 #make this rho
 
-# generate data
-dat <- data.frame(mvtnorm::rmvnorm(n_obs, mean = mu, sigma = sigma))
+varcov <- matrix(0, 2, 2)
+diag(varcov) <- 1
+varcov[1, 2] <- varcov[2, 1] <- corr 
 
-# ampute 
-amp.mcar.8 <- mice::ampute(dat[, c("X1", "X2")], prop = .5, mech = "MCAR")$amp
-amp.mcar.0 <- mice::ampute(dat[, c("X1", "X3")], prop = .5, mech = "MCAR")$amp
-amp.mar.8 <- mice::ampute(dat[, c("X1", "X2")], prop = .5, mech = "MAR")$amp
-amp.mar.0 <- mice::ampute(dat[, c("X1", "X3")], prop = .5, mech = "MAR")$amp
+# simulation function
+run <- function(
+  n_obs = 500,
+  corr = 0.4,
+  mis_mech = "MAR",
+  mis_type = "RIGHT",
[…]

### 2022-01-26 R/run_sim.R
failed to make the simulation function fully conditional on the parameters, try again with smaller functions

@@ -6,8 +6,8 @@ library(dplyr)
 source("R/utils.R")
 
 # parameters
-n_sim <- c(100, 1000)
-n_obs <- 500 #c(100, 500, 1000)
+n_sim <- 1000 #c(100, 1000)
+n_obs <- c(50, 500, 5000)
 n_var <- 2 #c(2, 5, 20)
 means <- c(10, 100)
 corr <- c(0, .8)
@@ -21,11 +21,11 @@ est <- c("mean(Y)", "var(Y)", "cov(X,Y)", "lm(Y~X)", "lm(X~Y)")
 # model vs design based
 # with or without sampling variance
 
-# data generating mechanism
-
-varcov <- matrix(0, 2, 2)
-diag(varcov) <- 1
-varcov[1, 2] <- varcov[2, 1] <- corr 
+# # data generating mechanism
+# means <- c(10, 100)
+# varcov <- matrix(0, 2, 2)
+# diag(varcov) <- 1
+# varcov[1, 2] <- varcov[2, 1] <- corr 
 
 # simulation function
 run <- function(
@@ -40,51 +40,74 @@ run <- function(
   ...){
     
     # generate data
-    dat <- data.frame(mvtnorm::rmvnorm(
-      n = n_obs,
+    dat <- purrr::map(n_obs,
+      ~{data.frame(mvtnorm::rmvnorm(
+      n = .x,
       mean = means,
       sigma = matrix(c(1, corr, corr, 1), 2, 2)
-    )) %>% setNames(c("Y", "X"))
+    )) %>% setNames(c("Y", "X"))}) %>% 
+      setNames(paste0("n_obs_", n_obs))
     
     # ampute
-    amp <- dat %>% 
-      mice::ampute(
+    amp <- purrr::map(
+      dat,
+      function(.d){
+        mice::ampute(.d,
         prop = mis_prop,
         mech = mis_mech,
         type = mis_type
         ) %>% 
-      .$amp
+      .$amp}) 
     
     # impute and fit analysis model
-    imp <- amp %>%
-      mice::mice(
+    imp <- purrr::map(
+      amp,
+      function(.a){
+        mice::mice(.a,
         m = n_imp,
         method = imp_meth,
         maxit = n_it,
         print = FALSE
       ) %>%
-      with(lm(Y ~ X))
+      with(lm(Y ~ X))})
 
     # pool results
-    res <- imp %>% 
-      mice::pool() %>% 
[…]

### 2022-01-26 R/run_sim.R
Split up simulation function into parts

@@ -9,7 +9,6 @@ source("R/utils.R")
 n_sim <- 1000 #c(100, 1000)
 n_obs <- c(50, 500, 5000)
 n_var <- 2 #c(2, 5, 20)
-means <- c(10, 100)
 corr <- c(0, .8)
 mis_mech <- c("MCAR", "MAR", "MNAR")
 mis_type <- "RIGHT" #c("left", "right", "mid", "tail")
@@ -27,6 +26,58 @@ est <- c("mean(Y)", "var(Y)", "cov(X,Y)", "lm(Y~X)", "lm(X~Y)")
 # diag(varcov) <- 1
 # varcov[1, 2] <- varcov[2, 1] <- corr 
 
+# generate complete data
+generation <- function(...){
+  dat <- data.frame(mvtnorm::rmvnorm(
+    n = n_obs,
+    mean = c(10, 100),
+    sigma = matrix(c(1, corr, corr, 1), 2, 2)
+  )) %>% setNames(c("Y", "X"))
+}
+
+# ampute the data
+amputation <- function(dat, ...){
+  amp <- dat %>% 
+  mice::ampute(
+    prop = mis_prop,
+    mech = mis_mech,
+    type = mis_type
+  ) %>% 
+  .$amp}
+
+# impute the data
+imputation <- function(amp, ...){
+  imp <- amp %>%
+  mice::mice(
+    m = n_imp,
+    method = imp_meth,
+    maxit = n_it,
+    print = FALSE
+  )}
+
+# evaluate the imputations
+evaluation <- function(imp, dat, ...){
+  fit <- imp %>% 
+    with(lm(Y ~ X))
+  
+  rmses <- purrr::map_dfr(1:imp$m, function(.m){
+    data.frame(
+      rmse_cell = rmse(imp$imp$Y[[.m]] - dat$Y[imp$where[, "Y"]]), 
+      rmse_pred = rmse(fit$analyses[[.m]]$residuals))    
+  }) %>% colMeans() %>% t()
+  
+  fit %>% 
+    mice::pool() %>% 
+    broom::tidy(conf.int = TRUE) %>% 
+    .[2, ] %>% 
+    mutate(
+      bias = estimate - corr,
+      ciw = conf.high - conf.low,
+      cr = conf.low <= corr && corr <= conf.high,
+      .keep = "none") %>% 
+    cbind(., rmses)
+}
+
 # simulation function
 run <- function(
   n_obs = 500,
[…]

### 2022-01-27 R/run_sim.R
Add default arguments to simulation function

@@ -7,47 +7,41 @@ source("R/utils.R")
 
 # parameters
 n_sim <- 1000 #c(100, 1000)
-n_obs <- c(50, 500, 5000)
+n_obs <- 500 #c(50, 500, 5000)
 n_var <- 2 #c(2, 5, 20)
-corr <- c(0, .8)
-mis_mech <- c("MCAR", "MAR", "MNAR")
+corr <- 0.4 #c(0, .8)
+mis_mech <- "MAR" #c("MCAR", "MAR", "MNAR")
 mis_type <- "RIGHT" #c("left", "right", "mid", "tail")
-mis_prop <- c(.1, .25, .5)
-n_imp <- c(1, 5)
-imp_meth <- c("mean", "norm.predict", "norm") #cca
-n_it <- c(1, 5, 10)
-est <- c("mean(Y)", "var(Y)", "cov(X,Y)", "lm(Y~X)", "lm(X~Y)")
+mis_prop <- 0.5 #c(.1, .25, .5)
+n_imp <- 5 #c(1, 5)
+imp_meth <- "norm" #c("mean", "norm.predict", "norm") #cca
+n_it <- 10 #c(1, 5, 10)
+# est <- c("mean(Y)", "var(Y)", "cov(X,Y)", "lm(Y~X)", "lm(X~Y)")
 # model vs design based
 # with or without sampling variance
 
-# # data generating mechanism
-# means <- c(10, 100)
-# varcov <- matrix(0, 2, 2)
-# diag(varcov) <- 1
-# varcov[1, 2] <- varcov[2, 1] <- corr 
 
-# generate complete data
-generation <- function(...){
+# generate data
+generation <- function(n_obs = 500, corr = 0.4){
   dat <- data.frame(mvtnorm::rmvnorm(
     n = n_obs,
-    mean = c(10, 100),
+    mean = c(10, 10),
     sigma = matrix(c(1, corr, corr, 1), 2, 2)
   )) %>% setNames(c("Y", "X"))
 }
 
 # ampute the data
-amputation <- function(dat, ...){
+amputation <- function(dat, mis_mech = "MAR", mis_type = "RIGHT", mis_prop = 0.5){
   amp <- dat %>% 
   mice::ampute(
     prop = mis_prop,
     mech = mis_mech,
     type = mis_type
-  ) %>% 
-  .$amp}
+  )}
 
 # impute the data
-imputation <- function(amp, ...){
-  imp <- amp %>%
[…]

### 2022-01-27 R/run_sim.R
Fix simulation function

@@ -5,24 +5,24 @@ set.seed(123)
 library(dplyr)
 source("R/utils.R")
 
-# parameters
-n_sim <- 1000 #c(100, 1000)
-n_obs <- 500 #c(50, 500, 5000)
-n_var <- 2 #c(2, 5, 20)
-corr <- 0.4 #c(0, .8)
-mis_mech <- "MAR" #c("MCAR", "MAR", "MNAR")
-mis_type <- "RIGHT" #c("left", "right", "mid", "tail")
-mis_prop <- 0.5 #c(.1, .25, .5)
-n_imp <- 5 #c(1, 5)
-imp_meth <- "norm" #c("mean", "norm.predict", "norm") #cca
-n_it <- 10 #c(1, 5, 10)
-# est <- c("mean(Y)", "var(Y)", "cov(X,Y)", "lm(Y~X)", "lm(X~Y)")
-# model vs design based
-# with or without sampling variance
+# # parameters
+# n_sim <- 1000 #c(100, 1000)
+# n_obs <- 500 #c(50, 500, 5000)
+# n_var <- 2 #c(2, 5, 20)
+# corr <- 0.4 #c(0, .8)
+# mis_mech <- "MAR" #c("MCAR", "MAR", "MNAR")
+# mis_type <- "RIGHT" #c("left", "right", "mid", "tail")
+# mis_prop <- 0.5 #c(.1, .25, .5)
+# n_imp <- 5 #c(1, 5)
+# imp_meth <- "norm" #c("mean", "norm.predict", "norm") #cca
+# n_it <- 10 #c(1, 5, 10)
+# # est <- c("mean(Y)", "var(Y)", "cov(X,Y)", "lm(Y~X)", "lm(X~Y)")
+# # model vs design based
+# # with or without sampling variance
 
 
 # generate data
-generation <- function(n_obs = 500, corr = 0.4){
+generation <- function(n_obs, corr){
   dat <- data.frame(mvtnorm::rmvnorm(
     n = n_obs,
     mean = c(10, 10),
@@ -31,7 +31,7 @@ generation <- function(n_obs = 500, corr = 0.4){
 }
 
 # ampute the data
-amputation <- function(dat, mis_mech = "MAR", mis_type = "RIGHT", mis_prop = 0.5){
+amputation <- function(dat, mis_mech, mis_type, mis_prop){
   amp <- dat %>% 
   mice::ampute(
     prop = mis_prop,
[…]

### 2022-01-27 Manuscript/StandardizedEvaluation.Rmd
Add small simulation to manuscript

@@
 output: pdf_document
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+library(dplyr)
+source(here::here("R/utils.R"))
+source(here::here("R/run_sim.R"))
 ```
@@
 @molenberghs2014handbook, part 5.
+
+```{r sim_mech, echo=FALSE, cache=TRUE, paged.print=TRUE}
+# setup
+set.seed(1)
+n_sim <- 10
+correlations <- c(0, .4, .8)
+mechanisms <- c("MCAR", "MAR", "MNAR")
+
+sim_mech <- purrr::map_dfr(1:n_sim, function(.i){
+   purrr::map_dfr(correlations, function(.c){
+     purrr::map_dfr(mechanisms, function(.m){
+       run(corr = .c, mis_mech = .m) %>% cbind(sim = .i, .)})})})
+sim_mech %>% 
+  group_by(mis_mech, corr) %>% 
+  summarise(
+    mean_cca = mean(mean_cca - 10),
+    mean_est = mean(mean_est - 10),
+    beta_cca = mean(beta_cca - corr),
+    beta_est = mean(beta_est - corr),
+    beta_ciw = mean(beta_ciw),
+    beta_cr = mean(beta_cr),
+    rmse_pred = mean(rmse_pred),
+    rmse_cell = mean(rmse_cell))
+```
 
@@
 
+```{r}
+sessionInfo()
+
+```
+
 # References

### 2022-01-27 R/run_sim.R

@@ -1,9 +1,9 @@
 # small simulation study for illustration
 
 # set-up 
-set.seed(123)
-library(dplyr)
-source("R/utils.R")
+# set.seed(123)
+# library(dplyr)
+# source("R/utils.R")
 
 # # parameters
 # n_sim <- 1000 #c(100, 1000)
@@ -22,35 +22,35 @@ source("R/utils.R")
 
 
 # generate data
-generation <- function(n_obs, corr){
+generation <- function(n, r){
   dat <- data.frame(mvtnorm::rmvnorm(
-    n = n_obs,
+    n = n,
     mean = c(10, 10),
-    sigma = matrix(c(1, corr, corr, 1), 2, 2)
+    sigma = matrix(c(1, r, r, 1), 2, 2)
   )) %>% setNames(c("Y", "X"))
 }
 
 # ampute the data
-amputation <- function(dat, mis_mech, mis_type, mis_prop){
+amputation <- function(dat, mech, type, prop){
   amp <- dat %>% 
   mice::ampute(
-    prop = mis_prop,
-    mech = mis_mech,
-    type = mis_type
+    prop = prop,
+    mech = mech,
+    type = type
   )}
 
 # impute the data
-imputation <- function(amp, imp_meth, n_imp, n_it){
+imputation <- function(amp, meth, m, it){
   imp <- amp$amp %>%
   mice::mice(
-    m = n_imp,
-    method = imp_meth,
-    maxit = n_it,
+    m = m,
+    method = meth,
+    maxit = it,
     print = FALSE
   )}
 
 # evaluate the imputations
-evaluation <- function(imp, amp, dat, corr){
+evaluation <- function(imp, amp, dat, r){
   # analysis model
   fit <- imp %>% 
     with(lm(Y ~ X))
@@ -78,11 +78,11 @@ evaluation <- function(imp, amp, dat, corr){
     mutate(
       beta_est = estimate,
       beta_ciw = conf.high - conf.low,
-      beta_cr = conf.low <= corr && corr <= conf.high,
+      beta_cr = conf.low <= r && r <= conf.high,
       .keep = "none") %>% 
     cbind(
       n_obs = nrow(dat), 
-      corr = corr, 
+      corr = r, 
       mis_mech = amp$mech,
       mis_type = amp$type,
       mis_prop = amp$prop,
[…]

### 2022-02-03 Manuscript/StandardizedEvaluation.Rmd
Add CI and RMSE for all methods in sim

@@
 
+```{r sim_nobs, echo=FALSE, cache=TRUE, paged.print=TRUE}
+# setup
+set.seed(1)
+n_sim <- 2
+
+sim_nobs <- purrr::map_dfr(1:n_sim, function(.i){
+  purrr::map_dfr(c(50, 500, 5000), ~{
+    run(n_obs = .x) %>% cbind(sim = .i, .)})})
+
+gtsummary::tbl_summary(sim_nobs, by = method)
+sim_mech %>% 
+  group_by(mis_mech, corr) %>% 
+  summarise(
+    mean_cca = mean(mean_cca - 10),
+    mean_est = mean(mean_est - 10),
+    beta_cca = mean(beta_cca - corr),
+    beta_est = mean(beta_est - corr),
+    beta_ciw = mean(beta_ciw),
+    beta_cr = mean(beta_cr),
+    rmse_pred = mean(rmse_pred),
+    rmse_cell = mean(rmse_cell))
+```
+
 ## Missingness generation problems
 
-In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng1994multiple.
+In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng94.
 The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
 Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable see e.g.
-@rubin1976inference, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
+@rubi76, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
 
@@
 
-```{r sim_mech, echo=FALSE, cache=TRUE, paged.print=TRUE}
+```{r sim_mech_corr, echo=FALSE, cache=TRUE, paged.print=TRUE}
 # setup
@@
 We are interested in the correct answer to the research question; not in the truth itself.
-This means that, given the framework provided by @rubin1987, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
-After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neyman1934two, p.
+This means that, given the framework provided by @rubi87, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
+After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neym34, p.
 591.
@@
 Third, sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
-However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink2014pooling.
+However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink14.
 After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
@@
 A straightforward technique for inducing different forms of univariate MAR missingness is described in @fimd, p.
-63; see Figure \ref{fig:MAR} and a generalization to multivariate MAR missingness can be found in @van2006fully, Appendix B, and @brand1999development, par.
+63; see Figure \ref{fig:MAR} and a generalization to multivariate MAR missingness can be found in @buur06, Appendix B, and @brand1999development, par.
 5.2.3.

### 2022-02-03 R/run_sim.R

@@ -51,35 +51,50 @@ imputation <- function(amp, meth, m, it){
 
 # evaluate the imputations
 evaluation <- function(imp, amp, dat, r){
-  # analysis model
-  fit <- imp %>% 
-    with(lm(Y ~ X))
+  # fit analysis model
+  fit_imps <- with(imp, lm(Y ~ X))
+  fits <- list(full = lm(Y~X, dat),
+               cca = lm(Y~X, amp$amp, na.action = na.omit),
+               imp = mice::pool(fit_imps))
+  # performance
+  out <- purrr::map_dfr(fits, ~{
+           broom::tidy(.x, conf.int = TRUE) %>% 
+             .[2, c("estimate", "conf.low", "conf.high")]}) %>% 
+    mutate(
+      dat = c("full", "cca", "imp"),
+      est = estimate,
+      ciw = conf.high - conf.low,
+      cov = conf.low <= r && r <= conf.high,
+      .keep = "none")
   
-  # reference performance
-  refs <- data.frame(
-    mean_obs = mean(dat$Y),
-    mean_cca = mean(amp$amp$Y, na.rm = TRUE),
-    mean_est = mean(mice::complete(imp, "long")$Y),
-    beta_obs = coef(lm(Y~X, dat))[2],
-    beta_cca = coef(lm(Y~X, amp$amp, na.action = na.omit))[2])
+  # refs <- data.frame(
+  #   mean_full = mean(dat$Y),
+  #   mean_cca = mean(amp$amp$Y, na.rm = TRUE),
+  #   mean_est = mean(mice::complete(imp, "long")$Y),
+  #   beta_full = coef(lm(Y~X, dat))[2],
+  #   beta_cca = coef(lm(Y~X, amp$amp, na.action = na.omit))[2])
   
   # root mean squared errors
-  rmses <- purrr::map_dfr(1:imp$m, function(.m){
+  rmses <- data.frame(
+    rmse_pred = c(rmse(fits$full$residuals), rmse(fits$cca$residuals)),
+    rmse_cell = c(0, NA)) %>% 
+    rbind(., purrr::map_dfr(1:imp$m, function(.m) {
     data.frame(
-      rmse_pred = rmse(fit$analyses[[.m]]$residuals),    
-    rmse_cell = rmse(imp$imp$Y[[.m]] - dat$Y[imp$where[, "Y"]])) 
-  }) %>% colMeans() %>% t()
[…]

### 2022-02-03 Manuscript/StandardizedEvaluation.Rmd
Add kable table

@@
 
-```{r sim_nobs, echo=FALSE, cache=TRUE, paged.print=TRUE}
+```{r sim_nobs, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, paged.print=TRUE}
 # setup
 set.seed(1)
-n_sim <- 2
+n_sim <- 100
 
@@
 
-gtsummary::tbl_summary(sim_nobs, by = method)
-sim_mech %>% 
-  group_by(mis_mech, corr) %>% 
+# gtsummary::tbl_summary(sim_nobs, by = dat)
+# https://haozhu233.github.io/kableExtra/awesome_table_in_html.html#Group_rows_via_labeling
+sim_nobs %>% 
+  group_by(n_obs, dat) %>% 
   summarise(
-    mean_cca = mean(mean_cca - 10),
-    mean_est = mean(mean_est - 10),
-    beta_cca = mean(beta_cca - corr),
-    beta_est = mean(beta_est - corr),
-    beta_ciw = mean(beta_ciw),
-    beta_cr = mean(beta_cr),
-    rmse_pred = mean(rmse_pred),
-    rmse_cell = mean(rmse_cell))
+    mean_Y = mean(ave),
+    beta_X = mean(est),
+    CIW = mean(ciw),
+    CR = mean(cov),
+    RMSE_pred = mean(rmse_pred),
+    RMSE_cell = mean(rmse_cell, na.rm = TRUE))  %>% 
+  .[, -1] %>% 
+  knitr::kable(
+    col.names = c("", "$\\mu_{Y} = 10$", "$\\beta_{X} = 0.4$", "CIW", "CR", "RMSE$_{p}$", "RMSE$_{c}$"), 
+    digits = 3,
+    escape = FALSE) %>% 
+  kableExtra::kable_classic(full_width = TRUE) %>% 
+  kableExtra::pack_rows(group_label = "$n = 50$", 1, 3) %>% 
+  kableExtra::pack_rows(group_label = "$n = 500$", 4, 6) %>% 
+  kableExtra::pack_rows(group_label = "$n = 5000$", 7, 9)
+
 ```
@@
 ```{r sim_mech_corr, echo=FALSE, cache=TRUE, paged.print=TRUE}
-# setup
-set.seed(1)
-n_sim <- 10
-correlations <- c(0, .4, .8)
-mechanisms <- c("MCAR", "MAR", "MNAR")
-
-sim_mech <- purrr::map_dfr(1:n_sim, function(.i){
-   purrr::map_dfr(correlations, function(.c){
-     purrr::map_dfr(mechanisms, function(.m){
-       run(corr = .c, mis_mech = .m) %>% cbind(sim = .i, .)})})})
-sim_mech %>% 
-  group_by(mis_mech, corr) %>% 
-  summarise(
-    mean_cca = mean(mean_cca - 10),
-    mean_est = mean(mean_est - 10),
-    beta_cca = mean(beta_cca - corr),
-    beta_est = mean(beta_est - corr),
-    beta_ciw = mean(beta_ciw),
-    beta_cr = mean(beta_cr),
-    rmse_pred = mean(rmse_pred),
-    rmse_cell = mean(rmse_cell))
+# # setup
+# set.seed(1)
+# n_sim <- 10
+# correlations <- c(0, .4, .8)
+# mechanisms <- c("MCAR", "MAR", "MNAR") 
+# 
+# sim_mech <- purrr::map_dfr(1:n_sim, function(.i){
+#    purrr::map_dfr(correlations, function(.c){
+#      purrr::map_dfr(mechanisms, function(.m){
+#        run(corr = .c, mis_mech = .m) %>% cbind(sim = .i, .)})})})
+# sim_mech %>% 
+#   group_by(mis_mech, corr) %>% 
+#   summarise(
+#     mean_cca = mean(mean_cca - 10),
+#     mean_est = mean(mean_est - 10),
+#     beta_cca = mean(beta_cca - corr),
+#     beta_est = mean(beta_est - corr),
+#     beta_ciw = mean(beta_ciw),
+#     beta_cr = mean(beta_cr),
+#     rmse_pred = mean(rmse_pred),
+#     rmse_cell = mean(rmse_cell))
 ```

### 2022-02-03 Manuscript/StandardizedEvaluation.tex
Add kable table

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Towards a standardized evaluation of multiple imputation routines},
  pdfauthor={Hanne Oberman and Gerko Vink},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Towards a standardized evaluation of multiple imputation
routines}
\author{Hanne Oberman and Gerko Vink}
\date{19-1-2022}

\begin{document}
\maketitle
\begin{abstract}
Developing new imputation methodology has become a very active field.
Unfortunately, there is no consensus on how to perform simulation
studies to evaluate the properties of imputation methods. In this paper
I propose a move towards a standardized evaluation of imputation
methods. To demonstrate the need for standardization, I highlight a set
of potential pitfalls that bring forth a chain of potential problems in
the objective assessment of the performance of imputation routines. This
may lead to suboptimal use of multiple imputation in practice.
Additionally, I suggest a course of action for simulating and evaluating
missing data problems.
\end{abstract}

\hypertarget{to-do}{%
\section{TO DO}\label{to-do}}

\begin{itemize}
\tightlist
\item
  add ADEMP paper (Morris, White, and Crowther 2019), diagnostic
  checking paper (Zhao 2022)
\item
  add PPC (refer to Mingyang, and Gelman \& King)
\item
  expand convergence section
\item
  add RMSE
\item
  add insights from literature review for chapter on missing data in EHR
\item
  add link to simulation script, and \(n\)
\item
  add simulation results
\end{itemize}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Multiple imputation Donald B. Rubin (1987) is a state-of-the-art
technique for drawing valid conclusions from incomplete data, and is
increasingly used in many substantive fields, such as medicine (see e.g.
Pulkki-Råback et al. 2015; Alkire et al. 2015; Stinesen Kollberg et al.
2015), psychology (e.g. Sorbi et al. 2014), economics (e.g. Wong et al.
2015; Dafoe, Oneal, and Russett 2013; Allan and Forder 2015), sociology
(e.g. Crowley and Kolenikov 2014; De Hoon and Van Tubergen 2014), and
official statistics (e.g. Roderick J. Little 2012; Klausch, Hox, and
Schouten 2015).

The technique has earned a permanent spot in research and policy making,
demonstrated e.g.~by the detailed manual created by the National
Research Council (Roderick J. Little et al. 2012). Although a top-down
enforcement of valid ways to handle missing data is not yet very
pronounced, an increasing amount of researchers are embracing multiple
imputation techniques. After all, the principle of multiple imputation
is very intuitive.

The idea behind multiple imputation is to impute (fill in) the missing
values multiple times to obtain a valid estimate of what could have
been. The multiple datasets that are thus obtained can be analyzed by
standard techniques and the analysis results can be combined into a
single inference. In contrast to ad hoc methods for dealing with missing
values (e.g.~list-wise deletion, mean imputation, regression imputation,
last observation carried forward), multiple imputation properly takes
the sources of uncertainty that are related to the missingness problem
into account.

The quality of a solution obtained by multiple imputation depends on the
statistical properties of the incomplete data and the degree to which an
imputation procedure is able to capture these properties when modeling
missing values. In general it holds that modeling missing data becomes
more challenging when the amount of missingness increases. However, when
(strong) relations in the data are present, the observed parts can hold
great predictive power for the models that estimate the missingness. In
that case, multiple imputation would be substantially more efficient
than the ubiquitous list-wise deletion.

When evaluating the statistical properties (and thereby the practical
applicability) of new imputation methodology, researchers most often
make use of simulation studies. In such studies a complete data set is
usually generated from a statistical model, another model is used to
induce missingness, and a set of evaluation criteria is postulated to
evaluate the performance of the imputation method. However, no golden
standard has been established and, as a result, validity of the
simulation and evaluation procedures may differ tremendously from one
developer to another.

The purpose of this paper is threefold: First, I raise some concerns
with respect to evaluating imputation methodology. These concerns stem
from careful consideration with fellow `imputers' and from encounters as
a reviewer for statistical journals. Second, I provide imputation
methodologists with a suggested course of action when simulating missing
data problems. This suggested approach should identify a common ground,
but is in no way intended as an absolute solution. This identifies the
third purpose of this paper: discussion. I hope to elicit critical
thinking with respect to the problems at hand. We are all convinced that
our methodology has some merit. But for sake of progress it would be
much more advantageous if the aim of our evaluations would go beyond
\emph{proving the point} and would legitimately consider the statistical
properties.

\hypertarget{why-some-evaluations-should-not-be-trusted}{%
\subsection{Why some evaluations should not be
trusted}\label{why-some-evaluations-should-not-be-trusted}}

As of today, there is no consensus on how to perform simulation studies
to evaluate the properties of imputation methods. Typically, the
developer of a new imputation routine does some tests by simulations,
but these tests differ across developers. This brings forth a chain of
potential problems in the objective assessment of the performance of
imputation routines that may lead to suboptimal use of multiple
imputation in practice. To demonstrate the broad impact of these
problems, I subdivide the problems in the following three distinct
categories: problems with data generation, problems with missingness
generation and problems with performance evaluation. I further detail
the impact these problems may have on the validity of the performance
evaluation of the imputation routines.

\hypertarget{data-generation-problems}{%
\subsection{Data generation problems}\label{data-generation-problems}}

To evaluate the ability of an imputation routine to handle the
non-response model, a form of truth has to be established. Those who
perform simulation studies are in the luxury position to establish the
truth beforehand, either by generating data from a statistical model, or
by sampling data from a sufficiently large observed set. However, when
generating data, two easily overlooked problems may arise.

First, data are often generated following the model (or distribution)
that is also used for imputing the data. The performance of the
imputation approach is then deemed good, which is no surprise as the
evaluated conditions are in favor of the problem that is studied.
Although such procedures may be statistically relevant, the approach
would be no good for even the simplest case. Such simulations pose
serious threats to the applicability of the method as real-life data
hardly ever follow a given theoretical distribution.

Second, data are often generated such that the problem that is being
studied is most pronounced. This results in simulated data that contains
very valuable information structures, i.e.~the correlations between
groups of variables may be very pronounced. In other words, no matter
what type of missingness is induced, the observed parts of the data may
still hold much, if not all, of the information about the missing part
and the performance of the imputation procedure is not surprisingly
evaluated as very good. Also, it may be that the relations in real-life
data are rarely as pronounced as the problem that is being studied.

\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\hline
 & $\mu_{Y} = 10$ & $\beta_{X} = 0.4$ & CIW & CR & RMSE$_{p}$ & RMSE$_{c}$\\
\hline
\multicolumn{7}{l}{\textbf{\$n = 50\$}}\\
\hline
\hspace{1em}CCA & 9.961 & 0.341 & 0.823 & 0.94 & 0.866 & NaN\\
\hline
\hspace{1em}Imp. & 10.007 & 0.386 & 0.877 & 0.99 & 0.903 & 1.330\\
\hline
\hspace{1em}True & 10.021 & 0.393 & 0.541 & 0.97 & 0.903 & 0.000\\
\hline
\multicolumn{7}{l}{\textbf{\$n = 500\$}}\\
\hline
\hspace{1em}CCA & 9.941 & 0.343 & 0.232 & 0.85 & 0.897 & NaN\\
\hline
\hspace{1em}Imp. & 9.994 & 0.392 & 0.249 & 0.94 & 0.917 & 1.296\\
\hline
\hspace{1em}True & 9.998 & 0.400 & 0.161 & 0.90 & 0.914 & 0.000\\
\hline
\multicolumn{7}{l}{\textbf{\$n = 5000\$}}\\
\hline
\hspace{1em}CCA & 9.945 & 0.344 & 0.074 & 0.09 & 0.899 & NaN\\
\hline
\hspace{1em}Imp. & 10.000 & 0.397 & 0.078 & 0.91 & 0.916 & 1.298\\
\hline
\hspace{1em}True & 10.002 & 0.398 & 0.051 & 0.95 & 0.917 & 0.000\\
\hline
\end{tabu}

\hypertarget{missingness-generation-problems}{%
\subsection{Missingness generation
problems}\label{missingness-generation-problems}}

In order to obtain valid imputation inference, the imputation model must
capture the essence of the true non-response mechanism Meng (1994). The
model - if any - that is used to generate the missingness is usually
assumed to be random (MAR) or completely random (MCAR). Missingness
mechanisms that are not random (MNAR) and mechanisms that are considered
non-ignorable see e.g. Donald B. Rubin (1976), are generally ignored
during evaluation of imputation methodologies, except for those methods
that are specifically targeted at non-ignorable applications.

With MCAR missingness mechanisms, the probability to be missing is the
same for all cases. This is a necessary simulation condition for
evaluating the performance of imputation procedures. If an imputation
method is not able to solve the problem (i.e.~yield valid inference)
under MCAR, the statistical properties of the procedure are not sound.
Sadly, the straightforward case of MCAR is often neglected from
simulation studies and focus is drawn to the evaluation of MAR
mechanisms only. Alternatively, some studies limit their evaluations to
MCAR mechanisms only, which in my view may be far too simplistic.

\begin{figure}
\centering
\includegraphics{img/plot_mmech.pdf}
\caption{\label{fig:mis_mech}Three missingness mechanisms that yield
approximately 50 percent missingness (\(N=500\)). Displayed are a MCAR
mechanism, a right-tailed MAR mechanism generated from bivariate normal
data with correlation \(\rho=.8\) and the same MAR mechanism, now
generated from bivariate normal data with correlation \(\rho= 0\).}
\end{figure}

With MAR missingness mechanisms, the probability to be missing is the
same within groups of cases defined by the observed data only
(e.g.~males are less likely to disclose their weight, but gender is
observed). The essence here is that the observed relations in the data
are used to induce missingness during simulation (e.g.~weight is made
incomplete based on gender), but these relations may be weak, or
non-existent.

When MAR is induced based on weaker relations in the data, claims for a
method's applicability to situations where the missingness is random
become less valid. When MAR is induced from data without multivariate
relations, the inferential implication of the missingness would mimic
those of MCAR. This problem is generally overlooked in simulation
studies and the procedure to generate missing data is often
insufficiently described in resulting publications. Figure
\ref{fig:mis_mech} demonstrates two realizations of missingness
mechanisms under MAR mechanisms derived from data with different
multivariate relations.

It is obvious that the MAR missingness generated from data with low
correlation would yield similar inference to the MCAR missingness, even
though the missingness is clearly random. The simulation results from
Table 1 confirm this. A detailed simulation setup can be found in
Appendix A.

As expected, under genuine MAR missingness, i.e.~the missingness
probabilities differ conditional on the level of the incomplete
variable, the performance of complete case analyses drops dramatically.
For \emph{spurious} MAR, i.e.~the conditional probabilities remain the
same over the level of the incomplete variable, the performance is
comparable to that under MCAR. In fact, this would amount to one of the
special cases under which complete case analysis would be more efficient
than multiple imputation Van Buuren (2012), p.~48: the missingness does
not depend on the incomplete variable. This property might be useful in
practice, but considering it as a condition to evaluate performance
under MAR missingness is useless.

Although one cannot definitively verify if the missingness is random -
after all, for every MNAR mechanism there is a MAR mechanism with equal
fit Molenberghs et al. (2008) - it can be argued that MNAR is the more
likely mechanism for real life missingness scenarios. In such cases, an
indication of the validity of the obtained inference, given that the
assumed missingness mechanism is suspected to be invalid, may be
obtained by performing sensitivity analysis see e.g. Molenberghs et al.
(2014), part 5.

The amount of missingness that is used to simulate the performance of
imputation methods differs between studies. Some studies use only 10\%
missing data where other studies push the limits to additionally
investigate the performance under at least 50\% missing data. This
inconsistent display of simulation results may impact the objectivity of
meta-evaluations over imputation methods, as one method's performance
may appear to be favorable because of the less stringent simulation
conditions. This ultimately may lead to statisticians recommending a
less efficient method to applied researchers, thereby limiting the
efficiency of the imputation approach and unnecessarily lowering the
statistical power.

\hypertarget{performance-evaluation}{%
\subsection{Performance evaluation}\label{performance-evaluation}}

The evaluation criteria used to assess imputation performance vary from
one developer to another. This is not surprising as people from
different fields could have a different focus on the problem at hand.
However, because multiple imputation was designed as a mode for
statistical inference, a minimal set of statistical properties should at
least be evaluated. I highlight four aspects of evaluation that should
be considered when evaluation imputation routines.

First, developers often only inquire about the `accuracy' (i.e.~how well
can the method reproduce the original data). The goal of multiple
imputation is not to reproduce the data, but to allow for obtaining
valid inference given that the data are incomplete. We are interested in
the correct answer to the research question; not in the truth itself.
This means that, given the framework provided by Donald B. Rubin (1987),
statistical properties such as bias, confidence intervals and the
coverage rate of the confidence intervals should be studied. After all,
the 95\% confidence interval should contain the `true' value at least 95
out of 100 times Neyman (1934), p.~591.

Second, the performance of imputation procedures on distributional
properties is often ignored in simulation studies, and even though the
estimates on the analysis level may be justified, some methods can yield
imputations that may seem completely invalid to applied researchers. For
example, one could very accurately estimate average human height by
filling in negative values and values that are unrealistically large.
While the obtained inference could still be valid under such
imputations, the plausibility of the imputed values given the observed
data should be under scrutiny. Rather, one would prefer an imputation
technique to yield both valid inference and plausible imputations. It
should be studied if an imputation method is prone to deliver such
impractical results, and if so, under what conditions.

Third, sampling variation has always been an essential part of the
evaluation of multiple imputation methodology. However, in order to
obtain information about a method's ability to handle the missing data
problem, or to objectively compare methods on their ability to correct
for missingness, it is not necessary to take sampling variation into
account Vink and van Buuren (2014). After all, we are interested only in
the missing data mechanism, and are not considering the noise induced by
the sampling mechanism for evaluation in such studies.

Last, when the evaluator is on the verge of drawing conclusions about
the performance of the imputation routine, the performance should be
carefully qualified. Comparing the performance of an imputation routine
given a population (or true) parameter allows for quantitative
evaluation. However, in order to pose qualitative statements about the
performance on simulated conditions, comparative methodology is
required. For example, when claiming that imputation performance is
unacceptable when deviations from normality become rather stringent,
such performance is highly dependent on the simulation conditions that
are used. For a well-balanced judgement about the severity of the
performance drop, comparative simulations with e.g.~nonparametric models
should be executed. A method may perform badly, but if it still
outperforms every other approach, it may yet be of great practical
relevance.

\hypertarget{suggested-course-of-action}{%
\section{Suggested course of action}\label{suggested-course-of-action}}

\hypertarget{step-1.-obtain-truth}{%
\subsection{Step 1. Obtain truth}\label{step-1.-obtain-truth}}

The evaluator must to decide whether there is a need for sampling
variance in the simulations scheme. If sampling variance cannot be
omitted from the simulation scheme, two approaches are possible:

\begin{itemize}
\tightlist
\item
  Model-based simulation: draw samples from a known probability
  distribution, such as the multivariate normal distribution. The
  theoretical parameters under which the samples are obtained will serve
  as the comparative truth during the simulations.
\item
  Design-based simulation: sample data from a sufficiently large set.
  The parameters of the sufficiently large set will serve as the
  comparative truth during the simulations. This approach is often used
  in situations where a probability distribution is not available, or
  where real-life data structures are of interest. Applications of
  design-based simulation are often found in official statistics. A
  benefit of design-based simulation is the ability to use real-life
  observed data structures.
\end{itemize}

For situations where sampling variance is not of interest, the sampling
process can be omitted and a single complete dataset can simply be
obtained by model- or design-based approaches. The parameters of that
single complete set will serve as comparative truth. This process is
computationally convenient, because only a single complete datasets has
to be considered during all of the simulations. One needs to realize,
however, that the conventional pooling rules cf. Donald B. Rubin (1987),
pp.~76-77, do not apply for finite population inference and that
alternative pooling rules need to be used Raghunathan, Reiter, and Rubin
(2003), Vink and van Buuren (2014).

\hypertarget{step-2.-induce-missingness}{%
\subsection{Step 2. Induce
missingness}\label{step-2.-induce-missingness}}

First, one should always consider MCAR missingness, i.e.~the scenario
where the missing values are missing at random and the observed values
are observed at random. Under MCAR, the statistical properties of the
observed data given the missing data are known and any imputation
routine that cannot at least mimic the performance of the observed data
inference, should be deemed inefficient in the scope of the simulation.

Next, missing data should be induced conform a model that is dependent
on the observed data. A straightforward technique for inducing different
forms of univariate MAR missingness is described in Van Buuren (2012),
p.~63; see Figure \ref{fig:MAR} and a generalization to multivariate MAR
missingness can be found in Buuren et al. (2006), Appendix B, and Brand
(1999), par. 5.2.3. If the missingness is to be induced in longitudinal
data, auto-regressive MAR models e.g.~cf. Shara et al. (2015), model 2
and model 3, can be useful.

\begin{figure}
\centering
\includegraphics{img/plot_mar.pdf}
\caption{\label{fig:MAR}Examples of four random missingness mechanisms
that yield approximately 50 percent missingness. For negatively
correlated missingness covariates the resulting mechanisms are
reversed.}
\end{figure}

Third, it is advisable to investigate varying shapes of MAR missingness
to achieve a more realistic indication of the robustness of the
imputation performance across the range of random missingness. Given the
simulated data-distributions, one random missingness model may be far
more disastrous to the observed information than another model. This may
influence the performance of some (but not necessarily all) imputation
routines For example, inference from hot-deck techniques such as
predictive mean matching Roderick JA Little (1988), Donald B. Rubin
(1986) may be more severely impacted by large amounts of one-tailed
missingness than inference from parametric techniques. It would be a
shame to overlook such results due to the focus on a single MAR
mechanism.

Fourth, the amount of missingness must be varied. Remember that
missingness is only ignorable under MAR when the parameter of the data
is distinct and a-priori independent from the parameter of the missing
data process. Under MAR missingness we assume that we may use the
observed data to make inferences about the joint (observed and
unobserved) data.

The dependency of the procedure on the assumption under which we obtain
inference is only influenced by the amount of missingness. If there is
no missingness - or if there is no data, for that matter - the inference
does not depend on the assumption. Alternatively, the validity of
assumptions become increasingly important when the missingness
increases. Since we control the MAR mechanism, the assumption under
which we may solve the missing data problem should hold and it is only
fair to assess performance under stringent missingness conditions. I
therefor propose, for all mechanisms, to evaluate at least the following
univariate missingness percentages when evaluating imputation routines:

\begin{itemize}
\tightlist
\item
  10\% missingness: Depending on the size of the data, this percentage
  can be considered as a lower bound of realistic evaluation. Anything
  less than 10\% may be of little influence on the true data inference.
  Performance of a missing data method should at least be acceptable for
  most missing data problems.
\item
  25\% missingness: This is a fair amount of missingness and will,
  depending on the observed data information, have a noticeable
  influence on the completed data inference. When compared to the
  condition with 10\% missingness, the inference obtained under 25\%
  missingness should be less certain (i.e.~confidence/credibility
  interval width should increase), but estimates should still be
  properly covered and the statistical properties of the missing data
  method should be sound. In practice, at least to my experience in
  social sciences and official statistics, 25\% univariate missingness
  can easily be considered as a realistic missingness percentage.
\item
  50\% missingness: Performance under 50\% percent simulated missingness
  will most likely be impacted severely. Depending on how the
  missingness mechanism interacts with the simulated data, some
  imputation techniques may yield estimates that are under-covered such
  that the completed data inference should not be deemed valid anymore.
  If a method yields acceptable inference under 50\% MAR missingness, we
  can determine that the statistical properties of the imputation
  methodology are sound.
\end{itemize}

Although I limit the focus here to ignorable non-response, the above
suggested proportions are equally applicable to simulations under
non-ignorable non-response.

Fifth, the used missingness mechanism should be detailed, either
graphically or written as a function of the data. Often, when inducing
missingness, authors remain vague about the actual missingness mechanism
under investigation and, even worse, some authors only report something
like

\begin{quote}
\emph{We generated missing data following a MAR missingness mechanism.}
\end{quote}

This should be considered unacceptable as claims about the validity of
the multiple imputation inference depends heavily on the simulated
missingness mechanism.

\hypertarget{step-3.-evaluate-performance}{%
\subsection{Step 3. Evaluate
performance}\label{step-3.-evaluate-performance}}

It is wise to evaluate the performance of complete case analysis (aka
list-wise deletion) in all simulated conditions. We know the theoretical
properties of complete case analysis, which makes the technique useful
as a point of departure when evaluating multiple imputation performance.
The evaluation criteria may depend on the estimand. When descriptive
statistics are the goal and when statistical inference would not be of
interest, bias of the estimates would still apply, but standard errors
are generally ignored. Chances are that for those who focus on
descriptive statistical applications, multiple imputation would not be
the mode of choice. In general, I would say that each multiple
imputation routine should be evaluated on at least the following points:

\begin{itemize}
\tightlist
\item
  \emph{Bias:} Results should preferable be unbiased. However, the way
  bias is considered can greatly influence the interpretation of the
  results. For example, a negligible absolute bias for a parameter for
  which the true value is zero, would yield infinite bias when relative
  bias is considered. The way bias is computed should therefore be
  carefully chosen and described.
\item
  \emph{Interval width:} The way the confidence interval is calculated
  should be described. Wider intervals are associated with more
  uncertainty and the more narrow interval that is still properly
  covered indicates a sharper inference. However, inference from a wider
  interval that is properly covered is to be considered more valid when
  compared to a more narrow interval that is not properly covered
  anymore.
\item
  \emph{Coverage:} Coverage of a 95\% interval should in theory be
  \(\geq 95\), where a coverage of 95\% would be most efficient.
  Under-coverage (when estimation is too liberal) may be an indication
  of invalid inference, while over-coverage (when estimation is too
  conservative) tells us that efficiency could still be gained.
\item
  \emph{Distributional characteristics:} In practice, the distribution
  of the incomplete data may differ greatly from the observed data.
  Under anything but the MCAR assumption, this can be expected. When
  evaluating imputations, the distributional shapes should be checked
  and diagnostic evaluations should be performed see Abayomi, Gelman,
  and Levy (2008) for an detailed overview of diagnostic evaluation for
  multivariate imputations. When anomalies are found, and if the
  imputation method is valid, there should be an explanation, especially
  in the controlled environment of a properly executed simulation study.
\item
  \emph{Plausibility of the imputed values:} Plausible imputations -
  imputations that could be real values if they had been observed - are
  not a necessary condition for obtaining valid inference. However, in
  practice, especially when the imputer and the analyst are different
  persons, plausibility of imputations may be a desired property. When
  evaluating imputation routines, the evaluator should mention whether
  the routine is prone to deliver implausible values.
\item
  \emph{Convergence of the algorithm:} Most contemporary imputation
  techniques rely on iterative algorithms, such as the Gibbs sampler,
  where some algorithms are critically considered to be possibly
  incompatible Gibbs samplers, PIGS, Li, Yu, and Rubin (2012). The
  convergence of all iterative algorithms should always be considered
  and if non-convergence is suspected, the inference resulting from the
  imputations should not be considered.
\end{itemize}

For evaluations of model-based simulations, it could be convincing to
demonstrate a method's applicability to real-world missing data
problems. This can, for example, be achieved by obtaining and imputing a
fully observed set of variables, wherein the missingness is mimicked
from a similar, incomplete set. Alternatively, an incomplete set could
be obtained and truth could be established by removing the incomplete
cases from the data. However, the real-world missingness would then be
omitted.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

This document is aimed at establishing a common ground for the
evaluation of imputation routines. Such a common ground would be the
basis of a standardized evaluation. This allows for fair and efficient
comparisons between imputation techniques. Ultimately, it would be
desirable to evaluate every imputation routine against the same
standardized set in order to quantify the statistical properties across
imputation routines. If properly executed, this would allow for careful
matching of imputation methodologies to new missing data problems.

\hypertarget{appendix-a-simulation-setup}{%
\section{Appendix A: Simulation
setup}\label{appendix-a-simulation-setup}}

Let \(Y=(Y_{obs},Y_{mis})\) be an incomplete semi-continuous variable,
where \(Y_{obs}\) and \(Y_{mis}\) denote the observed values and the
missing values in \(Y\), respectively. Further, \(X=(X_1,...,X_k)\) is a
set of \(k\) fully observed covariates, where \(X_{obs}\) and
\(X_{mis}\) correspond to the observed an missing parts in \(Y\) and
where \(X_j\) would indicate the \(j\)th variable in \(X\), with
\(j=1,\dots, k\). Finally, let \(R\) be a response indicator that is 1
if \(Y\) is observed and 0 if \(Y\) is missing.

To prove the point, I generated a single data set from the multivariate
normal distribution with means \[
\mu= \bordermatrix{&     \cr
 Y  &5  \cr
{ X}_1  &10 \cr
{ X}_2  &10 },
\] and covariance matrix \[
\Sigma = \bordermatrix{& Y &{ X}_1 &{ X}_2   \cr
 Y  &1  &0.8    &0\cr
{ X}_1  &0.8    &1  &0\cr
{ X}_2  &0  &0  &1}.
\]

Three incomplete sets were then generated, wherein random missingness
was imposed in \(Y\) in three ways. First, I left the observed data
\emph{observed at random} to impose MCAR missingness following the
mechanism \[
P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0)=.50.
\] Next, I created random missingness in \(Y\) according to the
following MAR missingness mechanism \[
P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0|Y_{obs}, X_j),
\] by using a random draw from a binomial distribution of the same
length as \(Y\) and of size 1 with missingness probability equal to the
inverse logit \[
P(R=0)=\frac{e^{a}}{(1+e^{a})}.
\] where \(a=(\mu_{X_j}-X_{ij})/\sqrt{\mathbb{E}[X_{ij} - \mu_{X_j}]}\)
in order to create approximately 50\% right-tailed MAR missingness. I
followed this procedure twice; once for each \(X_j\).

The three resulting incomplete data sets were then imputed with package
\texttt{mice} Van Buuren and Groothuis-Oudshoorn (2011), v2.25, in
\texttt{R} Team (2015), v3.2.3, with Bayesian linear regression
imputation (\texttt{mice.impute.norm}) as the imputation routine. The
completed-data analyses were combined into a single inference following
the rules described in Vink and van Buuren (2014).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{sessionInfo}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## R version 4.1.1 (2021-08-10)
## Platform: x86_64-w64-mingw32/x64 (64-bit)
## Running under: Windows 10 x64 (build 19042)
## 
## Matrix products: default
## 
## locale:
## [1] LC_COLLATE=Dutch_Netherlands.1252  LC_CTYPE=Dutch_Netherlands.1252   
## [3] LC_MONETARY=Dutch_Netherlands.1252 LC_NUMERIC=C                      
## [5] LC_TIME=Dutch_Netherlands.1252    
## 
## attached base packages:
## [1] stats     graphics  grDevices utils     datasets  methods   base     
## 
## other attached packages:
## [1] dplyr_1.0.7
## 
## loaded via a namespace (and not attached):
##  [1] Rcpp_1.0.7        highr_0.9         pillar_1.6.4      compiler_4.1.1   
##  [5] tools_4.1.1       digest_0.6.29     viridisLite_0.4.0 evaluate_0.14    
##  [9] lifecycle_1.0.1   tibble_3.1.6      lattice_0.20-44   pkgconfig_2.0.3  
## [13] rlang_0.4.12      rstudioapi_0.13   DBI_1.1.2         yaml_2.2.1       
## [17] mvtnorm_1.1-3     xfun_0.29         fastmap_1.1.0     kableExtra_1.3.4 
## [21] xml2_1.3.3        httr_1.4.2        withr_2.4.3       stringr_1.4.0    
## [25] knitr_1.37        systemfonts_1.0.3 generics_0.1.1    vctrs_0.3.8      
## [29] webshot_0.5.2     rprojroot_2.0.2   grid_4.1.1        tidyselect_1.1.1 
## [33] svglite_2.0.0     glue_1.6.0        mice_3.14.0       here_1.0.1       
## [37] R6_2.5.1          fansi_0.5.0       rmarkdown_2.11    purrr_0.3.4      
## [41] tidyr_1.1.4       magrittr_2.0.1    codetools_0.2-18  scales_1.1.1     
## [45] backports_1.4.1   ellipsis_0.3.2    htmltools_0.5.2   rvest_1.0.2      
## [49] assertthat_0.2.1  colorspace_2.0-2  utf8_1.2.2        stringi_1.7.6    
## [53] munsell_0.5.0     broom_0.7.12      crayon_1.4.2
\end{verbatim}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-abayomi2008diagnostics}{}}%
Abayomi, Kobi, Andrew Gelman, and Marc Levy. 2008. {``Diagnostics for
Multivariate Imputations.''} \emph{Journal of the Royal Statistical
Society: Series C (Applied Statistics)} 57 (3): 273--91.

\leavevmode\vadjust pre{\hypertarget{ref-alkire2015global}{}}%
Alkire, Blake C, Nakul P Raykar, Mark G Shrime, Thomas G Weiser, Stephen
W Bickler, John A Rose, Cameron T Nutt, et al. 2015. {``Global Access to
Surgical Care: A Modelling Study.''} \emph{The Lancet Global Health} 3
(6): e316--23.

\leavevmode\vadjust pre{\hypertarget{ref-allan2015determinants}{}}%
Allan, Stephen, and Julien Forder. 2015. {``The Determinants of Care
Home Closure.''} \emph{Health Economics} 24 (S1): 132--45.

\leavevmode\vadjust pre{\hypertarget{ref-brand1999development}{}}%
Brand, Jaap. 1999. {``Development, Implementation and Evaluation of
Multiple Imputation Strategies for the Statistical Analysis of
Incomplete Data Sets.''} PhD thesis, Erasmus University Rotterdam.

\leavevmode\vadjust pre{\hypertarget{ref-buur06}{}}%
Buuren, S. Van, J. P. L. Brand, C. G. M. Groothuis-Oudshoorn, and D. B.
Rubin. 2006. {``Fully Conditional Specification in Multivariate
Imputation.''} \emph{Journal of Statistical Computation and Simulation}
76 (12): 1049--64. \url{https://doi.org/10.1080/10629360600810434}.

\leavevmode\vadjust pre{\hypertarget{ref-crowley2014flexible}{}}%
Crowley, Jocelyn Elise, and Stanislav Kolenikov. 2014. {``Flexible Work
Options and Mothers' Perceptions of Career Harm.''} \emph{The
Sociological Quarterly} 55 (1): 168--95.

\leavevmode\vadjust pre{\hypertarget{ref-dafoe2013democratic}{}}%
Dafoe, Allan, John R Oneal, and Bruce Russett. 2013. {``The Democratic
Peace: {Weighing} the Evidence and Cautious Inference.''}
\emph{International Studies Quarterly} 57 (1): 201--14.

\leavevmode\vadjust pre{\hypertarget{ref-de2014religiosity}{}}%
De Hoon, Sean, and Frank Van Tubergen. 2014. {``The Religiosity of
Children of Immigrants and Natives in England, Germany, and the
Netherlands: {The} Role of Parents and Peers in Class.''} \emph{European
Sociological Review}, jcu038.

\leavevmode\vadjust pre{\hypertarget{ref-klausch2015selection}{}}%
Klausch, Thomas, Joop Hox, and Barry Schouten. 2015. {``Selection Error
in Single-and Mixed Mode Surveys of the {Dutch} General Population.''}
\emph{Journal of the Royal Statistical Society: Series A (Statistics in
Society)} 178 (4): 945--61.

\leavevmode\vadjust pre{\hypertarget{ref-li2012imputing}{}}%
Li, Fan, Yaming Yu, and Donald B Rubin. 2012. {``Imputing Missing Data
by Fully Conditional Models: {Some} Cautionary Examples and
Guidelines.''} \emph{Duke University Department of Statistical Science
Discussion Paper} 1124.

\leavevmode\vadjust pre{\hypertarget{ref-little2012calibrated}{}}%
Little, Roderick J. 2012. {``Calibrated {Bayes}, an Alternative
Inferential Paradigm for Official Statistics.''} \emph{Journal of
Official Statistics} 28 (3): 309.

\leavevmode\vadjust pre{\hypertarget{ref-little1988missing}{}}%
Little, Roderick JA. 1988. {``Missing-Data Adjustments in Large
Surveys.''} \emph{Journal of Business \& Economic Statistics} 6 (3):
287--96.

\leavevmode\vadjust pre{\hypertarget{ref-little2012prevention}{}}%
Little, Roderick J, Ralph D'Agostino, Michael L Cohen, Kay Dickersin,
Scott S Emerson, John T Farrar, Constantine Frangakis, et al. 2012.
{``The Prevention and Treatment of Missing Data in Clinical Trials.''}
\emph{New England Journal of Medicine} 367 (14): 1355--60.

\leavevmode\vadjust pre{\hypertarget{ref-meng94}{}}%
Meng, Xiao-Li. 1994. {``Multiple-{Imputation Inferences} with
{Uncongenial Sources} of {Input}.''} \emph{Statistical Science} 9 (4):
538--58. \url{https://doi.org/10.1214/ss/1177010269}.

\leavevmode\vadjust pre{\hypertarget{ref-molenberghs2008every}{}}%
Molenberghs, Geert, Caroline Beunckens, Cristina Sotto, and Michael G
Kenward. 2008. {``Every Missingness Not at Random Model Has a
Missingness at Random Counterpart with Equal Fit.''} \emph{Journal of
the Royal Statistical Society: Series B (Statistical Methodology)} 70
(2): 371--88.

\leavevmode\vadjust pre{\hypertarget{ref-molenberghs2014handbook}{}}%
Molenberghs, Geert, Garrett Fitzmaurice, Michael G Kenward, Anastasios
Tsiatis, and Geert Verbeke. 2014. \emph{Handbook of Missing Data
Methodology}. {CRC Press}.

\leavevmode\vadjust pre{\hypertarget{ref-morr18}{}}%
Morris, Tim P., Ian R. White, and Michael J. Crowther. 2019. {``Using
Simulation Studies to Evaluate Statistical Methods.''} \emph{Statistics
in Medicine} 38 (11): 2074--2102.
\url{https://doi.org/10.1002/sim.8086}.

\leavevmode\vadjust pre{\hypertarget{ref-neym34}{}}%
Neyman, Jerzy. 1934. {``On the {Two Different Aspects} of the
{Representative Method}: {The Method} of {Stratified Sampling} and the
{Method} of {Purposive Selection}.''} \emph{Journal of the Royal
Statistical Society} 97 (4): 558--625.
\url{https://doi.org/10.2307/2342192}.

\leavevmode\vadjust pre{\hypertarget{ref-pulkki2015cumulative}{}}%
Pulkki-Råback, Laura, Marko Elovainio, Christian Hakulinen, Jari
Lipsanen, Mirka Hintsanen, Markus Jokela, Laura D Kubzansky, et al.
2015. {``Cumulative Effect of Psychosocial Factors in Youth on Ideal
Cardiovascular Health in Adulthood the Cardiovascular Risk in Young
Finns Study.''} \emph{Circulation} 131 (3): 245--53.

\leavevmode\vadjust pre{\hypertarget{ref-raghunathan2003multiple}{}}%
Raghunathan, Trivellore E, Jerome P Reiter, and Donald B Rubin. 2003.
{``Multiple Imputation for Statistical Disclosure Limitation.''}
\emph{Journal of Official Statistics} 19 (1): 1.

\leavevmode\vadjust pre{\hypertarget{ref-rubin1986statistical}{}}%
Rubin, Donald B. 1986. {``Statistical Matching Using File Concatenation
with Adjusted Weights and Multiple Imputations.''} \emph{Journal of
Business \& Economic Statistics} 4 (1): 87--94.

\leavevmode\vadjust pre{\hypertarget{ref-rubi76}{}}%
Rubin, Donald B. 1976. {``Inference and {Missing Data}.''}
\emph{Biometrika} 63 (3): 581--92.
\url{https://doi.org/10.2307/2335739}.

\leavevmode\vadjust pre{\hypertarget{ref-rubi87}{}}%
---------. 1987. \emph{Multiple {Imputation} for Nonresponse in
Surveys}. Wiley Series in Probability and Mathematical Statistics
{Applied} Probability and Statistics. {New York, NY}: {Wiley}.

\leavevmode\vadjust pre{\hypertarget{ref-shara2015randomly}{}}%
Shara, Nawar, Sayf A Yassin, Eduardas Valaitis, Hong Wang, Barbara V
Howard, Wenyu Wang, Elisa T Lee, and Jason G Umans. 2015. {``Randomly
and Non-Randomly Missing Renal Function Data in the Strong Heart Study:
A Comparison of Imputation Methods.''} \emph{PloS One} 10 (9): e0138923.

\leavevmode\vadjust pre{\hypertarget{ref-sorbi2014medium}{}}%
Sorbi, MJ, AM Kleiboer, HG van Silfhout, G Vink, and J Passchier. 2014.
{``Medium-Term Effectiveness of Online Behavioral Training in Migraine
Self-Management: {A} Randomized Trial Controlled over 10 Months.''}
\emph{Cephalalgia : An International Journal of Headache},
0333102414547137.

\leavevmode\vadjust pre{\hypertarget{ref-stinesen2015reduced}{}}%
Stinesen Kollberg, Karin, Ann-Charlotte Waldenström, Karin Bergmark,
Gail Dunberger, Anna Rossander, Ulrica Wilderäng, Elisabeth
Åvall-Lundqvist, and Gunnar Steineck. 2015. {``Reduced Vaginal
Elasticity, Reduced Lubrication, and Deep and Superficial Dyspareunia in
Irradiated Gynecological Cancer Survivors.''} \emph{Acta Oncologica} 54
(5): 772--79.

\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
Team, R Core. 2015. \emph{R: {A} Language and Environment for
Statistical Computing}. Manual. {Vienna, Austria}: {R Foundation for
Statistical Computing}.

\leavevmode\vadjust pre{\hypertarget{ref-fimd}{}}%
Van Buuren, Stef. 2012. \emph{Flexible Imputation of Missing Data}. {CRC
press}.

\leavevmode\vadjust pre{\hypertarget{ref-mice}{}}%
Van Buuren, Stef, and Karin Groothuis-Oudshoorn. 2011. {``Mice:
{Multivariate Imputation} by {Chained Equations} in {R}.''}
\emph{Journal of Statistical Software} 45 (1): 1--67.
\url{https://doi.org/10.18637/jss.v045.i03}.

\leavevmode\vadjust pre{\hypertarget{ref-vink14}{}}%
Vink, Gerko, and Stef van Buuren. 2014. {``Pooling Multiple Imputations
When the Sample Happens to Be the Population,''} September.

\leavevmode\vadjust pre{\hypertarget{ref-wong2015economic}{}}%
Wong, Frances Kam Yuet, Ching So, June Chau, Antony Kwan Pui Law,
Stanley Ku Fu Tam, and Sarah McGhee. 2015. {``Economic Evaluation of the
Differential Benefits of Home Visits with Telephone Calls and Telephone
Calls Only in Transitional Discharge Support.''} \emph{Age and Ageing}
44 (1): 143--47.

\leavevmode\vadjust pre{\hypertarget{ref-zhao22}{}}%
Zhao, Yang. 2022. {``Diagnostic Checking of Multiple Imputation
Models.''} \emph{AStA Advances in Statistical Analysis}, January.
\url{https://doi.org/10.1007/s10182-021-00429-1}.

\end{CSLReferences}

\end{document}

### 2022-02-03 R/run_sim.R

@@ -61,10 +61,10 @@ evaluation <- function(imp, amp, dat, r){
            broom::tidy(.x, conf.int = TRUE) %>% 
              .[2, c("estimate", "conf.low", "conf.high")]}) %>% 
     mutate(
-      dat = c("full", "cca", "imp"),
+      dat = factor(c("True", "CCA", "Imp."), ordered = TRUE),
       est = estimate,
       ciw = conf.high - conf.low,
-      cov = conf.low <= r && r <= conf.high,
+      cov = conf.low <= r & r <= conf.high,
       .keep = "none")
   
   # refs <- data.frame(
@@ -80,7 +80,7 @@ evaluation <- function(imp, amp, dat, r){
     rmse_cell = c(0, NA)) %>% 
     rbind(., purrr::map_dfr(1:imp$m, function(.m) {
     data.frame(
-      rmse_pred = rmse(fit$analyses[[.m]]$residuals),
+      rmse_pred = rmse(fit_imps$analyses[[.m]]$residuals),
       rmse_cell = rmse(imp$imp$Y[[.m]] - dat$Y[imp$where[, "Y"]])
     )}) %>% colMeans() %>% t())
   

### 2022-02-09 Manuscript/StandardizedEvaluation.Rmd
Before full sim

@@
   kableExtra::kable_classic(full_width = TRUE) %>% 
-  kableExtra::pack_rows(group_label = "$n = 50$", 1, 3) %>% 
-  kableExtra::pack_rows(group_label = "$n = 500$", 4, 6) %>% 
-  kableExtra::pack_rows(group_label = "$n = 5000$", 7, 9)
+  kableExtra::pack_rows(group_label = "$n = 50$", 1, 3, escape = FALSE) %>% 
+  kableExtra::pack_rows(group_label = "$n = 500$", 4, 6, escape = FALSE) %>% 
+  kableExtra::pack_rows(group_label = "$n = 5000$", 7, 9, escape = FALSE)
 

### 2022-03-04 Manuscript/StandardizedEvaluation.Rmd
Split up manuscript and simulation in separate markdown files

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
 library(dplyr)
-source(here::here("R/utils.R"))
-source(here::here("R/run_sim.R"))
+# source(here::here("R/utils.R"))
+# source(here::here("R/run_sim.R"))
 ```
@@
 - add simulation results
+
+# Structure
+
+1. introduction
+
+- aim
+- number of simulations
+
+2. data generation
+
+- model vs design based
+- data types
+- sample vs population (vs bootstrapping??)
+- number of observations
+- number of variables
+- coherence between variables
+
+3. missingness generation
+
+- monotone/univariate vs multivariate
+- missingness mechanism
+- type of M(N)AR
+- missingness proportion (vs proportion of incomplete cases! 50% incomplete == 25% missing in bivariate data)
+
+4. missing data methods
+
+- imputation method
+- number of imputations
+- number of iterations
+
+5. performance evaluation
+
+- estimand
+- diagnostic
+- ppc
 

### 2022-03-04 Simulation/Simulation.Rmd
Split up manuscript and simulation in separate markdown files

---
title: "Simulation study"
author: "Hanne Oberman"
date: "4-3-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document contains the code and results for a simulation study accompanying the Standardized Evaluation manuscript.

# Overview

The simulation contains a whopping set of `r 3*3*3*4*3*2*4*3` methods/conditions. That's because we use different parameters for each level of the simulation, see pseudo-code:

```
for (data geneneration conditions) {
  generate complete data
  for (amputation conditions) {
    ampute complete data
    for (imputation conditions) {
      impute incomplete data
      analyze imputed data
      }}}
```

The conditions are as follows.

Data generation conditions:

- number of observations ($n = \{50, 500, 5000\}$) *only max required*
- number of variables ($p = \{2\}$)
- coherence between variables ($r = \{0, 0.4, 0.8\}$)

Amputation conditions:

- missingness mechanism (`mech` = {MCAR, MAR, MNAR})
- type of M(N)AR (`type` = {LEFT, RIGHT, MID, TAIL})
- missingness proportion (`prop` = {0.1, 0.25, 0.5})

Imputation conditions:

- imputation method (`meth` = {none, mean imputation, regression imputation, stochastic regression})
- number of imputations (`m` = {1, 5}) *only max required*
- number of iterations (`maxit` = 1, 5, 10) *use mice.mids to shorten runtime?*

Performance evaluation:
- estimand (average, regression coefficient)
- diagnostic (bias, coverage rate, confidence interval width, cell rmse, analysis model rmse)

# Simulation functions

For data generation we use the following:

```{r dat}
generation <- function(r, n = c(50, 500, 5000), p = 2) {
  # generate data based on variance-covariance matrix
  vcov <- matrix(r, p, p)
  diag(vcov) <- 1
  dat <- data.frame(mvtnorm::rmvnorm(
    n = max(n),
    mean = rep(0, p),
    sigma = vcov
  )) %>% setNames(paste0("X", 1:p))
  # split output by number of observations
  out <- purrr::map(n, ~{
    list(n_obs = .x,
         p = p,
         r = r,
         dat = dat[1:.x,])
  })
  return(out)
}
```

For amputation of the complete data:

```{r amp}
amputation <- function(dat, mech, type, prop) {
  # induce missingness
  amp <- dat %>%
    mice::ampute(prop = prop,
                 mech = mech,
                 type = type)
  # output
  out <- amp
  return(out)
}
```

To impute the incomplete data:

```{r imp}
imputation <- function(amp, meth, m = 5, it = 10) {
  imp <- amp$amp %>%
    mice::mice(
      m = m,
      method = meth,
      maxit = it,
      print = FALSE
    )
  # output
  out <- imp
  return(out)
}
```

And finally, for performance evaluation:

```{r perf}

```

### 2022-03-04 R/run_sim.R

@@ -1,25 +1,30 @@
-# small simulation study for illustration
+# simulation study for illustration
 
 # set-up 
 # set.seed(123)
 # library(dplyr)
 # source("R/utils.R")
+n_sim = 2
 
-# # parameters
-# n_sim <- 1000 #c(100, 1000)
-# n_obs <- 500 #c(50, 500, 5000)
-# n_var <- 2 #c(2, 5, 20)
-# corr <- 0.4 #c(0, .8)
-# mis_mech <- "MAR" #c("MCAR", "MAR", "MNAR")
-# mis_type <- "RIGHT" #c("left", "right", "mid", "tail")
-# mis_prop <- 0.5 #c(.1, .25, .5)
-# n_imp <- 5 #c(1, 5)
-# imp_meth <- "norm" #c("mean", "norm.predict", "norm") #cca
-# n_it <- 10 #c(1, 5, 10)
-# # est <- c("mean(Y)", "var(Y)", "cov(X,Y)", "lm(Y~X)", "lm(X~Y)")
-# # model vs design based
-# # with or without sampling variance
+## simulation parameters ##
 
+# conditions
+n_obs = c(50, 500, 5000) 
+corr = c(0, .4, .8)
+mis_mech = c("MCAR", "MAR", "MNAR")
+mis_type = c("LEFT", "RIGHT", "MID", "TAIL")
+mis_prop = c(.1, .25, .5)
+n_imp = c(1, 5)
+imp_meth = c("mean", "norm.predict", "norm") 
+n_it = c(1, 5, 10)
+
+# combine into lists for mapping
+dat_par <- expand.grid(n_obs = n_obs, corr = corr)
+amp_par <- expand.grid(dat_nr = 1:nrow(dat_par), mis_mech = mis_mech, mis_type = mis_type, mis_prop = mis_prop, stringsAsFactors = FALSE)
+imp_par <- expand.grid(amp_nr = 1:nrow(amp_par), imp_meth = imp_meth, n_imp = n_imp, n_it = n_it, stringsAsFactors = FALSE)
+
+
+## simulation functions ##
 
 # generate data
 generation <- function(n, r){
@@ -55,24 +60,17 @@ evaluation <- function(imp, amp, dat, r){
   fit_imps <- with(imp, lm(Y ~ X))
   fits <- list(full = lm(Y~X, dat),
                cca = lm(Y~X, amp$amp, na.action = na.omit),
-               imp = mice::pool(fit_imps))
[…]

### 2022-03-05 Manuscript/StandardizedEvaluation.Rmd
Edit intro and problems section

--- 2022-02-03 Manuscript/StandardizedEvaluation.Rmd
@@
 Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
-In this paper I propose a move towards a standardized evaluation of imputation methods.
-To demonstrate the need for standardization, I highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
+In this paper we propose a move towards a standardized evaluation of imputation methods.
+To demonstrate the need for standardization, we highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
 This may lead to suboptimal use of multiple imputation in practice.
-Additionally, I suggest a course of action for simulating and evaluating missing data problems."
+Additionally, we suggest a course of action for simulating and evaluating missing data problems."
 bibliography: StandardizedEvaluation.bib  
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
 library(dplyr)
-source(here::here("R/utils.R"))
-source(here::here("R/run_sim.R"))
+# source(here::here("R/utils.R"))
+# source(here::here("R/run_sim.R"))
 ```
 
-# TO DO
+<!-- # Structure -->
 
-- add ADEMP paper [@morr18], diagnostic checking paper [@zhao22]
-- add PPC (refer to Mingyang, and Gelman & King)
-- expand convergence section
-- add RMSE
-- add insights from literature review for chapter on missing data in EHR
-- add link to simulation script, and $n$
-- add simulation results
+<!-- 1. introduction -->
+
+<!-- - aim -->
+<!-- - number of simulations -->
+
+<!-- 2. data generation -->
+
+<!-- - model vs design based -->
+<!-- - data types -->
+<!-- - sample vs population (vs bootstrapping??) -->
+<!-- - number of observations -->
+<!-- - number of variables -->
+<!-- - coherence between variables -->
+
+<!-- 3. missingness generation -->
+
+<!-- - monotone/univariate vs multivariate -->
+<!-- - missingness mechanism -->
+<!-- - type of M(N)AR -->
+<!-- - missingness proportion (vs proportion of incomplete cases! 50% incomplete == 25% missing in bivariate data) -->
+
+<!-- 4. missing data methods -->
+
+<!-- - imputation method -->
+<!-- - number of imputations -->
+<!-- - number of iterations -->
+
+<!-- 5. performance evaluation -->
+
+<!-- - estimand -->
+<!-- - diagnostic -->
+<!-- - ppc -->
 
@@
 
-Multiple imputation @rubin1987 is a state-of-the-art technique for drawing valid conclusions from incomplete data, and is increasingly used in many substantive fields, such as medicine [see e.g.
-@pulkki2015cumulative; @alkire2015global; @stinesen2015reduced], psychology [e.g.
-@sorbi2014medium], economics [e.g.
-@wong2015economic; @dafoe2013democratic; @allan2015determinants], sociology [e.g.
-@crowley2014flexible; @de2014religiosity], and official statistics [e.g.
-@little2012calibrated; @klausch2015selection].
-
+Multiple imputation [@rubin1987] is a state-of-the-art technique for drawing valid conclusions from incomplete data.
 The technique has earned a permanent spot in research and policy making, demonstrated e.g.
@@
 
-The idea behind multiple imputation is to impute (fill in) the missing values multiple times to obtain a valid estimate of what could have been.
+The idea behind multiple imputation is to impute (fill in) the missing values in incomplete data multiple times, to obtain a valid estimate of what could have been observed.
 The multiple datasets that are thus obtained can be analyzed by standard techniques and the analysis results can be combined into a single inference.
 In contrast to ad hoc methods for dealing with missing values (e.g.
-list-wise deletion, mean imputation, regression imputation, last observation carried forward), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
+list-wise deletion, mean imputation, last observation carried forward), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
+[It has been empirically demonstrated time and again that the inferences on multiply imputed data yield unbiased and confidence valid results (through Rubin's rules).
+But how do we know that this is the case, if we do not have access to the true but unobserved values of the missing entries?
+And how can we be sure that extensions to the methodology yield valid inferences too?
+Especially in the field of ML/AI, there are promising new imputation methods, which may yield even sharper imputations than now-standard (semi-)parametric imputation methods.
+That's why we use simulation studies for the evaluation of imputation routines.]
 
-The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
+<!-- The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
 In general it holds that modeling missing data becomes more challenging when the amount of missingness increases.
@@
 In that case, multiple imputation would be substantially more efficient than the ubiquitous list-wise deletion.
+-->
 
-When evaluating the statistical properties (and thereby the practical applicability) of new imputation methodology, researchers most often make use of simulation studies.
-In such studies a complete data set is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of the imputation method.
-However, no golden standard has been established and, as a result, validity of the simulation and evaluation procedures may differ tremendously from one developer to another.
+When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies.
+In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, [then several missing data methods are applied] and a set of evaluation criteria is postulated to evaluate the performance of the imputation methods.
+However, no golden standard has been established to evaluate these imputation routines and, as a result, the validity of the simulation procedures may differ tremendously from one developer to another.
 
-The purpose of this paper is threefold: First, I raise some concerns with respect to evaluating imputation methodology.
-These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals.
-Second, I provide imputation methodologists with a suggested course of action when simulating missing data problems.
+The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology.
+These concerns stem from careful consideration with fellow 'imputers' and from encounters as a reviewer for statistical journals.
+Second, to provide imputation methodologists with a suggested course of action when simulating missing data problems.
 This suggested approach should identify a common ground, but is in no way intended as an absolute solution.
 This identifies the third purpose of this paper: discussion.
-I hope to elicit critical thinking with respect to the problems at hand.
+We hope to elicit critical thinking with respect to the problems at hand.
 We are all convinced that our methodology has some merit.
@@
 
-## Why some evaluations should not be trusted
+# Why some evaluations should not be trusted
 
@@
 Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers.
-This brings forth a chain of potential problems in the objective assessment of the performance of imputation routines that may lead to suboptimal use of multiple imputation in practice.
-To demonstrate the broad impact of these problems, I subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation.
-I further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.
+This brings forth a chain of potential problems in the objective assessment of the performance of imputation routines that may lead to sub-optimal use of multiple imputation in practice.
+To demonstrate the broad impact of these problems, we subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation.
+we further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.
 
@@
 
-To evaluate the ability of an imputation routine to handle the non-response model, a form of truth has to be established.
-Those who perform simulation studies are in the luxury position to establish the truth beforehand, either by generating data from a statistical model, or by sampling data from a sufficiently large observed set.
-However, when generating data, two easily overlooked problems may arise.
+To evaluate the ability of an imputation routine to handle ~~the non-response model~~ [missing data entries], a form of truth has to be established.
+Those who perform simulation studies are in the luxury position to establish the truth beforehand, by choosing a data generating mechanism to obtain a complete dataset.
+Complete data is typically either generated from a statistical model, or sampled from a sufficiently large ~~observed~~ set [model-based vs design-based simulation].
+~~However, when generating data, two easily overlooked problems may arise.~~
 
+The problem with model-based data generating mechanisms is that method performance may not translate to empirical data.
 First, data are often generated following the model (or distribution) that is also used for imputing the data.
-The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied.
-Although such procedures may be statistically relevant, the approach would be no good for even the simplest case.
+The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied [@morr18].
+Although such procedures may be statistically relevant, the approach would be no good for even the simplest empirical case.
 Such simulations pose serious threats to the applicability of the method as real-life data hardly ever follow a given theoretical distribution.
-
-Second, data are often generated such that the problem that is being studied is most pronounced.
+[Add that data types may be more messy.] Second, data are often generated such that the problem that is being studied is most pronounced.
 This results in simulated data that contains very valuable information structures, i.e.
@@
 
+The problem with design-based simulation is that it's difficult to find large datasets without missing entries.
+And imputing the missing data beforehand and then treating the once-imputed data as comparative truth may favor the used imputation method in further evaluations.
+A second issue is that SEs may be incorrect if the (re)sampling is not handled correctly.
+[Add pooling when the sample is the population [@vink14] and write something about bootstrapping SEs].
+
+Problems with either of these data generation approaches is that the number of observations, the number of variables, the variable type(s) and the coherence between variables may give some imputation methods a comparative advantage, if they are not a) varied across simulation conditions [when is this fair?
+is this even an option?], or b) the object of the study [add some examples of simulations specifically focused on small samples, categorical data, etc.].
+
+[Maybe add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
+I don't know where to introduce the difference between statistical inference and prediction.
+But this is an important third category, it seems.
+These studies apply different combinations of imputation and estimation methods on one or more benchmark datasets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
+In this context, there is no talk of retrieving the 'true' but unobserved values at all.
+Refer to missing data chapter [@liu21].]
+
 ```{r sim_nobs, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, paged.print=TRUE}
-# setup
-set.seed(1)
-n_sim <- 100
-
-sim_nobs <- purrr::map_dfr(1:n_sim, function(.i){
-  purrr::map_dfr(c(50, 500, 5000), ~{
-    run(n_obs = .x) %>% cbind(sim = .i, .)})})
-
-# gtsummary::tbl_summary(sim_nobs, by = dat)
-# https://haozhu233.github.io/kableExtra/awesome_table_in_html.html#Group_rows_via_labeling
-sim_nobs %>% 
-  group_by(n_obs, dat) %>% 
-  summarise(
-    mean_Y = mean(ave),
-    beta_X = mean(est),
-    CIW = mean(ciw),
-    CR = mean(cov),
-    RMSE_pred = mean(rmse_pred),
-    RMSE_cell = mean(rmse_cell, na.rm = TRUE))  %>% 
-  .[, -1] %>% 
-  knitr::kable(
-    col.names = c("", "$\\mu_{Y} = 10$", "$\\beta_{X} = 0.4$", "CIW", "CR", "RMSE$_{p}$", "RMSE$_{c}$"), 
-    digits = 3,
-    escape = FALSE) %>% 
-  kableExtra::kable_classic(full_width = TRUE) %>% 
-  kableExtra::pack_rows(group_label = "$n = 50$", 1, 3) %>% 
-  kableExtra::pack_rows(group_label = "$n = 500$", 4, 6) %>% 
-  kableExtra::pack_rows(group_label = "$n = 5000$", 7, 9)
-
+# # setup
+# set.seed(1)
+# n_sim <- 100
+# 
+# sim_nobs <- purrr::map_dfr(1:n_sim, function(.i){
+#   purrr::map_dfr(c(50, 500, 5000), ~{
+#     run(n_obs = .x) %>% cbind(sim = .i, .)})}) 
+# 
+# # gtsummary::tbl_summary(sim_nobs, by = dat)
+# # https://haozhu233.github.io/kableExtra/awesome_table_in_html.html#Group_rows_via_labeling
+# sim_nobs %>% 
+#   group_by(n_obs, dat) %>% 
+#   summarise(
+#     mean_Y = mean(ave),
+#     beta_X = mean(est),
+#     CIW = mean(ciw),
+#     CR = mean(cov),
+#     RMSE_pred = mean(rmse_pred),
+#     RMSE_cell = mean(rmse_cell, na.rm = TRUE))  %>% 
+#   .[, -1] %>% 
+#   knitr::kable(
+#     col.names = c("", "$\\mu_{Y} = 10$", "$\\beta_{X} = 0.4$", "CIW", "CR", "RMSE$_{p}$", "RMSE$_{c}$"), 
+#     digits = 3,
+#     escape = FALSE) %>% 
+#   kableExtra::kable_classic(full_width = TRUE) %>% 
+#   kableExtra::pack_rows(group_label = "$n = 50$", 1, 3, escape = FALSE) %>% 
+#   kableExtra::pack_rows(group_label = "$n = 500$", 4, 6, escape = FALSE) %>% 
+#   kableExtra::pack_rows(group_label = "$n = 5000$", 7, 9, escape = FALSE)
 ```
@@
 
-In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng94.
+Missingness comes in many shapes and forms [platitude, but how else to start?].
+The most obvious characteristics of missing data are the missingness pattern(s) and missingness mechanism(s).
+Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data [@litt20, p.
+8].
+
+<!-- L&R, p.
+8: "We find it useful to distinguish the missingness pattern, which describes which values are missing and observed in the data matrix and the missingness mechanism (or mechanisms), which concerns the relationship between missingness and the values of variables in the data matrix".-->
+
+The missing data pattern may inadvertently guide which imputation method prevails in simulation studies.
+Some ad hoc methods are known to yield valid inferences in very specific missing data patterns (e.g., list-wise deletion when there is missingness in the outcome variable of the analysis model exclusively), while they bias inferences in any other case.
+Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions [re-read this!
+and add monotone/univariate vs multivariate and sporadic vs systematic dichotomies].
+Related to the missingness pattern across variables is the missingness proportion.
+The terminology on this topic can be confusing.
+Does a missingness proportion of 50% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
+We will refer to the latter as the proportion of incomplete cases.
+But this is not a clear distinction in the literature, which may lead to incorrect conclusions [re-phrase this too].
+Moreover, the amount of missingness that is used to simulate the performance of imputation methods differs between studies.
+[Add ref to chapter [@liu21] and add info.] Some studies use only 10% missing data where other studies push the limits to additionally investigate the performance under at least 50% missing data.
+This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
+This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
+
+[Briefly introduce missingness mechanisms here and show potential problems, but without simulation results (moved to separate section).
+This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and sometimes realistic in practice, 3) that MAR can be generated *spuriously* if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
+And finally something about types of M(N)AR, but what are specific problems with that?]
+
+# 
+\hrule
+
+#
+[Until next hline: to be edited!] In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng94.
 The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
@@
 
-Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit @molenberghs2008every - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
+Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit [@molenberghs2008every] - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
 In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis see e.g.
@@
 
-The amount of missingness that is used to simulate the performance of imputation methods differs between studies.
-Some studies use only 10% missing data where other studies push the limits to additionally investigate the performance under at least 50% missing data.
-This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
-This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
+# 
+\hrule
 
@@
 
-The evaluation criteria used to assess imputation performance vary from one developer to another.
-This is not surprising as people from different fields could have a different focus on the problem at hand.
-However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated.
-I highlight four aspects of evaluation that should be considered when evaluation imputation routines.
+[Add bridge from last section. And explain how this part is subdivided: first general characteristics of the imputations such as convergence, then performance measures.]
 
-First, developers often only inquire about the 'accuracy' (i.e.
-how well can the method reproduce the original data).
-The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-We are interested in the correct answer to the research question; not in the truth itself.
-This means that, given the framework provided by @rubi87, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
-After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neym34, p.
-591.
+[Performance of different imputation methods may be influenced by preconditions of the methods themselves.
+All imputation models should be congenial, but this is difficult to test to often ignored even in single imputation endeavor, let alone in simulation studies with 100s of imputations.
+Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies.
+Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.]
 
-Second, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
+Moreover, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
 For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large.
@@
 
-Third, sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
+[Then onto the performance measure part.]
+
+The evaluation criteria used to assess imputation performance vary from one developer to another.
+This is not surprising as people from different fields could have a different focus on the problem at hand.
+But the choice of performance measure(s) may inadvertently distort statistical properties of the imputed data.
+
+<!-- ~~However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated.
+We highlight four aspects of evaluation that should be considered when evaluation imputation routines.~~ -->
+
+Developers often only inquire about the 'accuracy' (i.e.
+how well can the method reproduce the original data).
+[Add RMSE at cell level here?] The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
+We are interested in the correct answer to the research question; not in the truth itself.
+This means that, given the framework provided by @rubi87, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
+After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neym34, p.
+591.
+[So, what could go wrong by focusing on accuracy exclusively?]
+
+[Add choice of analysis model/estimand here, since performance measure are defined wrt the estimand, e.g., focusing on univariate estimates may favor invalid ad hoc methods.]
+
+[Add problems with RMSE here.]
+
+Sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
 However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink14.
@@
 A method may perform badly, but if it still outperforms every other approach, it may yet be of great practical relevance.
+
+# Illustration
+
+[Add simulation study + results here!]
 
@@
 This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations.
-One needs to realize, however, that the conventional pooling rules cf.
+One needs to realize, however, that the conventional pooling rules [cf.
 @rubin1987, pp.
-76-77, do not apply for finite population inference and that alternative pooling rules need to be used @raghunathan2003multiple, @vink2014pooling.
+76-77], do not apply for finite population inference and that alternative pooling rules need to be used [@raghunathan2003multiple; @vink2014pooling].
 
@@
 Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
-This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching @little1988missing, @rubin1986statistical may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching [@little1988missing; @rubin1986statistical] may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
 It would be a shame to overlook such results due to the focus on a single MAR mechanism.
@@
 Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
+- [Add RMSE 2x]
+
 - \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data.
@@
 When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
+[Add PPC]
 - \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers, PIGS, @li2012imputing.
 The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
+[Extend this part!]
+- [Add PPC (Mingyang) [@nguy17], [@zhao22]]
 

### 2022-03-07 Manuscript/StandardizedEvaluation.Rmd
Version to compare with latex diff or pdf diff

@@
 
-<!-- The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
+~~The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
 In general it holds that modeling missing data becomes more challenging when the amount of missingness increases.
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
-In that case, multiple imputation would be substantially more efficient than the ubiquitous list-wise deletion.
--->
+In that case, multiple imputation would be substantially more efficient than the ubiquitous list-wise deletion.~~
 
@@
 
-<!-- ~~However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated.
-We highlight four aspects of evaluation that should be considered when evaluation imputation routines.~~ -->
+~~However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated.
+We highlight four aspects of evaluation that should be considered when evaluation imputation routines.~~
 

### 2022-03-07 R/run_sim.R

@@ -1,9 +1,9 @@
 # simulation study for illustration
 
 # set-up 
-# set.seed(123)
-# library(dplyr)
-# source("R/utils.R")
+set.seed(123)
+library(dplyr)
+source("R/utils.R")
 n_sim = 2
 
 ## simulation parameters ##
@@ -85,14 +85,14 @@ evaluation <- function(imp, amp, dat, r){
   # performance
   out <- out %>% 
     cbind(
-      # n_obs = nrow(dat), 
-      # corr = r, 
-      # mis_mech = amp$mech,
-      # mis_type = amp$type,
-      # mis_prop = amp$prop,
-      # imp_meth = imp$method[1],
-      # n_imp = imp$m,
-      # n_it = imp$iteration,
+      n_obs = nrow(dat),
+      corr = r,
+      mis_mech = amp$mech,
+      mis_type = amp$type,
+      mis_prop = amp$prop,
+      imp_meth = imp$method[1],
+      n_imp = imp$m,
+      n_it = imp$iteration,
       .,
       ave = c(mean(dat$Y), mean(amp$amp$Y, na.rm = TRUE), mean(mice::complete(imp, "long")$Y)),
       rmses,
@@ -132,120 +132,3 @@ return(out)
 results <- purrr::map_dfr(1:n_sim, ~{
   cbind(sim = .x, simulation())
 })
-
-# all_par <- expand.grid(
-#   n_obs = n_obs, 
-#   corr = corr,
-#   mis_mech = mis_mech,
-#   mis_type = mis_type,
-#   mis_prop = mis_prop,
-#   imp_meth = imp_meth,
-#   n_imp = n_imp,
-#   n_it = n_it)
-
-# # simulate complete datasets
-# dat <- purrr::map(n_obs, function(n){
-#   purrr::map(corr, function(r){
-#     generation(n, r)
-#   }) %>% setNames(corr)
-# }) %>% setNames(n_obs)
-# 
-# # ampute all datasets
-# amp <- purrr::map(n_obs, function(n){
-#   purrr::map(corr, function(r){
-#     purrr::map(mis_mech, function(mech){
-#   amputation(dat[[n]][[r]], mech, type = mis_type, prop = mis_prop)})})})
-
-# impute all incomplete datasets
-
-
-
-# l2 <- list(
[…]

### 2022-03-22 Manuscript/StandardizedEvaluation.Rmd
recompile and save .tex

@@
 bibliography: StandardizedEvaluation.bib  
-output: pdf_document
+output: 
+  pdf_document:
+    keep_tex: true
 editor_options: 

### 2022-03-22 Manuscript/JournalTemplate/Semi_SEM_revise2.tex

\documentclass[bimj,fleqn]{w-art}
\usepackage{times}
\usepackage{w-thm}
\usepackage[authoryear]{natbib}
\setlength{\bibsep}{2pt}
\setlength{\bibhang}{2em}
\newcommand{\J}{J\"{o}reskog}
\newcommand{\So}{S\"{o}rbom}
\newcommand{\bcx}{{\bf X}}
\newcommand{\bcy}{{\bf Y}}
\newcommand{\bcz}{{\bf Z}}
\newcommand{\bcu}{{\bf U}}
\newcommand{\bcv}{{\bf V}}
\newcommand{\bcw}{{\bf W}}
\newcommand{\bci}{{\bf I}}
\newcommand{\bch}{{\bf H}}
\newcommand{\bcb}{{\bf B}}
\newcommand{\bcr}{{\bf R}}
\newcommand{\bcm}{{\bf M}}
\newcommand{\bcf}{{\bf F}}
\newcommand{\bcg}{{\bf G}}
\newcommand{\bcs}{{\bf S}}
\newcommand{\bca}{{\bf A}}
\newcommand{\bcd}{{\bf D}}
\newcommand{\bcc}{{\bf C}}
\newcommand{\bce}{{\bf E}}
\newcommand{\ba}{{\bf a}}
\newcommand{\bb}{{\bf b}}
\newcommand{\bc}{{\bf c}}
\newcommand{\bd}{{\bf d}}
\newcommand{\bx}{{\bf x}}
\newcommand{\by}{{\bf y}}
\newcommand{\bz}{{\bf z}}
\newcommand{\bu}{{\bf u}}
\newcommand{\bv}{{\bf v}}
\newcommand{\bh}{{\bf h}}
\newcommand{\bl}{{\bf l}}
\newcommand{\be}{{\bf e}}
\newcommand{\br}{{\bf r}}
\newcommand{\bw}{{\bf w}}
\newcommand{\de}{\stackrel{D}{=}}
\newcommand{\bt}{\bigtriangleup}
\newcommand{\bfequiv}{\mbox{\boldmath $\equiv$}}
\newcommand{\bmu}{\mbox{\boldmath $\mu$}}
\newcommand{\bnu}{\mbox{\boldmath $\nu$}}
\newcommand{\bxi}{\mbox{\boldmath $\xi$}}
\newcommand{\btau}{\mbox{\boldmath $\tau$}}
\newcommand{\bgamma}{\mbox{\boldmath $\Gamma$}}
\newcommand{\bphi}{\mbox{\boldmath $\Phi$}}
\newcommand{\bfphi}{\mbox{\boldmath $\varphi$}}
\newcommand{\bfeta}{\mbox{\boldmath $\eta$}}
\newcommand{\bpi}{\mbox{\boldmath $\Pi$}}
\newcommand{\bequiv}{\mbox{\boldmath $\equiv$}}
\newcommand{\bvarepsilon}{\mbox{\boldmath $\varepsilon$}}
\newcommand{\btriangle}{\mbox{\boldmath $\triangle$}}
\newcommand{\bdelta}{\mbox{\boldmath $\Delta$}}
\newcommand{\beps}{\mbox{\boldmath $\epsilon$}}
\newcommand{\btheta}{\mbox{\boldmath $\theta$}}
\newcommand{\balpha}{\mbox{\boldmath $\alpha$}}
\newcommand{\bsphi}{\mbox{\boldmath $\varphi$}}
\newcommand{\bsig}{\mbox{\boldmath $\sigma$}}
\newcommand{\bfpsi}{\mbox{\boldmath $\psi$}}
\newcommand{\bfdelta}{\mbox{\boldmath $\delta$}}
\newcommand{\bsigma}{{\bf \Sigma}}
\newcommand{\bzero}{{\bf 0}}
\newcommand{\bpsi}{\mbox{\boldmath $\Psi$}}
\newcommand{\bep}{\mbox{\boldmath $\epsilon$}}
\newcommand{\bomega}{\mbox{\boldmath $\Omega$}}
\newcommand{\bfomega}{\mbox{\boldmath $\omega$}}
\newcommand{\blambda}{\mbox{\boldmath $\Lambda$}}
\newcommand{\bflambda}{\mbox{\boldmath $\lambda$}}
\newcommand{\bfsigma}{\mbox{\boldmath $\sigma$}}
\newcommand{\bfpi}{{\mbox{\boldmath $\pi$}}}
\newcommand{\bupsilon}{\mbox{\boldmath $\upsilon$}}
\newcommand{\obs}{{\rm obs}}
\newcommand{\mis}{{\rm mis}}
\theoremstyle{plain}
\newtheorem{criterion}{Criterion}
\theoremstyle{definition}
\newtheorem{condition}[theorem]{Condition}
\usepackage[]{graphicx}
\chardef\bslash=`\\ % p. 424, TeXbook
\newcommand{\ntt}{\normalfont\ttfamily}
\newcommand{\cn}[1]{{\protect\ntt\bslash#1}}
\newcommand{\pkg}[1]{{\protect\ntt#1}}
\let\fn\pkg
\let\env\pkg
\let\opt\pkg
\hfuzz1pc % Don't bother to report overfull boxes if overage is < 1pc
\newcommand{\envert}[1]{\left\lvert#1\right\rvert}
\let\abs=\envert

\begin{document}
%\DOIsuffix{bimj.DOIsuffix}
\DOIsuffix{bimj.200100000}
\Volume{52}
\Issue{61}
\Year{2010}
\pagespan{1}{}
\keywords{Key word one; Key word two; Key word three; Key word four; Key word five;\\
\noindent \hspace*{-4pc} {\small\it (Up to five keywords are allowed and should be given in alphabetical order. Please capitalize the key}\\
\hspace*{-4pc} {\small\it words)}\\[1pc]
\noindent\hspace*{-4.2pc} Supporting Information for this article is available from the author or on the WWW under\break \hspace*{-4pc} \underline{http://dx.doi.org/10.1022/bimj.XXXXXXX} (please delete if not
applicable)
}  %%% semicolon and fullpoint added here for keyword style

\title[Running title]{Long title {\small\it (please use sentence case, e.g. Score tests for exploring complex models)}}
%% Information for the first author.
\author[First Author {\it{et al.}}]{First Author\footnote{Corresponding author: {\sf{e-mail: author@emailaddress.com}}, Phone: +00-999-999-999, Fax: +00-999-999-999}\inst{,1}} 
\address[\inst{1}]{First address {\it(please include the department and the postal address)}}
%%%%    Information for the second author
\author[dd]{Second Author\inst{1,2}}
\address[\inst{2}]{Second address {\it{(please include the department and the postal address)}}}
%%%%    Information for the third author
\author[]{Third Author\inst{2} {\small\it (please provide full author names. Middle names should be indicated by initials only, i.e. Henry J. James)}}
%%%%    \dedicatory{This is a dedicatory.}
\Receiveddate{zzz} \Reviseddate{zzz} \Accepteddate{zzz}

\begin{abstract}
Start. This is the abstract. You may use this LaTeX template to write your text.
The abstract should not contain multiple paragraphs, formulae NOR references and should ideally
not be longer than 250 words. This is the abstract. You may use this LaTeX template to write your text.
The abstract should not contain multiple paragraphs, formulae NOR references and should ideally
not be longer than 250 words.  This is the abstract. You may use this LaTeX template to write your text.
The abstract should not contain multiple paragraphs, formulae NOR references and should ideally
not be longer than 250 words.  This is the abstract. You may use this LaTeX template to write your text.
The abstract should not contain multiple paragraphs, formulae NOR references and should ideally
not be longer than 250 words.  This is the abstract. You may use this LaTeX template to write your text.
The abstract should not contain multiple paragraphs, formulae NOR references and should ideally
not be longer than 250 words. Start. This is the abstract. You may use this LaTeX template to write your text.
The abstract should not contain multiple paragraphs, formulae NOR references and should ideally
not be longer than 250 words. 
 Stop.
\end{abstract}

%% maketitle must follow the abstract.
\maketitle                   % Produces the title.

%% If there is not enough space inside the running head
%% for all authors including the title you may provide
%% the leftmark in one of the following three forms:

%% \renewcommand{\leftmark}
%% {First Author: A Short Title}

%% \renewcommand{\leftmark}
%% {First Author and Second Author: A Short Title}

%% \renewcommand{\leftmark}
%% {First Author et al.: A Short Title}

%% \tableofcontents  % Produces the table of contents.

\section{Introduction}

This is normal body text. Please note that the body text must be divided into numbered sections with
suitable short verbal titles. Sub-headings are allowed but not sub-subheadings. Please use sentence case in the
headings.

\subsection{Second level heading}

This is the body text. Please note that cross-references in the body text should be shown as follows:
(Miller, 1900), (Miller and Baker, 1900) or if three or more authors (Miller {\it{et al}}., 1900)
\vspace*{12pt}

\noindent Bullet lists are not allowed. Always use (i), (ii), etc.
\vspace*{12pt}

\noindent Sentences should never start with a symbol.
\vspace*{12pt}

\noindent Names of software packages and website addresses should be written in {\tt{Courier new, i.e. Stata, the R package
MASS, http://www.biometrical-journal.com.}}

\begin{table}[htb]
\begin{center}
\caption{The caption of a table.}
\begin{tabular}{lll}
\hline
Description 1 & Description 2 & Description 3\\
\hline
Row 1, Col 1 & Row 1, Col 2 & Row 1, Col 3\\
Row 2, Col 1 & Row 2, Col 2 & Row 2, Col 3\\
\hline
\end{tabular}
\end{center}
\end{table}
\begin{equation}
\left({\theta^{0}_{i}}\atop{\theta^{1}_{i}}\right) \sim N(\theta,\Sigma),\quad {\mathrm{with}}\ 
{{\theta}} = \left({\theta_{0}}\atop{\theta_{1}}\right)\ {\mathrm{and}}\ \Sigma =
\left(\begin{array}{cc}
\sigma^{2}_{0} & \rho\sigma_{0}\sigma_{1}\\
\rho\sigma_{0}\sigma_{2} & \sigma^{2}_{1}
\end{array}\right).
\end{equation}

\noindent This is the body text. Only number equations which are referred to in the text body. If equations
are numbered, these should be numbered continuously throughout the text. Not section wise! Please
carefully follow the rules for mathematical expressions in the ``Instructions to Authors''.

\begin{figure}[htb]
\begin{center}
\includegraphics[bb= 0 0 115 87]{empty.eps}
\caption{The figure caption ($b_{2}b_{2}a^{n}a^{n}$).}
\end{center}
\end{figure}
\begin{acknowledgement}
An acknowledgement may be placed at the end of the article.
\end{acknowledgement}
\vspace*{1pc}

\noindent {\bf{Conflict of Interest}}

\noindent {\it{The authors have declared no conflict of interest. (or please state any conflicts of interest)}}

\section*{Appendix {\it(please insert here, if applicable)}}

\subsection*{A.1.\enspace Second level heading}

Please insert appendices before the references.

\begin{thebibliography}{10}
\bibitem[Bauer and Bauer(1994)Bauer, P. and Bauer, M.M.]{bib1}Bauer, P. and  Bauer, M. M. (1994). Testing equivalence simultaneously for location and  dispersion of two normally distributed populations.  \textit{Biometrical  Journal} \textbf{36}, 643--660.
\bibitem[Farrington, C. P. and Andrews, N. (2003)]{bib2}Farrington, C.P. and Andrews, N. (2003). Outbreak detection:
Application to infectious disease surveillance. In: Monitoring the Health of Populations (eds. R. Brookmeyer and D. F. Stroup), Oxford University Press, Oxford,\break 203--231.
\bibitem[Rencher(1998)Rencher, A.C.]{bib3}Rencher, A. C. (1998).  \textit{Multivariate Statistical Inference and Applications}. Wiley, New  York. 
\end{thebibliography}
\newpage
\phantom{aaaa}
\end{document}

### 2022-03-22 Manuscript/StandardizedEvaluation.tex

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{graphicx}
@@
 \makeatother
+\usepackage[normalem]{ulem}
+% Avoid problems with \sout in headers with hyperref
+\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 \ifLuaTeX
@@
 studies to evaluate the properties of imputation methods. In this paper
-I propose a move towards a standardized evaluation of imputation
-methods. To demonstrate the need for standardization, I highlight a set
+we propose a move towards a standardized evaluation of imputation
+methods. To demonstrate the need for standardization, we highlight a set
 of potential pitfalls that bring forth a chain of potential problems in
@@
 may lead to suboptimal use of multiple imputation in practice.
-Additionally, I suggest a course of action for simulating and evaluating
-missing data problems.
+Additionally, we suggest a course of action for simulating and
+evaluating missing data problems.
 \end{abstract}
-
-\hypertarget{to-do}{%
-\section{TO DO}\label{to-do}}
-
-\begin{itemize}
-\tightlist
-\item
-  add ADEMP paper (Morris, White, and Crowther 2019), diagnostic
-  checking paper (Zhao 2022)
-\item
-  add PPC (refer to Mingyang, and Gelman \& King)
-\item
-  expand convergence section
-\item
-  add RMSE
-\item
-  add insights from literature review for chapter on missing data in EHR
-\item
-  add link to simulation script, and \(n\)
-\item
-  add simulation results
-\end{itemize}
 
@@
 
-Multiple imputation Donald B. Rubin (1987) is a state-of-the-art
-technique for drawing valid conclusions from incomplete data, and is
-increasingly used in many substantive fields, such as medicine (see e.g.
-Pulkki-Råback et al. 2015; Alkire et al. 2015; Stinesen Kollberg et al.
-2015), psychology (e.g. Sorbi et al. 2014), economics (e.g. Wong et al.
-2015; Dafoe, Oneal, and Russett 2013; Allan and Forder 2015), sociology
-(e.g. Crowley and Kolenikov 2014; De Hoon and Van Tubergen 2014), and
-official statistics (e.g. Roderick J. Little 2012; Klausch, Hox, and
-Schouten 2015).
-
-The technique has earned a permanent spot in research and policy making,
+Multiple imputation (Donald B. Rubin 1987) is a state-of-the-art
+technique for drawing valid conclusions from incomplete data. The
+technique has earned a permanent spot in research and policy making,
 demonstrated e.g.~by the detailed manual created by the National
@@
 The idea behind multiple imputation is to impute (fill in) the missing
-values multiple times to obtain a valid estimate of what could have
-been. The multiple datasets that are thus obtained can be analyzed by
-standard techniques and the analysis results can be combined into a
-single inference. In contrast to ad hoc methods for dealing with missing
-values (e.g.~list-wise deletion, mean imputation, regression imputation,
-last observation carried forward), multiple imputation properly takes
-the sources of uncertainty that are related to the missingness problem
-into account.
+values in incomplete data multiple times, to obtain a valid estimate of
+what could have been observed. The multiple datasets that are thus
+obtained can be analyzed by standard techniques and the analysis results
+can be combined into a single inference. In contrast to ad hoc methods
+for dealing with missing values (e.g.~list-wise deletion, mean
+imputation, last observation carried forward), multiple imputation
+properly takes the sources of uncertainty that are related to the
+missingness problem into account. {[}It has been empirically
+demonstrated time and again that the inferences on multiply imputed data
+yield unbiased and confidence valid results (through Rubin's rules). But
+how do we know that this is the case, if we do not have access to the
+true but unobserved values of the missing entries? And how can we be
+sure that extensions to the methodology yield valid inferences too?
+Especially in the field of ML/AI, there are promising new imputation
+methods, which may yield even sharper imputations than now-standard
+(semi-)parametric imputation methods. That's why we use simulation
+studies for the evaluation of imputation routines.{]}
 
-The quality of a solution obtained by multiple imputation depends on the
-statistical properties of the incomplete data and the degree to which an
-imputation procedure is able to capture these properties when modeling
-missing values. In general it holds that modeling missing data becomes
-more challenging when the amount of missingness increases. However, when
-(strong) relations in the data are present, the observed parts can hold
-great predictive power for the models that estimate the missingness. In
-that case, multiple imputation would be substantially more efficient
-than the ubiquitous list-wise deletion.
+\sout{The quality of a solution obtained by multiple imputation depends
+on the statistical properties of the incomplete data and the degree to
+which an imputation procedure is able to capture these properties when
+modeling missing values. In general it holds that modeling missing data
+becomes more challenging when the amount of missingness increases.
+However, when (strong) relations in the data are present, the observed
+parts can hold great predictive power for the models that estimate the
+missingness. In that case, multiple imputation would be substantially
+more efficient than the ubiquitous list-wise deletion.}
 
 When evaluating the statistical properties (and thereby the practical
-applicability) of new imputation methodology, researchers most often
-make use of simulation studies. In such studies a complete data set is
-usually generated from a statistical model, another model is used to
-induce missingness, and a set of evaluation criteria is postulated to
-evaluate the performance of the imputation method. However, no golden
-standard has been established and, as a result, validity of the
-simulation and evaluation procedures may differ tremendously from one
-developer to another.
+applicability) of imputation methodology, researchers most often make
+use of simulation studies. In such studies a complete dataset is usually
+generated from a statistical model, another model is used to induce
+missingness, {[}then several missing data methods are applied{]} and a
+set of evaluation criteria is postulated to evaluate the performance of
+the imputation methods. However, no golden standard has been established
+to evaluate these imputation routines and, as a result, the validity of
+the simulation procedures may differ tremendously from one developer to
+another.
 
-The purpose of this paper is threefold: First, I raise some concerns
+The purpose of this paper is threefold: First, to raise some concerns
 with respect to evaluating imputation methodology. These concerns stem
 from careful consideration with fellow `imputers' and from encounters as
-a reviewer for statistical journals. Second, I provide imputation
+a reviewer for statistical journals. Second, to provide imputation
 methodologists with a suggested course of action when simulating missing
@@
 but is in no way intended as an absolute solution. This identifies the
-third purpose of this paper: discussion. I hope to elicit critical
+third purpose of this paper: discussion. We hope to elicit critical
 thinking with respect to the problems at hand. We are all convinced that
@@
 \hypertarget{why-some-evaluations-should-not-be-trusted}{%
-\subsection{Why some evaluations should not be
+\section{Why some evaluations should not be
 trusted}\label{why-some-evaluations-should-not-be-trusted}}
@@
 potential problems in the objective assessment of the performance of
-imputation routines that may lead to suboptimal use of multiple
+imputation routines that may lead to sub-optimal use of multiple
 imputation in practice. To demonstrate the broad impact of these
-problems, I subdivide the problems in the following three distinct
+problems, we subdivide the problems in the following three distinct
 categories: problems with data generation, problems with missingness
-generation and problems with performance evaluation. I further detail
+generation and problems with performance evaluation. we further detail
 the impact these problems may have on the validity of the performance
@@
 
-To evaluate the ability of an imputation routine to handle the
-non-response model, a form of truth has to be established. Those who
-perform simulation studies are in the luxury position to establish the
-truth beforehand, either by generating data from a statistical model, or
-by sampling data from a sufficiently large observed set. However, when
-generating data, two easily overlooked problems may arise.
+To evaluate the ability of an imputation routine to handle \sout{the
+non-response model} {[}missing data entries{]}, a form of truth has to
+be established. Those who perform simulation studies are in the luxury
+position to establish the truth beforehand, by choosing a data
+generating mechanism to obtain a complete dataset. Complete data is
+typically either generated from a statistical model, or sampled from a
+sufficiently large \sout{observed} set {[}model-based vs design-based
+simulation{]}. \sout{However, when generating data, two easily
+overlooked problems may arise.}
 
-First, data are often generated following the model (or distribution)
-that is also used for imputing the data. The performance of the
-imputation approach is then deemed good, which is no surprise as the
-evaluated conditions are in favor of the problem that is studied.
+The problem with model-based data generating mechanisms is that method
+performance may not translate to empirical data. First, data are often
+generated following the model (or distribution) that is also used for
+imputing the data. The performance of the imputation approach is then
+deemed good, which is no surprise as the evaluated conditions are in
+favor of the problem that is studied (Morris, White, and Crowther 2019).
 Although such procedures may be statistically relevant, the approach
-would be no good for even the simplest case. Such simulations pose
-serious threats to the applicability of the method as real-life data
-hardly ever follow a given theoretical distribution.
+would be no good for even the simplest empirical case. Such simulations
+pose serious threats to the applicability of the method as real-life
+data hardly ever follow a given theoretical distribution. {[}Add that
+data types may be more messy.{]} Second, data are often generated such
+that the problem that is being studied is most pronounced. This results
+in simulated data that contains very valuable information structures,
+i.e.~the correlations between groups of variables may be very
+pronounced. In other words, no matter what type of missingness is
+induced, the observed parts of the data may still hold much, if not all,
+of the information about the missing part and the performance of the
+imputation procedure is not surprisingly evaluated as very good. Also,
+it may be that the relations in real-life data are rarely as pronounced
+as the problem that is being studied.
 
-Second, data are often generated such that the problem that is being
-studied is most pronounced. This results in simulated data that contains
-very valuable information structures, i.e.~the correlations between
-groups of variables may be very pronounced. In other words, no matter
-what type of missingness is induced, the observed parts of the data may
-still hold much, if not all, of the information about the missing part
-and the performance of the imputation procedure is not surprisingly
-evaluated as very good. Also, it may be that the relations in real-life
-data are rarely as pronounced as the problem that is being studied.
+The problem with design-based simulation is that it's difficult to find
+large datasets without missing entries. And imputing the missing data
+beforehand and then treating the once-imputed data as comparative truth
+may favor the used imputation method in further evaluations. A second
+issue is that SEs may be incorrect if the (re)sampling is not handled
+correctly. {[}Add pooling when the sample is the population (Vink and
+van Buuren 2014) and write something about bootstrapping SEs{]}.
 
-\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
-\hline
- & $\mu_{Y} = 10$ & $\beta_{X} = 0.4$ & CIW & CR & RMSE$_{p}$ & RMSE$_{c}$\\
-\hline
-\multicolumn{7}{l}{\textbf{\$n = 50\$}}\\
-\hline
-\hspace{1em}CCA & 9.961 & 0.341 & 0.823 & 0.94 & 0.866 & NaN\\
-\hline
-\hspace{1em}Imp. & 10.007 & 0.386 & 0.877 & 0.99 & 0.903 & 1.330\\
-\hline
-\hspace{1em}True & 10.021 & 0.393 & 0.541 & 0.97 & 0.903 & 0.000\\
-\hline
-\multicolumn{7}{l}{\textbf{\$n = 500\$}}\\
-\hline
-\hspace{1em}CCA & 9.941 & 0.343 & 0.232 & 0.85 & 0.897 & NaN\\
-\hline
-\hspace{1em}Imp. & 9.994 & 0.392 & 0.249 & 0.94 & 0.917 & 1.296\\
-\hline
-\hspace{1em}True & 9.998 & 0.400 & 0.161 & 0.90 & 0.914 & 0.000\\
-\hline
-\multicolumn{7}{l}{\textbf{\$n = 5000\$}}\\
-\hline
-\hspace{1em}CCA & 9.945 & 0.344 & 0.074 & 0.09 & 0.899 & NaN\\
-\hline
-\hspace{1em}Imp. & 10.000 & 0.397 & 0.078 & 0.91 & 0.916 & 1.298\\
-\hline
-\hspace{1em}True & 10.002 & 0.398 & 0.051 & 0.95 & 0.917 & 0.000\\
-\hline
-\end{tabu}
+Problems with either of these data generation approaches is that the
+number of observations, the number of variables, the variable type(s)
+and the coherence between variables may give some imputation methods a
+comparative advantage, if they are not a) varied across simulation
+conditions {[}when is this fair? is this even an option?{]}, or b) the
+object of the study {[}add some examples of simulations specifically
+focused on small samples, categorical data, etc.{]}.
+
+{[}Maybe add that some simulation studies do not use a `ground truth' at
+all, they just look for the method with the best predictive performance
+for a certain completely observed target variable. I don't know where to
+introduce the difference between statistical inference and prediction.
+But this is an important third category, it seems. These studies apply
+different combinations of imputation and estimation methods on one or
+more benchmark datasets, and assess the RMSE of the predicted values
+after imputation and estimation to say something about the comparative
+performance of the methods. In this context, there is no talk of
+retrieving the `true' but unobserved values at all. Refer to missing
+data chapter (Liu et al. 2021).{]}
 
@@
 
-In order to obtain valid imputation inference, the imputation model must
-capture the essence of the true non-response mechanism Meng (1994). The
-model - if any - that is used to generate the missingness is usually
-assumed to be random (MAR) or completely random (MCAR). Missingness
-mechanisms that are not random (MNAR) and mechanisms that are considered
-non-ignorable see e.g. Donald B. Rubin (1976), are generally ignored
-during evaluation of imputation methodologies, except for those methods
-that are specifically targeted at non-ignorable applications.
+Missingness comes in many shapes and forms {[}platitude, but how else to
+start?{]}. The most obvious characteristics of missing data are the
+missingness pattern(s) and missingness mechanism(s). Missingness
+patterns concern the location of missing entries across variables in an
+incomplete dataset, whereas missingness mechanisms describe the
+relationship between missingness and the values of variables in the data
+(Little and Rubin 2020, 8).
+
+The missing data pattern may inadvertently guide which imputation method
+prevails in simulation studies. Some ad hoc methods are known to yield
+valid inferences in very specific missing data patterns (e.g., list-wise
+deletion when there is missingness in the outcome variable of the
+analysis model exclusively), while they bias inferences in any other
+case. Extrapolating simulation results from such a specific missing data
+pattern to more intricate empirical missingness patterns could lead to
+false conclusions {[}re-read this! and add monotone/univariate vs
+multivariate and sporadic vs systematic dichotomies{]}. Related to the
+missingness pattern across variables is the missingness proportion. The
+terminology on this topic can be confusing. Does a missingness
+proportion of 50\% mean that half of the entries in an incomplete
+dataset are missing, or that half of the rows have at least one missing
+entry? We will refer to the latter as the proportion of incomplete
+cases. But this is not a clear distinction in the literature, which may
+lead to incorrect conclusions {[}re-phrase this too{]}. Moreover, the
+amount of missingness that is used to simulate the performance of
+imputation methods differs between studies. {[}Add ref to chapter (Liu
+et al. 2021) and add info.{]} Some studies use only 10\% missing data
+where other studies push the limits to additionally investigate the
+performance under at least 50\% missing data. This inconsistent display
+of simulation results may impact the objectivity of meta-evaluations
+over imputation methods, as one method's performance may appear to be
+favorable because of the less stringent simulation conditions. This
+ultimately may lead to statisticians recommending a less efficient
+method to applied researchers, thereby limiting the efficiency of the
+imputation approach and unnecessarily lowering the statistical power.
+
+{[}Briefly introduce missingness mechanisms here and show potential
+problems, but without simulation results (moved to separate section).
+This part should contain: 1) definitions of mechanisms, 2) that MCAR is
+often ignored, while it's the minimum requirement and sometimes
+realistic in practice, 3) that MAR can be generated \emph{spuriously} if
+the correlation between variables is low, and 4) that MNAR is often
+ignored, while it may be the most realistic in practice. And finally
+something about types of M(N)AR, but what are specific problems with
+that?{]}
+
+\hypertarget{section}{%
+\section{}\label{section}}
+
+\hrule
+
+\hypertarget{section-1}{%
+\section{}\label{section-1}}
+
+{[}Until next hline: to be edited!{]} In order to obtain valid
+imputation inference, the imputation model must capture the essence of
+the true non-response mechanism Meng (1994). The model - if any - that
+is used to generate the missingness is usually assumed to be random
+(MAR) or completely random (MCAR). Missingness mechanisms that are not
+random (MNAR) and mechanisms that are considered non-ignorable see e.g.
+Donald B. Rubin (1976), are generally ignored during evaluation of
+imputation methodologies, except for those methods that are specifically
+targeted at non-ignorable applications.
 
@@
 after all, for every MNAR mechanism there is a MAR mechanism with equal
-fit Molenberghs et al. (2008) - it can be argued that MNAR is the more
+fit (Molenberghs et al. 2008) - it can be argued that MNAR is the more
 likely mechanism for real life missingness scenarios. In such cases, an
@@
 
-The amount of missingness that is used to simulate the performance of
-imputation methods differs between studies. Some studies use only 10\%
-missing data where other studies push the limits to additionally
-investigate the performance under at least 50\% missing data. This
-inconsistent display of simulation results may impact the objectivity of
-meta-evaluations over imputation methods, as one method's performance
-may appear to be favorable because of the less stringent simulation
-conditions. This ultimately may lead to statisticians recommending a
-less efficient method to applied researchers, thereby limiting the
-efficiency of the imputation approach and unnecessarily lowering the
-statistical power.
+\hypertarget{section-2}{%
+\section{}\label{section-2}}
+
+\hrule
 
@@
 
-The evaluation criteria used to assess imputation performance vary from
-one developer to another. This is not surprising as people from
-different fields could have a different focus on the problem at hand.
-However, because multiple imputation was designed as a mode for
-statistical inference, a minimal set of statistical properties should at
-least be evaluated. I highlight four aspects of evaluation that should
-be considered when evaluation imputation routines.
+{[}Add bridge from last section. And explain how this part is
+subdivided: first general characteristics of the imputations such as
+convergence, then performance measures.{]}
 
-First, developers often only inquire about the `accuracy' (i.e.~how well
-can the method reproduce the original data). The goal of multiple
-imputation is not to reproduce the data, but to allow for obtaining
-valid inference given that the data are incomplete. We are interested in
-the correct answer to the research question; not in the truth itself.
-This means that, given the framework provided by Donald B. Rubin (1987),
-statistical properties such as bias, confidence intervals and the
-coverage rate of the confidence intervals should be studied. After all,
-the 95\% confidence interval should contain the `true' value at least 95
-out of 100 times Neyman (1934), p.~591.
+{[}Performance of different imputation methods may be influenced by
+preconditions of the methods themselves. All imputation models should be
+congenial, but this is difficult to test to often ignored even in single
+imputation endeavor, let alone in simulation studies with 100s of
+imputations. Same goes for convergence: FCS is iterative and requires
+algorithmic convergence, but this is typically evaluated through visual
+inspection which is unfeasible in simulation studies. Ignoring
+congeniality may yield sub-optimal imputations, which may disadvantage
+certain methods compared to others.{]}
 
-Second, the performance of imputation procedures on distributional
+Moreover, the performance of imputation procedures on distributional
 properties is often ignored in simulation studies, and even though the
@@
 
-Third, sampling variation has always been an essential part of the
-evaluation of multiple imputation methodology. However, in order to
-obtain information about a method's ability to handle the missing data
-problem, or to objectively compare methods on their ability to correct
-for missingness, it is not necessary to take sampling variation into
-account Vink and van Buuren (2014). After all, we are interested only in
-the missing data mechanism, and are not considering the noise induced by
-the sampling mechanism for evaluation in such studies.
+{[}Then onto the performance measure part.{]}
+
+The evaluation criteria used to assess imputation performance vary from
+one developer to another. This is not surprising as people from
+different fields could have a different focus on the problem at hand.
+But the choice of performance measure(s) may inadvertently distort
+statistical properties of the imputed data.
+
+\sout{However, because multiple imputation was designed as a mode for
+statistical inference, a minimal set of statistical properties should at
+least be evaluated. We highlight four aspects of evaluation that should
+be considered when evaluation imputation routines.}
+
+Developers often only inquire about the `accuracy' (i.e.~how well can
+the method reproduce the original data). {[}Add RMSE at cell level
+here?{]} The goal of multiple imputation is not to reproduce the data,
+but to allow for obtaining valid inference given that the data are
+incomplete. We are interested in the correct answer to the research
+question; not in the truth itself. This means that, given the framework
+provided by Donald B. Rubin (1987), statistical properties such as bias,
+confidence intervals and the coverage rate of the confidence intervals
+should be studied. After all, the 95\% confidence interval should
+contain the `true' value at least 95 out of 100 times Neyman (1934),
+p.~591. {[}So, what could go wrong by focusing on accuracy
+exclusively?{]}
+
+{[}Add choice of analysis model/estimand here, since performance measure
+are defined wrt the estimand, e.g., focusing on univariate estimates may
+favor invalid ad hoc methods.{]}
+
+{[}Add problems with RMSE here.{]}
+
+Sampling variation has always been an essential part of the evaluation
+of multiple imputation methodology. However, in order to obtain
+information about a method's ability to handle the missing data problem,
+or to objectively compare methods on their ability to correct for
+missingness, it is not necessary to take sampling variation into account
+Vink and van Buuren (2014). After all, we are interested only in the
+missing data mechanism, and are not considering the noise induced by the
+sampling mechanism for evaluation in such studies.
 
@@
 relevance.
+
+\hypertarget{illustration}{%
+\section{Illustration}\label{illustration}}
+
+{[}Add simulation study + results here!{]}
 
@@
 to be considered during all of the simulations. One needs to realize,
-however, that the conventional pooling rules cf. Donald B. Rubin (1987),
-pp.~76-77, do not apply for finite population inference and that
-alternative pooling rules need to be used Raghunathan, Reiter, and Rubin
-(2003), Vink and van Buuren (2014).
+however, that the conventional pooling rules (cf. Donald B. Rubin 1987,
+76--77), do not apply for finite population inference and that
+alternative pooling rules need to be used (Raghunathan, Reiter, and
+Rubin 2003; Vink and van Buuren 2014).
 
@@
 routines For example, inference from hot-deck techniques such as
-predictive mean matching Roderick JA Little (1988), Donald B. Rubin
-(1986) may be more severely impacted by large amounts of one-tailed
-missingness than inference from parametric techniques. It would be a
-shame to overlook such results due to the focus on a single MAR
-mechanism.
+predictive mean matching (Roderick JA Little 1988; Donald B. Rubin 1986)
+may be more severely impacted by large amounts of one-tailed missingness
+than inference from parametric techniques. It would be a shame to
+overlook such results due to the focus on a single MAR mechanism.
 
@@
 \begin{itemize}
-\tightlist
 \item
@@
 \item
+  {[}Add RMSE 2x{]}
+\item
   \emph{Distributional characteristics:} In practice, the distribution
@@
   evaluating imputation routines, the evaluator should mention whether
-  the routine is prone to deliver implausible values.
+  the routine is prone to deliver implausible values. {[}Add PPC{]}
 \item
@@
   and if non-convergence is suspected, the inference resulting from the
-  imputations should not be considered.
+  imputations should not be considered. {[}Extend this part!{]}
+\item
+  {[}Add PPC (Mingyang) (Nguyen, Carlin, and Lee 2017), (Zhao 2022){]}
 \end{itemize}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{sessionInfo}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
 \begin{verbatim}
-## R version 4.1.1 (2021-08-10)
-## Platform: x86_64-w64-mingw32/x64 (64-bit)
-## Running under: Windows 10 x64 (build 19042)
+## R version 4.1.3 (2022-03-10)
+## Platform: aarch64-apple-darwin20 (64-bit)
+## Running under: macOS Monterey 12.3
 ## 
 ## Matrix products: default
+## BLAS:   /Library/Frameworks/R.framework/Versions/4.1-arm64/Resources/lib/libRblas.0.dylib
+## LAPACK: /Library/Frameworks/R.framework/Versions/4.1-arm64/Resources/lib/libRlapack.dylib
 ## 
 ## locale:
-## [1] LC_COLLATE=Dutch_Netherlands.1252  LC_CTYPE=Dutch_Netherlands.1252   
-## [3] LC_MONETARY=Dutch_Netherlands.1252 LC_NUMERIC=C                      
-## [5] LC_TIME=Dutch_Netherlands.1252    
+## [1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
 ## 
@@
 ## other attached packages:
-## [1] dplyr_1.0.7
+## [1] dplyr_1.0.8
 ## 
 ## loaded via a namespace (and not attached):
-##  [1] Rcpp_1.0.7        highr_0.9         pillar_1.6.4      compiler_4.1.1   
-##  [5] tools_4.1.1       digest_0.6.29     viridisLite_0.4.0 evaluate_0.14    
-##  [9] lifecycle_1.0.1   tibble_3.1.6      lattice_0.20-44   pkgconfig_2.0.3  
-## [13] rlang_0.4.12      rstudioapi_0.13   DBI_1.1.2         yaml_2.2.1       
-## [17] mvtnorm_1.1-3     xfun_0.29         fastmap_1.1.0     kableExtra_1.3.4 
-## [21] xml2_1.3.3        httr_1.4.2        withr_2.4.3       stringr_1.4.0    
-## [25] knitr_1.37        systemfonts_1.0.3 generics_0.1.1    vctrs_0.3.8      
-## [29] webshot_0.5.2     rprojroot_2.0.2   grid_4.1.1        tidyselect_1.1.1 
-## [33] svglite_2.0.0     glue_1.6.0        mice_3.14.0       here_1.0.1       
-## [37] R6_2.5.1          fansi_0.5.0       rmarkdown_2.11    purrr_0.3.4      
-## [41] tidyr_1.1.4       magrittr_2.0.1    codetools_0.2-18  scales_1.1.1     
-## [45] backports_1.4.1   ellipsis_0.3.2    htmltools_0.5.2   rvest_1.0.2      
-## [49] assertthat_0.2.1  colorspace_2.0-2  utf8_1.2.2        stringi_1.7.6    
-## [53] munsell_0.5.0     broom_0.7.12      crayon_1.4.2
+##  [1] rstudioapi_0.13  knitr_1.37       magrittr_2.0.2   tidyselect_1.1.2
+##  [5] R6_2.5.1         rlang_1.0.2      fastmap_1.1.0    fansi_1.0.2     
+##  [9] highr_0.9        stringr_1.4.0    tools_4.1.3      xfun_0.30       
+## [13] utf8_1.2.2       DBI_1.1.2        cli_3.2.0        htmltools_0.5.2 
+## [17] ellipsis_0.3.2   assertthat_0.2.1 yaml_2.3.5       digest_0.6.29   
+## [21] tibble_3.1.6     lifecycle_1.0.1  crayon_1.5.0     purrr_0.3.4     
+## [25] vctrs_0.3.8      glue_1.6.2       evaluate_0.15    rmarkdown_2.13  
+## [29] stringi_1.7.6    compiler_4.1.3   pillar_1.7.0     generics_0.1.2  
+## [33] pkgconfig_2.0.3
 \end{verbatim}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-alkire2015global}{}}%
-Alkire, Blake C, Nakul P Raykar, Mark G Shrime, Thomas G Weiser, Stephen
-W Bickler, John A Rose, Cameron T Nutt, et al. 2015. {``Global Access to
-Surgical Care: A Modelling Study.''} \emph{The Lancet Global Health} 3
-(6): e316--23.
-
-\leavevmode\vadjust pre{\hypertarget{ref-allan2015determinants}{}}%
-Allan, Stephen, and Julien Forder. 2015. {``The Determinants of Care
-Home Closure.''} \emph{Health Economics} 24 (S1): 132--45.
-
 \leavevmode\vadjust pre{\hypertarget{ref-brand1999development}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-crowley2014flexible}{}}%
-Crowley, Jocelyn Elise, and Stanislav Kolenikov. 2014. {``Flexible Work
-Options and Mothers' Perceptions of Career Harm.''} \emph{The
-Sociological Quarterly} 55 (1): 168--95.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dafoe2013democratic}{}}%
-Dafoe, Allan, John R Oneal, and Bruce Russett. 2013. {``The Democratic
-Peace: {Weighing} the Evidence and Cautious Inference.''}
-\emph{International Studies Quarterly} 57 (1): 201--14.
-
-\leavevmode\vadjust pre{\hypertarget{ref-de2014religiosity}{}}%
-De Hoon, Sean, and Frank Van Tubergen. 2014. {``The Religiosity of
-Children of Immigrants and Natives in England, Germany, and the
-Netherlands: {The} Role of Parents and Peers in Class.''} \emph{European
-Sociological Review}, jcu038.
-
-\leavevmode\vadjust pre{\hypertarget{ref-klausch2015selection}{}}%
-Klausch, Thomas, Joop Hox, and Barry Schouten. 2015. {``Selection Error
-in Single-and Mixed Mode Surveys of the {Dutch} General Population.''}
-\emph{Journal of the Royal Statistical Society: Series A (Statistics in
-Society)} 178 (4): 945--61.
-
 \leavevmode\vadjust pre{\hypertarget{ref-li2012imputing}{}}%
@@
 Discussion Paper} 1124.
-
-\leavevmode\vadjust pre{\hypertarget{ref-little2012calibrated}{}}%
-Little, Roderick J. 2012. {``Calibrated {Bayes}, an Alternative
-Inferential Paradigm for Official Statistics.''} \emph{Journal of
-Official Statistics} 28 (3): 309.
 
@@
 \emph{New England Journal of Medicine} 367 (14): 1355--60.
+
+\leavevmode\vadjust pre{\hypertarget{ref-litt20}{}}%
+Little, and Rubin. 2020. \emph{Statistical {Analysis} with {Missing
+Data}, {Third Edition} \textbar{} {Wiley Series} in {Probability} and
+{Statistics}}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-liu21}{}}%
+Liu, Dawei, Hanne I. Oberman, Johanna Muñoz, Jeroen Hoogland, and Thomas
+P. A. Debray. 2021. {``Quality Control, Data Cleaning, Imputation.''} In
+\emph{Clinical Applications of Artificial Intelligence in Real-World
+Data}. \url{https://arxiv.org/abs/2110.15877}.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pulkki2015cumulative}{}}%
-Pulkki-Råback, Laura, Marko Elovainio, Christian Hakulinen, Jari
-Lipsanen, Mirka Hintsanen, Markus Jokela, Laura D Kubzansky, et al.
-2015. {``Cumulative Effect of Psychosocial Factors in Youth on Ideal
-Cardiovascular Health in Adulthood the Cardiovascular Risk in Young
-Finns Study.''} \emph{Circulation} 131 (3): 245--53.
+\leavevmode\vadjust pre{\hypertarget{ref-nguy17}{}}%
+Nguyen, Cattram D., John B. Carlin, and Katherine J. Lee. 2017. {``Model
+Checking in Multiple Imputation: An Overview and Case Study.''}
+\emph{Emerging Themes in Epidemiology} 14 (1): 8.
+\url{https://doi.org/10.1186/s12982-017-0062-6}.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-sorbi2014medium}{}}%
-Sorbi, MJ, AM Kleiboer, HG van Silfhout, G Vink, and J Passchier. 2014.
-{``Medium-Term Effectiveness of Online Behavioral Training in Migraine
-Self-Management: {A} Randomized Trial Controlled over 10 Months.''}
-\emph{Cephalalgia : An International Journal of Headache},
-0333102414547137.
-
-\leavevmode\vadjust pre{\hypertarget{ref-stinesen2015reduced}{}}%
-Stinesen Kollberg, Karin, Ann-Charlotte Waldenström, Karin Bergmark,
-Gail Dunberger, Anna Rossander, Ulrica Wilderäng, Elisabeth
-Åvall-Lundqvist, and Gunnar Steineck. 2015. {``Reduced Vaginal
-Elasticity, Reduced Lubrication, and Deep and Superficial Dyspareunia in
-Irradiated Gynecological Cancer Survivors.''} \emph{Acta Oncologica} 54
-(5): 772--79.
-
 \leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wong2015economic}{}}%
-Wong, Frances Kam Yuet, Ching So, June Chau, Antony Kwan Pui Law,
-Stanley Ku Fu Tam, and Sarah McGhee. 2015. {``Economic Evaluation of the
-Differential Benefits of Home Visits with Telephone Calls and Telephone
-Calls Only in Transitional Discharge Support.''} \emph{Age and Ageing}
-44 (1): 143--47.
-
 \leavevmode\vadjust pre{\hypertarget{ref-zhao22}{}}%

### 2022-03-22 Simulation/Simulation.Rmd

@@
 
-- number of observations ($n = \{50, 500, 5000\}$) *only max required*
+- number of observations ($n = \{50, 500, 5000\}$) *[for efficiency: only max required?]*
 - number of variables ($p = \{2\}$)
@@
 - imputation method (`meth` = {none, mean imputation, regression imputation, stochastic regression})
-- number of imputations (`m` = {1, 5}) *only max required*
-- number of iterations (`maxit` = 1, 5, 10) *use mice.mids to shorten runtime?*
+- number of imputations (`m` = {1, 5}) *[for efficiency: only max required?]*
+- number of iterations (`maxit` = 1, 5, 10) *[for efficiency: use mice.mids to shorten runtime?]*
 

### 2022-03-22 Manuscript/StandardizedEvaluation.Rmd
Manuscript after discussion on suggestions Hanne, including first version of the checklist.

@@
 ---
-title: "Towards a standardized evaluation of multiple imputation routines"
+title: "Towards a standardized evaluation of imputation routines"
 author: "Hanne Oberman and Gerko Vink"
@@
 
-<!-- # Structure -->
-
-<!-- 1. introduction -->
-
-<!-- - aim -->
-<!-- - number of simulations -->
-
-<!-- 2. data generation -->
-
-<!-- - model vs design based -->
-<!-- - data types -->
-<!-- - sample vs population (vs bootstrapping??) -->
-<!-- - number of observations -->
-<!-- - number of variables -->
-<!-- - coherence between variables -->
-
-<!-- 3. missingness generation -->
-
-<!-- - monotone/univariate vs multivariate -->
-<!-- - missingness mechanism -->
-<!-- - type of M(N)AR -->
-<!-- - missingness proportion (vs proportion of incomplete cases! 50% incomplete == 25% missing in bivariate data) -->
-
-<!-- 4. missing data methods -->
-
-<!-- - imputation method -->
-<!-- - number of imputations -->
-<!-- - number of iterations -->
-
-<!-- 5. performance evaluation -->
-
-<!-- - estimand -->
-<!-- - diagnostic -->
-<!-- - ppc -->
-
 # Introduction
 
-Multiple imputation [@rubin1987] is a state-of-the-art technique for drawing valid conclusions from incomplete data.
+Imputation is a state-of-the-art technique for drawing valid conclusions from incomplete data.
 The technique has earned a permanent spot in research and policy making, demonstrated e.g.
 by the detailed manual created by the National Research Council [@little2012prevention].
-Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing multiple imputation techniques.
-After all, the principle of multiple imputation is very intuitive.
+Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing imputation techniques.
+After all, the principle of imputation is very intuitive.
 
-The idea behind multiple imputation is to impute (fill in) the missing values in incomplete data multiple times, to obtain a valid estimate of what could have been observed.
-The multiple datasets that are thus obtained can be analyzed by standard techniques and the analysis results can be combined into a single inference.
+The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.
+A popular variant is multiple imputation [@rubi76], wherein missing values are imputed multiple times.
+The data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference [using Rubin's rules; @rubi87].
 In contrast to ad hoc methods for dealing with missing values (e.g.
-list-wise deletion, mean imputation, last observation carried forward), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
-[It has been empirically demonstrated time and again that the inferences on multiply imputed data yield unbiased and confidence valid results (through Rubin's rules).
+list-wise deletion, mean imputation, regression imputation, last observation carried forward, indicator method), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
+
+<!-- [It has been empirically demonstrated time and again that the inferences on multiply imputed data yield unbiased and confidence valid results (through Rubin's rules).
 But how do we know that this is the case, if we do not have access to the true but unobserved values of the missing entries?
@@
 Especially in the field of ML/AI, there are promising new imputation methods, which may yield even sharper imputations than now-standard (semi-)parametric imputation methods.
-That's why we use simulation studies for the evaluation of imputation routines.]
+That's why we use simulation studies for the evaluation of imputation routines.] -->
 
-~~The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
+The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
 In general it holds that modeling missing data becomes more challenging when the amount of missingness increases.
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
-In that case, multiple imputation would be substantially more efficient than the ubiquitous list-wise deletion.~~
+In that case, (multiple) imputation would be substantially more efficient than the ubiquitous list-wise deletion.
 
@@
 However, no golden standard has been established to evaluate these imputation routines and, as a result, the validity of the simulation procedures may differ tremendously from one developer to another.
+[Nowadays there are a lot of imp alg being propagated, esp ai/ml, whose perf often times seems quite promising.
+But may not be a fair comparison because of the simulation set-up/dev in ai/ml so fast that steps may be skipped]
 
@@
 
-To evaluate the ability of an imputation routine to handle ~~the non-response model~~ [missing data entries], a form of truth has to be established.
+To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established.
 Those who perform simulation studies are in the luxury position to establish the truth beforehand, by choosing a data generating mechanism to obtain a complete dataset.
-Complete data is typically either generated from a statistical model, or sampled from a sufficiently large ~~observed~~ set [model-based vs design-based simulation].
-~~However, when generating data, two easily overlooked problems may arise.~~
+Complete data is typically either generated from a statistical model, or sampled from a sufficiently large observed set [model-based vs design-based simulation].
 
@@
 Refer to missing data chapter [@liu21].]
+
+<!-- als je perf based sims doet, dan kijk je wel daarnaar. -->
 
@@
 [Briefly introduce missingness mechanisms here and show potential problems, but without simulation results (moved to separate section).
-This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and sometimes realistic in practice, 3) that MAR can be generated *spuriously* if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
-And finally something about types of M(N)AR, but what are specific problems with that?]
+This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated *spuriously* if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
+And finally something about types of M(N)AR, but what are specific problems with that?
+ref dance paper rianne]
 
@@
 
-~~However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated.
-We highlight four aspects of evaluation that should be considered when evaluation imputation routines.~~
-
 Developers often only inquire about the 'accuracy' (i.e.
@@
 [Add choice of analysis model/estimand here, since performance measure are defined wrt the estimand, e.g., focusing on univariate estimates may favor invalid ad hoc methods.]
+
+<!-- als inf het doel is, dan moet je de varianties juist krijgen dus kom je niet weg met m = 1 (want geen onderscheid tussen obs en imp) -->
 
@@
 A method may perform badly, but if it still outperforms every other approach, it may yet be of great practical relevance.
-
-# Illustration
-
-[Add simulation study + results here!]
 
@@
 ## Step 3. Evaluate performance
+
 It is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
@@
 However, the real-world missingness would then be omitted.
+
+## Step 4. Report
+
+<!-- [Checklist voor iedere sim studie met alle parameters die we bespreken] -->
+
+<!-- - doel -->
+
+<!-- - 11 punten van stef -->
+
+<!-- - andere punten uit dit paper -->
+
+<!-- - welke variables/features vormen imp model en welke analysis model (congenial, meng 94) -->
+
+<!-- - cross validation? -->
+
+<!-- als pfd met check boxes? -->
+
+Report at least the following:
+
+1. set-up
+
+- aim
+- number of simulations
+
+2. data generation
+
+- model vs design based
+- data types
+- sample vs population (vs bootstrapping??)
+- number of observations
+- number of variables
+- coherence between variables
+
+3. missingness generation
+
+- monotone/univariate vs multivariate
+- missingness mechanism
+- type of M(N)AR
+- missingness proportion
+
+4. missing data methods
+
+- imputation method
+- imputation model
+- number of imputations
+- number of iterations
+
+5. performance evaluation
+
+- analysis model
+- estimand
+- diagnostic
+- ppc
 

### 2022-03-22 Manuscript/StandardizedEvaluation.tex
Manuscript after discussion on suggestions Hanne, including first version of the checklist.

@@
 \hypersetup{
-  pdftitle={Towards a standardized evaluation of multiple imputation routines},
+  pdftitle={Towards a standardized evaluation of imputation routines},
   pdfauthor={Hanne Oberman and Gerko Vink},
@@
 \makeatother
-\usepackage[normalem]{ulem}
-% Avoid problems with \sout in headers with hyperref
-\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 
-\title{Towards a standardized evaluation of multiple imputation
-routines}
+\title{Towards a standardized evaluation of imputation routines}
 \author{Hanne Oberman and Gerko Vink}
@@
 
-Multiple imputation (Donald B. Rubin 1987) is a state-of-the-art
-technique for drawing valid conclusions from incomplete data. The
-technique has earned a permanent spot in research and policy making,
-demonstrated e.g.~by the detailed manual created by the National
-Research Council (Roderick J. Little et al. 2012). Although a top-down
-enforcement of valid ways to handle missing data is not yet very
-pronounced, an increasing amount of researchers are embracing multiple
-imputation techniques. After all, the principle of multiple imputation
+Imputation is a state-of-the-art technique for drawing valid conclusions
+from incomplete data. The technique has earned a permanent spot in
+research and policy making, demonstrated e.g.~by the detailed manual
+created by the National Research Council (Roderick J. Little et al.
+2012). Although a top-down enforcement of valid ways to handle missing
+data is not yet very pronounced, an increasing amount of researchers are
+embracing imputation techniques. After all, the principle of imputation
 is very intuitive.
 
-The idea behind multiple imputation is to impute (fill in) the missing
-values in incomplete data multiple times, to obtain a valid estimate of
-what could have been observed. The multiple datasets that are thus
-obtained can be analyzed by standard techniques and the analysis results
-can be combined into a single inference. In contrast to ad hoc methods
+The idea behind imputation is to impute (fill in) missing values, to
+obtain a valid estimate of what could have been. A popular variant is
+multiple imputation (Donald B. Rubin 1976), wherein missing values are
+imputed multiple times. The data that are thus obtained can be analyzed
+by standard techniques and, in the case of multiple imputation, the
+analysis results can be combined into a single inference {[}using
+Rubin's rules; Donald B. Rubin (1987){]}. In contrast to ad hoc methods
 for dealing with missing values (e.g.~list-wise deletion, mean
-imputation, last observation carried forward), multiple imputation
-properly takes the sources of uncertainty that are related to the
-missingness problem into account. {[}It has been empirically
-demonstrated time and again that the inferences on multiply imputed data
-yield unbiased and confidence valid results (through Rubin's rules). But
-how do we know that this is the case, if we do not have access to the
-true but unobserved values of the missing entries? And how can we be
-sure that extensions to the methodology yield valid inferences too?
-Especially in the field of ML/AI, there are promising new imputation
-methods, which may yield even sharper imputations than now-standard
-(semi-)parametric imputation methods. That's why we use simulation
-studies for the evaluation of imputation routines.{]}
+imputation, regression imputation, last observation carried forward,
+indicator method), multiple imputation properly takes the sources of
+uncertainty that are related to the missingness problem into account.
 
-\sout{The quality of a solution obtained by multiple imputation depends
-on the statistical properties of the incomplete data and the degree to
-which an imputation procedure is able to capture these properties when
-modeling missing values. In general it holds that modeling missing data
-becomes more challenging when the amount of missingness increases.
-However, when (strong) relations in the data are present, the observed
-parts can hold great predictive power for the models that estimate the
-missingness. In that case, multiple imputation would be substantially
-more efficient than the ubiquitous list-wise deletion.}
+The quality of a solution obtained by imputation depends on the
+statistical properties of the incomplete data and the degree to which an
+imputation procedure is able to capture these properties when modeling
+missing values. In general it holds that modeling missing data becomes
+more challenging when the amount of missingness increases. However, when
+(strong) relations in the data are present, the observed parts can hold
+great predictive power for the models that estimate the missingness. In
+that case, (multiple) imputation would be substantially more efficient
+than the ubiquitous list-wise deletion.
 
@@
 the simulation procedures may differ tremendously from one developer to
-another.
+another. {[}Nowadays there are a lot of imp alg being propagated, esp
+ai/ml, whose perf often times seems quite promising. But may not be a
+fair comparison because of the simulation set-up/dev in ai/ml so fast
+that steps may be skipped{]}
 
@@
 
-To evaluate the ability of an imputation routine to handle \sout{the
-non-response model} {[}missing data entries{]}, a form of truth has to
-be established. Those who perform simulation studies are in the luxury
-position to establish the truth beforehand, by choosing a data
-generating mechanism to obtain a complete dataset. Complete data is
-typically either generated from a statistical model, or sampled from a
-sufficiently large \sout{observed} set {[}model-based vs design-based
-simulation{]}. \sout{However, when generating data, two easily
-overlooked problems may arise.}
+To evaluate the ability of an imputation routine to handle missingness,
+a form of truth has to be established. Those who perform simulation
+studies are in the luxury position to establish the truth beforehand, by
+choosing a data generating mechanism to obtain a complete dataset.
+Complete data is typically either generated from a statistical model, or
+sampled from a sufficiently large observed set {[}model-based vs
+design-based simulation{]}.
 
@@
 This part should contain: 1) definitions of mechanisms, 2) that MCAR is
-often ignored, while it's the minimum requirement and sometimes
-realistic in practice, 3) that MAR can be generated \emph{spuriously} if
-the correlation between variables is low, and 4) that MNAR is often
-ignored, while it may be the most realistic in practice. And finally
-something about types of M(N)AR, but what are specific problems with
-that?{]}
+often ignored, while it's the minimum requirement and reference, and
+sometimes realistic in practice, 3) that MAR can be generated
+\emph{spuriously} if the correlation between variables is low, and 4)
+that MNAR is often ignored, while it may be the most realistic in
+practice. And finally something about types of M(N)AR, but what are
+specific problems with that? ref dance paper rianne{]}
 
@@
 
-\sout{However, because multiple imputation was designed as a mode for
-statistical inference, a minimal set of statistical properties should at
-least be evaluated. We highlight four aspects of evaluation that should
-be considered when evaluation imputation routines.}
-
 Developers often only inquire about the `accuracy' (i.e.~how well can
@@
 relevance.
-
-\hypertarget{illustration}{%
-\section{Illustration}\label{illustration}}
-
-{[}Add simulation study + results here!{]}
 
@@
 
+\hypertarget{step-4.-report}{%
+\subsection{Step 4. Report}\label{step-4.-report}}
+
+Report at least the following:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  set-up
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  aim
+\item
+  number of simulations
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  data generation
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  model vs design based
+\item
+  data types
+\item
+  sample vs population (vs bootstrapping??)
+\item
+  number of observations
+\item
+  number of variables
+\item
+  coherence between variables
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  missingness generation
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  monotone/univariate vs multivariate
+\item
+  missingness mechanism
+\item
+  type of M(N)AR
+\item
+  missingness proportion
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  missing data methods
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  imputation method
+\item
+  imputation model
+\item
+  number of imputations
+\item
+  number of iterations
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
+\tightlist
+\item
+  performance evaluation
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  analysis model
+\item
+  estimand
+\item
+  diagnostic
+\item
+  ppc
+\end{itemize}
+
 \hypertarget{discussion}{%
@@
 \begin{verbatim}
-## R version 4.1.3 (2022-03-10)
-## Platform: aarch64-apple-darwin20 (64-bit)
-## Running under: macOS Monterey 12.3
+## R version 4.1.2 (2021-11-01)
+## Platform: x86_64-w64-mingw32/x64 (64-bit)
+## Running under: Windows 10 x64 (build 19042)
 ## 
 ## Matrix products: default
-## BLAS:   /Library/Frameworks/R.framework/Versions/4.1-arm64/Resources/lib/libRblas.0.dylib
-## LAPACK: /Library/Frameworks/R.framework/Versions/4.1-arm64/Resources/lib/libRlapack.dylib
 ## 
 ## locale:
-## [1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
+## [1] LC_COLLATE=Dutch_Netherlands.1252  LC_CTYPE=Dutch_Netherlands.1252   
+## [3] LC_MONETARY=Dutch_Netherlands.1252 LC_NUMERIC=C                      
+## [5] LC_TIME=Dutch_Netherlands.1252    
 ## 
@@
 ##  [5] R6_2.5.1         rlang_1.0.2      fastmap_1.1.0    fansi_1.0.2     
-##  [9] highr_0.9        stringr_1.4.0    tools_4.1.3      xfun_0.30       
+##  [9] highr_0.9        stringr_1.4.0    tools_4.1.2      xfun_0.30       
 ## [13] utf8_1.2.2       DBI_1.1.2        cli_3.2.0        htmltools_0.5.2 
@@
 ## [25] vctrs_0.3.8      glue_1.6.2       evaluate_0.15    rmarkdown_2.13  
-## [29] stringi_1.7.6    compiler_4.1.3   pillar_1.7.0     generics_0.1.2  
+## [29] stringi_1.7.6    compiler_4.1.2   pillar_1.7.0     generics_0.1.2  
 ## [33] pkgconfig_2.0.3

### 2022-03-22 Simulation/Simulation.Rmd
Manuscript after discussion on suggestions Hanne, including first version of the checklist.

@@
 
-# Simulation functions
+# Simulation set-up
+
+Set-up environment:
+
+```{r env}
+# random seed
+set.seed(2022)
+# packages
+library(dplyr)
+# utils functions
+rmse <- function(obs, pred = 0){sqrt(mean((obs-pred)^2))}
+```
+
+## Simulation functions
 
@@
 ```{r dat}
-generation <- function(r, n = c(50, 500, 5000), p = 2) {
+generation <- function(r = 0.4, n = 500, p = 2) {
   # generate data based on variance-covariance matrix
@@
   diag(vcov) <- 1
-  dat <- data.frame(mvtnorm::rmvnorm(
-    n = max(n),
+  dat <- data.frame(
+    mvtnorm::rmvnorm(
+    n,
     mean = rep(0, p),
@@
   )) %>% setNames(paste0("X", 1:p))
-  # split output by number of observations
-  out <- purrr::map(n, ~{
-    list(n_obs = .x,
-         p = p,
-         r = r,
-         dat = dat[1:.x,])
-  })
+  # output
+  out <- list(
+    n_obs = n,
+    n_var = p,
+    corr = r,
+    dat = dat)
   return(out)
@@
 ```{r amp}
-amputation <- function(dat, mech, type, prop) {
+amputation <- function(datlist, mech = "MAR", type = "MID", prop = 0.25) {
   # induce missingness
-  amp <- dat %>%
-    mice::ampute(prop = prop,
+  amp <- mice::ampute(datlist$dat, 
+                 prop = prop,
                  mech = mech,
@@
   # output
-  out <- amp
+  out <- list(
+    n_obs = datlist$n_obs,
+    n_var = datlist$n_var,
+    corr = datlist$corr,
+    mis_mech = mech,
+    mis_type = type,
+    mis_prop = prop,
+    amp = amp)
   return(out)
@@
 ```{r imp}
-imputation <- function(amp, meth, m = 5, it = 10) {
-  imp <- amp$amp %>%
-    mice::mice(
+imputation <- function(amplist, meth = "norm", m = 5, it = 5) {
+  # for CCA
+  if(meth == "none") {
+    imp <- na.omit(amplist$amp$amp)
+    average <- mean(imp$X1)
+    rmse_cell <- NA
+  } 
+  # for actual imputation methods
+  if (meth != "none") {
+    if (meth == "mean" | meth == "norm.predict") {
+      imp <- mice::mice(
+        amplist$amp$amp,
+        m = 1,
+        method = meth,
+        maxit = 1,
+        print = FALSE
+      )
+    } else {
+    imp <- mice::mice(
+        amplist$amp$amp,
         m = m,
@@
       )
+    }
+    average <- mice::complete(imp, "long") %>% .$X1 %>% mean()
+    rmse_cell <- apply(imp$imp$X1 - amplist$amp$data$X1[imp$where[, "X1"]], 2, rmse) %>% mean()
+  }
   # output
-  out <- imp
+  out <- list(
+    n_obs = amplist$n_obs,
+    n_var = amplist$n_var,
+    corr = amplist$corr,
+    mis_mech = amplist$mis_mech,
+    mis_type = amplist$mis_type,
+    mis_prop = amplist$mis_prop,    
+    imp_meth = meth,
+    n_imp = m,
+    n_it = it,
+    average = average,
+    rmse_cell = rmse_cell,
+    imp = imp
+  )
   return(out)
@@
 ```{r perf}
+evaluation <- function(implist){
+  # for CCA
+  if (is.data.frame(implist$imp)) {
+    fit <- lm(X1 ~ ., implist$imp)
+    rmse_pred <- rmse(fit$residuals)
+  }
+  # for actual imputation methods
+  if (mice::is.mids(implist$imp)) {
+    mira <- mice::complete(implist$imp, "all") %>% 
+      purrr::map(~lm(X1 ~ ., .))
+    fit <- suppressWarnings(mice::pool(mira))
+    rmse_pred <- purrr::map_dbl(mira, ~rmse(.$residuals)) %>% mean()
+  }
+  # get estimated regression coefficients
+  est <- fit %>% broom::tidy(conf.int = TRUE) %>%
+    .[2, c("estimate", "conf.low", "conf.high")] 
+  # output
+  out <- data.frame(
+    n_obs = implist$n_obs,
+    n_var = implist$n_var,
+    corr = implist$corr,
+    mis_mech = implist$mis_mech,
+    mis_type = implist$mis_type,
+    mis_prop = implist$mis_prop,    
+    imp_meth = implist$imp_meth,
+    n_imp = implist$n_imp,
+    n_it = implist$n_it,
+    average = implist$average, 
+    coeff = est$estimate, 
+    ci_ll = est$conf.low,
+    ci_ul = est$conf.high,
+    rmse_pred = rmse_pred, 
+    rmse_cell = implist$rmse_cell
+  )
+  return(out)
+}
+```
 
+The combined function will consist of these steps:
+
+```{r test_default}
+dat <- generation()
+amp <- amputation(dat)
+imp <- imputation(amp)
+est <- evaluation(imp)
 ```
+
+## Simulation conditions
+
+The full simulation function will be applied to these conditions:
+
+```{r cond}
+# data generation parameters
+n_obs = c(50, 500, 5000) 
+corr = c(0, .4, .8)
+dat_par <- expand.grid(n_obs = n_obs, corr = corr)
+# amputation parameters
+mis_mech = c("MCAR", "MAR", "MNAR")
+mis_type = c("LEFT", "RIGHT", "MID", "TAIL")
+mis_prop = c(.1, .25, .5)
+amp_par <- expand.grid(dat_nr = 1:nrow(dat_par), mis_mech = mis_mech, mis_type = mis_type, mis_prop = mis_prop, stringsAsFactors = FALSE)
+# imputation parameters
+imp_meth = c("none", "mean", "norm.predict", "norm") 
+n_imp = c(1, 5)
+n_it = c(1, 5, 10)
+imp_par <- expand.grid(amp_nr = 1:nrow(amp_par), imp_meth = imp_meth, n_imp = n_imp, n_it = n_it, stringsAsFactors = FALSE)
+```
+
+To generate the incomplete data for each condition we use:
+
+```{r test_cond, eval=FALSE, include=FALSE}
+all_dat <- purrr::pmap(dat_par, ~{generation(n = ..1, r = ..2)})
+all_amp <- purrr::pmap(amp_par, ~{amputation(all_dat[[..1]], mech = ..2, type = ..3, prop = ..4)})
+all_imp <- purrr::pmap_dfr(imp_par, ~{imputation(all_amp[[..1]], meth = ..2, m = ..3, it = ..4) %>% 
+    evaluation()})
+```

### 2022-03-22 OldTex/Standardized evaluation.tex

--- 2022-03-07 Manuscript/StandardizedEvaluation.Rmd
@@
----
-title: "Towards a standardized evaluation of multiple imputation routines"
-author: "Hanne Oberman and Gerko Vink"
-date: "19-1-2022"
-abstract: "Developing new imputation methodology has become a very active field.
+\documentclass[12pt, a4paper]{article}
+\usepackage{amssymb, amsmath, amsthm}
+\usepackage{multirow}
+\usepackage{color}
+\usepackage[sort&compress]{natbib}
+\usepackage{hyperref}
+\usepackage{graphicx}
+\usepackage{setspace}
+\hypersetup{
+    pdftitle={Towards a standardized evaluation of multiple imputation routines}
+    pdfauthor={Gerko Vink and Andrew Gelman},
+    pdfsubject={Evaluating imputation routines},
+    pdfkeywords={Multiple imputation, evaluation, simulation, finite population},
+    bookmarksnumbered=true,     
+    bookmarksopen=true,         
+    bookmarksopenlevel=1,       
+    colorlinks=false,
+    citecolor=blue,
+    linkcolor=blue,
+    urlcolor=blue,                 
+    pdfstartview=Fit,           
+    pdfpagemode=UseOutlines,      
+    pdfpagelayout=TwoPageRight
+}
+
+\singlespacing
+
+\begin{document}
+\title{Towards a standardized evaluation of multiple imputation routines}
+
+\author{Gerko Vink}
+%\author{Gerko Vink$^{1}$\thanks{Correspondence to: Gerko Vink, Department of Methodology and Statistics, Utrecht University, PO Box 80140, 3508 TC Utrecht, the Netherlands.
+E-mail: \texttt{G.Vink@uu.nl}}\; and Andrew Gelman$^{2}$\\
+%\newline\\
+%\small{\textsl{$^1$Department of Methodology and Statistics, Utrecht University}\protect}\\
+%\small{\textsl{$^2$Department of Statistics, Columbia University}\protect}\\
+%\newline}
+\date{}
+\maketitle
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ABSTRACT
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{abstract}
+Developing new imputation methodology has become a very active field.
 Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
-In this paper we propose a move towards a standardized evaluation of imputation methods.
-To demonstrate the need for standardization, we highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
+In this paper I propose a move towards a standardized evaluation of imputation methods.
+To demonstrate the need for standardization, I highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
 This may lead to suboptimal use of multiple imputation in practice.
-Additionally, we suggest a course of action for simulating and evaluating missing data problems."
-bibliography: StandardizedEvaluation.bib  
-output: pdf_document
-editor_options: 
-  chunk_output_type: inline
----
+Additionally, I suggest a course of action for simulating and evaluating missing data problems.
+\end{abstract}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
-library(dplyr)
-# source(here::here("R/utils.R"))
-# source(here::here("R/run_sim.R"))
-```
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% INTRODUCTION
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Introduction}
+\label{intro}
 
-<!-- # Structure -->
+Multiple imputation \citep{rubin1987} is a state-of-the-art technique for drawing valid conclusions from incomplete data, and is increasingly used in many substantive fields, such as medicine \citep[see e.g.][]{pulkki2015cumulative,alkire2015global,stinesen2015reduced}, psychology \citep[e.g.][]{sorbi2014medium}, economics \citep[e.g.][]{wong2015economic,dafoe2013democratic,allan2015determinants}, sociology \citep[e.g.][]{crowley2014flexible,de2014religiosity} and official statistics \citep[e.g.][]{little2012calibrated,klausch2015selection}.
 
-<!-- 1. introduction -->
-
-<!-- - aim -->
-<!-- - number of simulations -->
-
-<!-- 2. data generation -->
-
-<!-- - model vs design based -->
-<!-- - data types -->
-<!-- - sample vs population (vs bootstrapping??) -->
-<!-- - number of observations -->
-<!-- - number of variables -->
-<!-- - coherence between variables -->
-
-<!-- 3. missingness generation -->
-
-<!-- - monotone/univariate vs multivariate -->
-<!-- - missingness mechanism -->
-<!-- - type of M(N)AR -->
-<!-- - missingness proportion (vs proportion of incomplete cases! 50% incomplete == 25% missing in bivariate data) -->
-
-<!-- 4. missing data methods -->
-
-<!-- - imputation method -->
-<!-- - number of imputations -->
-<!-- - number of iterations -->
-
-<!-- 5. performance evaluation -->
-
-<!-- - estimand -->
-<!-- - diagnostic -->
-<!-- - ppc -->
-
-# Introduction
-
-Multiple imputation [@rubin1987] is a state-of-the-art technique for drawing valid conclusions from incomplete data.
 The technique has earned a permanent spot in research and policy making, demonstrated e.g.
-by the detailed manual created by the National Research Council [@little2012prevention].
+by the detailed manual created by the National Research Council \citep*{little2012prevention}.
 Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing multiple imputation techniques.
@@
 
-The idea behind multiple imputation is to impute (fill in) the missing values in incomplete data multiple times, to obtain a valid estimate of what could have been observed.
+The idea behind multiple imputation is to impute (fill in) the missing values multiple times to obtain a valid estimate of what could have been.
 The multiple datasets that are thus obtained can be analyzed by standard techniques and the analysis results can be combined into a single inference.
 In contrast to ad hoc methods for dealing with missing values (e.g.
-list-wise deletion, mean imputation, last observation carried forward), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
-[It has been empirically demonstrated time and again that the inferences on multiply imputed data yield unbiased and confidence valid results (through Rubin's rules).
-But how do we know that this is the case, if we do not have access to the true but unobserved values of the missing entries?
-And how can we be sure that extensions to the methodology yield valid inferences too?
-Especially in the field of ML/AI, there are promising new imputation methods, which may yield even sharper imputations than now-standard (semi-)parametric imputation methods.
-That's why we use simulation studies for the evaluation of imputation routines.]
+listwise deletion, mean imputation, regression imputation, last observation carried forward), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
 
-~~The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
+The quality of a solution obtained by multiple imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
 In general it holds that modeling missing data becomes more challenging when the amount of missingness increases.
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
-In that case, multiple imputation would be substantially more efficient than the ubiquitous list-wise deletion.~~
+In that case, multiple imputation would be substantially more efficient than the ubiquitous listwise deletion.
 
-When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies.
-In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, [then several missing data methods are applied] and a set of evaluation criteria is postulated to evaluate the performance of the imputation methods.
-However, no golden standard has been established to evaluate these imputation routines and, as a result, the validity of the simulation procedures may differ tremendously from one developer to another.
+When evaluating the statistical properties (and thereby the practical applicability) of new imputation methodology, researchers most often make use of simulation studies.
+In such studies a complete data set is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of the imputation method.
+However, no golden standard has been established and, as a result, validity of the simulation and evaluation procedures may differ tremendously from one developer to another.
 
-The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology.
-These concerns stem from careful consideration with fellow 'imputers' and from encounters as a reviewer for statistical journals.
-Second, to provide imputation methodologists with a suggested course of action when simulating missing data problems.
+The purpose of this paper is threefold: First, I raise some concerns with respect to evaluating imputation methodology.
+These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals.
+Second, I provide imputation methodologists with a suggested course of action when simulating missing data problems.
 This suggested approach should identify a common ground, but is in no way intended as an absolute solution.
 This identifies the third purpose of this paper: discussion.
-We hope to elicit critical thinking with respect to the problems at hand.
+I hope to elicit critical thinking with respect to the problems at hand.
 We are all convinced that our methodology has some merit.
@@
 
-# Why some evaluations should not be trusted
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% INTRODUCTION
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Why some evaluations should not be trusted}
 As of today, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
 Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers.
-This brings forth a chain of potential problems in the objective assessment of the performance of imputation routines that may lead to sub-optimal use of multiple imputation in practice.
-To demonstrate the broad impact of these problems, we subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation.
-we further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.
+This brings forth a chain of potential problems in the objective assessment of the performance of imputation routines that may lead to suboptimal use of multiple imputation in practice.
+To demonstrate the broad impact of these problems, I subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation.
+I further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.
 
-## Data generation problems
+\subsection*{Data generation problems}
+To evaluate the ability of an imputation routine to handle the nonresponse model, a form of truth has to be established.
+Those who perform simulation studies are in the luxury position to establish the truth beforehand, either by generating data from a statistical model, or by sampling data from a sufficiently large observed set.
+However, when generating data, two easily overlooked problems may arise.
 
-To evaluate the ability of an imputation routine to handle ~~the non-response model~~ [missing data entries], a form of truth has to be established.
-Those who perform simulation studies are in the luxury position to establish the truth beforehand, by choosing a data generating mechanism to obtain a complete dataset.
-Complete data is typically either generated from a statistical model, or sampled from a sufficiently large ~~observed~~ set [model-based vs design-based simulation].
-~~However, when generating data, two easily overlooked problems may arise.~~
+First, data are often generated following the model (or distribution) that is also used for imputing the data.
+The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied.
+Although such procedures may be statistically relevant, the approach would be no good for even the simplest case.
+Such simulations pose serious threats to the applicability of the method as real-life data hardly ever follow a given theoretical distribution.
 
-The problem with model-based data generating mechanisms is that method performance may not translate to empirical data.
-First, data are often generated following the model (or distribution) that is also used for imputing the data.
-The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied [@morr18].
-Although such procedures may be statistically relevant, the approach would be no good for even the simplest empirical case.
-Such simulations pose serious threats to the applicability of the method as real-life data hardly ever follow a given theoretical distribution.
-[Add that data types may be more messy.] Second, data are often generated such that the problem that is being studied is most pronounced.
+Second, data are often generated such that the problem that is being studied is most pronounced.
 This results in simulated data that contains very valuable information structures, i.e.
@@
 
-The problem with design-based simulation is that it's difficult to find large datasets without missing entries.
-And imputing the missing data beforehand and then treating the once-imputed data as comparative truth may favor the used imputation method in further evaluations.
-A second issue is that SEs may be incorrect if the (re)sampling is not handled correctly.
-[Add pooling when the sample is the population [@vink14] and write something about bootstrapping SEs].
-
-Problems with either of these data generation approaches is that the number of observations, the number of variables, the variable type(s) and the coherence between variables may give some imputation methods a comparative advantage, if they are not a) varied across simulation conditions [when is this fair?
-is this even an option?], or b) the object of the study [add some examples of simulations specifically focused on small samples, categorical data, etc.].
-
-[Maybe add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
-I don't know where to introduce the difference between statistical inference and prediction.
-But this is an important third category, it seems.
-These studies apply different combinations of imputation and estimation methods on one or more benchmark datasets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
-In this context, there is no talk of retrieving the 'true' but unobserved values at all.
-Refer to missing data chapter [@liu21].]
-
-```{r sim_nobs, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, paged.print=TRUE}
-# # setup
-# set.seed(1)
-# n_sim <- 100
-# 
-# sim_nobs <- purrr::map_dfr(1:n_sim, function(.i){
-#   purrr::map_dfr(c(50, 500, 5000), ~{
-#     run(n_obs = .x) %>% cbind(sim = .i, .)})}) 
-# 
-# # gtsummary::tbl_summary(sim_nobs, by = dat)
-# # https://haozhu233.github.io/kableExtra/awesome_table_in_html.html#Group_rows_via_labeling
-# sim_nobs %>% 
-#   group_by(n_obs, dat) %>% 
-#   summarise(
-#     mean_Y = mean(ave),
-#     beta_X = mean(est),
-#     CIW = mean(ciw),
-#     CR = mean(cov),
-#     RMSE_pred = mean(rmse_pred),
-#     RMSE_cell = mean(rmse_cell, na.rm = TRUE))  %>% 
-#   .[, -1] %>% 
-#   knitr::kable(
-#     col.names = c("", "$\\mu_{Y} = 10$", "$\\beta_{X} = 0.4$", "CIW", "CR", "RMSE$_{p}$", "RMSE$_{c}$"), 
-#     digits = 3,
-#     escape = FALSE) %>% 
-#   kableExtra::kable_classic(full_width = TRUE) %>% 
-#   kableExtra::pack_rows(group_label = "$n = 50$", 1, 3, escape = FALSE) %>% 
-#   kableExtra::pack_rows(group_label = "$n = 500$", 4, 6, escape = FALSE) %>% 
-#   kableExtra::pack_rows(group_label = "$n = 5000$", 7, 9, escape = FALSE)
-```
-
-## Missingness generation problems
-
-Missingness comes in many shapes and forms [platitude, but how else to start?].
-The most obvious characteristics of missing data are the missingness pattern(s) and missingness mechanism(s).
-Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data [@litt20, p.
-8].
-
-<!-- L&R, p.
-8: "We find it useful to distinguish the missingness pattern, which describes which values are missing and observed in the data matrix and the missingness mechanism (or mechanisms), which concerns the relationship between missingness and the values of variables in the data matrix".-->
-
-The missing data pattern may inadvertently guide which imputation method prevails in simulation studies.
-Some ad hoc methods are known to yield valid inferences in very specific missing data patterns (e.g., list-wise deletion when there is missingness in the outcome variable of the analysis model exclusively), while they bias inferences in any other case.
-Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions [re-read this!
-and add monotone/univariate vs multivariate and sporadic vs systematic dichotomies].
-Related to the missingness pattern across variables is the missingness proportion.
-The terminology on this topic can be confusing.
-Does a missingness proportion of 50% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
-We will refer to the latter as the proportion of incomplete cases.
-But this is not a clear distinction in the literature, which may lead to incorrect conclusions [re-phrase this too].
-Moreover, the amount of missingness that is used to simulate the performance of imputation methods differs between studies.
-[Add ref to chapter [@liu21] and add info.] Some studies use only 10% missing data where other studies push the limits to additionally investigate the performance under at least 50% missing data.
-This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
-This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
-
-[Briefly introduce missingness mechanisms here and show potential problems, but without simulation results (moved to separate section).
-This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and sometimes realistic in practice, 3) that MAR can be generated *spuriously* if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
-And finally something about types of M(N)AR, but what are specific problems with that?]
-
-# 
-\hrule
-
-#
-[Until next hline: to be edited!] In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng94.
+\subsection*{Missingness generation problems}
+In order to obtain valid imputation inference, the imputation model must capture the essence of the true nonresponse mechanism \citep{meng1994multiple}.
 The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
-Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable see e.g.
-@rubi76, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
+Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubin1976inference}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
 
@@
 Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in my view may be far too simplistic.
-
-```{r mis_mech, echo=FALSE, fig.cap="\\label{fig:mis_mech}Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
-Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\\rho= 0$."}
-knitr::include_graphics("img/plot_mmech.pdf")
-```
+\begin{figure}[t!]
+  \resizebox{\textwidth}{!}{
+    \includegraphics[scale=1]{figures/plot_mmech.pdf}
+  }%
+  \caption{Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
+Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$.}
+  \label{fig:mis.mech}
+\end{figure}
 
@@
 This problem is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.
-Figure \ref{fig:mis_mech} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
+Figure \ref{fig:mis.mech} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
 
@@
 the conditional probabilities remain the same over the level of the incomplete variable, the performance is comparable to that under MCAR.
-In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation @fimd, p.
-48: the missingness does not depend on the incomplete variable.
+In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p.
+48]{fimd}: the missingness does not depend on the incomplete variable.
 This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.
 
-Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit [@molenberghs2008every] - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
-In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis see e.g.
-@molenberghs2014handbook, part 5.
+Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every} - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
+In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis \citep[see e.g.][part 5]{molenberghs2014handbook}.
 
-```{r sim_mech_corr, echo=FALSE, cache=TRUE, paged.print=TRUE}
-# # setup
-# set.seed(1)
-# n_sim <- 10
-# correlations <- c(0, .4, .8)
-# mechanisms <- c("MCAR", "MAR", "MNAR") 
-# 
-# sim_mech <- purrr::map_dfr(1:n_sim, function(.i){
-#    purrr::map_dfr(correlations, function(.c){
-#      purrr::map_dfr(mechanisms, function(.m){
-#        run(corr = .c, mis_mech = .m) %>% cbind(sim = .i, .)})})})
-# sim_mech %>% 
-#   group_by(mis_mech, corr) %>% 
-#   summarise(
-#     mean_cca = mean(mean_cca - 10),
-#     mean_est = mean(mean_est - 10),
-#     beta_cca = mean(beta_cca - corr),
-#     beta_est = mean(beta_est - corr),
-#     beta_ciw = mean(beta_ciw),
-#     beta_cr = mean(beta_cr),
-#     rmse_pred = mean(rmse_pred),
-#     rmse_cell = mean(rmse_cell))
-```
+% latex table generated in R 3.2.2 by xtable 1.8-0 package
+% Thu Feb  4 16:27:54 2016
+\begin{table}[t]
+\caption{Inferences obtained by complete case analysis (CCA) and multiple imputation over 1000 simulations.
+Displayed are the bias, coverage rate (cov) of the 95\% confidence interval and the confidence interval width (ciw).
+Imputations are generated by Bayesian linear regression imputation.}
+\centering
+\begin{tabular}{rlccccccc}
+  \hline
+ 	& &\multicolumn{3}{c}{CCA} &&\multicolumn{3}{c}{imputation}	\\\cline{3-5}\cline{7-9}
+ \multicolumn{2}{c}{mechanism}	& bias & cov & ciw && bias & cov & ciw \\ 
+  \hline
+  $[\rho = .8]$ &mcar & 0.001 & 0.959 & 0.179 && -0.001 & 0.946 & 0.156 \\ 
+  $[\rho = .8]$ &mar & {\color{red} 0.334} & {\color{red} 0.000} & 0.169 && 0.004 & 0.955 & 0.195 \\ 
+  $[\rho = .0]$ &mar & 0.003 & 0.973 &{\color{red} 0.180} && 0.013 & 0.947 & 0.311 \\    \hline
+\end{tabular}
+\label{tab:mis.mech}
+\end{table}
 
-<!-- table -->
-<!-- Inferences obtained by complete case analysis (CCA) and multiple imputation over 1000 simulations.
-Displayed are the bias, coverage rate (cov) of the 95% confidence interval and the confidence interval width (ciw).
-Imputations are generated by Bayesian linear regression imputation.
--->
+The amount of missingness that is used to simulate the performance of imputation methods differs between studies.
+Some studies use only 10\% missing data where other studies push the limits to additionally investigate the performance under at least 50\% missing data.
+This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
+This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
 
-# 
-\hrule
+\subsection*{Performance evaluation}
+The evaluation criteria used to assess imputation performance vary from one developer to another.
+This is not surprising as people from different fields could have a different focus on the problem at hand.
+However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated.
+I highlight four aspects of evaluation that should be considered when evaluation imputation routines.
 
-## Performance evaluation
+First, developers often only inquire about the `accuracy' (i.e.
+how well can the method reproduce the original data).
+The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
+We are interested in the correct answer to the research question; not in the truth itself.
+This means that, given the framework provided by \citep{rubin1987}, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
+After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p.
+591]{neyman1934two}.
 
-[Add bridge from last section. And explain how this part is subdivided: first general characteristics of the imputations such as convergence, then performance measures.]
-
-[Performance of different imputation methods may be influenced by preconditions of the methods themselves.
-All imputation models should be congenial, but this is difficult to test to often ignored even in single imputation endeavor, let alone in simulation studies with 100s of imputations.
-Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies.
-Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.]
-
-Moreover, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
+Second, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
 For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large.
@@
 
-[Then onto the performance measure part.]
-
-The evaluation criteria used to assess imputation performance vary from one developer to another.
-This is not surprising as people from different fields could have a different focus on the problem at hand.
-But the choice of performance measure(s) may inadvertently distort statistical properties of the imputed data.
-
-~~However, because multiple imputation was designed as a mode for statistical inference, a minimal set of statistical properties should at least be evaluated.
-We highlight four aspects of evaluation that should be considered when evaluation imputation routines.~~
-
-Developers often only inquire about the 'accuracy' (i.e.
-how well can the method reproduce the original data).
-[Add RMSE at cell level here?] The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-We are interested in the correct answer to the research question; not in the truth itself.
-This means that, given the framework provided by @rubi87, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
-After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neym34, p.
-591.
-[So, what could go wrong by focusing on accuracy exclusively?]
-
-[Add choice of analysis model/estimand here, since performance measure are defined wrt the estimand, e.g., focusing on univariate estimates may favor invalid ad hoc methods.]
-
-[Add problems with RMSE here.]
-
-Sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
-However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink14.
+Third, sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
+However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink2014pooling}.
 After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
@@
 
-# Illustration
-
-[Add simulation study + results here!]
-
-# Suggested course of action
-
-## Step 1. Obtain truth
-
+\section*{Suggested course of action}
+\subsection*{Step 1. Obtain truth}
 The evaluator must to decide whether there is a need for sampling variance in the simulations scheme. If sampling variance cannot be omitted from the simulation scheme, two approaches are possible:
-
-- Model-based simulation: draw samples from a known probability distribution, such as the multivariate normal distribution.
+\begin{enumerate}
+  \item Model-based simulation: draw samples from a known probability distribution, such as the multivariate normal distribution.
 The theoretical parameters under which the samples are obtained will serve as the comparative truth during the simulations.
-- Design-based simulation: sample data from a sufficiently large set.
+  \item Design-based simulation: sample data from a sufficiently large set.
 The parameters of the sufficiently large set will serve as the comparative truth during the simulations.
@@
 A benefit of design-based simulation is the ability to use real-life observed data structures.
-
+\end{enumerate}
 For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches.
@@
 This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations.
-One needs to realize, however, that the conventional pooling rules [cf.
-@rubin1987, pp.
-76-77], do not apply for finite population inference and that alternative pooling rules need to be used [@raghunathan2003multiple; @vink2014pooling].
+One needs to realize, however, that the conventional pooling rules \citep[cf.][p.
+76-77]{rubin1987} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink2014pooling}.
 
-## Step 2. Induce missingness
-
+\subsection*{Step 2. Induce missingness}
 First, one should always consider MCAR missingness, i.e.
@@
 Next, missing data should be induced conform a model that is dependent on the observed data.
-A straightforward technique for inducing different forms of univariate MAR missingness is described in @fimd, p.
-63; see Figure \ref{fig:MAR} and a generalization to multivariate MAR missingness can be found in @buur06, Appendix B, and @brand1999development, par.
-5.2.3.
-If the missingness is to be induced in longitudinal data, auto-regressive MAR models e.g.
-cf.
-@shara2015randomly, model 2 and model 3, can be useful.
+A straightforward technique for inducing different forms of univariate MAR missingness is described in \citet[][p.
+63; see Figure \ref{fig:mar}]{fimd} and a generalization to multivariate MAR missingness can be found in \citet[][Appendix B]{van2006fully} and \citet[][\S 5.2.3]{brand1999development}.
+If the missingness is to be induced in longitudinal data, autoregressive MAR models \citep[e.g.
+cf.][model 2 and model 3]{shara2015randomly} can be useful.
 
-```{r MAR, echo=FALSE, fig.cap="\\label{fig:MAR}Examples of four random missingness mechanisms that yield approximately 50 percent missingness.
-For negatively correlated missingness covariates the resulting mechanisms are reversed."}
-knitr::include_graphics("img/plot_mar.pdf")
-```
-
+\begin{figure}[t!]
+\begin{center}
+  \resizebox{\textwidth}{!}{
+    \includegraphics[scale=1]{figures/plot_mar.pdf}
+  }%
+  \caption{Examples of four random missingness mechanisms that yield approximately 50 percent missingness. For negatively correlated missingness covariates the resulting mechanisms are reversed. }
+  \label{fig:mar}
+  \end{center}
+\end{figure}
 Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
 Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
-This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching [@little1988missing; @rubin1986statistical] may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
 It would be a shame to overlook such results due to the focus on a single MAR mechanism.
@@
 
-- 10% missingness: Depending on the size of the data, this percentage can be considered as a lower bound of realistic evaluation.
-Anything less than 10% may be of little influence on the true data inference.
+\begin{itemize}
+  \item 10\% missingness: Depending on the size of the data, this percentage can be considered as a lower bound of realistic evaluation.
+Anything less than 10\% may be of little influence on the true data inference.
 Performance of a missing data method should at least be acceptable for most missing data problems.
-- 25% missingness: This is a fair amount of missingness and will, depending on the observed data information, have a noticeable influence on the completed data inference.
-When compared to the condition with 10% missingness, the inference obtained under 25% missingness should be less certain (i.e.
+  \item 25\% missingness: This is a fair amount of missingness and will, depending on the observed data information, have a noticeable influence on the completed data inference.
+When compared to the condition with 10\% missingness, the inference obtained under 25\% missingness should be less certain (i.e.
 confidence/credibility interval width should increase), but estimates should still be properly covered and the statistical properties of the missing data method should be sound.
-In practice, at least to my experience in social sciences and official statistics, 25% univariate missingness can easily be considered as a realistic missingness percentage.
-- 50% missingness: Performance under 50% percent simulated missingness will most likely be impacted severely.
-Depending on how the missingness mechanism interacts with the simulated data, some imputation techniques may yield estimates that are under-covered such that the completed data inference should not be deemed valid anymore.
-If a method yields acceptable inference under 50% MAR missingness, we can determine that the statistical properties of the imputation methodology are sound.
-
-Although I limit the focus here to ignorable non-response, the above suggested proportions are equally applicable to simulations under non-ignorable non-response.
+In practice, at least to my experience in social sciences and official statistics, 25\% univariate missingness can easily be considered as a realistic missingness percentage.
+  \item 50\% missingness: Performance under 50\% percent simulated missingness will most likely be impacted severely.
+Depending on how the missingness mechanism interacts with the simulated data, some imputation techniques may yield estimates that are undercovered such that the completed data inference should not be deemed valid anymore.
+If a method yields acceptable inference under 50\% MAR missingness, we can determine that the statistical properties of the imputation methodology are sound.
+\end{itemize}
+Although I limit the focus here to ignorable nonresponse, the above suggested proportions are equally applicable to simulations under non-ignorable nonresponse.
 
@@
 Often, when inducing missingness, authors remain vague about the actual missingness mechanism under investigation and, even worse, some authors only report something like
-
-> *We generated missing data following a MAR missingness mechanism.*
-
+\begin{quote}
+\emph{We generated missing data following a MAR missingness mechanism.}
+\end{quote}
 This should be considered unacceptable as claims about the validity of the multiple imputation inference depends heavily on the simulated missingness mechanism.
 
-## Step 3. Evaluate performance
-It is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
+\subsection*{Step 3. Evaluate performance}
+It is wise to evaluate the performance of complete case analysis (aka listwise deletion) in all simulated conditions.
 We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance.
@@
 In general, I would say that each multiple imputation routine should be evaluated on at least the following points:
-
-- \emph{Bias:} Results should preferable be unbiased.
+\begin{itemize}
+  \item \emph{Bias:} Results should preferable be unbiased.
 However, the way bias is considered can greatly influence the interpretation of the results.
@@
 The way bias is computed should therefore be carefully chosen and described.
-- \emph{Interval width:} The way the confidence interval is calculated should be described.
+  \item \emph{Interval width:} The way the confidence interval is calculated should be described.
 Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference.
 However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.
-- \emph{Coverage:} Coverage of a 95% interval should in theory be $\geq 95$, where a coverage of 95% would be most efficient.
-Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
-- [Add RMSE 2x]
-
-- \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data.
+  \item \emph{Coverage:} Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage of 95\% would be most efficient.
+Undercoverage (when estimation is too liberal) may be an indication of invalid inference, while overcoverage (when estimation is too conservative) tells us that efficiency could still be gained.
+  \item \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data.
 Under anything but the MCAR assumption, this can be expected.
-When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed see @abayomi2008diagnostics for an detailed overview of diagnostic evaluation for multivariate imputations.
+When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for an detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}.
 When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.
-- \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference.
+  \item \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference.
 However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property.
 When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
-[Add PPC]
-- \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers, PIGS, @li2012imputing.
+  \item \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
 The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
-[Extend this part!]
-- [Add PPC (Mingyang) [@nguy17], [@zhao22]]
-
+  \end{itemize}
 For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
@@
 
-# Discussion
-
+\section*{Discussion}
 This document is aimed at establishing a common ground for the evaluation of imputation routines.
@@
 
-# Appendix A: Simulation setup
-
-Let $Y=(Y_{obs},Y_{mis})$ be an incomplete semi-continuous variable, where $Y_{obs}$ and $Y_{mis}$ denote the observed values and the missing values in $Y$, respectively.
+\section*{Appendix A: Simulation setup}
+Let $Y=(Y_{obs},Y_{mis})$ be an incomplete semicontinuous variable, where $Y_{obs}$ and $Y_{mis}$ denote the observed values and the missing values in $Y$, respectively.
 Further, $X=(X_1,...,X_k)$ is a set of $k$ fully observed covariates, where $X_{obs}$ and $X_{mis}$ correspond to the observed an missing parts in $Y$ and where $X_j$ would indicate the $j$th variable in $X$, with $j=1,\dots, k$.
@@
 To prove the point, I generated a single data set from the multivariate normal distribution with means
-$$
+\begin{equation*}\label{meanmatrix}
 \mu= \bordermatrix{&	 \cr
@@
 { X}_2	&10	},
-$$
+\end{equation*}
 and covariance matrix
-$$
+\begin{equation*}\label{covmatrix}
 \Sigma = \bordermatrix{& Y &{ X}_1 &{ X}_2	 \cr
@@
 { X}_2	&0	&0	&1}.
-$$
+\end{equation*}
 
@@
 First, I left the observed data \emph{observed at random} to impose MCAR missingness following the mechanism
-$$
+\begin{equation*}
 P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0)=.50.
-$$
+\end{equation*}
 Next, I created random missingness in $Y$ according to the following MAR missingness mechanism
-$$
+\begin{equation*}
 P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0|Y_{obs}, X_j),
-$$
+\end{equation*}
 by using a random draw from a binomial distribution of the same length as $Y$ and of size 1 with missingness probability equal to the inverse logit
-$$
+\begin{equation*}
 P(R=0)=\frac{e^{a}}{(1+e^{a})}.
-$$
-where $a=(\mu_{X_j}-X_{ij})/\sqrt{\mathbb{E}[X_{ij} - \mu_{X_j}]}$ in order to create approximately 50% right-tailed MAR missingness. I followed this procedure twice; once for each $X_j$.
+\end{equation*}
+where $a=(\mu_{X_j}-X_{ij})/\sqrt{\mathbb{E}[X_{ij} - \mu_{X_j}]}$ in order to create approximately 50\% right-tailed MAR missingness. I followed this procedure twice; once for each $X_j$.
 
-The three resulting incomplete data sets were then imputed with package `mice` @mice, v2.25, in `R` @R, v3.2.3, with Bayesian linear regression imputation (`mice.impute.norm`) as the imputation routine.
-The completed-data analyses were combined into a single inference following the rules described in @vink2014pooling.
+The three resulting incomplete data sets were then imputed with package \texttt{mice} \citep[][v2.25]{mice} in \texttt{R} \citep[][v3.2.3]{R}, with Bayesian linear regression imputation (\texttt{mice.impute.norm}) as the imputation routine.
+The completed-data analyses were combined into a single inference following the rules described in \citet{vink2014pooling}.
 
-```{r}
-sessionInfo()
-
-```
-
-# References
+\bibliography{bibliography}
+\bibliographystyle{apalike}
+\end{document}

### 2022-03-22 R/run_sim.R
Clean up repo

@@ -1,134 +1,0 @@
-# simulation study for illustration
-
-# set-up 
-set.seed(123)
-library(dplyr)
-source("R/utils.R")
-n_sim = 2
-
-## simulation parameters ##
-
-# conditions
-n_obs = c(50, 500, 5000) 
-corr = c(0, .4, .8)
-mis_mech = c("MCAR", "MAR", "MNAR")
-mis_type = c("LEFT", "RIGHT", "MID", "TAIL")
-mis_prop = c(.1, .25, .5)
-n_imp = c(1, 5)
-imp_meth = c("mean", "norm.predict", "norm") 
-n_it = c(1, 5, 10)
-
-# combine into lists for mapping
-dat_par <- expand.grid(n_obs = n_obs, corr = corr)
-amp_par <- expand.grid(dat_nr = 1:nrow(dat_par), mis_mech = mis_mech, mis_type = mis_type, mis_prop = mis_prop, stringsAsFactors = FALSE)
-imp_par <- expand.grid(amp_nr = 1:nrow(amp_par), imp_meth = imp_meth, n_imp = n_imp, n_it = n_it, stringsAsFactors = FALSE)
-
-
-## simulation functions ##
-
-# generate data
-generation <- function(n, r){
-  dat <- data.frame(mvtnorm::rmvnorm(
-    n = n,
-    mean = c(10, 10),
-    sigma = matrix(c(1, r, r, 1), 2, 2)
-  )) %>% setNames(c("Y", "X"))
-}
-
-# ampute the data
-amputation <- function(dat, mech, type, prop){
-  amp <- dat %>% 
-  mice::ampute(
-    prop = prop,
-    mech = mech,
-    type = type
-  )}
-
-# impute the data
-imputation <- function(amp, meth, m, it){
-  imp <- amp$amp %>%
-  mice::mice(
-    m = m,
-    method = meth,
-    maxit = it,
-    print = FALSE
-  )}
-
-# evaluate the imputations
-evaluation <- function(imp, amp, dat, r){
-  # fit analysis model
-  fit_imps <- with(imp, lm(Y ~ X))
-  fits <- list(full = lm(Y~X, dat),
-               cca = lm(Y~X, amp$amp, na.action = na.omit),
-               imp = suppressWarnings(mice::pool(fit_imps)))
-  # performance
-  out <- purrr::map_dfr(fits, ~{
[…]

### 2022-03-22 R/utils.R

@@ -1,6 +1,0 @@
-# utils functions
-
-`%nin%` <- Negate(`%in%`)
-logistic <- function(x){exp(x)/(1+exp(x))} 
-rmse <- function(obs, pred = 0){sqrt(mean((obs-pred)^2))}
-

### 2022-03-22 R/visualize.R

@@ -1,55 +1,0 @@
-# create figures for the manuscript
-
-# set-up
-set.seed(123)
-library(dplyr)
-library(ggplot2)
-library(patchwork)
-source("R/utils.R")
-
-# missingness mechanism and correlation 
-# data generating mechanism
-sigma <- matrix(0, 3, 3)
-diag(sigma) <- 1
-sigma[1, 2] <- sigma[2, 1] <- 0.8 
-dat <- data.frame(mvtnorm::rmvnorm(500, mean = c(5, 10, 10), sigma = sigma)) %>% setNames(c("Y", "X1", "X2"))
-amp.mcar.8 <- mice::ampute(dat[, c("Y", "X1")], prop = .5, mech = "MCAR")$amp %>% setNames(c("Y", "X"))
-amp.mcar.0 <- mice::ampute(dat[, c("Y", "X2")], prop = .5, mech = "MCAR")$amp %>% setNames(c("Y", "X"))
-amp.mar.8 <- mice::ampute(dat[, c("Y", "X1")], prop = .5, mech = "MAR")$amp %>% setNames(c("Y", "X"))
-amp.mar.0 <- mice::ampute(dat[, c("Y", "X2")], prop = .5, mech = "MAR")$amp %>% setNames(c("Y", "X"))
-
-a <- dat[dat$Y != is.na(amp.mcar.8$Y), ] %>% 
-  ggplot() +
-  geom_point(aes(x = X1, y = Y), color = mice:::mdc(2), alpha = 0.25) +
-  geom_point(aes(x = X, y = Y), color = mice:::mdc(1), alpha = 0.5, data = amp.mcar.8) +
-  theme_classic()
-b <- dat[dat$Y != is.na(amp.mcar.0$Y), ] %>% 
-  ggplot() +
-  geom_point(aes(x = X2, y = Y), color = mice:::mdc(2), alpha = 0.25) +
-  geom_point(aes(x = X, y = Y), color = mice:::mdc(1), alpha = 0.5, data = amp.mcar.0) +
-  theme_classic()
-c <- dat[dat$Y != is.na(amp.mar.8$Y), ] %>% 
-  ggplot() +
-  geom_point(aes(x = X1, y = Y), color = mice:::mdc(2), alpha = 0.25) +
-  geom_point(aes(x = X, y = Y), color = mice:::mdc(1), alpha = 0.5, data = amp.mar.8) +
-  theme_classic()
-d <- dat[dat$Y != is.na(amp.mar.0$Y), ] %>% 
[…]

### 2022-03-22 Formatted/Formatted.tex

--- 2022-03-22 Manuscript/StandardizedEvaluation.Rmd
@@
----
-title: "Towards a standardized evaluation of imputation routines"
-author: "Hanne Oberman and Gerko Vink"
-date: "19-1-2022"
-abstract: "Developing new imputation methodology has become a very active field.
+\documentclass[bimj,fleqn]{w-art}
+\usepackage{times}
+\usepackage{w-thm}
+\usepackage[authoryear]{natbib}
+\setlength{\bibsep}{2pt}
+\setlength{\bibhang}{2em}
+\newcommand{\J}{J\"{o}reskog}
+\newcommand{\So}{S\"{o}rbom}
+\newcommand{\bcx}{{\bf X}}
+\newcommand{\bcy}{{\bf Y}}
+\newcommand{\bcz}{{\bf Z}}
+\newcommand{\bcu}{{\bf U}}
+\newcommand{\bcv}{{\bf V}}
+\newcommand{\bcw}{{\bf W}}
+\newcommand{\bci}{{\bf I}}
+\newcommand{\bch}{{\bf H}}
+\newcommand{\bcb}{{\bf B}}
+\newcommand{\bcr}{{\bf R}}
+\newcommand{\bcm}{{\bf M}}
+\newcommand{\bcf}{{\bf F}}
+\newcommand{\bcg}{{\bf G}}
+\newcommand{\bcs}{{\bf S}}
+\newcommand{\bca}{{\bf A}}
+\newcommand{\bcd}{{\bf D}}
+\newcommand{\bcc}{{\bf C}}
+\newcommand{\bce}{{\bf E}}
+\newcommand{\ba}{{\bf a}}
+\newcommand{\bb}{{\bf b}}
+\newcommand{\bc}{{\bf c}}
+\newcommand{\bd}{{\bf d}}
+\newcommand{\bx}{{\bf x}}
+\newcommand{\by}{{\bf y}}
+\newcommand{\bz}{{\bf z}}
+\newcommand{\bu}{{\bf u}}
+\newcommand{\bv}{{\bf v}}
+\newcommand{\bh}{{\bf h}}
+\newcommand{\bl}{{\bf l}}
+\newcommand{\be}{{\bf e}}
+\newcommand{\br}{{\bf r}}
+\newcommand{\bw}{{\bf w}}
+\newcommand{\de}{\stackrel{D}{=}}
+\newcommand{\bt}{\bigtriangleup}
+\newcommand{\bfequiv}{\mbox{\boldmath $\equiv$}}
+\newcommand{\bmu}{\mbox{\boldmath $\mu$}}
+\newcommand{\bnu}{\mbox{\boldmath $\nu$}}
+\newcommand{\bxi}{\mbox{\boldmath $\xi$}}
+\newcommand{\btau}{\mbox{\boldmath $\tau$}}
+\newcommand{\bgamma}{\mbox{\boldmath $\Gamma$}}
+\newcommand{\bphi}{\mbox{\boldmath $\Phi$}}
+\newcommand{\bfphi}{\mbox{\boldmath $\varphi$}}
+\newcommand{\bfeta}{\mbox{\boldmath $\eta$}}
+\newcommand{\bpi}{\mbox{\boldmath $\Pi$}}
+\newcommand{\bequiv}{\mbox{\boldmath $\equiv$}}
+\newcommand{\bvarepsilon}{\mbox{\boldmath $\varepsilon$}}
+\newcommand{\btriangle}{\mbox{\boldmath $\triangle$}}
+\newcommand{\bdelta}{\mbox{\boldmath $\Delta$}}
+\newcommand{\beps}{\mbox{\boldmath $\epsilon$}}
+\newcommand{\btheta}{\mbox{\boldmath $\theta$}}
+\newcommand{\balpha}{\mbox{\boldmath $\alpha$}}
+\newcommand{\bsphi}{\mbox{\boldmath $\varphi$}}
+\newcommand{\bsig}{\mbox{\boldmath $\sigma$}}
+\newcommand{\bfpsi}{\mbox{\boldmath $\psi$}}
+\newcommand{\bfdelta}{\mbox{\boldmath $\delta$}}
+\newcommand{\bsigma}{{\bf \Sigma}}
+\newcommand{\bzero}{{\bf 0}}
+\newcommand{\bpsi}{\mbox{\boldmath $\Psi$}}
+\newcommand{\bep}{\mbox{\boldmath $\epsilon$}}
+\newcommand{\bomega}{\mbox{\boldmath $\Omega$}}
+\newcommand{\bfomega}{\mbox{\boldmath $\omega$}}
+\newcommand{\blambda}{\mbox{\boldmath $\Lambda$}}
+\newcommand{\bflambda}{\mbox{\boldmath $\lambda$}}
+\newcommand{\bfsigma}{\mbox{\boldmath $\sigma$}}
+\newcommand{\bfpi}{{\mbox{\boldmath $\pi$}}}
+\newcommand{\bupsilon}{\mbox{\boldmath $\upsilon$}}
+\newcommand{\obs}{{\rm obs}}
+\newcommand{\mis}{{\rm mis}}
+\theoremstyle{plain}
+\newtheorem{criterion}{Criterion}
+\theoremstyle{definition}
+\newtheorem{condition}[theorem]{Condition}
+\usepackage[]{graphicx}
+\chardef\bslash=`\\ % p. 424, TeXbook
+\newcommand{\ntt}{\normalfont\ttfamily}
+\newcommand{\cn}[1]{{\protect\ntt\bslash#1}}
+\newcommand{\pkg}[1]{{\protect\ntt#1}}
+\let\fn\pkg
+\let\env\pkg
+\let\opt\pkg
+\hfuzz1pc % Don't bother to report overfull boxes if overage is < 1pc
+\newcommand{\envert}[1]{\left\lvert#1\right\rvert}
+\let\abs=\envert
+
+\begin{document}
+%\DOIsuffix{bimj.DOIsuffix}
+\DOIsuffix{bimj.200100000}
+\Volume{52}
+\Issue{61}
+\Year{2010}
+\pagespan{1}{}
+\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\[1pc]
+\noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
+}  %%% semicolon and fullpoint added here for keyword style
+
+\title[Standardized evaluation]{Towards a standardized evaluation of imputation routines}
+%% Information for the first author.
+\author[Oberman]{Hanne I. Oberman\footnote{Corresponding author: {\sf{e-mail: h.i.oberman@uu.nl}}}\inst{,1}} 
+\address[\inst{1}]{Departement of Methodology \& Statistics, Padualaan 14, 3584 CH Utrecht, The Netherlands}
+%%%%    Information for the second author
+\author[Vink]{Gerko Vink\inst{1}}
+%%%%    \dedicatory{This is a dedicatory.}
+\Receiveddate{zzz} \Reviseddate{zzz} \Accepteddate{zzz}
+
+\begin{abstract}
+Developing new imputation methodology has become a very active field.
 Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
@@
 This may lead to suboptimal use of multiple imputation in practice.
-Additionally, we suggest a course of action for simulating and evaluating missing data problems."
-bibliography: StandardizedEvaluation.bib  
-output: 
-  pdf_document:
-    keep_tex: true
-editor_options: 
-  chunk_output_type: inline
----
+Additionally, we suggest a course of action for simulating and evaluating missing data problems.
+\end{abstract}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
-library(dplyr)
-# source(here::here("R/utils.R"))
-# source(here::here("R/run_sim.R"))
-```
+%% maketitle must follow the abstract.
+\maketitle                   % Produces the title.
 
-# Introduction
+%% If there is not enough space inside the running head
+%% for all authors including the title you may provide
+%% the leftmark in one of the following three forms:
+
+%% \renewcommand{\leftmark}
+%% {First Author: A Short Title}
+
+%% \renewcommand{\leftmark}
+%% {First Author and Second Author: A Short Title}
+
+%% \renewcommand{\leftmark}
+%% {First Author et al.: A Short Title}
+
+%% \tableofcontents  % Produces the table of contents.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% INTRODUCTION
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Introduction}
 
@@
 The technique has earned a permanent spot in research and policy making, demonstrated e.g.
-by the detailed manual created by the National Research Council [@little2012prevention].
+by the detailed manual created by the National Research Council \citep*{little2012prevention}.
 Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing imputation techniques.
@@
 The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.
-A popular variant is multiple imputation [@rubi76], wherein missing values are imputed multiple times.
-The data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference [using Rubin's rules; @rubi87].
+A popular variant is multiple imputation \citep{rubi76}, wherein missing values are imputed multiple times.
+The data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference \citep[using Rubin's rules][]{rubi87}.
 In contrast to ad hoc methods for dealing with missing values (e.g.
 list-wise deletion, mean imputation, regression imputation, last observation carried forward, indicator method), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
-
-<!-- [It has been empirically demonstrated time and again that the inferences on multiply imputed data yield unbiased and confidence valid results (through Rubin's rules).
-But how do we know that this is the case, if we do not have access to the true but unobserved values of the missing entries?
-And how can we be sure that extensions to the methodology yield valid inferences too?
-Especially in the field of ML/AI, there are promising new imputation methods, which may yield even sharper imputations than now-standard (semi-)parametric imputation methods.
-That's why we use simulation studies for the evaluation of imputation routines.] -->
 
@@
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
-In that case, (multiple) imputation would be substantially more efficient than the ubiquitous list-wise deletion.
+In that case,  imputation would be substantially more efficient than the ubiquitous list-wise deletion.
 
@@
 
-# Why some evaluations should not be trusted
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% PROBLEMS
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Why some evaluations should not be trusted}
 
@@
 To demonstrate the broad impact of these problems, we subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation.
-we further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.
+We further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.
 
-## Data generation problems
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Data generation problems}
 
@@
 The problem with model-based data generating mechanisms is that method performance may not translate to empirical data.
+
 First, data are often generated following the model (or distribution) that is also used for imputing the data.
-The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied [@morr18].
+The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied.
 Although such procedures may be statistically relevant, the approach would be no good for even the simplest empirical case.
@@
 A second issue is that SEs may be incorrect if the (re)sampling is not handled correctly.
-[Add pooling when the sample is the population [@vink14] and write something about bootstrapping SEs].
+[Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs].
+[Add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
+These studies apply different combinations of imputation and estimation methods on one or more benchmark datasets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
+In this context, there is no talk of retrieving the 'true' but unobserved values at all.
+Refer to missing data chapter \citep{liu21}.]
 
@@
 
-[Maybe add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
-I don't know where to introduce the difference between statistical inference and prediction.
-But this is an important third category, it seems.
-These studies apply different combinations of imputation and estimation methods on one or more benchmark datasets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
-In this context, there is no talk of retrieving the 'true' but unobserved values at all.
-Refer to missing data chapter [@liu21].]
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-<!-- als je perf based sims doet, dan kijk je wel daarnaar. -->
-
-```{r sim_nobs, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, paged.print=TRUE}
-# # setup
-# set.seed(1)
-# n_sim <- 100
-# 
-# sim_nobs <- purrr::map_dfr(1:n_sim, function(.i){
-#   purrr::map_dfr(c(50, 500, 5000), ~{
-#     run(n_obs = .x) %>% cbind(sim = .i, .)})}) 
-# 
-# # gtsummary::tbl_summary(sim_nobs, by = dat)
-# # https://haozhu233.github.io/kableExtra/awesome_table_in_html.html#Group_rows_via_labeling
-# sim_nobs %>% 
-#   group_by(n_obs, dat) %>% 
-#   summarise(
-#     mean_Y = mean(ave),
-#     beta_X = mean(est),
-#     CIW = mean(ciw),
-#     CR = mean(cov),
-#     RMSE_pred = mean(rmse_pred),
-#     RMSE_cell = mean(rmse_cell, na.rm = TRUE))  %>% 
-#   .[, -1] %>% 
-#   knitr::kable(
-#     col.names = c("", "$\\mu_{Y} = 10$", "$\\beta_{X} = 0.4$", "CIW", "CR", "RMSE$_{p}$", "RMSE$_{c}$"), 
-#     digits = 3,
-#     escape = FALSE) %>% 
-#   kableExtra::kable_classic(full_width = TRUE) %>% 
-#   kableExtra::pack_rows(group_label = "$n = 50$", 1, 3, escape = FALSE) %>% 
-#   kableExtra::pack_rows(group_label = "$n = 500$", 4, 6, escape = FALSE) %>% 
-#   kableExtra::pack_rows(group_label = "$n = 5000$", 7, 9, escape = FALSE)
-```
-
-## Missingness generation problems
+\subsection{Missingness generation problems}
 
@@
 The most obvious characteristics of missing data are the missingness pattern(s) and missingness mechanism(s).
-Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data [@litt20, p.
-8].
+Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data \citep[][p.
+8]{litt20}.
 
-<!-- L&R, p.
-8: "We find it useful to distinguish the missingness pattern, which describes which values are missing and observed in the data matrix and the missingness mechanism (or mechanisms), which concerns the relationship between missingness and the values of variables in the data matrix".-->
+%  L&R, p.
+8: "We find it useful to distinguish the missingness pattern, which describes which values are missing and observed in the data matrix and the missingness mechanism (or mechanisms), which concerns the relationship between missingness and the values of variables in the data matrix".
 
@@
 The terminology on this topic can be confusing.
-Does a missingness proportion of 50% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
+Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
 We will refer to the latter as the proportion of incomplete cases.
@@
 Moreover, the amount of missingness that is used to simulate the performance of imputation methods differs between studies.
-[Add ref to chapter [@liu21] and add info.] Some studies use only 10% missing data where other studies push the limits to additionally investigate the performance under at least 50% missing data.
+[Add ref to chapter \citep{liu21} and add info.] Some studies use only 10\% missing data where other studies push the limits to additionally investigate the performance under at least 50\% missing data.
 This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
@@
 [Briefly introduce missingness mechanisms here and show potential problems, but without simulation results (moved to separate section).
-This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated *spuriously* if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
+This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \emph{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
 And finally something about types of M(N)AR, but what are specific problems with that?
@@
 
-# 
-\hrule
-
-#
-[Until next hline: to be edited!] In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism @meng94.
+In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}.
 The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
-Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable see e.g.
-@rubi76, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
+Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
 
@@
 Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in my view may be far too simplistic.
-
-```{r mis_mech, echo=FALSE, fig.cap="\\label{fig:mis_mech}Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
-Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\\rho= 0$."}
-knitr::include_graphics("img/plot_mmech.pdf")
-```
 
@@
 This problem is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.
-Figure \ref{fig:mis_mech} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
+Figure \ref{fig:MAR} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
 
-It is obvious that the MAR missingness generated from data with low correlation would yield similar inference to the MCAR missingness, even though the missingness is clearly random.
-The simulation results from Table 1 confirm this.
-A detailed simulation setup can be found in Appendix A.
+\begin{figure}[htb]
+\begin{center}
+\resizebox{\textwidth}{!}{
+\includegraphics[]{figures/spurious_MAR.pdf}}
+\caption{Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
+Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$.}
+\end{center}
+\label{fig:MAR}
+\end{figure}
+\vspace*{1pc}
 
-As expected, under genuine MAR missingness, i.e.
-the missingness probabilities differ conditional on the level of the incomplete variable, the performance of complete case analyses drops dramatically.
-For \emph{spurious} MAR, i.e.
-the conditional probabilities remain the same over the level of the incomplete variable, the performance is comparable to that under MCAR.
-In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation @fimd, p.
-48: the missingness does not depend on the incomplete variable.
+It is obvious that the MAR missingness generated from data with low correlation would yield similar inference to the MCAR missingness, even though the missingness is random.
+In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p.
+48]{fimd}: the missingness does not depend on the incomplete variable.
 This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.
 
-Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit [@molenberghs2008every] - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
-In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis see e.g.
-@molenberghs2014handbook, part 5.
+Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every} - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
+In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis \citep[see e.g.][part 5]{molenberghs2014handbook}.
 
-```{r sim_mech_corr, echo=FALSE, cache=TRUE, paged.print=TRUE}
-# # setup
-# set.seed(1)
-# n_sim <- 10
-# correlations <- c(0, .4, .8)
-# mechanisms <- c("MCAR", "MAR", "MNAR") 
-# 
-# sim_mech <- purrr::map_dfr(1:n_sim, function(.i){
-#    purrr::map_dfr(correlations, function(.c){
-#      purrr::map_dfr(mechanisms, function(.m){
-#        run(corr = .c, mis_mech = .m) %>% cbind(sim = .i, .)})})})
-# sim_mech %>% 
-#   group_by(mis_mech, corr) %>% 
-#   summarise(
-#     mean_cca = mean(mean_cca - 10),
-#     mean_est = mean(mean_est - 10),
-#     beta_cca = mean(beta_cca - corr),
-#     beta_est = mean(beta_est - corr),
-#     beta_ciw = mean(beta_ciw),
-#     beta_cr = mean(beta_cr),
-#     rmse_pred = mean(rmse_pred),
-#     rmse_cell = mean(rmse_cell))
-```
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-<!-- table -->
-<!-- Inferences obtained by complete case analysis (CCA) and multiple imputation over 1000 simulations.
-Displayed are the bias, coverage rate (cov) of the 95% confidence interval and the confidence interval width (ciw).
-Imputations are generated by Bayesian linear regression imputation.
--->
-
-# 
-\hrule
-
-## Performance evaluation
+\subsection{Performance evaluation problems}
 
@@
 We are interested in the correct answer to the research question; not in the truth itself.
-This means that, given the framework provided by @rubi87, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
-After all, the 95% confidence interval should contain the `true' value at least 95 out of 100 times @neym34, p.
-591.
+This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
+After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p.
+591]{neym34}.
 [So, what could go wrong by focusing on accuracy exclusively?]
@@
 
-<!-- als inf het doel is, dan moet je de varianties juist krijgen dus kom je niet weg met m = 1 (want geen onderscheid tussen obs en imp) -->
-
-[Add problems with RMSE here.]
+[Add problems with RMSE here. And that if the goal is inference (not prediction) then you need appropriate uncertainty, so m = 1 does not suffice.]
 
 Sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
-However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account @vink14.
+However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink14}.
 After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
@@
 
-# Suggested course of action
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SOLUTIONS
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-## Step 1. Obtain truth
+\section{Suggested course of action}
+
+[In geveral, adhere to ADEMP \citep{morr18}.]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Step 1. Obtain truth}
 
 The evaluator must to decide whether there is a need for sampling variance in the simulations scheme. If sampling variance cannot be omitted from the simulation scheme, two approaches are possible:
-
-- Model-based simulation: draw samples from a known probability distribution, such as the multivariate normal distribution.
+\begin{enumerate}
+  \item Model-based simulation: draw samples from a known probability distribution, such as the multivariate normal distribution.
 The theoretical parameters under which the samples are obtained will serve as the comparative truth during the simulations.
-- Design-based simulation: sample data from a sufficiently large set.
+  \item Design-based simulation: sample data from a sufficiently large set.
 The parameters of the sufficiently large set will serve as the comparative truth during the simulations.
@@
 This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations.
-One needs to realize, however, that the conventional pooling rules [cf.
-@rubin1987, pp.
-76-77], do not apply for finite population inference and that alternative pooling rules need to be used [@raghunathan2003multiple; @vink2014pooling].
+One needs to realize, however, that the conventional pooling rules \citep[cf.][p.
+76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
 
-## Step 2. Induce missingness
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Step 2. Induce missingness}
 
@@
 Next, missing data should be induced conform a model that is dependent on the observed data.
-A straightforward technique for inducing different forms of univariate MAR missingness is described in @fimd, p.
-63; see Figure \ref{fig:MAR} and a generalization to multivariate MAR missingness can be found in @buur06, Appendix B, and @brand1999development, par.
-5.2.3.
-If the missingness is to be induced in longitudinal data, auto-regressive MAR models e.g.
-cf.
-@shara2015randomly, model 2 and model 3, can be useful.
-
-```{r MAR, echo=FALSE, fig.cap="\\label{fig:MAR}Examples of four random missingness mechanisms that yield approximately 50 percent missingness.
-For negatively correlated missingness covariates the resulting mechanisms are reversed."}
-knitr::include_graphics("img/plot_mar.pdf")
-```
+A straightforward technique for inducing different forms of univariate MAR missingness is described in \citet[][p.
+63]{fimd} and a generalization to multivariate MAR missingness can be found in \citet[][Appendix B]{buur06} and \citet[][\S 5.2.3]{brand1999development}.
+If the missingness is to be induced in longitudinal data, autoregressive MAR models \citep[e.g.
+cf.][model 2 and model 3]{shara2015randomly} can be useful.
 
@@
 Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
-This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching [@little1988missing; @rubin1986statistical] may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
 It would be a shame to overlook such results due to the focus on a single MAR mechanism.
@@
 
-- 10% missingness: Depending on the size of the data, this percentage can be considered as a lower bound of realistic evaluation.
-Anything less than 10% may be of little influence on the true data inference.
+\begin{itemize}
+  \item 10\% missingness: Depending on the size of the data, this percentage can be considered as a lower bound of realistic evaluation.
+Anything less than 10\% may be of little influence on the true data inference.
 Performance of a missing data method should at least be acceptable for most missing data problems.
-- 25% missingness: This is a fair amount of missingness and will, depending on the observed data information, have a noticeable influence on the completed data inference.
-When compared to the condition with 10% missingness, the inference obtained under 25% missingness should be less certain (i.e.
+  \item 25\% missingness: This is a fair amount of missingness and will, depending on the observed data information, have a noticeable influence on the completed data inference.
+When compared to the condition with 10\% missingness, the inference obtained under 25\% missingness should be less certain (i.e.
 confidence/credibility interval width should increase), but estimates should still be properly covered and the statistical properties of the missing data method should be sound.
-In practice, at least to my experience in social sciences and official statistics, 25% univariate missingness can easily be considered as a realistic missingness percentage.
-- 50% missingness: Performance under 50% percent simulated missingness will most likely be impacted severely.
+In practice, at least to my experience in social sciences and official statistics, 25\% univariate missingness can easily be considered as a realistic missingness percentage.
+  \item 50\% missingness: Performance under 50\% percent simulated missingness will most likely be impacted severely.
 Depending on how the missingness mechanism interacts with the simulated data, some imputation techniques may yield estimates that are under-covered such that the completed data inference should not be deemed valid anymore.
-If a method yields acceptable inference under 50% MAR missingness, we can determine that the statistical properties of the imputation methodology are sound.
-
+If a method yields acceptable inference under 50\% MAR missingness, we can determine that the statistical properties of the imputation methodology are sound.
+\end{itemize}
 Although I limit the focus here to ignorable non-response, the above suggested proportions are equally applicable to simulations under non-ignorable non-response.
@@
 Often, when inducing missingness, authors remain vague about the actual missingness mechanism under investigation and, even worse, some authors only report something like
-
-> *We generated missing data following a MAR missingness mechanism.*
-
+\begin{quote}
+\emph{We generated missing data following a MAR missingness mechanism.}
+\end{quote}
 This should be considered unacceptable as claims about the validity of the multiple imputation inference depends heavily on the simulated missingness mechanism.
 
-## Step 3. Evaluate performance
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Step 3. Evaluate performance}
 
@@
 Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
-In general, I would say that each multiple imputation routine should be evaluated on at least the following points:
-
-- \emph{Bias:} Results should preferable be unbiased.
+In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
+\begin{itemize}
+  \item \emph{Bias:} Results should preferable be unbiased.
 However, the way bias is considered can greatly influence the interpretation of the results.
@@
 The way bias is computed should therefore be carefully chosen and described.
-- \emph{Interval width:} The way the confidence interval is calculated should be described.
+  \item \emph{Interval width:} The way the confidence interval is calculated should be described.
 Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference.
 However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.
-- \emph{Coverage:} Coverage of a 95% interval should in theory be $\geq 95$, where a coverage of 95% would be most efficient.
+  \item \emph{Coverage:} Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage of 95\% would be most efficient.
 Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
-- [Add RMSE 2x]
-
-- \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data.
+  \item \emph{RMSE:} [RMSE at cell level and at prediction level (if applicable).]
+  \item \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data.
 Under anything but the MCAR assumption, this can be expected.
-When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed see @abayomi2008diagnostics for an detailed overview of diagnostic evaluation for multivariate imputations.
+When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for an detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}.
 When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.
-- \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference.
+  \item \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference.
 However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property.
 When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
-[Add PPC]
-- \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers, PIGS, @li2012imputing.
+  \item \emph{Fit of the imputation model:} [PPC, ref Mingyang, \citep{nguy17, zhao22}.]
+  \item \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
 The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
-[Extend this part!]
-- [Add PPC (Mingyang) [@nguy17], [@zhao22]]
-
+  \end{itemize}
 For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
@@
 
-## Step 4. Report
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-<!-- [Checklist voor iedere sim studie met alle parameters die we bespreken] -->
+\subsection{Step 4. Report}
 
-<!-- - doel -->
+Report the following.
 
-<!-- - 11 punten van stef -->
+[Put this in a table, like this one]
 
-<!-- - andere punten uit dit paper -->
-
-<!-- - welke variables/features vormen imp model en welke analysis model (congenial, meng 94) -->
-
-<!-- - cross validation? -->
-
-<!-- als pfd met check boxes? -->
-
-Report at least the following:
+\begin{table}[htb]
+\begin{center}
+\caption{The caption of a table.}
+\begin{tabular}{lll}
+\hline
+Description 1 & Description 2 & Description 3\\
+\hline
+Row 1, Col 1 & Row 1, Col 2 & Row 1, Col 3\\
+Row 2, Col 1 & Row 2, Col 2 & Row 2, Col 3\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 - aim
+
 - number of simulations
@@
 - model vs design based
+
 - data types
+
 - sample vs population (vs bootstrapping??)
+
 - number of observations
+
 - number of variables
+
 - coherence between variables
@@
 - monotone/univariate vs multivariate
+
 - missingness mechanism
+
 - type of M(N)AR
+
 - missingness proportion
@@
 - imputation method
+
 - imputation model
+
 - number of imputations
+
 - number of iterations
@@
 - analysis model
+
 - estimand
+
 - diagnostic
+
 - ppc
 
-# Discussion
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% DISCUSSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-This document is aimed at establishing a common ground for the evaluation of imputation routines.
-Such a common ground would be the basis of a standardized evaluation.
-This allows for fair and efficient comparisons between imputation techniques.
-Ultimately, it would be desirable to evaluate every imputation routine against the same standardized set in order to quantify the statistical properties across imputation routines.
-If properly executed, this would allow for careful matching of imputation methodologies to new missing data problems.
+\section{Discussion}
 
-# Appendix A: Simulation setup
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% APPENDICES
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Let $Y=(Y_{obs},Y_{mis})$ be an incomplete semi-continuous variable, where $Y_{obs}$ and $Y_{mis}$ denote the observed values and the missing values in $Y$, respectively.
-Further, $X=(X_1,...,X_k)$ is a set of $k$ fully observed covariates, where $X_{obs}$ and $X_{mis}$ correspond to the observed an missing parts in $Y$ and where $X_j$ would indicate the $j$th variable in $X$, with $j=1,\dots, k$.
-Finally, let $R$ be a response indicator that is 1 if $Y$ is observed and 0 if $Y$ is missing.
+\begin{acknowledgement}
+[Maybe add Gelman here?]
+\end{acknowledgement}
+\vspace*{1pc}
 
-To prove the point, I generated a single data set from the multivariate normal distribution with means
-$$
-\mu= \bordermatrix{&	 \cr
- Y	&5	\cr
-{ X}_1	&10	\cr
-{ X}_2	&10	},
-$$
-and covariance matrix
-$$
-\Sigma = \bordermatrix{& Y &{ X}_1 &{ X}_2	 \cr
- Y	&1	&0.8	&0\cr
-{ X}_1	&0.8	&1	&0\cr
-{ X}_2	&0	&0	&1}.
-$$
+\noindent {\bf{Conflict of Interest}}
 
-Three incomplete sets were then generated, wherein random missingness was imposed in $Y$ in three ways.
-First, I left the observed data \emph{observed at random} to impose MCAR missingness following the mechanism
-$$
-P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0)=.50.
-$$
-Next, I created random missingness in $Y$ according to the following MAR missingness mechanism
-$$
-P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0|Y_{obs}, X_j),
-$$
-by using a random draw from a binomial distribution of the same length as $Y$ and of size 1 with missingness probability equal to the inverse logit
-$$
-P(R=0)=\frac{e^{a}}{(1+e^{a})}.
-$$
-where $a=(\mu_{X_j}-X_{ij})/\sqrt{\mathbb{E}[X_{ij} - \mu_{X_j}]}$ in order to create approximately 50% right-tailed MAR missingness. I followed this procedure twice; once for each $X_j$.
+\noindent {\it{The authors have declared no conflict of interest. (or please state any conflicts of interest)}}
 
-The three resulting incomplete data sets were then imputed with package `mice` @mice, v2.25, in `R` @R, v3.2.3, with Bayesian linear regression imputation (`mice.impute.norm`) as the imputation routine.
-The completed-data analyses were combined into a single inference following the rules described in @vink2014pooling.
+% \section*{Appendix {\it(please insert here, if applicable)}}
+% 
+% \subsection*{A.1.\enspace Second level heading}
+% 
+% Please insert appendices before the references.
 
-```{r}
-sessionInfo()
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% REFERENCES
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliography{bibliography}
+\bibliographystyle{apalike}
 
-```
-
-# References
+% \begin{thebibliography}{10}
+% \bibitem[Bauer and Bauer(1994)Bauer, P.
+and Bauer, M.M.]{bib1}Bauer, P.
+and  Bauer, M.
+M.
+(1994).
+Testing equivalence simultaneously for location and  dispersion of two normally distributed populations.
+\textit{Biometrical  Journal} \textbf{36}, 643--660.
+% \bibitem[Farrington, C. P. and Andrews, N. (2003)]{bib2}Farrington, C.P. and Andrews, N. (2003). Outbreak detection:
+% Application to infectious disease surveillance. In: Monitoring the Health of Populations (eds. R. Brookmeyer and D. F. Stroup), Oxford University Press, Oxford,\break 203--231.
+% \bibitem[Rencher(1998)Rencher, A.C.]{bib3}Rencher, A. C. (1998).  \textit{Multivariate Statistical Inference and Applications}. Wiley, New  York. 
+% \end{thebibliography}
+% \newpage
+% \phantom{aaaa}
+\end{document}

### 2022-03-23 Formatted/Formatted.tex
Some textual changes in the intro and problems section (especially wrt data generation), add sub-steps in suggestions section (commented out for now)

@@
 
-\title[Standardized evaluation]{Towards a standardized evaluation of imputation routines}
+\title[Standardized evaluation]{Towards a standardized evaluation of imputation methodology}
 %% Information for the first author.
@@
 
-The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.
-A popular variant is multiple imputation \citep{rubi76}, wherein missing values are imputed multiple times.
-The data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference \citep[using Rubin's rules][]{rubi87}.
+The idea behind imputation is to impute (fill in) missing data, to obtain a valid estimate of what could have been.
+A popular variant is multiple imputation \citep{rubi76}, whereby each missing value is imputed multiple times.
+The completed data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference \citep[using Rubin's rules][]{rubi87}.
 In contrast to ad hoc methods for dealing with missing values (e.g.
-list-wise deletion, mean imputation, regression imputation, last observation carried forward, indicator method), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
+list-wise deletion, mean imputation, regression imputation, last observation carried forward, indicator method), multiple imputation properly takes into account the sources of uncertainty that are related to the missingness problem.
 
 The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
-In general it holds that modeling missing data becomes more challenging when the amount of missingness increases.
+In general it holds that modeling missing data becomes more challenging when the amount of missingness increases [ref?].
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
@@
 When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies.
-In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, [then several missing data methods are applied] and a set of evaluation criteria is postulated to evaluate the performance of the imputation methods.
-However, no golden standard has been established to evaluate these imputation routines and, as a result, the validity of the simulation procedures may differ tremendously from one developer to another.
-[Nowadays there are a lot of imp alg being propagated, esp ai/ml, whose perf often times seems quite promising.
-But may not be a fair comparison because of the simulation set-up/dev in ai/ml so fast that steps may be skipped]
+In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods.
+However, no golden standard has been established to evaluate imputation routines and, as a result, the validity of the simulation procedures may differ tremendously from one developer to another.
+Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced.
+Although these promising new methods seem to yield even sharper imputations than now-standard (semi-)parametric imputation methods, the comparison may not be fair due to the simulation setup.
+
+% [Nowadays there are a lot of imp alg being propagated, esp ai/ml, whose perf often times seems quite promising.
+But may not be a fair comparison because of the simulation set-up/dev in ai/ml so fast that steps may be skipped.
+Especially in the field of ML/AI, there are promising new imputation methods, which may yield even sharper imputations than now-standard (semi-)parametric imputation methods]
 
@@
 Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers.
-This brings forth a chain of potential problems in the objective assessment of the performance of imputation routines that may lead to sub-optimal use of multiple imputation in practice.
+This brings forth a chain of potential problems in the objective assessment of imputation method performance, within and across studies, which may lead to sub-optimal use of imputation in practice.
 To demonstrate the broad impact of these problems, we subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation.
-We further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.
+We further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines [this sentance could be smoother].
 
@@
 To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established.
-Those who perform simulation studies are in the luxury position to establish the truth beforehand, by choosing a data generating mechanism to obtain a complete dataset.
-Complete data is typically either generated from a statistical model, or sampled from a sufficiently large observed set [model-based vs design-based simulation].
+Those who perform simulation studies are in the luxury position to establish the truth beforehand through a data generating mechanism which yields complete data.
+The data generating mechanism is typically either in line with model-based or design-based simulation.
+With the former, data are generatd from a statistical model.
+With the latter, data are sampled from a sufficiently large observed set.
 
-The problem with model-based data generating mechanisms is that method performance may not translate to empirical data.
+The problem with model-based data generating mechanisms is that the performance of imputation methods may not translate to empirical data.
+With this approach, the complete data are drawn from a known (parametric) model, whereas real-life data hardly ever follow a given theoretical distribution.
+The generalizibility of performance outside of the simulation setting is at an even greater risk when data are generated following a model that is also used for imputing the data.
+The performance of this specific imputation approach is then deemed good (or better than other methods), purely due to the evaluated conditions being in favor of the problem that is studied.
+~~Although such procedures may be statistically relevant, the approach would be no good for even the simplest empirical case.~~ Moreover, data are often generated such that the problem that is being studied is most pronounced.
+This results in simulated data that contain very valuable information structures, e.g.
+consistently high correlations between groups of variables.
+Then, no matter what type of missingness is induced, the observed parts of the data still hold much, if not all, of the information about the missing part and the performance of the imputation procedure(s) is not surprisingly evaluated as very good.
+~~Also, it may be that the relations in real-life data are rarely as pronounced as the problem that is being studied.~~ [maybe end in something positive?
+or some other bridge to the next paragraph?]
 
-First, data are often generated following the model (or distribution) that is also used for imputing the data.
-The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied.
-Although such procedures may be statistically relevant, the approach would be no good for even the simplest empirical case.
-Such simulations pose serious threats to the applicability of the method as real-life data hardly ever follow a given theoretical distribution.
-[Add that data types may be more messy.] Second, data are often generated such that the problem that is being studied is most pronounced.
-This results in simulated data that contains very valuable information structures, i.e.
-the correlations between groups of variables may be very pronounced.
-In other words, no matter what type of missingness is induced, the observed parts of the data may still hold much, if not all, of the information about the missing part and the performance of the imputation procedure is not surprisingly evaluated as very good.
-Also, it may be that the relations in real-life data are rarely as pronounced as the problem that is being studied.
+An obvious problem with design-based simulation is that obtaining a large dataset without missing entries is very challenging.
+The idea behind this data generating mechanisim is to resample repeatedly from an empirical dataset which presents certain relevant characteristics but has an unknown true data generating mechanism.
+Since most real-world data contains at least one missing entry, and the non-response model underlying this missingness is by definition unknown, the simulator needs to deal with missingness in \emph{some} way before empirical data can serve as comparative truth in the simulations.
+It may seem as an intuitive solution to only draw complete cases from the incomplete dataset, which would indeed yield complete samples.
+However, there may be inherent differences between cases with and cases without any missing values in the incomplete dataset, due to the unknown non-response model.
+Only sampling complete cases would thus result in a simulation which fails to capture all relevant real-world conditions [which refutes the main reason for using a design-based data generating mechanism in the first place].
+Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset from which samples may be drawn.
+Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.
 
-The problem with design-based simulation is that it's difficult to find large datasets without missing entries.
-And imputing the missing data beforehand and then treating the once-imputed data as comparative truth may favor the used imputation method in further evaluations.
-A second issue is that SEs may be incorrect if the (re)sampling is not handled correctly.
-[Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs].
+% For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
+This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
+Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
+However, the real-world missingness would then be omitted.
+
 [Add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
@@
 
+% Irrespective of the data generating mechanism,
+
 Problems with either of these data generation approaches is that the number of observations, the number of variables, the variable type(s) and the coherence between variables may give some imputation methods a comparative advantage, if they are not a) varied across simulation conditions [when is this fair?
 is this even an option?], or b) the object of the study [add some examples of simulations specifically focused on small samples, categorical data, etc.].
+
+Something else to consider is that if missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly (whether it may be from a known model or empirical data).
+[Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs and that SEs may be incorrect if the (re)sampling is not handled correctly.]
+
+For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches.
+The parameters of that single complete set will serve as comparative truth.
+This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations.
+One needs to realize, however, that the conventional pooling rules \citep[cf.][p.
+76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
+[in our own experience it may be difficult to get true parameter estimates of transformed data (e.g., the mean of a variable with skewness).]
 
@@
 
-[In geveral, adhere to ADEMP \citep{morr18}.]
+[In geveral, adhere to ADEMP \citep{morr18}.
+Make aim clear.
+Choose scope (type of DGM, target) and imputation methods to evaluate.
+Ideally, perform fully factorial design, see \citet{morr18}.
+Be explicit in choices, better even to share the script directly.]
 
@@
 
-The evaluator must to decide whether there is a need for sampling variance in the simulations scheme. If sampling variance cannot be omitted from the simulation scheme, two approaches are possible:
+% 1a. choose DGM (if model-based: choose n_obs, n_var, relations between var, var types, etc; if design-based: choose n_obs, missing data handling strategy)
+% 1b. decide if sampling variance is needed
+% 1c. draw complete sample(s)
+
+The simulator must to decide whether there is a need for sampling variance in the simulations scheme. If sampling variance cannot be omitted from the simulation scheme, two approaches are possible:
 \begin{enumerate}
@@
 
+% we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar.
+we verplaatsen liever de monte carlo naar het niveau van de missingness.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Step 2. Induce missingness}
+
+% 2a. choose missingness mechanism(s): at least consider MCAR, vary MAR types
+% 2b. choose missing data pattern(s): introduce non-monotone multivariate missingness
+% 2c. choose missingness proportion(s): at least consider 10, 25 and 50\%
 
@@
 
-\subsection{Step 3. Evaluate performance}
+\subsection{Step 3. Apply missing data method(s) and analysis model(s)}
+
+% 3a. Define missing data methods: at least consider complete case analysis as benchmark
+% 3b. Choose imputation model parameters: appropriate n_imp and n_it (if applicable), maybe vary predictor matrices?
+% 3c. Impute the data
+% 3d. Obtain estimates
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Step 4. Evaluate performance}
+
+% 4a. Evaluate imputations: check failed methods, algorithmic convergence (if applicable)
+% 4b. Check imputation model fit: anomalies in distributionsal characteristics, plausibility of imputed values (if requested), PPC
+% 4c. Apply performance measures: at least consider bias, coverage, CI length, if applicable also RMSE (at prediction and/or cell level)
 
@@
 
-\subsection{Step 4. Report}
+\subsection{Step 5. Report}
 

### 2022-03-23 Formatted/Formatted.tex
Merge duplicate info about data generation from problems and solutions sections

@@
 To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established.
-Those who perform simulation studies are in the luxury position to establish the truth beforehand through a data generating mechanism which yields complete data.
-The data generating mechanism is typically either in line with model-based or design-based simulation.
-With the former, data are generatd from a statistical model.
-With the latter, data are sampled from a sufficiently large observed set.
+Those who perform simulation studies are in the luxury position to establish the truth beforehand by choosing a data generating mechanism.
+Data generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
+There are two general approaches to generaring complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as from official statistics records.
 
-The problem with model-based data generating mechanisms is that the performance of imputation methods may not translate to empirical data.
-With this approach, the complete data are drawn from a known (parametric) model, whereas real-life data hardly ever follow a given theoretical distribution.
-The generalizibility of performance outside of the simulation setting is at an even greater risk when data are generated following a model that is also used for imputing the data.
-The performance of this specific imputation approach is then deemed good (or better than other methods), purely due to the evaluated conditions being in favor of the problem that is studied.
-~~Although such procedures may be statistically relevant, the approach would be no good for even the simplest empirical case.~~ Moreover, data are often generated such that the problem that is being studied is most pronounced.
-This results in simulated data that contain very valuable information structures, e.g.
-consistently high correlations between groups of variables.
-Then, no matter what type of missingness is induced, the observed parts of the data still hold much, if not all, of the information about the missing part and the performance of the imputation procedure(s) is not surprisingly evaluated as very good.
-~~Also, it may be that the relations in real-life data are rarely as pronounced as the problem that is being studied.~~ [maybe end in something positive?
-or some other bridge to the next paragraph?]
+The problem with model-based data generating mechanisms is that a method's performance on simulated data may not translate to empirical data.
+Real-life data hardly ever follow a given theoretical distribution, so there is no guarantee that simulation results are generalizable.
+Moreover, data are often generated such that the problem that is being studied is most pronounced, e.g.
+with consistently high correlations between groups of variables.
+This results in simulated data that contain such valuable information structures that, no matter what type of missingness is subsequently induced, the observed parts of the data still hold much (if not all) of the information about the missing part.
+Unsurprisingly, the performance of any imputation method will then be evaluated as good.
+Another threat to the generalizibility of model-based simulations is the use of a single model for both data generation and imputation.
+If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied.
+Other (unfair) comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s) and the coherence between variables.
+In contrast to design-based studies, such characteristics are not always explicit conditions of the data, which may give a false sense of objectivity.
 
@@
 
-% For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
-This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
-Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
-However, the real-world missingness would then be omitted.
-
 [Add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
@@
 
-% Irrespective of the data generating mechanism,
-
-Problems with either of these data generation approaches is that the number of observations, the number of variables, the variable type(s) and the coherence between variables may give some imputation methods a comparative advantage, if they are not a) varied across simulation conditions [when is this fair?
-is this even an option?], or b) the object of the study [add some examples of simulations specifically focused on small samples, categorical data, etc.].
-
 Something else to consider is that if missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly (whether it may be from a known model or empirical data).
-[Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs and that SEs may be incorrect if the (re)sampling is not handled correctly.]
-
-For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches.
-The parameters of that single complete set will serve as comparative truth.
-This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations.
-One needs to realize, however, that the conventional pooling rules \citep[cf.][p.
+[Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs and that SEs may be incorrect if the (re)sampling is not handled correctly.]  One needs to realize that the conventional pooling rules \citep[cf.][p.
 76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
-[in our own experience it may be difficult to get true parameter estimates of transformed data (e.g., the mean of a variable with skewness).]
+[in our own experience it may be difficult to get true parameter estimates of transformed data (e.g., the mean of a variable with skewness).] % we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar.
+we verplaatsen liever de monte carlo naar het niveau van de missingness.
 
@@
 [In geveral, adhere to ADEMP \citep{morr18}.
+But there are some differences for the specific case of imputation methodology.
+All of the parts in ADEMP are used, but a bit scrambled up.
+Table 1 outlines some of the steps to consider in simulation studies focused on evaluating imputation methodology.
+The rest of this section is ordered based on these steps.]
+
+\begin{table}[htb]
+\begin{center}
+\caption{Steps to consider in imputation simulation studies.}
+\begin{tabular}{lll}
+\hline
+Step                  & Aspects under consideration \\
+\hline
+1. Set scope          & Aim(s), missing data method(s) to evaluate, number of simulation \\
+                      & repetitions \\
+2. Obtain truth       & Data generating mechanism(s), estimand(s), sampling variance \\
+3. Induce missingness & Missingness mechanism(s), missing data pattern(s), missingness \\
+                      & proportion(s) \\
+4. Apply methods      & Imputation model(s), analysis model(s) \\
+% Benchmarking, imputation model parameter, estimand
+5. Evaluate methods   & Imputation model fit, performance measures \\
+6. Report             & Text, visualization, checklist, simulation script \\
+\hline
+\end{tabular}
+\end{center}
+% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data generating mechanisms and missingness proportions).}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Set scope}
+
 Make aim clear.
@@
 Ideally, perform fully factorial design, see \citet{morr18}.
-Be explicit in choices, better even to share the script directly.]
+Think about tolerable level of uncertainty to pick a number of simulation repetitions.
+Think about general set-up: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?
 
@@
 
-\subsection{Step 1. Obtain truth}
+\subsection{Obtain truth}
 
@@
 
-The simulator must to decide whether there is a need for sampling variance in the simulations scheme. If sampling variance cannot be omitted from the simulation scheme, two approaches are possible:
-\begin{enumerate}
-  \item Model-based simulation: draw samples from a known probability distribution, such as the multivariate normal distribution.
-The theoretical parameters under which the samples are obtained will serve as the comparative truth during the simulations.
-  \item Design-based simulation: sample data from a sufficiently large set.
-The parameters of the sufficiently large set will serve as the comparative truth during the simulations.
-This approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
+[Decide which data genetation mechanism to use and obtain true value of the estimand(s).] In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations.
+In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations.
+
+The design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
 Applications of design-based simulation are often found in official statistics.
 A benefit of design-based simulation is the ability to use real-life observed data structures.
-\end{enumerate}
-For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches.
+
+For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
+This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
+Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
+However, the real-world missingness would then be omitted.
+
+The simulator must to decide whether there is a need for sampling variance in the simulations scheme.
+For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism.
 The parameters of that single complete set will serve as comparative truth.
 This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations.
-One needs to realize, however, that the conventional pooling rules \citep[cf.][p.
-76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
-
-% we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar.
-we verplaatsen liever de monte carlo naar het niveau van de missingness.
+Use alternative pooling rules for finite population inference \citep{raghunathan2003multiple,vink14}.
+If sampling variance cannot be omitted from the simulation scheme, draw data repeatedly and apply conventional pooling rules \citep[cf.][p.
+76-77]{rubi87}.
 
@@
 
-\subsection{Step 2. Induce missingness}
+\subsection{Induce missingness}
 
@@
 
-\subsection{Step 3. Apply missing data method(s) and analysis model(s)}
+\subsection{Apply missing data method(s) and analysis model(s)}
 
@@
 
-Report the following.
+Be explicit in choices, better even to share the script directly.
 
-[Put this in a table, like this one]
+Describe the simulation textually, use pseudo-code, a flowchart, visualize missingness and simulation results.
 
-\begin{table}[htb]
-\begin{center}
-\caption{The caption of a table.}
-\begin{tabular}{lll}
-\hline
-Description 1 & Description 2 & Description 3\\
-\hline
-Row 1, Col 1 & Row 1, Col 2 & Row 1, Col 3\\
-Row 2, Col 1 & Row 2, Col 2 & Row 2, Col 3\\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
+Fill in the following checklist.
 
@@
 
+This document is aimed at establishing a common ground for the evaluation of imputation routines.
+Such a common ground would be the basis of a standardized evaluation.
+This allows for fair and efficient comparisons between imputation techniques.
+Ultimately, it would be desirable to evaluate every imputation routine against the same standardized set in order to quantify the statistical properties across imputation routines.
+If properly executed, this would allow for careful matching of imputation methodologies to new missing data problems.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-03-24 Formatted/Formatted.tex
Re-order missingness problems section

@@
 \pagespan{1}{}
-\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\[1pc]
-\noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
+\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\ [1pc]
+% \noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
+% } %%% semicolon and fullpoint added here for keyword style
+\noindent \hspace*{-4pc} Supporting Information for this article is available online from \\
+\hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}. 
 }  %%% semicolon and fullpoint added here for keyword style
@@
 We are all convinced that our methodology has some merit.
-But for sake of progress it would be much more advantageous if the aim of our evaluations would go beyond \emph{proving the point} and would legitimately consider the statistical properties.
+But for sake of progress it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties.
 
@@
 The idea behind this data generating mechanisim is to resample repeatedly from an empirical dataset which presents certain relevant characteristics but has an unknown true data generating mechanism.
-Since most real-world data contains at least one missing entry, and the non-response model underlying this missingness is by definition unknown, the simulator needs to deal with missingness in \emph{some} way before empirical data can serve as comparative truth in the simulations.
+Since most real-world data contains at least one missing entry, and the non-response model underlying this missingness is by definition unknown, the simulator needs to deal with missingness in \textit{some} way before empirical data can serve as comparative truth in the simulations.
 It may seem as an intuitive solution to only draw complete cases from the incomplete dataset, which would indeed yield complete samples.
@@
 Something else to consider is that if missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly (whether it may be from a known model or empirical data).
-[Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs and that SEs may be incorrect if the (re)sampling is not handled correctly.]  One needs to realize that the conventional pooling rules \citep[cf.][p.
-76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
-[in our own experience it may be difficult to get true parameter estimates of transformed data (e.g., the mean of a variable with skewness).] % we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar.
+Sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
+However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink14}.
+After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
+Note that the conventional pooling rules \citep[cf.][p.
+76-77]{rubi87} do not apply for finite population inference.
+Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
+[An example from our own experience is that it may be difficult to get true parameter estimates of transformed data (in this case, the mean of a variable with a skew normal distribution), If we wouldn't have used a simulation set-up without sampling variance, it may have led to imprecise reference values of the estimands in our simulation.
+So, on the one hand, the typical set-up obfuscates the process that we're interested in (i.e., varying the missing data problem over simulation repetitions), while on the other hand, SEs may be incorrect if the sampling variance is not handled correctly in subsequent inferences.] % we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar.
 we verplaatsen liever de monte carlo naar het niveau van de missingness.
@@
 
-Missingness comes in many shapes and forms [platitude, but how else to start?].
-The most obvious characteristics of missing data are the missingness pattern(s) and missingness mechanism(s).
+In order to evaluate the effectivity of imputation methodology to recover the truth established from a chosen data generating mechanism, some form of missingness has to be induced.
+Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like
+\begin{quote}
+\textit{We generated missing data following a missing at random missingness mechanism.}
+\end{quote}
+This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness patterns and missingness mechanisms.
 Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data \citep[][p.
 8]{litt20}.
+Under this definition, the missingness pattern also encompasses the proportion of missingness in a given dataset.
 
@@
 
-The missing data pattern may inadvertently guide which imputation method prevails in simulation studies.
-Some ad hoc methods are known to yield valid inferences in very specific missing data patterns (e.g., list-wise deletion when there is missingness in the outcome variable of the analysis model exclusively), while they bias inferences in any other case.
-Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions [re-read this!
-and add monotone/univariate vs multivariate and sporadic vs systematic dichotomies].
-Related to the missingness pattern across variables is the missingness proportion.
-The terminology on this topic can be confusing.
+Even the terminology on missingness generation can be confusing.
 Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
-We will refer to the latter as the proportion of incomplete cases.
-But this is not a clear distinction in the literature, which may lead to incorrect conclusions [re-phrase this too].
-Moreover, the amount of missingness that is used to simulate the performance of imputation methods differs between studies.
-[Add ref to chapter \citep{liu21} and add info.] Some studies use only 10\% missing data where other studies push the limits to additionally investigate the performance under at least 50\% missing data.
+In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term 'missingness proportion' restricted to the variable-by-variable interpretation.
+This distinction, however, is not always clear in the existing literature.
+Convoluting the two concepts may lead to incorrect generalizations from simulation studies, since the two are rarely equal (e.g., a proportion of 50\% incomplete cases could translate to a missingness proportion of just 25\% in bivariate data, or even 10\% in data with five variables).
+Moreover, the value of the actual missingness proportion could be diffusing too.
+Some studies use only 10\% missing data where other studies push the limits to additionally investigate the performance under at least 50\% missing data.
 This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
@@
 
-[Briefly introduce missingness mechanisms here and show potential problems, but without simulation results (moved to separate section).
-This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \emph{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
+% Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
+The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
+If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
+Alternatively, the validity of assumptions become increasingly important when the missingness increases.
+Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
+
+Other problems with missingness patterns entail the distribution of missing values across cases (as opposed to variables).
+[Add: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivatiate patterns.] Specific missing data patterns may inadvertently guide which imputation method prevails in performance.
+Some ad hoc methods are known to yield valid inferences under certain restricted conditions (e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively), while they bias inferences in any other case.
+Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions.
+
+[Missingness mechanisms: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
 And finally something about types of M(N)AR, but what are specific problems with that?
 ref dance paper rianne]
+
+% MCAR missingness: The scenario where the missing values are missing at random and the observed values are observed at random.
+Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
+
+% Missingness type: Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
+This may influence the performance of some (but not necessarily all) imputation routines.
+For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+It would be a shame to overlook such results due to the focus on a single MAR mechanism.
+
+%  the validity of assumptions become increasingly important when the missingness increases.
 
@@
 Sadly, the straightforward case of MCAR is often neglected from simulation studies and focus is drawn to the evaluation of MAR mechanisms only.
-Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in my view may be far too simplistic.
+Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in our view may be far too simplistic.
 
@@
 This problem is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.
-Figure \ref{fig:MAR} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
 
-\begin{figure}[htb]
-\begin{center}
-\resizebox{\textwidth}{!}{
-\includegraphics[]{figures/spurious_MAR.pdf}}
-\caption{Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
+% Figure \ref{fig:MAR} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
+
+% \begin{figure}[htb]
+% \begin{center}
+% \resizebox{\textwidth}{!}{
+% \includegraphics[]{figures/spurious_MAR.pdf}}
+% \caption{Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
 Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$.}
-\end{center}
-\label{fig:MAR}
-\end{figure}
-\vspace*{1pc}
+% \end{center}
+% \label{fig:MAR}
+% \end{figure}
+% \vspace*{1pc}
 
-It is obvious that the MAR missingness generated from data with low correlation would yield similar inference to the MCAR missingness, even though the missingness is random.
+A MAR missingness generated from data with low correlation would yield similar inference to the MCAR missingness, even though the missingness is random.
 In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p.
@@
 
-\subsection{Performance evaluation problems}
+\subsection{Evaluation problems}
 
-[Add bridge from last section. And explain how this part is subdivided: first general characteristics of the imputations such as convergence, then performance measures.]
+% The evaluation criteria may depend on the estimand.
+When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
+Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
 
-[Performance of different imputation methods may be influenced by preconditions of the methods themselves.
-All imputation models should be congenial, but this is difficult to test to often ignored even in single imputation endeavor, let alone in simulation studies with 100s of imputations.
+% For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.
+
+[Add bridge from last section.
+And explain how this part is subdivided: first problems with evaluation of imputations (general characteristics of the imputations such as convergence and plausibility), then evaluation of performance (in terms of actual performance measures).]
+
+[All imputation models should be congenial, but this is difficult to test to often ignored even in single imputation endeavor, let alone in simulation studies with hundreds of imputations.
+Checks of imputation model fit are often omitted.
+Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.
 Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies.
-Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.]
+Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]
 
@@
 
-[Then onto the performance measure part.]
-
+[Add segway between evaluation of imputation and actual performance measures.]
 The evaluation criteria used to assess imputation performance vary from one developer to another.
@@
 But the choice of performance measure(s) may inadvertently distort statistical properties of the imputed data.
+
+[Add choice of analysis model/estimand here, since performance measure are defined wrt the estimand, e.g., focusing on univariate estimates may favor invalid ad hoc methods.]
 
@@
 591]{neym34}.
-[So, what could go wrong by focusing on accuracy exclusively?]
-
-[Add choice of analysis model/estimand here, since performance measure are defined wrt the estimand, e.g., focusing on univariate estimates may favor invalid ad hoc methods.]
+[So, what could go wrong by focusing on accuracy exclusively?
+See problems with RMSE of predictions.]
 
 [Add problems with RMSE here. And that if the goal is inference (not prediction) then you need appropriate uncertainty, so m = 1 does not suffice.]
-
-Sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
-However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink14}.
-After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
-
-Last, when the evaluator is on the verge of drawing conclusions about the performance of the imputation routine, the performance should be carefully qualified.
-Comparing the performance of an imputation routine given a population (or true) parameter allows for quantitative evaluation.
-However, in order to pose qualitative statements about the performance on simulated conditions, comparative methodology is required.
-For example, when claiming that imputation performance is unacceptable when deviations from normality become rather stringent, such performance is highly dependent on the simulation conditions that are used.
-For a well-balanced judgement about the severity of the performance drop, comparative simulations with e.g.
-nonparametric models should be executed.
-A method may perform badly, but if it still outperforms every other approach, it may yet be of great practical relevance.
 
@@
 4. Apply methods        & Imputation model(s), analysis model(s) \\
-% Benchmarking, imputation model parameter, estimand
-5. Evaluate methods   & Imputation model fit, performance measures \\
-6. Report             & Text, visualization, checklist, simulation script \\
+5. Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
+                        & model fit \\
+6. Evaluate performance & Performance measures,  \\
+7. Report               & Text, visualization, checklist, simulation script \\
 \hline
@@
 
-Make aim clear.
+[Make aim clear.
 Choose scope (type of DGM, target) and imputation methods to evaluate.
@@
 Think about tolerable level of uncertainty to pick a number of simulation repetitions.
-Think about general set-up: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?
+Think about general set-up: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?]
 
@@
 
-[Decide which data genetation mechanism to use and obtain true value of the estimand(s).] In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations.
+[Decide which data genetation mechanism to use and obtain true value of the estimand(s).]
+
+In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations.
 In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations.
@@
 
+[Missingness mechanisms]
 First, one should always consider MCAR missingness, i.e.
@@
 
+[Missingness type]
 Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
 Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
-This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+This may influence the performance of some (but not necessarily all) imputation routines.
+For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
 It would be a shame to overlook such results due to the focus on a single MAR mechanism.
 
+[Missingness proportion]
 Fourth, the amount of missingness must be varied.
@@
 Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
-
 The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
-If there is no missingness - or if there is no data, for that matter - the inference does not depend on the assumption.
+If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
 Alternatively, the validity of assumptions become increasingly important when the missingness increases.
 Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
-I therefor propose, for all mechanisms, to evaluate at least the following univariate missingness percentages when evaluating imputation routines:
+We therefore propose, for all mechanisms, to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness see Table 2.
 
-\begin{itemize}
-  \item 10\% missingness: Depending on the size of the data, this percentage can be considered as a lower bound of realistic evaluation.
-Anything less than 10\% may be of little influence on the true data inference.
-Performance of a missing data method should at least be acceptable for most missing data problems.
-  \item 25\% missingness: This is a fair amount of missingness and will, depending on the observed data information, have a noticeable influence on the completed data inference.
-When compared to the condition with 10\% missingness, the inference obtained under 25\% missingness should be less certain (i.e.
-confidence/credibility interval width should increase), but estimates should still be properly covered and the statistical properties of the missing data method should be sound.
-In practice, at least to my experience in social sciences and official statistics, 25\% univariate missingness can easily be considered as a realistic missingness percentage.
-  \item 50\% missingness: Performance under 50\% percent simulated missingness will most likely be impacted severely.
-Depending on how the missingness mechanism interacts with the simulated data, some imputation techniques may yield estimates that are under-covered such that the completed data inference should not be deemed valid anymore.
-If a method yields acceptable inference under 50\% MAR missingness, we can determine that the statistical properties of the imputation methodology are sound.
-\end{itemize}
-Although I limit the focus here to ignorable non-response, the above suggested proportions are equally applicable to simulations under non-ignorable non-response.
+\begin{table}[htb]
+\begin{center}
+\caption{Suggested univariate missingness proportions.}
+\begin{tabular}{lll}
+\hline
+Prop. & Reasoning for suggested missingness proportion \\
+\hline  
+10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
+      & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
+      & the true data inference. Performance of a missing data method should at least be \\
+      & acceptable for most missing data problems. \\ & \\
+25\%  & This is a fair amount of missingness and will, depending on the observed data \\
+      & information, have a noticeable influence on the completed data inference. When \\
+      & compared to the condition with 10\% missingness, the inference obtained under \\
+      & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
+      & should increase), but estimates should still be properly covered and the statistical \\
+      & properties of the missing data method should be sound. In practice, at least to our \\
+      & at least to my experience in social sciences and official statistics, 25\% \\
+      & univariate missingness can easily be considered as a realistic missingness percentage. \\ & \\
+50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
+      & severely. Depending on how the missingness mechanism interacts with the simulated \\
+      & data, some imputation techniques may yield estimates that are under-covered such that \\
+      & the completed data inference should not be deemed valid anymore. If a method yields \\
+      & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
+      & properties of the imputation methodology are sound. \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
-Fifth, the used missingness mechanism should be detailed, either graphically or written as a function of the data.
-Often, when inducing missingness, authors remain vague about the actual missingness mechanism under investigation and, even worse, some authors only report something like
-\begin{quote}
-\emph{We generated missing data following a MAR missingness mechanism.}
-\end{quote}
-This should be considered unacceptable as claims about the validity of the multiple imputation inference depends heavily on the simulated missingness mechanism.
+Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.
 
@@
 
+[Use CCA as benchmark method: the imputation method should outperform CCA.
+Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matix, regularization if applicable, etc.).
+Estimate estimand(s) using the analysis model, use correct pooling rules.]
+
+It is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
+We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsection{Step 4. Evaluate performance}
+\subsection{Evaluate imputations}
 
@@
 % 4b. Check imputation model fit: anomalies in distributionsal characteristics, plausibility of imputed values (if requested), PPC
+
+Convergence of the algorithm: Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
+The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
+
+Distributional characteristics: In practice, the distribution of the incomplete data may differ greatly from the observed data.
+Under anything but the MCAR assumption, this can be expected.
+When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for an detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}.
+When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.
+
+Fit of the imputation model: [PPC, ref Mingyang, \citep{nguy17, zhao22}.]
+
+Plausibility of the imputed values: Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference.
+However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property.
+When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Evaluate performance}
+
 % 4c. Apply performance measures: at least consider bias, coverage, CI length, if applicable also RMSE (at prediction and/or cell level)
 
-It is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
-We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance.
 The evaluation criteria may depend on the estimand.
@@
 Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
+
 In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
-\begin{itemize}
-  \item \emph{Bias:} Results should preferable be unbiased.
+
+Bias: Results should preferable be unbiased.
 However, the way bias is considered can greatly influence the interpretation of the results.
@@
 The way bias is computed should therefore be carefully chosen and described.
-  \item \emph{Interval width:} The way the confidence interval is calculated should be described.
+
+Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage of 95\% would be most efficient.
+Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
+
+Interval width: The way the confidence interval is calculated should be described.
 Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference.
 However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.
-  \item \emph{Coverage:} Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage of 95\% would be most efficient.
-Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
-  \item \emph{RMSE:} [RMSE at cell level and at prediction level (if applicable).]
-  \item \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data.
-Under anything but the MCAR assumption, this can be expected.
-When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for an detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}.
-When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.
-  \item \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference.
-However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property.
-When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
-  \item \emph{Fit of the imputation model:} [PPC, ref Mingyang, \citep{nguy17, zhao22}.]
-  \item \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
-The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
-  \end{itemize}
-For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
-This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
-Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
-However, the real-world missingness would then be omitted.
+
+RMSE: [RMSE at cell level and at prediction level (if applicable).]
+
+Last, when the evaluator is on the verge of drawing conclusions about the performance of the imputation routine, the performance should be carefully qualified.
+Comparing the performance of an imputation routine given a population (or true) parameter allows for quantitative evaluation.
+However, in order to pose qualitative statements about the performance on simulated conditions, comparative methodology is required.
+For example, when claiming that imputation performance is unacceptable when deviations from normality become rather stringent, such performance is highly dependent on the simulation conditions that are used.
+For a well-balanced judgement about the severity of the performance drop, comparative simulations with e.g.
+nonparametric models should be executed.
+A method may perform badly, but if it still outperforms every other approach, it may yet be of great practical relevance.
 
@@
 
-\subsection{Step 5. Report}
+\subsection{Report}
 
@@
 Describe the simulation textually, use pseudo-code, a flowchart, visualize missingness and simulation results.
+
+The used missingness mechanism should be detailed, either graphically or written as a function of the data.
 

### 2022-03-25 Formatted/Formatted.tex
Add last version that we discussed to run latexdiff on

@@
 Other (unfair) comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s) and the coherence between variables.
-In contrast to design-based studies, such characteristics are not always explicit conditions of the data, which may give a false sense of objectivity.
+In contrast to design-based studies, such characteristics are not always explicit simulation conditions, which may give a false sense of objectivity.
 
-An obvious problem with design-based simulation is that obtaining a large dataset without missing entries is very challenging.
-The idea behind this data generating mechanisim is to resample repeatedly from an empirical dataset which presents certain relevant characteristics but has an unknown true data generating mechanism.
-Since most real-world data contains at least one missing entry, and the non-response model underlying this missingness is by definition unknown, the simulator needs to deal with missingness in \textit{some} way before empirical data can serve as comparative truth in the simulations.
-It may seem as an intuitive solution to only draw complete cases from the incomplete dataset, which would indeed yield complete samples.
-However, there may be inherent differences between cases with and cases without any missing values in the incomplete dataset, due to the unknown non-response model.
-Only sampling complete cases would thus result in a simulation which fails to capture all relevant real-world conditions [which refutes the main reason for using a design-based data generating mechanism in the first place].
-Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset from which samples may be drawn.
+An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be very challenging.
+Most real-world data contains at least one missing entry, for which the true underlying non-response model is by default unknown.
+Therefore, the simulator needs to deal with missingness in \textit{some} way before incomplete empirical data can serve as comparative truth in the simulations.
+It may seem as an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples.
+However, there may be inherent differences between cases with and cases without any missing values, due to the unknown non-response model.
+Only sampling complete cases would thus result in a simulation that fails to capture all relevant real-world conditions [which refutes the main reason for using a design-based data generating mechanism in the first place].
+Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset to draw samples from.
 Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.
-
+Nowadays, ...
 [Add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
@@
 
-Something else to consider is that if missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly (whether it may be from a known model or empirical data).
+Something else to consider is the number of samples that is drawn from the data generating mechanism.
+If evaulating missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly.
 Sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
@@
 Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
-[An example from our own experience is that it may be difficult to get true parameter estimates of transformed data (in this case, the mean of a variable with a skew normal distribution), If we wouldn't have used a simulation set-up without sampling variance, it may have led to imprecise reference values of the estimands in our simulation.
+[An example from our own experience is that it may be difficult to get true parameter estimates of transformed data (in this case, the mean of a variable with a skew normal distribution).
+If we wouldn't have used a simulation set-up without sampling variance, it may have led to imprecise reference values of the estimands in our simulation.
 So, on the one hand, the typical set-up obfuscates the process that we're interested in (i.e., varying the missing data problem over simulation repetitions), while on the other hand, SEs may be incorrect if the sampling variance is not handled correctly in subsequent inferences.] % we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar.
@@
 
-In order to evaluate the effectivity of imputation methodology to recover the truth established from a chosen data generating mechanism, some form of missingness has to be induced.
+In order to evaluate the effectivity of imputation methodology in recovering the established truthfrom the data generating mechanism, some form of missingness has to be induced.
 Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like
@@
 In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term 'missingness proportion' restricted to the variable-by-variable interpretation.
-This distinction, however, is not always clear in the existing literature.
+This distinction, however, is not always clear in the literature.
 Convoluting the two concepts may lead to incorrect generalizations from simulation studies, since the two are rarely equal (e.g., a proportion of 50\% incomplete cases could translate to a missingness proportion of just 25\% in bivariate data, or even 10\% in data with five variables).
-Moreover, the value of the actual missingness proportion could be diffusing too.
-Some studies use only 10\% missing data where other studies push the limits to additionally investigate the performance under at least 50\% missing data.
+Moreover, the value of the actual missingness proportion may be diffusing too.
+Some studies use only 10\% missing data [rephrase to use missingness proportion?] where other studies push the limits to additionally investigate the performance under at least 50\% missing data.
 This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
@@
 
-Other problems with missingness patterns entail the distribution of missing values across cases (as opposed to variables).
-[Add: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivatiate patterns.] Specific missing data patterns may inadvertently guide which imputation method prevails in performance.
-Some ad hoc methods are known to yield valid inferences under certain restricted conditions (e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively), while they bias inferences in any other case.
+Other problems with missingness patterns entail the distribution of missing values across cases (as opposed to the proportion over variables).
+[Add: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivatiate patterns.
+For example, iterative imputation algorithms converge instantly if there is only 1 variable with missingness or monotone pattern, but require iteration in all other situations.] Specific missing data patterns may inadvertently guide which imputation method prevails in performance.
+Some ad hoc methods are known to yield valid inferences under certain restricted conditions (e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively), while biasing inferences in any other case.
 Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions.
 
-[Missingness mechanisms: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
-And finally something about types of M(N)AR, but what are specific problems with that?
-ref dance paper rianne]
+[Missingness mechanisms: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice, 5) and finally something about types of M(N)AR (ref dance paper rianne).]
 
 % MCAR missingness: The scenario where the missing values are missing at random and the observed values are observed at random. 
-Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
-
-% Missingness type: Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
-This may influence the performance of some (but not necessarily all) imputation routines.
-For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
-It would be a shame to overlook such results due to the focus on a single MAR mechanism.
-
 %  the validity of assumptions become increasingly important when the missingness increases.
 
+\begin{table}[htb]
+\begin{center}
+\caption{Missingness mechanisms.}
+\begin{tabular}{lll}
+\hline
+Mech. & Interpretation and consequences of the mechanism \\
+\hline  
+MCAR  & The probability to be missing is the same for all cases. In other words, the missing \\
+      & values are missing at random and the observed values are observed at random. \\
+MAR   & The probability to be missing is the same within groups of cases defined by the \\
+      & observed data only. The essence is that observed relations in the data inform the \\
+      & missingness, such that MCAR may be assumed within the observed groups. \\
+MNAR  & The probability to be missing dependend on unobserved aspects of the data. The \\
+      & cause of the missingness is unknown and cannot be inferred from the observed data. \\
+      & This is considered non-ignorable missingness \citep[see e.g.][]{rubi76}. \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}.
-The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
-Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
+The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
 
-With MCAR missingness mechanisms, the probability to be missing is the same for all cases.
-This is a necessary simulation condition for evaluating the performance of imputation procedures.
+[MCAR is important but often ignored:] MCAR is a necessary simulation condition for evaluating the performance of imputation procedures.
+Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
 If an imputation method is not able to solve the problem (i.e.
@@
 
-With MAR missingness mechanisms, the probability to be missing is the same within groups of cases defined by the observed data only (e.g.
+[MAR may be spurious:] With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g.
+weight is made incomplete based on gender) such that the probability to be missing is the same within groups of cases defined by the observed data (e.g.
 males are less likely to disclose their weight, but gender is observed).
-The essence here is that the observed relations in the data are used to induce missingness during simulation (e.g.
-weight is made incomplete based on gender), but these relations may be weak, or non-existent.
-
-When MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid.
-When MAR is induced from data without multivariate relations, the inferential implication of the missingness would mimic those of MCAR.
-This problem is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.
+The relations in the observed data may, however, be weak or non-existent.
+If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid.
+In the most extreme case, MAR is induced from data without multivariate relations.
+The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random.
+In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p.
+48]{fimd}: the missingness does not depend on the incomplete variable.
+This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.
 
@@
 
-A MAR missingness generated from data with low correlation would yield similar inference to the MCAR missingness, even though the missingness is random.
-In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p.
-48]{fimd}: the missingness does not depend on the incomplete variable.
-This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.
+[MNAR is often ignored:] Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
+Although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
 
-Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every} - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
-In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis \citep[see e.g.][part 5]{molenberghs2014handbook}.
+[Missingness type is often ignored (duplicated text from solutions section): Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
+This may influence the performance of some (but not necessarily all) imputation routines.
+For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+It would be a shame to overlook such results due to the focus on a single MAR mechanism.
 
@@
 While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.
-Rather, one would prefer an imputation technique to yield both valid inference and plausible imputations.
-It should be studied if an imputation method is prone to deliver such impractical results, and if so, under what conditions.
 
@@
 
+[Although often ignored, it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
+It would be wise to include the mechanism in your simulation study, or perform sensitivity analyses.] An indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis \citep[see e.g.][part 5]{molenberghs2014handbook}.
+
 [Missingness type]
@@
 The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
-
 Distributional characteristics: In practice, the distribution of the incomplete data may differ greatly from the observed data.
@@
 
-Plausibility of the imputed values: Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference.
+Plausibility of the imputed values: Plausible imputations--imputations that could be real values if they had been observed--are not a necessary condition for obtaining valid inference.
 However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property.
+One would prefer an imputation technique to yield both valid inference and plausible imputations.
+It should be studied if an imputation method is prone to deliver such impractical results, and if so, under what conditions.
 When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
@@
 The used missingness mechanism should be detailed, either graphically or written as a function of the data.
+
+The problem of spurious MAR is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.
 

### 2022-03-25 Formatted/Formatted.tex
Run spell check

@@
 To demonstrate the need for standardization, we highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
-This may lead to suboptimal use of multiple imputation in practice.
+This may lead to sub-optimal use of multiple imputation in practice.
 Additionally, we suggest a course of action for simulating and evaluating missing data problems.
@@
 Imputation is a state-of-the-art technique for drawing valid conclusions from incomplete data.
-The technique has earned a permanent spot in research and policy making, demonstrated e.g.
+The technique has earned a permanent spot in research and policymaking, demonstrated e.g.
 by the detailed manual created by the National Research Council \citep*{little2012prevention}.
@@
 The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
-In general it holds that modeling missing data becomes more challenging when the amount of missingness increases [ref?].
+In general, it holds that modeling missing data becomes more challenging when the amount of missingness increases [ref?].
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
@@
 The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology.
-These concerns stem from careful consideration with fellow 'imputers' and from encounters as a reviewer for statistical journals.
+These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals.
 Second, to provide imputation methodologists with a suggested course of action when simulating missing data problems.
-This suggested approach should identify a common ground, but is in no way intended as an absolute solution.
+This suggested approach should identify common ground but is in no way intended as an absolute solution.
 This identifies the third purpose of this paper: discussion.
-We hope to elicit critical thinking with respect to the problems at hand.
+We hope to elicit critical thinking regarding the problems at hand.
 We are all convinced that our methodology has some merit.
-But for sake of progress it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties.
+But for sake of progress, it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties.
 
@@
 This brings forth a chain of potential problems in the objective assessment of imputation method performance, within and across studies, which may lead to sub-optimal use of imputation in practice.
-To demonstrate the broad impact of these problems, we subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation.
-We further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines [this sentance could be smoother].
+To demonstrate the broad impact of these problems, we subdivide the problems into the following three distinct categories: problems with data generation, problems with missingness generation, and problems with performance evaluation.
+We further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines [this sentence could be smoother].
 
@@
 Data generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
-There are two general approaches to generaring complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as from official statistics records.
+There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as from official statistics records.
 
@@
 Unsurprisingly, the performance of any imputation method will then be evaluated as good.
-Another threat to the generalizibility of model-based simulations is the use of a single model for both data generation and imputation.
+Another threat to the generalizability of model-based simulations is the use of a single model for both data generation and imputation.
 If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied.
-Other (unfair) comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s) and the coherence between variables.
+Other (unfair) comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s), and the coherence between variables.
 In contrast to design-based studies, such characteristics are not always explicit simulation conditions, which may give a false sense of objectivity.
@@
 An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be very challenging.
-Most real-world data contains at least one missing entry, for which the true underlying non-response model is by default unknown.
+Most real-world data contains at least one missing entry, for which the true underlying missing data model is by definition unknown.
 Therefore, the simulator needs to deal with missingness in \textit{some} way before incomplete empirical data can serve as comparative truth in the simulations.
-It may seem as an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples.
+It may seem like an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples.
 However, there may be inherent differences between cases with and cases without any missing values, due to the unknown non-response model.
@@
 Nowadays, ...
-[Add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
-These studies apply different combinations of imputation and estimation methods on one or more benchmark datasets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
-In this context, there is no talk of retrieving the 'true' but unobserved values at all.
+[Add that some simulation studies do not use a `ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
+These studies apply different combinations of imputation and estimation methods on one or more benchmark data sets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
+In this context, there is no talk of retrieving the `true' but unobserved values at all.
 Refer to missing data chapter \citep{liu21}.]
 
-Something else to consider is the number of samples that is drawn from the data generating mechanism.
-If evaulating missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly.
+[Design-based model simulation??] Something else to consider is the number of samples that are drawn from the data generating mechanism.
+If evaluating missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly.
 Sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
 However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink14}.
-After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
+After all, we are interested only in the missing data mechanism and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
 Note that the conventional pooling rules \citep[cf.][p.
@@
 If we wouldn't have used a simulation set-up without sampling variance, it may have led to imprecise reference values of the estimands in our simulation.
-So, on the one hand, the typical set-up obfuscates the process that we're interested in (i.e., varying the missing data problem over simulation repetitions), while on the other hand, SEs may be incorrect if the sampling variance is not handled correctly in subsequent inferences.] % we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar.
+So, on the one hand, the typical set-up obfuscates the process that we're interested in (i.e., varying the missing data problem over simulation repetitions), while on the other hand, SEs may be incorrect if the sampling variance is not handled correctly in subsequent inferences.] % we hebben al sampling variance tussen de incomplete data sets, extra sampling variance op het complete data level vertroebelt het proces alleen maar.
 we verplaatsen liever de monte carlo naar het niveau van de missingness.
@@
 
-In order to evaluate the effectivity of imputation methodology in recovering the established truthfrom the data generating mechanism, some form of missingness has to be induced.
+Evaluating imputation methodology requires a missing data problem to be solved.
+After establishing a comparative truth from a data generating mechanism, some form of missingness has to be induced.
 Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like
@@
 Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
-In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term 'missingness proportion' restricted to the variable-by-variable interpretation.
+In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation.
 This distinction, however, is not always clear in the literature.
-Convoluting the two concepts may lead to incorrect generalizations from simulation studies, since the two are rarely equal (e.g., a proportion of 50\% incomplete cases could translate to a missingness proportion of just 25\% in bivariate data, or even 10\% in data with five variables).
+Convoluting the two concepts may lead to incorrect generalizations from simulation studies since the two are rarely equal (e.g., a proportion of 50\% incomplete cases could translate to a missingness proportion of just 25\% in bivariate data, or even 10\% in data with five variables).
 Moreover, the value of the actual missingness proportion may be diffusing too.
@@
 Other problems with missingness patterns entail the distribution of missing values across cases (as opposed to the proportion over variables).
-[Add: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivatiate patterns.
+[Add: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.
 For example, iterative imputation algorithms converge instantly if there is only 1 variable with missingness or monotone pattern, but require iteration in all other situations.] Specific missing data patterns may inadvertently guide which imputation method prevails in performance.
@@
 \begin{center}
-\caption{Missingness mechanisms.}
+\caption{Missingness mechanisms. [Add PMAR]}
 \begin{tabular}{lll}
@@
       & missingness, such that MCAR may be assumed within the observed groups. \\
-MNAR  & The probability to be missing dependend on unobserved aspects of the data. The \\
+MNAR  & The probability to be missing depends on unobserved aspects of the data. The \\
       & cause of the missingness is unknown and cannot be inferred from the observed data. \\
@@
 
-In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}.
-The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
-
 [MCAR is important but often ignored:] MCAR is a necessary simulation condition for evaluating the performance of imputation procedures.
-Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
+Under MCAR, the statistical properties of the observed data given the missing data are known.
+Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation.
 If an imputation method is not able to solve the problem (i.e.
@@
 [MNAR is often ignored:] Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
-Although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
+Although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
 
-[Missingness type is often ignored (duplicated text from solutions section): Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
+[Missingness type is often ignored (duplicated text from solutions section): Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
 This may influence the performance of some (but not necessarily all) imputation routines.
@@
 
-[All imputation models should be congenial, but this is difficult to test to often ignored even in single imputation endeavor, let alone in simulation studies with hundreds of imputations.
+To obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}.
+The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
+
+[All imputation models should be congenial, but this is difficult to test so often ignored even in single imputation endeavor, let alone in simulation studies with hundreds of imputations.
 Checks of imputation model fit are often omitted.
@@
 We are interested in the correct answer to the research question; not in the truth itself.
-This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
+This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.
 After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p.
@@
 
-[In geveral, adhere to ADEMP \citep{morr18}.
+[In general, adhere to ADEMP \citep{morr18}.
 But there are some differences for the specific case of imputation methodology.
@@
                         & model fit \\
-6. Evaluate performance & Performance measures,  \\
+6. Evaluate performance & Performance measures \\
 7. Report               & Text, visualization, checklist, simulation script \\
@@
 
-[Decide which data genetation mechanism to use and obtain true value of the estimand(s).]
+[Decide which data generation mechanism to use and obtain true value of the estimand(s).]
 
@@
 
-The simulator must to decide whether there is a need for sampling variance in the simulations scheme.
+The simulator must decide whether there is a need for sampling variance in the simulations scheme.
 For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism.
 The parameters of that single complete set will serve as comparative truth.
-This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations.
+This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
 Use alternative pooling rules for finite population inference \citep{raghunathan2003multiple,vink14}.
@@
 the scenario where the missing values are missing at random and the observed values are observed at random.
-Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
+% Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
 
@@
 
-[Although often ignored, it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
+[Although often ignored, it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
 It would be wise to include the mechanism in your simulation study, or perform sensitivity analyses.] An indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis \citep[see e.g.][part 5]{molenberghs2014handbook}.
@@
 Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
-Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
+% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
 This may influence the performance of some (but not necessarily all) imputation routines.
 For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
-It would be a shame to overlook such results due to the focus on a single MAR mechanism.
+It would be a shame to overlook results due to the focus on a single MAR mechanism.
 
@@
 Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
-We therefore propose, for all mechanisms, to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness see Table 2.
+We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness see Table 2.
 
@@
       & the true data inference. Performance of a missing data method should at least be \\
-      & acceptable for most missing data problems. \\ & \\
+      & acceptable for most missing data problems. \\ % & \\
 25\%  & This is a fair amount of missingness and will, depending on the observed data \\
@@
       & at least to my experience in social sciences and official statistics, 25\% \\
-      & univariate missingness can easily be considered as a realistic missingness percentage. \\ & \\
+      & univariate missingness can easily be considered as a realistic missingness percentage. \\ % & \\
 50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
@@
 [Use CCA as benchmark method: the imputation method should outperform CCA.
-Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matix, regularization if applicable, etc.).
+Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matrix, regularization if applicable, etc.).
 Estimate estimand(s) using the analysis model, use correct pooling rules.]
@@
 % 4a. Evaluate imputations: check failed methods, algorithmic convergence (if applicable)
-% 4b. Check imputation model fit: anomalies in distributionsal characteristics, plausibility of imputed values (if requested), PPC
+% 4b. Check imputation model fit: anomalies in distributional characteristics, plausibility of imputed values (if requested), PPC
 
+% = evaluation of imputation generating process
 Convergence of the algorithm: Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
@@
 Under anything but the MCAR assumption, this can be expected.
-When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for an detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}.
+When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for a detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}.
 When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.
@@
 Plausibility of the imputed values: Plausible imputations--imputations that could be real values if they had been observed--are not a necessary condition for obtaining valid inference.
-However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property.
+However, in practice, especially when the imputer and the analyst are different persons, plausible imputations may be a desired property.
 One would prefer an imputation technique to yield both valid inference and plausible imputations.
@@
 % 4c. Apply performance measures: at least consider bias, coverage, CI length, if applicable also RMSE (at prediction and/or cell level)
+% fraction of missing information (= model parameter), variance inflation
 
+% sim perf and model perf
 The evaluation criteria may depend on the estimand.
-When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
+When descriptive statistics are the goal and when statistical inference would not be of interest, the bias of the estimates would still apply, but standard errors are generally ignored.
 Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
@@
 However, the way bias is considered can greatly influence the interpretation of the results.
-For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.
+For example, a negligible absolute bias for a parameter for which the true value is zero would yield infinite bias when relative bias is considered.
 The way bias is computed should therefore be carefully chosen and described.
 
-Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage of 95\% would be most efficient.
+Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
 Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
@@
 For example, when claiming that imputation performance is unacceptable when deviations from normality become rather stringent, such performance is highly dependent on the simulation conditions that are used.
-For a well-balanced judgement about the severity of the performance drop, comparative simulations with e.g.
+For a well-balanced judgment about the severity of the performance drop, comparative simulations with e.g.
 nonparametric models should be executed.
@@
 
-Fill in the following checklist.
+Fill in the following checklist (make appendix).
 
@@
 \begin{acknowledgement}
-[Maybe add Gelman here?]
+[Maybe add Gelman here? Mice team]
 \end{acknowledgement}

### 2022-03-25 Formatted/LastDiscussed.tex
Add latexdiff file

--- 2022-03-22 Manuscript/StandardizedEvaluation.Rmd
@@
----
-title: "Towards a standardized evaluation of imputation routines"
-author: "Hanne Oberman and Gerko Vink"
-date: "19-1-2022"
-abstract: "Developing new imputation methodology has become a very active field.
+\documentclass[bimj,fleqn]{w-art}
+\usepackage{times}
+\usepackage{w-thm}
+\usepackage[authoryear]{natbib}
+% \setlength{\bibsep}{2pt}
+% \setlength{\bibhang}{2em}
+% \newcommand{\J}{J\"{o}reskog}
+% \newcommand{\So}{S\"{o}rbom}
+% \newcommand{\bcx}{{\bf X}}
+% \newcommand{\bcy}{{\bf Y}}
+% \newcommand{\bcz}{{\bf Z}}
+% \newcommand{\bcu}{{\bf U}}
+% \newcommand{\bcv}{{\bf V}}
+% \newcommand{\bcw}{{\bf W}}
+% \newcommand{\bci}{{\bf I}}
+% \newcommand{\bch}{{\bf H}}
+% \newcommand{\bcb}{{\bf B}}
+% \newcommand{\bcr}{{\bf R}}
+% \newcommand{\bcm}{{\bf M}}
+% \newcommand{\bcf}{{\bf F}}
+% \newcommand{\bcg}{{\bf G}}
+% \newcommand{\bcs}{{\bf S}}
+% \newcommand{\bca}{{\bf A}}
+% \newcommand{\bcd}{{\bf D}}
+% \newcommand{\bcc}{{\bf C}}
+% \newcommand{\bce}{{\bf E}}
+% \newcommand{\ba}{{\bf a}}
+% \newcommand{\bb}{{\bf b}}
+% \newcommand{\bc}{{\bf c}}
+% \newcommand{\bd}{{\bf d}}
+% \newcommand{\bx}{{\bf x}}
+% \newcommand{\by}{{\bf y}}
+% \newcommand{\bz}{{\bf z}}
+% \newcommand{\bu}{{\bf u}}
+% \newcommand{\bv}{{\bf v}}
+% \newcommand{\bh}{{\bf h}}
+% \newcommand{\bl}{{\bf l}}
+% \newcommand{\be}{{\bf e}}
+% \newcommand{\br}{{\bf r}}
+% \newcommand{\bw}{{\bf w}}
+% \newcommand{\de}{\stackrel{D}{=}}
+% \newcommand{\bt}{\bigtriangleup}
+% \newcommand{\bfequiv}{\mbox{\boldmath $\equiv$}}
+% \newcommand{\bmu}{\mbox{\boldmath $\mu$}}
+% \newcommand{\bnu}{\mbox{\boldmath $\nu$}}
+% \newcommand{\bxi}{\mbox{\boldmath $\xi$}}
+% \newcommand{\btau}{\mbox{\boldmath $\tau$}}
+% \newcommand{\bgamma}{\mbox{\boldmath $\Gamma$}}
+% \newcommand{\bphi}{\mbox{\boldmath $\Phi$}}
+% \newcommand{\bfphi}{\mbox{\boldmath $\varphi$}}
+% \newcommand{\bfeta}{\mbox{\boldmath $\eta$}}
+% \newcommand{\bpi}{\mbox{\boldmath $\Pi$}}
+% \newcommand{\bequiv}{\mbox{\boldmath $\equiv$}}
+% \newcommand{\bvarepsilon}{\mbox{\boldmath $\varepsilon$}}
+% \newcommand{\btriangle}{\mbox{\boldmath $\triangle$}}
+% \newcommand{\bdelta}{\mbox{\boldmath $\Delta$}}
+% \newcommand{\beps}{\mbox{\boldmath $\epsilon$}}
+% \newcommand{\btheta}{\mbox{\boldmath $\theta$}}
+% \newcommand{\balpha}{\mbox{\boldmath $\alpha$}}
+% \newcommand{\bsphi}{\mbox{\boldmath $\varphi$}}
+% \newcommand{\bsig}{\mbox{\boldmath $\sigma$}}
+% \newcommand{\bfpsi}{\mbox{\boldmath $\psi$}}
+% \newcommand{\bfdelta}{\mbox{\boldmath $\delta$}}
+% \newcommand{\bsigma}{{\bf \Sigma}}
+% \newcommand{\bzero}{{\bf 0}}
+% \newcommand{\bpsi}{\mbox{\boldmath $\Psi$}}
+% \newcommand{\bep}{\mbox{\boldmath $\epsilon$}}
+% \newcommand{\bomega}{\mbox{\boldmath $\Omega$}}
+% \newcommand{\bfomega}{\mbox{\boldmath $\omega$}}
+% \newcommand{\blambda}{\mbox{\boldmath $\Lambda$}}
+% \newcommand{\bflambda}{\mbox{\boldmath $\lambda$}}
+% \newcommand{\bfsigma}{\mbox{\boldmath $\sigma$}}
+% \newcommand{\bfpi}{{\mbox{\boldmath $\pi$}}}
+% \newcommand{\bupsilon}{\mbox{\boldmath $\upsilon$}}
+% \newcommand{\obs}{{\rm obs}}
+% \newcommand{\mis}{{\rm mis}}
+% \theoremstyle{plain}
+% \newtheorem{criterion}{Criterion}
+% \theoremstyle{definition}
+% \newtheorem{condition}[theorem]{Condition}
+% \usepackage[]{graphicx}
+% \chardef\bslash=`\\ % p. 424, TeXbook
+% \newcommand{\ntt}{\normalfont\ttfamily}
+% \newcommand{\cn}[1]{{\protect\ntt\bslash#1}}
+% \newcommand{\pkg}[1]{{\protect\ntt#1}}
+% \let\fn\pkg
+% \let\env\pkg
+% \let\opt\pkg
+% \hfuzz1pc % Don't bother to report overfull boxes if overage is < 1pc
+% \newcommand{\envert}[1]{\left\lvert#1\right\rvert}
+% \let\abs=\envert
+
+\begin{document}
+%\DOIsuffix{bimj.DOIsuffix}
+\DOIsuffix{bimj.200100000}
+\Volume{52}
+\Issue{61}
+\Year{2010}
+\pagespan{1}{}
+\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\[1pc]
+\noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
+}  %%% semicolon and fullpoint added here for keyword style
+
+\title[Standardized evaluation]{Towards a standardized evaluation of imputation routines}
+%% Information for the first author.
+\author[Oberman]{Hanne I. Oberman\footnote{Corresponding author: {\sf{e-mail: h.i.oberman@uu.nl}}}\inst{,1}} 
+\address[\inst{1}]{Departement of Methodology \& Statistics, Padualaan 14, 3584 CH Utrecht, The Netherlands}
+%%%%    Information for the second author
+\author[Vink]{Gerko Vink\inst{1}}
+%%%%    \dedicatory{This is a dedicatory.}
+\Receiveddate{zzz} \Reviseddate{zzz} \Accepteddate{zzz}
+
+\begin{abstract}
+Developing new imputation methodology has become a very active field.
 Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.

### 2022-03-25 Formatted/Formatted.tex

@@
 \usepackage[authoryear]{natbib}
-\setlength{\bibsep}{2pt}
-\setlength{\bibhang}{2em}
-\newcommand{\J}{J\"{o}reskog}
-\newcommand{\So}{S\"{o}rbom}
-\newcommand{\bcx}{{\bf X}}
-\newcommand{\bcy}{{\bf Y}}
-\newcommand{\bcz}{{\bf Z}}
-\newcommand{\bcu}{{\bf U}}
-\newcommand{\bcv}{{\bf V}}
-\newcommand{\bcw}{{\bf W}}
-\newcommand{\bci}{{\bf I}}
-\newcommand{\bch}{{\bf H}}
-\newcommand{\bcb}{{\bf B}}
-\newcommand{\bcr}{{\bf R}}
-\newcommand{\bcm}{{\bf M}}
-\newcommand{\bcf}{{\bf F}}
-\newcommand{\bcg}{{\bf G}}
-\newcommand{\bcs}{{\bf S}}
-\newcommand{\bca}{{\bf A}}
-\newcommand{\bcd}{{\bf D}}
-\newcommand{\bcc}{{\bf C}}
-\newcommand{\bce}{{\bf E}}
-\newcommand{\ba}{{\bf a}}
-\newcommand{\bb}{{\bf b}}
-\newcommand{\bc}{{\bf c}}
-\newcommand{\bd}{{\bf d}}
-\newcommand{\bx}{{\bf x}}
-\newcommand{\by}{{\bf y}}
-\newcommand{\bz}{{\bf z}}
-\newcommand{\bu}{{\bf u}}
-\newcommand{\bv}{{\bf v}}
-\newcommand{\bh}{{\bf h}}
-\newcommand{\bl}{{\bf l}}
-\newcommand{\be}{{\bf e}}
-\newcommand{\br}{{\bf r}}
-\newcommand{\bw}{{\bf w}}
-\newcommand{\de}{\stackrel{D}{=}}
-\newcommand{\bt}{\bigtriangleup}
-\newcommand{\bfequiv}{\mbox{\boldmath $\equiv$}}
-\newcommand{\bmu}{\mbox{\boldmath $\mu$}}
-\newcommand{\bnu}{\mbox{\boldmath $\nu$}}
-\newcommand{\bxi}{\mbox{\boldmath $\xi$}}
-\newcommand{\btau}{\mbox{\boldmath $\tau$}}
-\newcommand{\bgamma}{\mbox{\boldmath $\Gamma$}}
-\newcommand{\bphi}{\mbox{\boldmath $\Phi$}}
-\newcommand{\bfphi}{\mbox{\boldmath $\varphi$}}
-\newcommand{\bfeta}{\mbox{\boldmath $\eta$}}
-\newcommand{\bpi}{\mbox{\boldmath $\Pi$}}
-\newcommand{\bequiv}{\mbox{\boldmath $\equiv$}}
-\newcommand{\bvarepsilon}{\mbox{\boldmath $\varepsilon$}}
-\newcommand{\btriangle}{\mbox{\boldmath $\triangle$}}
-\newcommand{\bdelta}{\mbox{\boldmath $\Delta$}}
-\newcommand{\beps}{\mbox{\boldmath $\epsilon$}}
-\newcommand{\btheta}{\mbox{\boldmath $\theta$}}
-\newcommand{\balpha}{\mbox{\boldmath $\alpha$}}
-\newcommand{\bsphi}{\mbox{\boldmath $\varphi$}}
-\newcommand{\bsig}{\mbox{\boldmath $\sigma$}}
-\newcommand{\bfpsi}{\mbox{\boldmath $\psi$}}
-\newcommand{\bfdelta}{\mbox{\boldmath $\delta$}}
-\newcommand{\bsigma}{{\bf \Sigma}}
-\newcommand{\bzero}{{\bf 0}}
-\newcommand{\bpsi}{\mbox{\boldmath $\Psi$}}
-\newcommand{\bep}{\mbox{\boldmath $\epsilon$}}
-\newcommand{\bomega}{\mbox{\boldmath $\Omega$}}
-\newcommand{\bfomega}{\mbox{\boldmath $\omega$}}
-\newcommand{\blambda}{\mbox{\boldmath $\Lambda$}}
-\newcommand{\bflambda}{\mbox{\boldmath $\lambda$}}
-\newcommand{\bfsigma}{\mbox{\boldmath $\sigma$}}
-\newcommand{\bfpi}{{\mbox{\boldmath $\pi$}}}
-\newcommand{\bupsilon}{\mbox{\boldmath $\upsilon$}}
-\newcommand{\obs}{{\rm obs}}
-\newcommand{\mis}{{\rm mis}}
-\theoremstyle{plain}
-\newtheorem{criterion}{Criterion}
-\theoremstyle{definition}
-\newtheorem{condition}[theorem]{Condition}
-\usepackage[]{graphicx}
-\chardef\bslash=`\\ % p. 424, TeXbook
-\newcommand{\ntt}{\normalfont\ttfamily}
-\newcommand{\cn}[1]{{\protect\ntt\bslash#1}}
-\newcommand{\pkg}[1]{{\protect\ntt#1}}
-\let\fn\pkg
-\let\env\pkg
-\let\opt\pkg
-\hfuzz1pc % Don't bother to report overfull boxes if overage is < 1pc
-\newcommand{\envert}[1]{\left\lvert#1\right\rvert}
-\let\abs=\envert
+% \setlength{\bibsep}{2pt}
+% \setlength{\bibhang}{2em}
+% \newcommand{\J}{J\"{o}reskog}
+% \newcommand{\So}{S\"{o}rbom}
+% \newcommand{\bcx}{{\bf X}}
+% \newcommand{\bcy}{{\bf Y}}
+% \newcommand{\bcz}{{\bf Z}}
+% \newcommand{\bcu}{{\bf U}}
+% \newcommand{\bcv}{{\bf V}}
+% \newcommand{\bcw}{{\bf W}}
+% \newcommand{\bci}{{\bf I}}
+% \newcommand{\bch}{{\bf H}}
+% \newcommand{\bcb}{{\bf B}}
+% \newcommand{\bcr}{{\bf R}}
+% \newcommand{\bcm}{{\bf M}}
+% \newcommand{\bcf}{{\bf F}}
+% \newcommand{\bcg}{{\bf G}}
+% \newcommand{\bcs}{{\bf S}}
+% \newcommand{\bca}{{\bf A}}
+% \newcommand{\bcd}{{\bf D}}
+% \newcommand{\bcc}{{\bf C}}
+% \newcommand{\bce}{{\bf E}}
+% \newcommand{\ba}{{\bf a}}
+% \newcommand{\bb}{{\bf b}}
+% \newcommand{\bc}{{\bf c}}
+% \newcommand{\bd}{{\bf d}}
+% \newcommand{\bx}{{\bf x}}
+% \newcommand{\by}{{\bf y}}
+% \newcommand{\bz}{{\bf z}}
+% \newcommand{\bu}{{\bf u}}
+% \newcommand{\bv}{{\bf v}}
+% \newcommand{\bh}{{\bf h}}
+% \newcommand{\bl}{{\bf l}}
+% \newcommand{\be}{{\bf e}}
+% \newcommand{\br}{{\bf r}}
+% \newcommand{\bw}{{\bf w}}
+% \newcommand{\de}{\stackrel{D}{=}}
+% \newcommand{\bt}{\bigtriangleup}
+% \newcommand{\bfequiv}{\mbox{\boldmath $\equiv$}}
+% \newcommand{\bmu}{\mbox{\boldmath $\mu$}}
+% \newcommand{\bnu}{\mbox{\boldmath $\nu$}}
+% \newcommand{\bxi}{\mbox{\boldmath $\xi$}}
+% \newcommand{\btau}{\mbox{\boldmath $\tau$}}
+% \newcommand{\bgamma}{\mbox{\boldmath $\Gamma$}}
+% \newcommand{\bphi}{\mbox{\boldmath $\Phi$}}
+% \newcommand{\bfphi}{\mbox{\boldmath $\varphi$}}
+% \newcommand{\bfeta}{\mbox{\boldmath $\eta$}}
+% \newcommand{\bpi}{\mbox{\boldmath $\Pi$}}
+% \newcommand{\bequiv}{\mbox{\boldmath $\equiv$}}
+% \newcommand{\bvarepsilon}{\mbox{\boldmath $\varepsilon$}}
+% \newcommand{\btriangle}{\mbox{\boldmath $\triangle$}}
+% \newcommand{\bdelta}{\mbox{\boldmath $\Delta$}}
+% \newcommand{\beps}{\mbox{\boldmath $\epsilon$}}
+% \newcommand{\btheta}{\mbox{\boldmath $\theta$}}
+% \newcommand{\balpha}{\mbox{\boldmath $\alpha$}}
+% \newcommand{\bsphi}{\mbox{\boldmath $\varphi$}}
+% \newcommand{\bsig}{\mbox{\boldmath $\sigma$}}
+% \newcommand{\bfpsi}{\mbox{\boldmath $\psi$}}
+% \newcommand{\bfdelta}{\mbox{\boldmath $\delta$}}
+% \newcommand{\bsigma}{{\bf \Sigma}}
+% \newcommand{\bzero}{{\bf 0}}
+% \newcommand{\bpsi}{\mbox{\boldmath $\Psi$}}
+% \newcommand{\bep}{\mbox{\boldmath $\epsilon$}}
+% \newcommand{\bomega}{\mbox{\boldmath $\Omega$}}
+% \newcommand{\bfomega}{\mbox{\boldmath $\omega$}}
+% \newcommand{\blambda}{\mbox{\boldmath $\Lambda$}}
+% \newcommand{\bflambda}{\mbox{\boldmath $\lambda$}}
+% \newcommand{\bfsigma}{\mbox{\boldmath $\sigma$}}
+% \newcommand{\bfpi}{{\mbox{\boldmath $\pi$}}}
+% \newcommand{\bupsilon}{\mbox{\boldmath $\upsilon$}}
+% \newcommand{\obs}{{\rm obs}}
+% \newcommand{\mis}{{\rm mis}}
+% \theoremstyle{plain}
+% \newtheorem{criterion}{Criterion}
+% \theoremstyle{definition}
+% \newtheorem{condition}[theorem]{Condition}
+% \usepackage[]{graphicx}
+% \chardef\bslash=`\\ % p. 424, TeXbook
+% \newcommand{\ntt}{\normalfont\ttfamily}
+% \newcommand{\cn}[1]{{\protect\ntt\bslash#1}}
+% \newcommand{\pkg}[1]{{\protect\ntt#1}}
+% \let\fn\pkg
+% \let\env\pkg
+% \let\opt\pkg
+% \hfuzz1pc % Don't bother to report overfull boxes if overage is < 1pc
+% \newcommand{\envert}[1]{\left\lvert#1\right\rvert}
+% \let\abs=\envert
 
@@
 It may seem like an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples.
-However, there may be inherent differences between cases with and cases without any missing values, due to the unknown non-response model.
-Only sampling complete cases would thus result in a simulation that fails to capture all relevant real-world conditions [which refutes the main reason for using a design-based data generating mechanism in the first place].
+However, there may be inherent differences between cases with and cases without any missing values, due to the unknown missing data model.
+Only sampling complete cases from the data set may thus result in samples that fail to capture all relevant real-world conditions [which in turn refutes the main reason for using design-based simulation].
 Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset to draw samples from.
 Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.
-Nowadays, ...
-[Add that some simulation studies do not use a `ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
-These studies apply different combinations of imputation and estimation methods on one or more benchmark data sets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
-In this context, there is no talk of retrieving the `true' but unobserved values at all.
-Refer to missing data chapter \citep{liu21}.]
+[Just to be clear: leaving the missingness as-is and inducing additional missing values to impute doesn't work here, because we wouldn't have a real comparative truth (other than CCA estimates from the large data set, which may be biased).]
 
-[Design-based model simulation??] Something else to consider is the number of samples that are drawn from the data generating mechanism.
-If evaluating missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly.
-Sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
-However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink14}.
-After all, we are interested only in the missing data mechanism and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
-Note that the conventional pooling rules \citep[cf.][p.
+After deciding upon a data generating mechanism, another influential--but often overlooked--decision awaits.
+Simulation studies on missing data methodolology have the unique option to exclude sampling variance from their evaluations, and only use the missing data generation procedure as source of Monte Carlo variation.
+After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
+Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep{vink14}.
+A lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
+[An example from our own experience is that it may be difficult to get true parameter estimates of transformed data (in this case, the mean of a variable with a skew normal distribution).
+If we would have used a simulation set-up with sampling variance, it may have led to imprecise reference values of the estimands in our simulation.
+Table \ref{table:dgm_truth} displays the source of the comparative truth in simulation studies with and without sampling variance.] There is a downside, however, in deviating from the standard simulation workflow by adapting a set-up withoput sampling variance, namely an increased potential for errors in pooling multiple imputation inferences.
+The conventional pooling rules \citep[cf.][p.
 76-77]{rubi87} do not apply for finite population inference.
 Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
-[An example from our own experience is that it may be difficult to get true parameter estimates of transformed data (in this case, the mean of a variable with a skew normal distribution).
-If we wouldn't have used a simulation set-up without sampling variance, it may have led to imprecise reference values of the estimands in our simulation.
-So, on the one hand, the typical set-up obfuscates the process that we're interested in (i.e., varying the missing data problem over simulation repetitions), while on the other hand, SEs may be incorrect if the sampling variance is not handled correctly in subsequent inferences.] % we hebben al sampling variance tussen de incomplete data sets, extra sampling variance op het complete data level vertroebelt het proces alleen maar.
-we verplaatsen liever de monte carlo naar het niveau van de missingness.
+If the sampling variance is not handled correctly, subsequent inferences may be invalid and the simulation results cannot be trusted.
 
@@
 This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness patterns and missingness mechanisms.
-Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data \citep[][p.
+Missingness patterns concern the location of missing entries across incomplete data, whereas missingness mechanisms describe the relationship between missing entries and observed data values \citep[][p.
 8]{litt20}.
-Under this definition, the missingness pattern also encompasses the proportion of missingness in a given dataset.
+[Under this definition, there is a row-wise element to the missingness pattern describing which variables are jointly observed, and a column-wise element encompassing the proportion of missingness in the data.]
 
-% L&R, p.
-8: "We find it useful to distinguish the missingness pattern, which describes which values are missing and observed in the data matrix and the missingness mechanism (or mechanisms), which concerns the relationship between missingness and the values of variables in the data matrix".
-
-Even the terminology on missingness generation can be confusing.
-Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
-In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation.
-This distinction, however, is not always clear in the literature.
-Convoluting the two concepts may lead to incorrect generalizations from simulation studies since the two are rarely equal (e.g., a proportion of 50\% incomplete cases could translate to a missingness proportion of just 25\% in bivariate data, or even 10\% in data with five variables).
+[Even the terminology on missingness generation can be confusing.
+Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?] In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation.
+The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion 25\% in both variables, or one completely observed variable and one with 50\% missingness).
 Moreover, the value of the actual missingness proportion may be diffusing too.
-Some studies use only 10\% missing data [rephrase to use missingness proportion?] where other studies push the limits to additionally investigate the performance under at least 50\% missing data.
-This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
+Some studies use only 10\% missing data [rephrase to use missingness proportion?] where other studies push the limits to additionally investigate the performance under missingness proportions of at least 50\%.
+The inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
 This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
@@
 
-Other problems with missingness patterns entail the distribution of missing values across cases (as opposed to the proportion over variables).
-[Add: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.
-For example, iterative imputation algorithms converge instantly if there is only 1 variable with missingness or monotone pattern, but require iteration in all other situations.] Specific missing data patterns may inadvertently guide which imputation method prevails in performance.
-Some ad hoc methods are known to yield valid inferences under certain restricted conditions (e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively), while biasing inferences in any other case.
+[Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.
+For example, iterative imputation algorithms converge instantly if there is only 1 variable with missingness or monotone pattern, but require iteration in all other situations.
+And very specific missing data patterns may misguide which imputation method prevails in performance: some ad hoc methods are known to yield valid inferences under certain restricted conditions (e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively), while biasing inferences in any other case.
 Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions.
+]
 
-[Missingness mechanisms: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice, 5) and finally something about types of M(N)AR (ref dance paper rianne).]
+[Missingness mechanisms paragraphs: 1) definitions of mechanisms (now in table), 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low,  4) that MNAR is often ignored, while it may be the most realistic in practice, and 5) finally add something about types of M(N)AR (ref Rianne's Dance paper).]
 
@@
 [MAR may be spurious:] With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g.
-weight is made incomplete based on gender) such that the probability to be missing is the same within groups of cases defined by the observed data (e.g.
-males are less likely to disclose their weight, but gender is observed).
-The relations in the observed data may, however, be weak or non-existent.
+weight is made incomplete based on gender to emulate a situation in which male participants are less likely to disclose their weight, but gender is observed).
+These relations in the observed data may, however, be weak or non-existent.
 If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid.
@@
 This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.
+%The problem of spurious MAR is generally overlooked in simulation studies
 
@@
 
-[MNAR is often ignored:] Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
-Although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
+[MNAR is often ignored:] Although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
+Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
 
-[Missingness type is often ignored (duplicated text from solutions section): Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
+[Missingness type is often ignored (text duplicated from solutions section):] Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
 This may influence the performance of some (but not necessarily all) imputation routines.
@@
 It would be a shame to overlook such results due to the focus on a single MAR mechanism.
+
+% Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
+The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
+If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
+Alternatively, the validity of assumptions become increasingly important when the missingness increases.
+Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
 
@@
 
-To obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}.
-The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
-
 [All imputation models should be congenial, but this is difficult to test so often ignored even in single imputation endeavor, let alone in simulation studies with hundreds of imputations.
@@
 Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies.
-Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]
+Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.] To obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}.
+% The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
 
@@
 This is not surprising as people from different fields could have a different focus on the problem at hand.
-But the choice of performance measure(s) may inadvertently distort statistical properties of the imputed data.
+[Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]
 
-[Add choice of analysis model/estimand here, since performance measure are defined wrt the estimand, e.g., focusing on univariate estimates may favor invalid ad hoc methods.]
+The choice of performance measures may inadvertently distort statistical properties of the imputed data.
+Developers often only inquire about the `accuracy' (i.e.
+how well can the method reproduce the original data).
+[Add RMSE at cell level here?]
 
-Developers often only inquire about the 'accuracy' (i.e.
-how well can the method reproduce the original data).
-[Add RMSE at cell level here?] The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
+If the goal is inference (not prediction) then you need appropriate uncertainty, for which a single imputation does not suffice.
+The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
 We are interested in the correct answer to the research question; not in the truth itself.
@@
 [So, what could go wrong by focusing on accuracy exclusively?
-See problems with RMSE of predictions.]
-
-[Add problems with RMSE here. And that if the goal is inference (not prediction) then you need appropriate uncertainty, so m = 1 does not suffice.]
+Add problems with RMSE of predictions.]
 
@@
 [In general, adhere to ADEMP \citep{morr18}.
-But there are some differences for the specific case of imputation methodology.
+But there are some differences for the specific case of evaluating imputation methodology.
 All of the parts in ADEMP are used, but a bit scrambled up.
-Table 1 outlines some of the steps to consider in simulation studies focused on evaluating imputation methodology.
+Table \ref{table:steps} outlines some of the steps to consider for the evaluation of imputation methodology.
 The rest of this section is ordered based on these steps.]
@@
 \caption{Steps to consider in imputation simulation studies.}
+\label{table:steps}
 \begin{tabular}{lll}
@@
 \hline  
-1. Set scope            & Aim(s), missing data method(s) to evaluate, number of simulation \\
-                        & repetitions \\
+1. Set scope            & Aim(s), missing data method(s) to evaluate, precision level, number \\
+                        & of simulation repetitions \\
 2. Obtain truth         & Data generating mechanism(s), estimand(s), sampling variance \\
@@
 
-[Make aim clear.
-Choose scope (type of DGM, target) and imputation methods to evaluate.
+[Substeps for the simulator include: Define aim.
+Choose scope (e.g.
+data source, target) and imputation methods to evaluate.
 Ideally, perform fully factorial design, see \citet{morr18}.
-Think about tolerable level of uncertainty to pick a number of simulation repetitions.
-Think about general set-up: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?]
+Set tolerable level of uncertainty and use this to pick a number of simulation repetitions.
+Choose general set-up/simulation flow: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?]
+
+[Add that some simulation studies do not use a `ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
+These studies apply different combinations of imputation and estimation methods on one or more benchmark data sets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
+Refer to missing data chapter \citep{liu21}.
+In this context, there is no talk of retrieving the `true' but unobserved values at all.
+The DGM is by definition design-based and there is no sampling variance.
+The missingness generation step is skipped and the only performance measure that can be used between methods is RMSE and maybe comparison with repect to CCA in terms of bias.]
 
@@
 
-In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations.
-In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations.
+[Choose sampling variance:] The simulator must decide whether there is a need for sampling variance in the simulations scheme.
+For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism.
+The parameters of that single complete set will serve as comparative truth.
+This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
+Be carefull to use alternative pooling rules for finite population inference \citep{raghunathan2003multiple,vink14}.
+If sampling variance cannot be omitted from the simulation scheme, draw data repeatedly and apply conventional pooling rules \citep[cf.][p.
+76-77]{rubi87}.
 
-The design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
+[Choose DGM:] The design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
 Applications of design-based simulation are often found in official statistics.
 A benefit of design-based simulation is the ability to use real-life observed data structures.
-
 For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
@@
 
-The simulator must decide whether there is a need for sampling variance in the simulations scheme.
-For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism.
-The parameters of that single complete set will serve as comparative truth.
-This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
-Use alternative pooling rules for finite population inference \citep{raghunathan2003multiple,vink14}.
-If sampling variance cannot be omitted from the simulation scheme, draw data repeatedly and apply conventional pooling rules \citep[cf.][p.
-76-77]{rubi87}.
+[Obtain truth: In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations.
+In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations.
+But this is not a complete picture, see Table \ref{dgm_truth}.]
+
+\begin{table}[htb]
+\begin{center}
+\caption{Comparative truth in model-based and design-based simulation.}
+\label{table:dgm_truth}
+\begin{tabular}{lll}
+\hline
+               & With sampling variance      & Without sampling variance \\
+               & (multiple samples drawn)    & (one sample drawn) \\
+\hline  
+Model-based    & Truth from model            & Truth from sample \\
+Design-based   & Truth from large data set   & Truth from sample \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-[Missingness mechanisms]
+[Missingness mechanisms:] 
 First, one should always consider MCAR missingness, i.e.
@@
 % Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
-
 Next, missing data should be induced conform a model that is dependent on the observed data.
@@
 cf.][model 2 and model 3]{shara2015randomly} can be useful.
-
 [Although often ignored, it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
@@
 
-[Missingness type]
+[Missingness type:] 
 Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
@@
 
-[Missingness proportion]
+[Missingness proportion: ]
 Fourth, the amount of missingness must be varied.
@@
 We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness see Table 2.
+Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.
 
@@
 \end{table}
-
-Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.
 
@@
 
+Some suggested minimal requirements for imputation methods: the absence of algorithmic non-convergence, the imputation model fits the observed parts of the data (PPC), under a MCAR mechanism the imputed values are distributed equivalent to the observed values.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
-
-Bias: Results should preferable be unbiased.
+ (i) Bias: Results should preferable be unbiased.
 However, the way bias is considered can greatly influence the interpretation of the results.
@@
 The way bias is computed should therefore be carefully chosen and described.
-
-Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
+(ii) Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
 Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
-
-Interval width: The way the confidence interval is calculated should be described.
+(iii) Interval width: The way the confidence interval is calculated should be described.
 Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference.
@@
 
-RMSE: [RMSE at cell level and at prediction level (if applicable).]
+[Add RMSE at cell level and at prediction level (if applicable).]
+
+Some suggested minimal requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis (i.e.
+performance after list-wise deletion).
 
@@
 
-Be explicit in choices, better even to share the script directly.
-
+[Be explicit in choices.
 Describe the simulation textually, use pseudo-code, a flowchart, visualize missingness and simulation results.
-
-The used missingness mechanism should be detailed, either graphically or written as a function of the data.
-
-The problem of spurious MAR is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.
+For example, the used missingness mechanism should be detailed, either graphically or written as a function of the data.
+And the procedure to generate missing data is often insufficiently described in resulting publications.
+]
 
@@
 
-1. set-up
+1. Set-up
 
@@
 
-- number of simulations
+- target/estimand + analysis model
 
-2. data generation
+- precision level + number of simulations
+
+2. Data generation
 
@@
 
-- data types
-
-- sample vs population (vs bootstrapping??)
+- sampling variance
 
@@
 
+- data types + structures (e.g. clustering/timeseries)
+
 - coherence between variables
 
-3. missingness generation
+3. Missingness generation
 
-- monotone/univariate vs multivariate
+- missingness mechanism + type (if applicable)
 
-- missingness mechanism
+- missingness patterns (rows = univariate/multivariate; colums = proportion)
 
-- type of M(N)AR
+4. Missing data methods
 
-- missingness proportion
+- imputation methods + list-wise deletion
 
-4. missing data methods
+- imputation method parameters (e.g. number of imputations)
 
-- imputation method
+- imputation model parameters (e.g. predictor matrix)
 
-- imputation model
+5. Imputation evaluation
 
-- number of imputations
+- imputation process (e.g. non-convergence)
 
-- number of iterations
+- imputation model fit (PPC)
 
-5. performance evaluation
+- distribution of imputations (e.g. plausibility)
 
-- analysis model
+- imputation accuracy (e.g. RMSE at cell level)
 
-- estimand
+6. Performance evaluation
 
-- diagnostic
+- statistical properties (e.g. confidence validity)
 
-- ppc
+- comparative performance (e.g. RMSE of predictions)
+
+7. Report
 

### 2022-03-29 Formatted/Formatted.tex
Integrated feedback Gerko 27/3

@@
 by the detailed manual created by the National Research Council \citep*{little2012prevention}.
-Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing imputation techniques.
+Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers and data scientists are embracing imputation techniques.
 After all, the principle of imputation is very intuitive.
 
-The idea behind imputation is to impute (fill in) missing data, to obtain a valid estimate of what could have been.
+The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.
 A popular variant is multiple imputation \citep{rubi76}, whereby each missing value is imputed multiple times.
@@
 The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
-In general, it holds that modeling missing data becomes more challenging when the amount of missingness increases [ref?].
+In general, it holds that modeling missing data becomes more challenging when the amount of missingness increases.
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
-In that case, imputation would be substantially more efficient than the ubiquitous list-wise deletion.
+In that case, imputation would be substantially more efficient than the ubiquitous complete case analysis.
 
@@
 In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods.
-However, no golden standard has been established to evaluate imputation routines and, as a result, the validity of the simulation procedures may differ tremendously from one developer to another.
+However, no golden standard has been established to evaluate imputation routines and, as a result, the validity and aim of the simulation set-ups may differ tremendously from one developer to another.
 Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced.
@@
 To demonstrate the broad impact of these problems, we subdivide the problems into the following three distinct categories: problems with data generation, problems with missingness generation, and problems with performance evaluation.
-We further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines [this sentence could be smoother].
+We further detail the impact each of these problems may have on the validity of our evaluations.
 
@@
 Data generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
-There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as from official statistics records.
+There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as from official registers.
 
@@
 with consistently high correlations between groups of variables.
-This results in simulated data that contain such valuable information structures that, no matter what type of missingness is subsequently induced, the observed parts of the data still hold much (if not all) of the information about the missing part.
+This results in simulated data that contain such valuable information structures that, no matter what type of missingness would subsequently be induced, the observed parts of the data will still hold much (if not all) of the information about the missing part.
 Unsurprisingly, the performance of any imputation method will then be evaluated as good.
@@
 An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be very challenging.
-Most real-world data contains at least one missing entry, for which the true underlying missing data model is by definition unknown.
+Most real-world data contains at least some missing entries, for which the true underlying missing data model is--by definition--unknown.
 Therefore, the simulator needs to deal with missingness in \textit{some} way before incomplete empirical data can serve as comparative truth in the simulations.
@@
 However, there may be inherent differences between cases with and cases without any missing values, due to the unknown missing data model.
-Only sampling complete cases from the data set may thus result in samples that fail to capture all relevant real-world conditions [which in turn refutes the main reason for using design-based simulation].
+Only sampling complete cases from the data set may thus result in samples that fail to capture all relevant real-world conditions, which in turn refutes the main reason for using design-based simulation.
 Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset to draw samples from.
 Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.
-[Just to be clear: leaving the missingness as-is and inducing additional missing values to impute doesn't work here, because we wouldn't have a real comparative truth (other than CCA estimates from the large data set, which may be biased).]
+Just to be clear: leaving the missingness as-is and inducing additional missing values to impute could not work here, because we wouldn't have a real and unbiased comparative truth.
 
@@
 After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
-Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep{vink14}.
+Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}.
 A lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
-[An example from our own experience is that it may be difficult to get true parameter estimates of transformed data (in this case, the mean of a variable with a skew normal distribution).
-If we would have used a simulation set-up with sampling variance, it may have led to imprecise reference values of the estimands in our simulation.
-Table \ref{table:dgm_truth} displays the source of the comparative truth in simulation studies with and without sampling variance.] There is a downside, however, in deviating from the standard simulation workflow by adapting a set-up withoput sampling variance, namely an increased potential for errors in pooling multiple imputation inferences.
-The conventional pooling rules \citep[cf.][p.
-76-77]{rubi87} do not apply for finite population inference.
+The only required change to simulation set-ups is to remove sampling variance from performance evaluation.
+We highly value the additional flexibility that such a simulation set-up offers.
+Especially in the case of data transformations it can be challenging to derive the true parametric references to evaluate against.
+Simply using a single generated complete set in which simulated missingness is induced and evaluating against that generated set awards a plethora of procedural problems and computational challenges.
+
+% Table \ref{table:dgm_truth} displays the source of the comparative truth in simulation studies with and without sampling variance.When deviating conventional pooling rules \citep[cf.][p.
+76-77]{rubi87} do not apply.
 Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
@@
 Evaluating imputation methodology requires a missing data problem to be solved.
-After establishing a comparative truth from a data generating mechanism, some form of missingness has to be induced.
+After establishing a comparative truth from a data generating mechanism, some form of missingness therefore has to be induced.
 Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like
@@
 8]{litt20}.
-[Under this definition, there is a row-wise element to the missingness pattern describing which variables are jointly observed, and a column-wise element encompassing the proportion of missingness in the data.]
+Under this definition, there is a row-wise element to the missingness pattern describing which variables are jointly observed, and a column-wise element encompassing the proportion of missingness in the data.
 
-[Even the terminology on missingness generation can be confusing.
-Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?] In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation.
-The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion 25\% in both variables, or one completely observed variable and one with 50\% missingness).
+Even the terminology on missingness generation can be confusing.
+Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
+In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation.
+The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).
 Moreover, the value of the actual missingness proportion may be diffusing too.
-Some studies use only 10\% missing data [rephrase to use missingness proportion?] where other studies push the limits to additionally investigate the performance under missingness proportions of at least 50\%.
+Some studies use only 10\% missingness where other studies push the limits to additionally investigate the performance under missingness proportions of at least 50\%.
 The inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
@@
 
-[Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.
-For example, iterative imputation algorithms converge instantly if there is only 1 variable with missingness or monotone pattern, but require iteration in all other situations.
-And very specific missing data patterns may misguide which imputation method prevails in performance: some ad hoc methods are known to yield valid inferences under certain restricted conditions (e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively), while biasing inferences in any other case.
-Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions.
-]
+[Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.] For example, iterative imputation algorithms converge instantly if there is only one variable with missingness or when the pattern is monotone, but require iteration in all other situations.
+And very specific missing data patterns may misguide which imputation method prevails in performance: some ad hoc methods are known to yield valid inferences under certain restricted conditions \citep[e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively][]{fimd}, while biasing inferences in other cases.
+Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to suboptimal advice in practice.
 
-[Missingness mechanisms paragraphs: 1) definitions of mechanisms (now in table), 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low,  4) that MNAR is often ignored, while it may be the most realistic in practice, and 5) finally add something about types of M(N)AR (ref Rianne's Dance paper).]
+% [Missingness mechanisms paragraphs: 1) definitions of mechanisms (now in table), 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low,  4) that MNAR is often ignored, while it may be the most realistic in practice, and 5) finally add something about types of M(N)AR (ref Rianne's Dance paper).]
 
@@
 
+The missingness mechanism is usually assumed to be random (MAR) or completely random (MCAR). The validity of the assumptions becomes increasingly important when the missingness proportion increases.
+
 [MCAR is important but often ignored:] MCAR is a necessary simulation condition for evaluating the performance of imputation procedures.
@@
 [MAR may be spurious:] With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g.
-weight is made incomplete based on gender to emulate a situation in which male participants are less likely to disclose their weight, but gender is observed).
+weight is made incomplete based on observed gender to emulate a situation wherin one gender is less likely to disclose their weight).
 These relations in the observed data may, however, be weak or non-existent.
 If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid.
-In the most extreme case, MAR is induced from data without multivariate relations.
+The most extreme example would be when MAR is induced from data without multivariate relations.
 The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random.
@@
 Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
+% The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
 
@@
 
-[Add bridge from last section.
-And explain how this part is subdivided: first problems with evaluation of imputations (general characteristics of the imputations such as convergence and plausibility), then evaluation of performance (in terms of actual performance measures).]
+The evaulation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e.
+the complete data model used to estimate a target or estimand).
+When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
 
-[All imputation models should be congenial, but this is difficult to test so often ignored even in single imputation endeavor, let alone in simulation studies with hundreds of imputations.
+All imputation models should be congenial, but this is difficult to test and often ignored.
 Checks of imputation model fit are often omitted.
@@
 Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies.
-Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.] To obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}.
-% The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
+Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]
 
@@
 
-[Add segway between evaluation of imputation and actual performance measures.]
-The evaluation criteria used to assess imputation performance vary from one developer to another.
+% [Add segway between evaluation of imputation and actual performance measures.]
+The evaluation criteria used to assess imputation performance vary from one simulator to another.
 This is not surprising as people from different fields could have a different focus on the problem at hand.
-[Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]
+Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.
+For many outcome-driven imputation taks a single valid imputation may suffice.
+Measures of predictive accuracy or the RMSE (root mean squared error) of predicted values could then serve as sensible evaluation criteria.
 
-The choice of performance measures may inadvertently distort statistical properties of the imputed data.
+% [Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]
+% The choice of performance measures may inadvertently distort statistical properties of the imputed data.
 Developers often only inquire about the `accuracy' (i.e.
@@
 
-If the goal is inference (not prediction) then you need appropriate uncertainty, for which a single imputation does not suffice.
-The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-We are interested in the correct answer to the research question; not in the truth itself.
+If the goal is inference--not prediction--then the uncertainty about estimates needs to be quantified.
+A single imputation would then not suffice.
+The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
 This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.
@@
 591]{neym34}.
-[So, what could go wrong by focusing on accuracy exclusively?
-Add problems with RMSE of predictions.]
+Merely focussing on distance measures would open the door for invalid inferences.
 
@@
 
-[In general, adhere to ADEMP \citep{morr18}.
-But there are some differences for the specific case of evaluating imputation methodology.
-All of the parts in ADEMP are used, but a bit scrambled up.
-Table \ref{table:steps} outlines some of the steps to consider for the evaluation of imputation methodology.
-The rest of this section is ordered based on these steps.]
+Standardizing the evaluation of imputation methodology requires simulators--including ourselves--to consider a couple of key aspects in their simulation workflows.
+Table \ref{table:steps} outlines some suggested steps to adopt.
+Please note that there is--of course--no 'silver bullet' and that we do not claim to present a universally applicable approach.
+Our suggestions are the accumulated result of scientific literature, research experience, and discussion.
+For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}.
+We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology.
 
@@
 \hline
-Step                    & Aspects under consideration \\
+Step                    & Simulation aspects under consideration \\
 \hline  
@@
 [Choose sampling variance:] The simulator must decide whether there is a need for sampling variance in the simulations scheme.
-For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism.
-The parameters of that single complete set will serve as comparative truth.
+If sampling variance is not of interest, the procedure outlined in section XYZ can be used to simulate missingness in a single complete data set.
+% For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism. 
+The parameters of the complete sample will serve as comparative truth during the simulation.
 This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
-Be carefull to use alternative pooling rules for finite population inference \citep{raghunathan2003multiple,vink14}.
+[insert part about transformed data here?].
+When deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
+76-77]{rubi87} do not apply.
+Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
+Table \ref{table:dgm_truth} displays the source of the comparative truth in simulation studies with and without sampling variance.
 If sampling variance cannot be omitted from the simulation scheme, draw data repeatedly and apply conventional pooling rules \citep[cf.][p.
@@
 
-This document is aimed at establishing a common ground for the evaluation of imputation routines.
+This manuscript aims to establish a common ground for the evaluation of imputation routines.
 Such a common ground would be the basis of a standardized evaluation.
-This allows for fair and efficient comparisons between imputation techniques.
+This would allow for fairer and more efficient comparisons between imputation techniques.
 Ultimately, it would be desirable to evaluate every imputation routine against the same standardized set in order to quantify the statistical properties across imputation routines.

### 2022-03-29 Formatted/Formatted.tex
Write 'set scope' subsection

@@
 
-[Substeps for the simulator include: Define aim.
-Choose scope (e.g.
-data source, target) and imputation methods to evaluate.
-Ideally, perform fully factorial design, see \citet{morr18}.
-Set tolerable level of uncertainty and use this to pick a number of simulation repetitions.
-Choose general set-up/simulation flow: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?]
-
+Before setting up their simulation study, the simulator should clearly define the scope of their evaluations.
+Evaluating imputation methodology may entail e.g.
+comparisons between several imputation methods, investigations of the validity of a single (novel) method, or an application to varying characteristics of a specific use case.
+The choice of imputation method(s) under evaluation and the simulation study's aims are highly intertwined.
 [Add that some simulation studies do not use a `ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
@@
 The missingness generation step is skipped and the only performance measure that can be used between methods is RMSE and maybe comparison with repect to CCA in terms of bias.]
+
+The scope of the study also encompasses the general simulation design.
+Ideally, one should perform fully factorial design, where each simulation condition is varied against all other conditions.
+This is important because there are known interaction effects between e.g.
+missingness proportions and missingness mechanisms.
+If multiple imputation methods are under evaluation (and/or parallel computing is used), it may be wise to split-up the simulation workflow into separate parts, such that each imputation method is applied to the same incomplete dataset within every simulation repetition.
+
+Other aspects to consider at this stage are the required precision level of the study and the number of simulation repetitions.
+The tolerable level of uncertainty in terms of Monte Carlo error for each of the performance measures may be used to pick a number of simulation repetitions \citep[see e.g.][]{morr18}.
 

### 2022-03-29 Formatted/Formatted.tex
Write 'obtain truth' subsection

@@
 Especially in the case of data transformations it can be challenging to derive the true parametric references to evaluate against.
-Simply using a single generated complete set in which simulated missingness is induced and evaluating against that generated set awards a plethora of procedural problems and computational challenges.
+Simply using a single generated complete set in which simulated missingness is induced and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
 
@@
 \end{quote}
-This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness patterns and missingness mechanisms.
-Missingness patterns concern the location of missing entries across incomplete data, whereas missingness mechanisms describe the relationship between missing entries and observed data values \citep[][p.
+This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness mechanisms and missingness patterns.
+Missingness mechanisms describe the relationship between missing entries and observed data values \citep[generally categorized into MCAR, MAR and MNAR mechanisms cf][see Table \ref{table:mech}]{rubi76}.
+Missingness patterns concern the location of missing entries across incomplete data \citep[][p.
 8]{litt20}.
@@
 \caption{Missingness mechanisms. [Add PMAR]}
+\label{table:mech}
 \begin{tabular}{lll}
@@
 
-[Missingness type is often ignored (text duplicated from solutions section):] Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
+[Missingness type is often ignored:] Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
 This may influence the performance of some (but not necessarily all) imputation routines.
@@
 Table \ref{table:steps} outlines some suggested steps to adopt.
-Please note that there is--of course--no 'silver bullet' and that we do not claim to present a universally applicable approach.
+Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach.
 Our suggestions are the accumulated result of scientific literature, research experience, and discussion.
@@
 The choice of imputation method(s) under evaluation and the simulation study's aims are highly intertwined.
+
 [Add that some simulation studies do not use a `ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
@@
 
-[Decide which data generation mechanism to use and obtain true value of the estimand(s).]
+The simulator should choose their data generating mechanism(s) and estimand(s) in line with the study scope.
+Model-based data-generating mechanisms have the advantage of flexibility and precision, since data are generated from a probability distribution or statistical model and the true theoretical parameters can be known.
+A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
+A benefit of design-based simulation is the ability to use real-life observed data structures.
 
-[Choose sampling variance:] The simulator must decide whether there is a need for sampling variance in the simulations scheme.
-If sampling variance is not of interest, the procedure outlined in section XYZ can be used to simulate missingness in a single complete data set.
-% For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism. 
-The parameters of the complete sample will serve as comparative truth during the simulation.
+The comparative truth of the estimand(s) is determined by the choice of data generating mechanism and whether sampling variance is included in the simulation set-up, see Table \ref{table:dgm_truth}.
+If sampling variance cannot be omitted from the simulation scheme, multiple samples are drawn from the data generating mechanism (i.e., one sample per simulation repetition, a standard simulation set-up).
+If sampling variance is not of interest, a single complete data set can be obtained from the data generating mechanism (i.e., one sample for all simulation repetitions).
 This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
-[insert part about transformed data here?].
+We highly value the additional flexibility that such a simulation set-up offers.
+Especially in the case of data transformations it can be challenging to derive the true parametric references to evaluate against.
+Simply using a single generated complete set in which simulated missingness is induced and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
+% The simulator should first choose an appropriate data generating mechanism and then decide if sampling variance can be omitted from the simulation scheme to draw one or more samples from the data generating mechanism, respectively.
 When deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
@@
 Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
-Table \ref{table:dgm_truth} displays the source of the comparative truth in simulation studies with and without sampling variance.
-If sampling variance cannot be omitted from the simulation scheme, draw data repeatedly and apply conventional pooling rules \citep[cf.][p.
-76-77]{rubi87}.
-
-[Choose DGM:] The design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
-Applications of design-based simulation are often found in official statistics.
-A benefit of design-based simulation is the ability to use real-life observed data structures.
-For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
-This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
-Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
-However, the real-world missingness would then be omitted.
-
-[Obtain truth: In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations.
-In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations.
-But this is not a complete picture, see Table \ref{dgm_truth}.]
 
@@
 \begin{center}
-\caption{Comparative truth in model-based and design-based simulation.}
+\caption{Source of the comparative truth in simulation studies.}
 \label{table:dgm_truth}
@@
 \hline  
-Model-based    & Truth from model            & Truth from sample \\
-Design-based   & Truth from large data set   & Truth from sample \\
+Model-based simulation   & data-generating model         & single sample \\
+Design-based simulation  & sufficiently large data set   & single sample \\
 \hline
@@
 A straightforward technique for inducing different forms of univariate MAR missingness is described in \citet[][p.
-63]{fimd} and a generalization to multivariate MAR missingness can be found in \citet[][Appendix B]{buur06} and \citet[][\S 5.2.3]{brand1999development}.
+63]{fimd} and a generalization to multivariate MAR missingness can be found in \citet{ampute}.
 If the missingness is to be induced in longitudinal data, autoregressive MAR models \citep[e.g.
@@
 Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
-% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
+The eefects of different forms of MAR mechanisms are described in \citet{scho18}.% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
 This may influence the performance of some (but not necessarily all) imputation routines.
@@
 
+It could be convincing to demonstrate a method's applicability to real-world missing data problems.
+This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
+Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
+However, the real-world missingness would then be omitted.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-03-29 Formatted/Formatted.tex
remove some duplicate info from missingness paragraphs in problems and solutions sections

@@
 Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}.
+The only required change to simulation set-ups is to remove sampling variance from performance evaluation.
 A lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
-The only required change to simulation set-ups is to remove sampling variance from performance evaluation.
-We highly value the additional flexibility that such a simulation set-up offers.
-Especially in the case of data transformations it can be challenging to derive the true parametric references to evaluate against.
-Simply using a single generated complete set in which simulated missingness is induced and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
 
@@
 
+The missingness mechanism is usually assumed to be random (MAR) or completely random (MCAR).
+
 \begin{table}[htb]
@@
 \hline  
-MCAR  & The probability to be missing is the same for all cases. In other words, the missing \\
-      & values are missing at random and the observed values are observed at random. \\
-MAR   & The probability to be missing is the same within groups of cases defined by the \\
-      & observed data only. The essence is that observed relations in the data inform the \\
-      & missingness, such that MCAR may be assumed within the observed groups. \\
-MNAR  & The probability to be missing depends on unobserved aspects of the data. The \\
-      & cause of the missingness is unknown and cannot be inferred from the observed data. \\
-      & This is considered non-ignorable missingness \citep[see e.g.][]{rubi76}. \\
+MCAR  & Missing completely at random: The probability to be missing is the same for all \\
+      & cases. In other words, the missing values are missing at random and the observed \\
+      & values are observed at random. \\
+MAR   & Missing at random: The probability to be missing is the same within groups of \\
+      & cases defined by the observed data only. The essence is that observed relations in \\
+      & the data inform the missingness, such that MCAR may be assumed within the \\ 
+      & observed groups. \\
+MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
+      & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
+      & inferred from the observed data. This is considered non-ignorable missingness. \\
 \hline
@@
 \end{table}
-
-The missingness mechanism is usually assumed to be random (MAR) or completely random (MCAR). The validity of the assumptions becomes increasingly important when the missingness proportion increases.
 
@@
 
-All imputation models should be congenial, but this is difficult to test and often ignored.
+[All imputation models should be congenial, but this is difficult to test and often ignored.
 Checks of imputation model fit are often omitted.
@@
 Ideally, one should perform fully factorial design, where each simulation condition is varied against all other conditions.
-This is important because there are known interaction effects between e.g.
-missingness proportions and missingness mechanisms.
+This may be important because of known interaction effects, e.g.
+the validity of the missingness mechanism assumptions becomes increasingly important when the missingness proportion increases.
 If multiple imputation methods are under evaluation (and/or parallel computing is used), it may be wise to split-up the simulation workflow into separate parts, such that each imputation method is applied to the same incomplete dataset within every simulation repetition.
@@
 The simulator should choose their data generating mechanism(s) and estimand(s) in line with the study scope.
-Model-based data-generating mechanisms have the advantage of flexibility and precision, since data are generated from a probability distribution or statistical model and the true theoretical parameters can be known.
+Model-based data-generating mechanisms have the advantages of flexibility and precision since data are generated from a known statistical model so the true theoretical parameters can be derived.
 A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
@@
 
-[Missingness mechanisms:] 
-First, one should always consider MCAR missingness, i.e.
+Missingness should be induced to generate different missing data conditions.
+[Discussed are missingness mechanisms incl different types (see Table \ref{table:mech} for definitions), missingness patterns incl different missingness proportions (see Table \ref{table:prop} for suggested proportions).]
+
+First, one should always consider MCAR missingness [as reference, because all imputation methods should yield valid inference under MCAR and we can use this mechanism to check imputation model fit] %, i.e.
 the scenario where the missing values are missing at random and the observed values are observed at random.
@@
 cf.][model 2 and model 3]{shara2015randomly} can be useful.
+Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
+The effects of different forms of MAR mechanisms are described in \citet{scho18}.% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
+This may influence the performance of some (but not necessarily all) imputation routines.
+For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+It would be a shame to overlook results due to the focus on a single MAR mechanism.
 [Although often ignored, it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
@@
 
-[Missingness type:] 
-Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
-The eefects of different forms of MAR mechanisms are described in \citet{scho18}.% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
-This may influence the performance of some (but not necessarily all) imputation routines.
-For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
-It would be a shame to overlook results due to the focus on a single MAR mechanism.
-
-[Missingness proportion: ]
 Fourth, the amount of missingness must be varied.
@@
 \caption{Suggested univariate missingness proportions.}
+\label{table:prop}
 \begin{tabular}{lll}

### 2022-03-30 Formatted/Formatted.tex
Write out bullet point sentences from the problems with missingness, set scope and induce missingness subsections

@@
 8]{litt20}.
-Under this definition, there is a row-wise element to the missingness pattern describing which variables are jointly observed, and a column-wise element encompassing the proportion of missingness in the data.
+Under this definition, there is a row-wise element to the missingness pattern describing which variables are jointly observed, and a column-wise element encompassing the amount of missingness in the data.
 
@@
 
-[Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.] For example, iterative imputation algorithms converge instantly if there is only one variable with missingness or when the pattern is monotone, but require iteration in all other situations.
+Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables.
+It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways.
+Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to suboptimal advice in practice.
+For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that complete case analysis outperforms their imputation method(s).
+Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described by][]{fimd}, while biasing inferences in other cases.
+Another example of a missingness pattern that may inadverted impose assumptions on the evaluations is a monotone pattern \citep[i.e.
+a pattern with uniformly increasing missingness proportions along variables][]{litt20}.
+If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration in all other situations.
+A final example is multivariate missingness that is generated using stepwise univariate missingness induction.
+The resulting incomplete data may not have the intended or expected statistical properties \citep[e.g.
+skewness][]{ampute}, which complicates definitive conclusions about imputation method performance in relation to the data generating mechanism.
+
+% is the   [Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.] For example, iterative imputation algorithms converge instantly if there is only one variable with missingness or when the pattern is monotone, but require iteration in all other situations.
 And very specific missing data patterns may misguide which imputation method prevails in performance: some ad hoc methods are known to yield valid inferences under certain restricted conditions \citep[e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively][]{fimd}, while biasing inferences in other cases.
-Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to suboptimal advice in practice.
-
 % [Missingness mechanisms paragraphs: 1) definitions of mechanisms (now in table), 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low,  4) that MNAR is often ignored, while it may be the most realistic in practice, and 5) finally add something about types of M(N)AR (ref Rianne's Dance paper).]
-
 % MCAR missingness: The scenario where the missing values are missing at random and the observed values are observed at random. 
@@
 
-\begin{table}[htb]
+\begin{table}[tb]
 \begin{center}
@@
 
-Standardizing the evaluation of imputation methodology requires simulators--including ourselves--to consider a couple of key aspects in their simulation workflows.
+Standardizing the evaluation of imputation methodology requires simulators--like ourselves--to consider a couple of key aspects in their simulation workflows.
 Table \ref{table:steps} outlines some suggested steps to adopt.
@@
 \hline
-Step                    & Simulation aspects under consideration \\
+&Step                    & Simulation aspects under consideration \\
 \hline  
-1. Set scope            & Aim(s), missing data method(s) to evaluate, precision level, number \\
-                        & of simulation repetitions \\
-2. Obtain truth         & Data generating mechanism(s), estimand(s), sampling variance \\
-3. Induce missingness   & Missingness mechanism(s), missing data pattern(s), missingness \\
-                        & proportion(s) \\
-4. Apply methods        & Imputation model(s), analysis model(s) \\
-5. Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
-                        & model fit \\
-6. Evaluate performance & Performance measures \\
-7. Report               & Text, visualization, checklist, simulation script \\
+1&Set scope            & Aim(s), missing data method(s) to evaluate, simulation design \\
+2&Obtain truth         & Data generating mechanism(s), estimand(s), sampling variance \\
+3&Induce missingness   & Missingness mechanism(s), missing data pattern(s) \\
+4&Apply methods        & Imputation model(s), analysis model(s) \\
+5&Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
+ &                      & model fit \\
+6&Evaluate performance & Performance measures, qualitative judgement \\
+7&Report               & Text, visualization, checklist, simulation script \\
 \hline
@@
 
-Before setting up their simulation study, the simulator should clearly define the scope of their evaluations.
-Evaluating imputation methodology may entail e.g.
-comparisons between several imputation methods, investigations of the validity of a single (novel) method, or an application to varying characteristics of a specific use case.
-The choice of imputation method(s) under evaluation and the simulation study's aims are highly intertwined.
+Before setting up a simulation study, the simulator should clearly define the scope of their evaluations.
+A simulation study aimed at imputation model comparison will typically have a  different design than one aimed at establishing the inferential validity of a single (novel) method.
+The choice of imputation method(s) under evaluation is naturally intertwined with the simulation study's aims and set-up.
 
-[Add that some simulation studies do not use a `ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
-These studies apply different combinations of imputation and estimation methods on one or more benchmark data sets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
-Refer to missing data chapter \citep{liu21}.
+One specific simulation study aim (that we have thus far ignored) is to evaluate the predictive performance of pairs of imputation methods and estimation methods in empirical, incomplete data.
+These kinds of simulation studies do not start out from a complete data set in which missingness is induced by the simulator.
+Rather, the evaluations use the observed values of an outcome variable (or target) as the estimand, which is estimated from the incomplete predictor (or feature) space by first imputing the missingness and then applying a prediction method.
+Pairs of imputation and prediction methods with high predictive accuracy in one or more benchmark data sets are deemed as `good' \citep{liu21}.
+Note that such a simulation design does not have a `ground truth', so only the comparative performance of methods may be established.
+We do not recommend this approach if the inferential validity of the imputations is of interest.
+
+% RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
+Refer to missing data chapter.
 In this context, there is no talk of retrieving the `true' but unobserved values at all.
@@
 
-The scope of the study also encompasses the general simulation design.
-Ideally, one should perform fully factorial design, where each simulation condition is varied against all other conditions.
-This may be important because of known interaction effects, e.g.
-the validity of the missingness mechanism assumptions becomes increasingly important when the missingness proportion increases.
-If multiple imputation methods are under evaluation (and/or parallel computing is used), it may be wise to split-up the simulation workflow into separate parts, such that each imputation method is applied to the same incomplete dataset within every simulation repetition.
-
-Other aspects to consider at this stage are the required precision level of the study and the number of simulation repetitions.
-The tolerable level of uncertainty in terms of Monte Carlo error for each of the performance measures may be used to pick a number of simulation repetitions \citep[see e.g.][]{morr18}.
+Some general choices in the simulation set-up to consider at this stage are the simulation design and the number of simulation repetitions.
+The simulation design should ideally be fully factorial, i.e.
+varying each simulation condition against all other conditions.
+There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions).
+If there are several imputation method under evaluation, the simulation design should accomodate that each method is applied to each incomplete data set.
+This is in contrast to a design with as many incomplete data sets as imputation methods in every simulation repetition.
+Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation which makes for fairer comparisons.
+The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
+as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
 
@@
 
-The simulator should choose their data generating mechanism(s) and estimand(s) in line with the study scope.
-Model-based data-generating mechanisms have the advantages of flexibility and precision since data are generated from a known statistical model so the true theoretical parameters can be derived.
+The simulator should choose their data generating mechanism(s) in line with the study scope.
+Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived.
 A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
-A benefit of design-based simulation is the ability to use real-life observed data structures.
+The benefit of design-based simulation is the ability to use real-life observed data structures.
+
+Estimand(s) or other simulation targets should defined in the context of the study aim and data-generating mechanism.
+[As a general rule, it would be wise to include one or more multivariate estimands.
+It is a realistic expectation for imputation methods to yield valid univariate estimates but also to capture multivariable relations in the data.
+Imputation methods that fail to do so, should not be considered general purpose methods (e.g., mean imputation).]
 
 The comparative truth of the estimand(s) is determined by the choice of data generating mechanism and whether sampling variance is included in the simulation set-up, see Table \ref{table:dgm_truth}.
-If sampling variance cannot be omitted from the simulation scheme, multiple samples are drawn from the data generating mechanism (i.e., one sample per simulation repetition, a standard simulation set-up).
+If sampling variance cannot be omitted from the simulation scheme, multiple samples should be drawn from the data generating mechanism (i.e., one sample per simulation repetition, a standard simulation set-up).
 If sampling variance is not of interest, a single complete data set can be obtained from the data generating mechanism (i.e., one sample for all simulation repetitions).
@@
 Especially in the case of data transformations it can be challenging to derive the true parametric references to evaluate against.
-Simply using a single generated complete set in which simulated missingness is induced and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
+Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
 % The simulator should first choose an appropriate data generating mechanism and then decide if sampling variance can be omitted from the simulation scheme to draw one or more samples from the data generating mechanism, respectively.
@@
 
-Missingness should be induced to generate different missing data conditions.
-[Discussed are missingness mechanisms incl different types (see Table \ref{table:mech} for definitions), missingness patterns incl different missingness proportions (see Table \ref{table:prop} for suggested proportions).]
+Missingness should be induced according to several sets of missing data conditions.
+We encourage simulators to consider different missingness mechanisms (including missingness types), and different missingness patterns (including missingness proportions).
+Although not every missingness mechanism is realistically assumed in practice, they can all offer valuable insights as simulation condition.
 
-First, one should always consider MCAR missingness [as reference, because all imputation methods should yield valid inference under MCAR and we can use this mechanism to check imputation model fit] %, i.e.
+First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions).
+This mechanism may be used as a reference condition.
+Valid imputations under MCAR is a minimal requirement for any [multipurpose] imputation method.
+We also know that, under MCAR, the distribution of imputed values should be equivalent to the distribution of observed values for a given variable, which offers an indication of imputation model (mis)fit.
+% First, one should always consider MCAR missingness [as reference, because all imputation methods should yield valid inference under MCAR and we can use this mechanism to check imputation model fit] %, i.e.
 the scenario where the missing values are missing at random and the observed values are observed at random.
 % Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
-Next, missing data should be induced conform a model that is dependent on the observed data.
-A straightforward technique for inducing different forms of univariate MAR missingness is described in \citet[][p.
-63]{fimd} and a generalization to multivariate MAR missingness can be found in \citet{ampute}.
-If the missingness is to be induced in longitudinal data, autoregressive MAR models \citep[e.g.
+
+Next, missing data should be induced conform a model that is dependent on the observed data (i.e.
+a MAR mechanism).
+Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully.
+A straightforward technique for inducing univariate MAR missingness is described in \citet[][p.
+63]{fimd}, generalizations to multivariate MAR missingness can be found in \citet{ampute}, and if the missingness is to be induced in longitudinal data autoregressive MAR models \citep[e.g.
 cf.][model 2 and model 3]{shara2015randomly} can be useful.
-Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
-The effects of different forms of MAR mechanisms are described in \citet{scho18}.% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
+It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
+The effects of different types of MAR mechanisms are described in \citet{scho18}.% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
 This may influence the performance of some (but not necessarily all) imputation routines.
@@
 It would be a shame to overlook results due to the focus on a single MAR mechanism.
-[Although often ignored, it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
-It would be wise to include the mechanism in your simulation study, or perform sensitivity analyses.] An indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis \citep[see e.g.][part 5]{molenberghs2014handbook}.
 
-Fourth, the amount of missingness must be varied.
+Third, a non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice.
+Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time.
+It would, therefore, be wise to include MNAR missingness in the simulation study just in case, or to perform sensitivity analyses to get an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
+
+Apart from the missingness mechanisms, the missingness pattern must be varied--in particular the missingness proportion.
 Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
@@
 Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
-We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness see Table 2.
-Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.
+We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness (see Table \ref{table:prop}).
+% Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.
 
@@
 ]
+Fill in the following checklist (in appendix).
 
-Fill in the following checklist (make appendix).
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% DISCUSSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Discussion}
+
+This manuscript aims to establish a common ground for the evaluation of imputation routines.
+Such a common ground would be the basis of a standardized evaluation.
+This would allow for fairer and more efficient comparisons between imputation techniques.
+Ultimately, it would be desirable to evaluate every imputation routine against the same standardized set in order to quantify the statistical properties across imputation routines.
+If properly executed, this would allow for careful matching of imputation methodologies to new missing data problems.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% APPENDICES
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{acknowledgement}
+We thank the Amices team for the fruitful discussions.
+\end{acknowledgement}
+\vspace*{1pc}
+
+\noindent {\bf{Conflict of Interest}}
+
+\noindent {\it{The authors have declared no conflict of interest.}}
+
+\section*{Appendix}
+
+\subsection*{A.1.\enspace Checklist for the evaluation of imputation methodology}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%
-%% DISCUSSION
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Discussion}
-
-This manuscript aims to establish a common ground for the evaluation of imputation routines.
-Such a common ground would be the basis of a standardized evaluation.
-This would allow for fairer and more efficient comparisons between imputation techniques.
-Ultimately, it would be desirable to evaluate every imputation routine against the same standardized set in order to quantify the statistical properties across imputation routines.
-If properly executed, this would allow for careful matching of imputation methodologies to new missing data problems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDICES
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{acknowledgement}
-[Maybe add Gelman here? Mice team]
-\end{acknowledgement}
-\vspace*{1pc}
-
-\noindent {\bf{Conflict of Interest}}
-
-\noindent {\it{The authors have declared no conflict of interest. (or please state any conflicts of interest)}}
-
-% \section*{Appendix {\it(please insert here, if applicable)}}
-% 
-% \subsection*{A.1.\enspace Second level heading}
-% 
-% Please insert appendices before the references.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% REFERENCES

### 2022-03-30 Formatted/Formatted.tex
Start writing out the missingness mechanisms part

@@
 A popular variant is multiple imputation \citep{rubi76}, whereby each missing value is imputed multiple times.
-The completed data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference \citep[using Rubin's rules][]{rubi87}.
+The completed data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference \citep[using `Rubin's rules'][]{rubi87}.
 In contrast to ad hoc methods for dealing with missing values (e.g.
@@
 This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness mechanisms and missingness patterns.
-Missingness mechanisms describe the relationship between missing entries and observed data values \citep[generally categorized into MCAR, MAR and MNAR mechanisms cf][see Table \ref{table:mech}]{rubi76}.
+Missingness mechanisms describe the relationship between missing entries and observed data values \citep[generally categorized into MCAR, MAR and MNAR mechanisms c.f.][see Table \ref{table:mech}]{rubi76}.
 Missingness patterns concern the location of missing entries across incomplete data \citep[][p.
@@
 For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that complete case analysis outperforms their imputation method(s).
-Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described by][]{fimd}, while biasing inferences in other cases.
-Another example of a missingness pattern that may inadverted impose assumptions on the evaluations is a monotone pattern \citep[i.e.
-a pattern with uniformly increasing missingness proportions along variables][]{litt20}.
-If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration in all other situations.
+Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described in][\S 2.7]{fimd}, while biasing inferences in other cases.
+Another example of a missingness pattern that may inadverted impose assumptions on generalizations is a monotone pattern \citep[i.e.
+a pattern with uniformly increasing missingness proportions along variables;][]{litt20}.
+If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations.
 A final example is multivariate missingness that is generated using stepwise univariate missingness induction.
-The resulting incomplete data may not have the intended or expected statistical properties \citep[e.g.
-skewness][]{ampute}, which complicates definitive conclusions about imputation method performance in relation to the data generating mechanism.
+The resulting incomplete data may then not have the intended or expected statistical properties \citep[e.g.
+skewness;][]{ampute}.
+This complicates any definitive conclusions about imputation method performance in relation to the data generating mechanism.
 
@@
 
-The missingness mechanism is usually assumed to be random (MAR) or completely random (MCAR).
-
 \begin{table}[tb]
 \begin{center}
-\caption{Missingness mechanisms. [Add PMAR]}
+\caption{Missingness mechanisms. [Add PMAR?]}
 \label{table:mech}
@@
 \end{table}
+
+The missingness mechanisms that are induced in simulation studies do not necessarily match the true underlying mechanisms in incomplete empirical data, nor the mechanisms that are assumed when imputing such data.
+Out of the three mechanisms (see Table \ref{table:mech}), MAR missingness is usually investigated in simulations and assumed in practice.
+Although MCAR might not reasonably be the true missingness generating mechanism and naturally not typically assumed, the mechanism  can be very insightful in the context of imputation methodology evaluation.
+By contrast, MNAR
+
+% usually assumed to be random (MAR) or completely random (MCAR).
 
@@
 
-Fit of the imputation model: [PPC, ref Mingyang, \citep{nguy17, zhao22}.]
+The fit of the imputation model may be verified using posterior predictive checking. [PPC, ref Mingyang, \citep{nguy17, zhao22}.]
 

### 2022-03-30 Formatted/Formatted.tex
Write convergence issues paragraph

@@
 After deciding upon a data generating mechanism, another influential--but often overlooked--decision awaits.
-Simulation studies on missing data methodolology have the unique option to exclude sampling variance from their evaluations, and only use the missing data generation procedure as source of Monte Carlo variation.
+Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations, and only use the missing data generation procedure as source of Monte Carlo variation.
 After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
@@
 It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways.
-Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to suboptimal advice in practice.
-For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that complete case analysis outperforms their imputation method(s).
+Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to sub-optimal advice in practice.
+For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that list-wise deletion outperforms their imputation method(s).
 Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described in][\S 2.7]{fimd}, while biasing inferences in other cases.
-Another example of a missingness pattern that may inadverted impose assumptions on generalizations is a monotone pattern \citep[i.e.
+Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e.
 a pattern with uniformly increasing missingness proportions along variables;][]{litt20}.
 If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations.
-A final example is multivariate missingness that is generated using stepwise univariate missingness induction.
+A final example is multivariate missingness that is generated using step-wise univariate missingness induction.
 The resulting incomplete data may then not have the intended or expected statistical properties \citep[e.g.
@@
       & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
-      & inferred from the observed data. This is considered non-ignorable missingness. \\
+      & inferred from the observed data. This is considered non-ignorable missingness \\
+      & \citep[see e.g.][]{rubi76}. \\
 \hline
@@
 
-The missingness mechanisms that are induced in simulation studies do not necessarily match the true underlying mechanisms in incomplete empirical data, nor the mechanisms that are assumed when imputing such data.
-Out of the three mechanisms (see Table \ref{table:mech}), MAR missingness is usually investigated in simulations and assumed in practice.
+Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions).
+However, the missingness mechanisms that are induced in simulation studies do not necessarily match any true missingness generating mechanisms in incomplete empirical data, nor the mechanisms that are typically assumed when imputing such data.
+On the one hand, there is MCAR missingness, which may be unlikely as the true missingness generating mechanism, but would yield insightful evaluations in a simulation setting.
+On the other hand, there is MNAR, which is arguably a quite reasonable source of many real-world missingness scenarios, but is generally ignored except for those evaluations that are specifically targeted at non-ignorable applications.
+A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.
+
+% Out of the three mechanisms , random missingness (MAR) is usually generated for the evaluation of imputation methodology.
+Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored, except for those methods that are specifically targeted at non-ignorable applications.
+
+% MAR missingness is usually investigated in simulations and assumed in practice.
 Although MCAR might not reasonably be the true missingness generating mechanism and naturally not typically assumed, the mechanism  can be very insightful in the context of imputation methodology evaluation.
@@
 
-[MCAR is important but often ignored:] MCAR is a necessary simulation condition for evaluating the performance of imputation procedures.
+MCAR is a necessary simulation condition for evaluating the performance of imputation procedures, yet often omitted.
 Under MCAR, the statistical properties of the observed data given the missing data are known.
@@
 
-[MAR may be spurious:] With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g.
-weight is made incomplete based on observed gender to emulate a situation wherin one gender is less likely to disclose their weight).
+Although MAR missingness is often considered as simulation condition, the problem of spurious MAR is generally overlooked.
+With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g.
+weight is made incomplete based on observed gender to emulate a situation wherein one gender is less likely to disclose their weight).
 These relations in the observed data may, however, be weak or non-existent.
@@
 This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.
-%The problem of spurious MAR is generally overlooked in simulation studies
+
+Inducing a MAR mechanism comes with another issue, namely the type or functional form of the missingness model.
+Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}.
+This may influence the performance of some (but not necessarily all) imputation routines.
+For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+It would be a shame to overlook such results due to the focus on a single MAR mechanism.
 
@@
 
-[MNAR is often ignored:] Although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
-Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
+Finally, although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
+As stated, this mechanism is not usually included in evaluations.
 % The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
-
-[Missingness type is often ignored:] Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
-This may influence the performance of some (but not necessarily all) imputation routines.
-For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
-It would be a shame to overlook such results due to the focus on a single MAR mechanism.
 
@@
 
-The evaulation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e.
+The evaluation criteria used to assess imputation performance vary from one simulator to another.
+This is not surprising as people from different fields could have a different focus on the problem at hand.
+There are, however, some over-arching issues with assessing imputation method performance.
+In the first place, evaluations of the imputation-generating process (e.g., an iterative imputation algorithm) are generally left out of simulation study results.
+In the second place, there are performance measures pitfalls to consider.
+
+Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations.
+As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated.
+Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{fimd} is neither efficient nor failproof.
+As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method.
+Moreover, failing to produce stable imputations is not an exclusive quality iterative aimputation algorithms.
+For example if a method does not yield any imputations at all in a certain simulation repetition (e.g.
+due to over-parameterization), this can be seen as failure in the imputation-generating process too.
+
+The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e.
 the complete data model used to estimate a target or estimand).
 When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
-
-[All imputation models should be congenial, but this is difficult to test and often ignored.
+Congeniality is difficult to diagnose, but ignoring it may yield sub-optimal imputations.
+Methods to asses the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}, which is an infeasible endeavor in many simulation studies.
+% [All imputation models should be congenial, but this is difficult to test and often ignored.
 Checks of imputation model fit are often omitted.
@@
 Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]
-
-Moreover, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
+Therefore, the performance of imputation procedures on distributional properties is often ignored in simulation studies.
+Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
 For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large.
@@
 % [Add segway between evaluation of imputation and actual performance measures.]
-The evaluation criteria used to assess imputation performance vary from one simulator to another.
-This is not surprising as people from different fields could have a different focus on the problem at hand.
-Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.
-For many outcome-driven imputation taks a single valid imputation may suffice.
-Measures of predictive accuracy or the RMSE (root mean squared error) of predicted values could then serve as sensible evaluation criteria.
-
 % [Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]
@@
 591]{neym34}.
-Merely focussing on distance measures would open the door for invalid inferences.
+Merely focusing on distance measures would open the door for invalid inferences.
+
+Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.
+For many outcome-driven imputation task a single valid imputation may suffice.
+Measures of predictive accuracy or the RMSE (root mean squared error) of predicted values could then serve as sensible evaluation criteria.
 
@@
 The DGM is by definition design-based and there is no sampling variance.
-The missingness generation step is skipped and the only performance measure that can be used between methods is RMSE and maybe comparison with repect to CCA in terms of bias.]
+The missingness generation step is skipped and the only performance measure that can be used between methods is RMSE and maybe comparison with respect to CCA in terms of bias.]
 
@@
 There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions).
-If there are several imputation method under evaluation, the simulation design should accomodate that each method is applied to each incomplete data set.
+If there are several imputation method under evaluation, the simulation design should accommodate that each method is applied to each incomplete data set.
 This is in contrast to a design with as many incomplete data sets as imputation methods in every simulation repetition.
@@
 
-[Use CCA as benchmark method: the imputation method should outperform CCA.
+The imputation method(s) under evaluation should be applied to the generated incomplete data.
+It is important to choose the imputation model parameters carefully (e.g., number of imputations, number of iterations, regularization, etc.).
+After imputation, an analysis model may be fitted to estimate the estimand(s).
+
+%[Use CCA as benchmark method: the imputation method should outperform CCA.
 Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matrix, regularization if applicable, etc.).
@@
 
-It is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
+Apart from the imputation method(s), list-wise deletion should be considered as missing data method too.
 We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance.
+List-wise deletion may therefore serve as a benchmark method: out-performing it should be a minimal requirement for the imputation method(s).
+%it is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
 
@@
 
+[Some suggested minimal requirements for imputation methods: the absence of algorithmic non-convergence, the imputation model fits the observed parts of the data (PPC), under a MCAR mechanism the imputed values are distributed equivalent to the observed values.]
+
 % = evaluation of imputation generating process
-Convergence of the algorithm: Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
+The absence of non-convergence in the imputation models is a minimal requirement for imputation methods.
+Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
 The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
-Distributional characteristics: In practice, the distribution of the incomplete data may differ greatly from the observed data.
+
+[Distributional characteristics] In practice, the distribution of the incomplete data may differ greatly from the observed data.
 Under anything but the MCAR assumption, this can be expected.
@@
 
-The fit of the imputation model may be verified using posterior predictive checking. [PPC, ref Mingyang, \citep{nguy17, zhao22}.]
+The fit of the imputation model may be verified with the help of a posterior predictive check \citep[PPC][]{nguy17, zhao22}.
+A PPC in the context imputation methodology could be the over-imputation of observed data values.
+The reasoning is that if the over-imputed values have  equivalent statistical properties to the true data values, the imputation model fits the observed part of the incomplete data well.
+We might then assume that the imputation model would be able to produce good imputations for missing values as well.
 
-Plausibility of the imputed values: Plausible imputations--imputations that could be real values if they had been observed--are not a necessary condition for obtaining valid inference.
+[Plausibility of the imputed values] Plausible imputations--imputations that could be real values if they had been observed--are not a necessary condition for obtaining valid inference.
 However, in practice, especially when the imputer and the analyst are different persons, plausible imputations may be a desired property.
@@
 When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
-
-Some suggested minimal requirements for imputation methods: the absence of algorithmic non-convergence, the imputation model fits the observed parts of the data (PPC), under a MCAR mechanism the imputed values are distributed equivalent to the observed values.
 
@@
 
+[Some suggested minimal requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis (i.e.
+performance after list-wise deletion).]
+
 % sim perf and model perf
-The evaluation criteria may depend on the estimand.
+[The evaluation criteria may depend on the estimand.
 When descriptive statistics are the goal and when statistical inference would not be of interest, the bias of the estimates would still apply, but standard errors are generally ignored.
-Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
+Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.]
 
@@
 However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.
-
-[Add RMSE at cell level and at prediction level (if applicable).]
-
-Some suggested minimal requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis (i.e.
-performance after list-wise deletion).
+% [Add RMSE at cell level and at prediction level (if applicable).]
 
@@
 And the procedure to generate missing data is often insufficiently described in resulting publications.
-]
-Fill in the following checklist (in appendix).
+Fill in the following checklist (in appendix).]
 
@@
 
-- data types + structures (e.g. clustering/timeseries)
+- data types + structures (e.g. clustering/time-series)
 
@@
 
-- missingness patterns (rows = univariate/multivariate; colums = proportion)
+- missingness patterns (rows = univariate/multivariate; columns = proportion)
 

### 2022-03-31 Formatted/Formatted.tex
Apply suggested changes (feedback Gerko 30-03), save result as LastDiscussed.tex for latexdiff

@@
 A popular variant is multiple imputation \citep{rubi76}, whereby each missing value is imputed multiple times.
-The completed data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference \citep[using `Rubin's rules'][]{rubi87}.
+The completed data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference using Rubin's rules \citep[][p.
+76]{rubi87}.
 In contrast to ad hoc methods for dealing with missing values (e.g.
@@
 To demonstrate the broad impact of these problems, we subdivide the problems into the following three distinct categories: problems with data generation, problems with missingness generation, and problems with performance evaluation.
-We further detail the impact each of these problems may have on the validity of our evaluations.
+We further detail the impact each of these problems may have on the validity of evaluations.
 
@@
 After establishing a comparative truth from a data generating mechanism, some form of missingness therefore has to be induced.
-Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like
+Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like:
 \begin{quote}
@@
 This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness mechanisms and missingness patterns.
-Missingness mechanisms describe the relationship between missing entries and observed data values \citep[generally categorized into MCAR, MAR and MNAR mechanisms c.f.][see Table \ref{table:mech}]{rubi76}.
-Missingness patterns concern the location of missing entries across incomplete data \citep[][p.
+Missingness mechanisms describe the relationship between missing entries and observed data values and are generally categorized into MCAR, MAR and MNAR mechanisms \citep[][see Table \ref{table:mech}]{rubi76}.
+[TODO: maybe remove mechanisms here?] Missingness patterns concern the location of missing entries across incomplete data \citep[][p.
 8]{litt20}.
@@
 A final example is multivariate missingness that is generated using step-wise univariate missingness induction.
-The resulting incomplete data may then not have the intended or expected statistical properties \citep[e.g.
+The resulting incomplete data may then not have the desired statistical properties \citep[e.g.
 skewness;][]{ampute}.
@@
 \begin{center}
-\caption{Missingness mechanisms. [Add PMAR?]}
+\caption{Missingness mechanisms. [TODO: add PMAR?]}
 \label{table:mech}
@@
 48]{fimd}: the missingness does not depend on the incomplete variable.
-This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.
+This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is pointless.
 
-Inducing a MAR mechanism comes with another issue, namely the type or functional form of the missingness model.
+Inducing a MAR mechanism presents the simulator with another choice, namely the type or functional form of the missingness model.
 Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}.
@@
 
-% The evaluation criteria may depend on the estimand.
-When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
-Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
-
-% For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.
-
 The evaluation criteria used to assess imputation performance vary from one simulator to another.
@@
 Methods to asses the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}, which is an infeasible endeavor in many simulation studies.
-% [All imputation models should be congenial, but this is difficult to test and often ignored.
-Checks of imputation model fit are often omitted.
-Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.
-Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies.
-Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]
 Therefore, the performance of imputation procedures on distributional properties is often ignored in simulation studies.
@@
 While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.
+
+If the goal is inference--not prediction--then the uncertainty about estimates needs to be properly quantified.
+A single imputation would then not suffice.
+The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
+This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.
+After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p.
+591]{neym34}.
+Merely focusing on the discrepancy between the true and imputed data would open the door for invalid inferences.
+
+Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.
+For many outcome-driven imputation task a single valid imputation may perhaps suffice.
+We would like to note, however, that using measures of predictive accuracy or the RMSE (root mean squared error) of predicted values as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{fimd}.
+
+% The evaluation criteria may depend on the estimand.
+When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
+Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
+
+% For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.
 
@@
 
-If the goal is inference--not prediction--then the uncertainty about estimates needs to be quantified.
-A single imputation would then not suffice.
-The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.
-After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p.
-591]{neym34}.
-Merely focusing on distance measures would open the door for invalid inferences.
-
-Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.
-For many outcome-driven imputation task a single valid imputation may suffice.
-Measures of predictive accuracy or the RMSE (root mean squared error) of predicted values could then serve as sensible evaluation criteria.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Standardizing the evaluation of imputation methodology requires simulators--like ourselves--to consider a couple of key aspects in their simulation workflows.
+Standardizing the evaluation of imputation methodology requires simulators to consider a couple of key aspects in their simulation workflows.
 Table \ref{table:steps} outlines some suggested steps to adopt.
@@
 Before setting up a simulation study, the simulator should clearly define the scope of their evaluations.
-A simulation study aimed at imputation model comparison will typically have a  different design than one aimed at establishing the inferential validity of a single (novel) method.
-The choice of imputation method(s) under evaluation is naturally intertwined with the simulation study's aims and set-up.
+A simulation study aimed at imputation method comparison will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
+The choice of imputation method(s) under evaluation is naturally intertwined with the simulation aims and set-up.
 
@@
 There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions).
-If there are several imputation method under evaluation, the simulation design should accommodate that each method is applied to each incomplete data set.
-This is in contrast to a design with as many incomplete data sets as imputation methods in every simulation repetition.
-Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation which makes for fairer comparisons.
+If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
+Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation, which makes for fairer comparisons.
 The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
@@
 A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
-The benefit of design-based simulation is the ability to use real-life observed data structures.
 
-Estimand(s) or other simulation targets should defined in the context of the study aim and data-generating mechanism.
-[As a general rule, it would be wise to include one or more multivariate estimands.
-It is a realistic expectation for imputation methods to yield valid univariate estimates but also to capture multivariable relations in the data.
-Imputation methods that fail to do so, should not be considered general purpose methods (e.g., mean imputation).]
+Estimand(s) or other simulation targets should be defined in the context of the study aim and data-generating mechanism.
+As a general rule, it would be wise to include one or more multivariate estimands.
+It is a realistic expectation for imputation methods to preserve both marginal and conditional distributions with respect to the comparative truth.
+Imputation methods that fail to do so, should not be considered general purpose methods (e.g., mean imputation).
 
@@
 First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions).
-This mechanism may be used as a reference condition.
-Valid imputations under MCAR is a minimal requirement for any [multipurpose] imputation method.
-We also know that, under MCAR, the distribution of imputed values should be equivalent to the distribution of observed values for a given variable, which offers an indication of imputation model (mis)fit.
+This mechanism should be used as a reference condition.
+Every imputation method should yield valid imputations under MCAR.
+We also know that, under MCAR, the distribution of imputed values should be equivalent to the distribution of observed values for a given variable, which may offer an indication of imputation model (mis)fit.
+[TODO: leg uit!]
 % First, one should always consider MCAR missingness [as reference, because all imputation methods should yield valid inference under MCAR and we can use this mechanism to check imputation model fit] %, i.e.
@@
 A straightforward technique for inducing univariate MAR missingness is described in \citet[][p.
-63]{fimd}, generalizations to multivariate MAR missingness can be found in \citet{ampute}, and if the missingness is to be induced in longitudinal data autoregressive MAR models \citep[e.g.
-cf.][model 2 and model 3]{shara2015randomly} can be useful.
+63]{fimd}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
+If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly} can be useful.
 It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
@@
 Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time.
-It would, therefore, be wise to include MNAR missingness in the simulation study just in case, or to perform sensitivity analyses to get an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
+It would, therefore, be wise to include MNAR missingness in the simulation study just in case.
+Alternatively, sensitivity analyses may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
 
-Apart from the missingness mechanisms, the missingness pattern must be varied--in particular the missingness proportion.
-Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Apart from the missingness mechanisms, the missingness pattern must be varied.
+[TODO: beter bruggetje?] Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
 Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
@@
 
-\subsection{Apply missing data method(s) and analysis model(s)}
+\subsection{Apply methods}
 
@@
 
-The imputation method(s) under evaluation should be applied to the generated incomplete data.
-It is important to choose the imputation model parameters carefully (e.g., number of imputations, number of iterations, regularization, etc.).
+The imputation methods under evaluation should be applied to the generated incomplete data, after which an analysis model may be fitted to estimate the estimand(s).
+% It is important to choose the imputation model parameters carefully (e.g., number of imputations, number of iterations, regularization, etc.).
 After imputation, an analysis model may be fitted to estimate the estimand(s).
-
 %[Use CCA as benchmark method: the imputation method should outperform CCA.
@@
 Estimate estimand(s) using the analysis model, use correct pooling rules.]
-
-Apart from the imputation method(s), list-wise deletion should be considered as missing data method too.
-We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance.
-List-wise deletion may therefore serve as a benchmark method: out-performing it should be a minimal requirement for the imputation method(s).
+In addition, the simulator should always perform complete case analysis.
+We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance.
+Complete case analysis may therefore serve as a benchmark method against which performance should be evaluated.
+Moreover, pairing complete case analysis with MCAR and MAR mechanisms allows the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.% out-performing it should be a minimal requirement for the imputation method(s).
 %it is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
@@
 % = evaluation of imputation generating process
-The absence of non-convergence in the imputation models is a minimal requirement for imputation methods.
+The absence of non-convergence in the imputation models is a minimum requirement for imputation methods.
 Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
-The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
+The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations might be invalid.
 
@@
 The fit of the imputation model may be verified with the help of a posterior predictive check \citep[PPC][]{nguy17, zhao22}.
-A PPC in the context imputation methodology could be the over-imputation of observed data values.
+A PPC in the context of imputation methodology could be the multiple over-imputation of observed data values.
 The reasoning is that if the over-imputed values have equivalent statistical properties to the true data values, the imputation model fits the observed part of the incomplete data well.
-We might then assume that the imputation model would be able to produce good imputations for missing values as well.
+We might then assume that the imputation model would be able to produce good imputations for missing values too.
 
@@
 
-[Some suggested minimal requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis (i.e.
-performance after list-wise deletion).]
+[Some suggested minimum requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis.]
 
@@
 However, the real-world missingness would then be omitted.
+[TODO: check waar dit al eerder staat!]
 

### 2022-03-31 Formatted/LastDiscussed.tex
Apply suggested changes (feedback Gerko 30-03), save result as LastDiscussed.tex for latexdiff

@@
 \pagespan{1}{}
-\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\[1pc]
-\noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
+\keywords{Evaluation; Imputation; Missing data; Simulation studies;\\ [1pc]
+% \noindent \hspace*{-4.2pc} Supporting Information for this article is available from Github, \break \hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}.
+% } %%% semicolon and fullpoint added here for keyword style
+\noindent \hspace*{-4pc} Supporting Information for this article is available online from \\
+\hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}. 
 }  %%% semicolon and fullpoint added here for keyword style
 
-\title[Standardized evaluation]{Towards a standardized evaluation of imputation routines}
+\title[Standardized evaluation]{Towards a standardized evaluation of imputation methodology}
 %% Information for the first author.
@@
 Imputation is a state-of-the-art technique for drawing valid conclusions from incomplete data.
-The technique has earned a permanent spot in research and policy making, demonstrated e.g.
+The technique has earned a permanent spot in research and policymaking, demonstrated e.g.
 by the detailed manual created by the National Research Council \citep*{little2012prevention}.
-Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers are embracing imputation techniques.
+Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers and data scientists are embracing imputation techniques.
 After all, the principle of imputation is very intuitive.
@@
 The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.
-A popular variant is multiple imputation \citep{rubi76}, wherein missing values are imputed multiple times.
-The data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference \citep[using Rubin's rules][]{rubi87}.
+A popular variant is multiple imputation \citep{rubi76}, whereby each missing value is imputed multiple times.
+The completed data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference using Rubin's rules \citep[][p.
+76]{rubi87}.
 In contrast to ad hoc methods for dealing with missing values (e.g.
-list-wise deletion, mean imputation, regression imputation, last observation carried forward, indicator method), multiple imputation properly takes the sources of uncertainty that are related to the missingness problem into account.
+list-wise deletion, mean imputation, regression imputation, last observation carried forward, indicator method), multiple imputation properly takes into account the sources of uncertainty that are related to the missingness problem.
 
 The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
-In general it holds that modeling missing data becomes more challenging when the amount of missingness increases.
+In general, it holds that modeling missing data becomes more challenging when the amount of missingness increases.
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
-In that case,  imputation would be substantially more efficient than the ubiquitous list-wise deletion.
+In that case, imputation would be substantially more efficient than the ubiquitous complete case analysis.
 
 When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies.
-In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, [then several missing data methods are applied] and a set of evaluation criteria is postulated to evaluate the performance of the imputation methods.
-However, no golden standard has been established to evaluate these imputation routines and, as a result, the validity of the simulation procedures may differ tremendously from one developer to another.
-[Nowadays there are a lot of imp alg being propagated, esp ai/ml, whose perf often times seems quite promising.
-But may not be a fair comparison because of the simulation set-up/dev in ai/ml so fast that steps may be skipped]
+In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods.
+However, no golden standard has been established to evaluate imputation routines and, as a result, the validity and aim of the simulation set-ups may differ tremendously from one developer to another.
+Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced.
+Although these promising new methods seem to yield even sharper imputations than now-standard (semi-)parametric imputation methods, the comparison may not be fair due to the simulation setup.
+
+% [Nowadays there are a lot of imp alg being propagated, esp ai/ml, whose perf often times seems quite promising.
+But may not be a fair comparison because of the simulation set-up/dev in ai/ml so fast that steps may be skipped.
+Especially in the field of ML/AI, there are promising new imputation methods, which may yield even sharper imputations than now-standard (semi-)parametric imputation methods]
 
 The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology.
-These concerns stem from careful consideration with fellow 'imputers' and from encounters as a reviewer for statistical journals.
+These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals.
 Second, to provide imputation methodologists with a suggested course of action when simulating missing data problems.
-This suggested approach should identify a common ground, but is in no way intended as an absolute solution.
+This suggested approach should identify common ground but is in no way intended as an absolute solution.
 This identifies the third purpose of this paper: discussion.
-We hope to elicit critical thinking with respect to the problems at hand.
+We hope to elicit critical thinking regarding the problems at hand.
 We are all convinced that our methodology has some merit.
-But for sake of progress it would be much more advantageous if the aim of our evaluations would go beyond \emph{proving the point} and would legitimately consider the statistical properties.
+But for sake of progress, it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties.
 
@@
 Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers.
-This brings forth a chain of potential problems in the objective assessment of the performance of imputation routines that may lead to sub-optimal use of multiple imputation in practice.
-To demonstrate the broad impact of these problems, we subdivide the problems in the following three distinct categories: problems with data generation, problems with missingness generation and problems with performance evaluation.
-We further detail the impact these problems may have on the validity of the performance evaluation of the imputation routines.
+This brings forth a chain of potential problems in the objective assessment of imputation method performance, within and across studies, which may lead to sub-optimal use of imputation in practice.
+To demonstrate the broad impact of these problems, we subdivide the problems into the following three distinct categories: problems with data generation, problems with missingness generation, and problems with performance evaluation.
+We further detail the impact each of these problems may have on the validity of evaluations.
 
@@
 To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established.
-Those who perform simulation studies are in the luxury position to establish the truth beforehand, by choosing a data generating mechanism to obtain a complete dataset.
-Complete data is typically either generated from a statistical model, or sampled from a sufficiently large observed set [model-based vs design-based simulation].
+Those who perform simulation studies are in the luxury position to establish the truth beforehand by choosing a data generating mechanism.
+Data generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
+There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as from official registers.
 
-The problem with model-based data generating mechanisms is that method performance may not translate to empirical data.
+The problem with model-based data generating mechanisms is that a method's performance on simulated data may not translate to empirical data.
+Real-life data hardly ever follow a given theoretical distribution, so there is no guarantee that simulation results are generalizable.
+Moreover, data are often generated such that the problem that is being studied is most pronounced, e.g.
+with consistently high correlations between groups of variables.
+This results in simulated data that contain such valuable information structures that, no matter what type of missingness would subsequently be induced, the observed parts of the data will still hold much (if not all) of the information about the missing part.
+Unsurprisingly, the performance of any imputation method will then be evaluated as good.
+Another threat to the generalizability of model-based simulations is the use of a single model for both data generation and imputation.
+If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied.
+Other (unfair) comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s), and the coherence between variables.
+In contrast to design-based studies, such characteristics are not always explicit simulation conditions, which may give a false sense of objectivity.
 
-First, data are often generated following the model (or distribution) that is also used for imputing the data.
-The performance of the imputation approach is then deemed good, which is no surprise as the evaluated conditions are in favor of the problem that is studied.
-Although such procedures may be statistically relevant, the approach would be no good for even the simplest empirical case.
-Such simulations pose serious threats to the applicability of the method as real-life data hardly ever follow a given theoretical distribution.
-[Add that data types may be more messy.] Second, data are often generated such that the problem that is being studied is most pronounced.
-This results in simulated data that contains very valuable information structures, i.e.
-the correlations between groups of variables may be very pronounced.
-In other words, no matter what type of missingness is induced, the observed parts of the data may still hold much, if not all, of the information about the missing part and the performance of the imputation procedure is not surprisingly evaluated as very good.
-Also, it may be that the relations in real-life data are rarely as pronounced as the problem that is being studied.
+An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be very challenging.
+Most real-world data contains at least some missing entries, for which the true underlying missing data model is--by definition--unknown.
+Therefore, the simulator needs to deal with missingness in \textit{some} way before incomplete empirical data can serve as comparative truth in the simulations.
+It may seem like an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples.
+However, there may be inherent differences between cases with and cases without any missing values, due to the unknown missing data model.
+Only sampling complete cases from the data set may thus result in samples that fail to capture all relevant real-world conditions, which in turn refutes the main reason for using design-based simulation.
+Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset to draw samples from.
+Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.
+Just to be clear: leaving the missingness as-is and inducing additional missing values to impute could not work here, because we wouldn't have a real and unbiased comparative truth.
 
-The problem with design-based simulation is that it's difficult to find large datasets without missing entries.
-And imputing the missing data beforehand and then treating the once-imputed data as comparative truth may favor the used imputation method in further evaluations.
-A second issue is that SEs may be incorrect if the (re)sampling is not handled correctly.
-[Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs].
-[Add that some simulation studies do not use a 'ground truth' at all, they just look for the method with the best predictive performance for a certain completely observed target variable.
-These studies apply different combinations of imputation and estimation methods on one or more benchmark datasets, and assess the RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
-In this context, there is no talk of retrieving the 'true' but unobserved values at all.
-Refer to missing data chapter \citep{liu21}.]
+After deciding upon a data generating mechanism, another influential--but often overlooked--decision awaits.
+Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations, and only use the missing data generation procedure as source of Monte Carlo variation.
+After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
+Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}.
+The only required change to simulation set-ups is to remove sampling variance from performance evaluation.
+A lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
 
-Problems with either of these data generation approaches is that the number of observations, the number of variables, the variable type(s) and the coherence between variables may give some imputation methods a comparative advantage, if they are not a) varied across simulation conditions [when is this fair?
-is this even an option?], or b) the object of the study [add some examples of simulations specifically focused on small samples, categorical data, etc.].
+% Table \ref{table:dgm_truth} displays the source of the comparative truth in simulation studies with and without sampling variance.When deviating conventional pooling rules \citep[cf.][p.
+76-77]{rubi87} do not apply.
+Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
+If the sampling variance is not handled correctly, subsequent inferences may be invalid and the simulation results cannot be trusted.
 
@@
 
-Missingness comes in many shapes and forms [platitude, but how else to start?].
-The most obvious characteristics of missing data are the missingness pattern(s) and missingness mechanism(s).
-Missingness patterns concern the location of missing entries across variables in an incomplete dataset, whereas missingness mechanisms describe the relationship between missingness and the values of variables in the data \citep[][p.
+Evaluating imputation methodology requires a missing data problem to be solved.
+After establishing a comparative truth from a data generating mechanism, some form of missingness therefore has to be induced.
+Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like:
+\begin{quote}
+\textit{We generated missing data following a missing at random missingness mechanism.}
+\end{quote}
+This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness mechanisms and missingness patterns.
+Missingness mechanisms describe the relationship between missing entries and observed data values and are generally categorized into MCAR, MAR and MNAR mechanisms \citep[][see Table \ref{table:mech}]{rubi76}.
+[TODO: maybe remove mechanisms here?] Missingness patterns concern the location of missing entries across incomplete data \citep[][p.
 8]{litt20}.
+Under this definition, there is a row-wise element to the missingness pattern describing which variables are jointly observed, and a column-wise element encompassing the amount of missingness in the data.
 
-%  L&R, p.
-8: "We find it useful to distinguish the missingness pattern, which describes which values are missing and observed in the data matrix and the missingness mechanism (or mechanisms), which concerns the relationship between missingness and the values of variables in the data matrix".
-
-The missing data pattern may inadvertently guide which imputation method prevails in simulation studies.
-Some ad hoc methods are known to yield valid inferences in very specific missing data patterns (e.g., list-wise deletion when there is missingness in the outcome variable of the analysis model exclusively), while they bias inferences in any other case.
-Extrapolating simulation results from such a specific missing data pattern to more intricate empirical missingness patterns could lead to false conclusions [re-read this!
-and add monotone/univariate vs multivariate and sporadic vs systematic dichotomies].
-Related to the missingness pattern across variables is the missingness proportion.
-The terminology on this topic can be confusing.
+Even the terminology on missingness generation can be confusing.
 Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
-We will refer to the latter as the proportion of incomplete cases.
-But this is not a clear distinction in the literature, which may lead to incorrect conclusions [re-phrase this too].
-Moreover, the amount of missingness that is used to simulate the performance of imputation methods differs between studies.
-[Add ref to chapter \citep{liu21} and add info.] Some studies use only 10\% missing data where other studies push the limits to additionally investigate the performance under at least 50\% missing data.
-This inconsistent display of simulation results may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of the less stringent simulation conditions.
+In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation.
+The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).
+Moreover, the value of the actual missingness proportion may be diffusing too.
+Some studies use only 10\% missingness where other studies push the limits to additionally investigate the performance under missingness proportions of at least 50\%.
+The inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
 This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
 
-[Briefly introduce missingness mechanisms here and show potential problems, but without simulation results (moved to separate section).
-This part should contain: 1) definitions of mechanisms, 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \emph{spuriously} if the correlation between variables is low, and 4) that MNAR is often ignored, while it may be the most realistic in practice.
-And finally something about types of M(N)AR, but what are specific problems with that?
-ref dance paper rianne]
+% Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
+The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
+If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
+Alternatively, the validity of assumptions become increasingly important when the missingness increases.
+Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
 
-In order to obtain valid imputation inference, the imputation model must capture the essence of the true non-response mechanism \citep{meng94}.
-The model - if any - that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
-Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.
+Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables.
+It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways.
+Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to sub-optimal advice in practice.
+For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that list-wise deletion outperforms their imputation method(s).
+Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described in][\S 2.7]{fimd}, while biasing inferences in other cases.
+Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e.
+a pattern with uniformly increasing missingness proportions along variables;][]{litt20}.
+If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations.
+A final example is multivariate missingness that is generated using step-wise univariate missingness induction.
+The resulting incomplete data may then not have the desired statistical properties \citep[e.g.
+skewness;][]{ampute}.
+This complicates any definitive conclusions about imputation method performance in relation to the data generating mechanism.
 
-With MCAR missingness mechanisms, the probability to be missing is the same for all cases.
-This is a necessary simulation condition for evaluating the performance of imputation procedures.
+% is the   [Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.] For example, iterative imputation algorithms converge instantly if there is only one variable with missingness or when the pattern is monotone, but require iteration in all other situations.
+And very specific missing data patterns may misguide which imputation method prevails in performance: some ad hoc methods are known to yield valid inferences under certain restricted conditions \citep[e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively][]{fimd}, while biasing inferences in other cases.
+% [Missingness mechanisms paragraphs: 1) definitions of mechanisms (now in table), 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low,  4) that MNAR is often ignored, while it may be the most realistic in practice, and 5) finally add something about types of M(N)AR (ref Rianne's Dance paper).]
+% MCAR missingness: The scenario where the missing values are missing at random and the observed values are observed at random. 
+%  the validity of assumptions become increasingly important when the missingness increases.
+
+\begin{table}[tb]
+\begin{center}
+\caption{Missingness mechanisms. [TODO: add PMAR?]}
+\label{table:mech}
+\begin{tabular}{lll}
+\hline
+Mech. & Interpretation and consequences of the mechanism \\
+\hline  
+MCAR  & Missing completely at random: The probability to be missing is the same for all \\
+      & cases. In other words, the missing values are missing at random and the observed \\
+      & values are observed at random. \\
+MAR   & Missing at random: The probability to be missing is the same within groups of \\
+      & cases defined by the observed data only. The essence is that observed relations in \\
+      & the data inform the missingness, such that MCAR may be assumed within the \\ 
+      & observed groups. \\
+MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
+      & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
+      & inferred from the observed data. This is considered non-ignorable missingness \\
+      & \citep[see e.g.][]{rubi76}. \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions).
+However, the missingness mechanisms that are induced in simulation studies do not necessarily match any true missingness generating mechanisms in incomplete empirical data, nor the mechanisms that are typically assumed when imputing such data.
+On the one hand, there is MCAR missingness, which may be unlikely as the true missingness generating mechanism, but would yield insightful evaluations in a simulation setting.
+On the other hand, there is MNAR, which is arguably a quite reasonable source of many real-world missingness scenarios, but is generally ignored except for those evaluations that are specifically targeted at non-ignorable applications.
+A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.
+
+% Out of the three mechanisms , random missingness (MAR) is usually generated for the evaluation of imputation methodology.
+Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored, except for those methods that are specifically targeted at non-ignorable applications.
+
+% MAR missingness is usually investigated in simulations and assumed in practice.
+Although MCAR might not reasonably be the true missingness generating mechanism and naturally not typically assumed, the mechanism  can be very insightful in the context of imputation methodology evaluation.
+By contrast, MNAR
+
+% usually assumed to be random (MAR) or completely random (MCAR).
+
+MCAR is a necessary simulation condition for evaluating the performance of imputation procedures, yet often omitted.
+Under MCAR, the statistical properties of the observed data given the missing data are known.
+Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation.
 If an imputation method is not able to solve the problem (i.e.
@@
 Sadly, the straightforward case of MCAR is often neglected from simulation studies and focus is drawn to the evaluation of MAR mechanisms only.
-Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in my view may be far too simplistic.
+Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in our view may be far too simplistic.
 
-With MAR missingness mechanisms, the probability to be missing is the same within groups of cases defined by the observed data only (e.g.
-males are less likely to disclose their weight, but gender is observed).
-The essence here is that the observed relations in the data are used to induce missingness during simulation (e.g.
-weight is made incomplete based on gender), but these relations may be weak, or non-existent.
-
-When MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid.
-When MAR is induced from data without multivariate relations, the inferential implication of the missingness would mimic those of MCAR.
-This problem is generally overlooked in simulation studies and the procedure to generate missing data is often insufficiently described in resulting publications.
-Figure \ref{fig:MAR} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
-
-\begin{figure}[htb]
-\begin{center}
-\resizebox{\textwidth}{!}{
-\includegraphics[]{figures/spurious_MAR.pdf}}
-\caption{Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
-Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$.}
-\end{center}
-\label{fig:MAR}
-\end{figure}
-\vspace*{1pc}
-
-It is obvious that the MAR missingness generated from data with low correlation would yield similar inference to the MCAR missingness, even though the missingness is random.
+Although MAR missingness is often considered as simulation condition, the problem of spurious MAR is generally overlooked.
+With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g.
+weight is made incomplete based on observed gender to emulate a situation wherein one gender is less likely to disclose their weight).
+These relations in the observed data may, however, be weak or non-existent.
+If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid.
+The most extreme example would be when MAR is induced from data without multivariate relations.
+The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random.
 In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p.
 48]{fimd}: the missingness does not depend on the incomplete variable.
-This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.
+This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is pointless.
 
-Although one cannot definitively verify if the missingness is random - after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every} - it can be argued that MNAR is the more likely mechanism for real life missingness scenarios.
-In such cases, an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid, may be obtained by performing sensitivity analysis \citep[see e.g.][part 5]{molenberghs2014handbook}.
+Inducing a MAR mechanism presents the simulator with another choice, namely the type or functional form of the missingness model.
+Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}.
+This may influence the performance of some (but not necessarily all) imputation routines.
+For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+It would be a shame to overlook such results due to the focus on a single MAR mechanism.
+
+% Figure \ref{fig:MAR} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
+
+% \begin{figure}[htb]
+% \begin{center}
+% \resizebox{\textwidth}{!}{
+% \includegraphics[]{figures/spurious_MAR.pdf}}
+% \caption{Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
+Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$.}
+% \end{center}
+% \label{fig:MAR}
+% \end{figure}
+% \vspace*{1pc}
+
+Finally, although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
+As stated, this mechanism is not usually included in evaluations.
+% The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
+
+% Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
+The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
+If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
+Alternatively, the validity of assumptions become increasingly important when the missingness increases.
+Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
 
@@
 
-\subsection{Performance evaluation problems}
-
-[Add bridge from last section. And explain how this part is subdivided: first general characteristics of the imputations such as convergence, then performance measures.]
-
-[Performance of different imputation methods may be influenced by preconditions of the methods themselves.
-All imputation models should be congenial, but this is difficult to test to often ignored even in single imputation endeavor, let alone in simulation studies with 100s of imputations.
-Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies.
-Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.]
-
-Moreover, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
+% \subsection{Evaluation problems}
+% 
+% The evaluation criteria used to assess imputation performance vary from one simulator to another.
+This is not surprising as people from different fields could have a different focus on the problem at hand.
+There are, however, some over-arching issues with assessing imputation method performance.
+In the first place, evaluations of the imputation-generating process (e.g., an iterative imputation algorithm) are generally left out of simulation study results.
+In the second place, there are performance measures pitfalls to consider.
+% 
+% Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations.
+As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated.
+Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{fimd} is neither efficient nor failproof.
+As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method.
+Moreover, failing to produce stable imputations is not an exclusive quality iterative aimputation algorithms.
+For example if a method does not yield any imputations at all in a certain simulation repetition (e.g.
+due to over-parameterization), this can be seen as failure in the imputation-generating process too.
+% 
+% The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e.
+the complete data model used to estimate a target or estimand).
+When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
+Congeniality is difficult to diagnose, but ignoring it may yield sub-optimal imputations.
+Methods to asses the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}, which is an infeasible endeavor in many simulation studies.
+Therefore, the performance of imputation procedures on distributional properties is often ignored in simulation studies.
+Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
 For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large.
 While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.
-Rather, one would prefer an imputation technique to yield both valid inference and plausible imputations.
-It should be studied if an imputation method is prone to deliver such impractical results, and if so, under what conditions.
-
-[Then onto the performance measure part.]
-
-The evaluation criteria used to assess imputation performance vary from one developer to another.
-This is not surprising as people from different fields could have a different focus on the problem at hand.
-But the choice of performance measure(s) may inadvertently distort statistical properties of the imputed data.
-
-Developers often only inquire about the 'accuracy' (i.e.
-how well can the method reproduce the original data).
-[Add RMSE at cell level here?] The goal of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-We are interested in the correct answer to the research question; not in the truth itself.
-This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals and the coverage rate of the confidence intervals should be studied.
+% 
+% If the goal is inference--not prediction--then the uncertainty about estimates needs to be properly quantified.
+A single imputation would then not suffice.
+The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
+This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.
 After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p.
 591]{neym34}.
-[So, what could go wrong by focusing on accuracy exclusively?]
+Merely focusing on the discrepancy between the true and imputed data would open the door for invalid inferences.
+% 
+% Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.
+For many outcome-driven imputation task a single valid imputation may perhaps suffice.
+We would like to note, however, that using measures of predictive accuracy or the RMSE (root mean squared error) of predicted values as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{fimd}.
 
-[Add choice of analysis model/estimand here, since performance measure are defined wrt the estimand, e.g., focusing on univariate estimates may favor invalid ad hoc methods.]
+%%%%%%%%%%%%%
+\subsection{Evaluation problems}
 
-[Add problems with RMSE here. And that if the goal is inference (not prediction) then you need appropriate uncertainty, so m = 1 does not suffice.]
+% The evaluation criteria may depend on the estimand.
+When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
+Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
 
-Sampling variation has always been an essential part of the evaluation of multiple imputation methodology.
-However, in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness, it is not necessary to take sampling variation into account \citep{vink14}.
-After all, we are interested only in the missing data mechanism, and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
+% For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.
 
-Last, when the evaluator is on the verge of drawing conclusions about the performance of the imputation routine, the performance should be carefully qualified.
-Comparing the performance of an imputation routine given a population (or true) parameter allows for quantitative evaluation.
-However, in order to pose qualitative statements about the performance on simulated conditions, comparative methodology is required.
-For example, when claiming that imputation performance is unacceptable when deviations from normality become rather stringent, such performance is highly dependent on the simulation conditions that are used.
-For a well-balanced judgement about the severity of the performance drop, comparative simulations with e.g.
-nonparametric models should be executed.
-A method may perform badly, but if it still outperforms every other approach, it may yet be of great practical relevance.
+The evaluation criteria used to assess imputation performance vary from one simulator to another.
+This is not surprising as people from different fields could have a different focus on the problem at hand.
+There are, however, some over-arching issues with assessing imputation method performance.
+In the first place, evaluations of the imputation-generating process (e.g., an iterative imputation algorithm) are generally left out of simulation study results.
+In the second place, there are performance measures pitfalls to consider.
+
+Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations.
+As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated.
+Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{fimd} is neither efficient nor failproof.
+As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method.
+Moreover, failing to produce stable imputations is not an exclusive quality iterative aimputation algorithms.
+For example if a method does not yield any imputations at all in a certain simulation repetition (e.g.
+due to over-parameterization), this can be seen as failure in the imputation-generating process too.
+
+The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e.
+the complete data model used to estimate a target or estimand).
+When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
+Congeniality is difficult to diagnose, but ignoring it may yield sub-optimal imputations.
+Methods to asses the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}, which is an infeasible endeavor in many simulation studies.
+% [All imputation models should be congenial, but this is difficult to test and often ignored.
+Checks of imputation model fit are often omitted.
+Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.
+Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies.
+Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]
+Therefore, the performance of imputation procedures on distributional properties is often ignored in simulation studies.
+Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
+For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large.
+While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.
+
+% [Add segway between evaluation of imputation and actual performance measures.]
+% [Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]
+% The choice of performance measures may inadvertently distort statistical properties of the imputed data.
+Developers often only inquire about the `accuracy' (i.e.
+how well can the method reproduce the original data).
+[Add RMSE at cell level here?]
+
+If the goal is inference--not prediction--then the uncertainty about estimates needs to be quantified.
+A single imputation would then not suffice.
+The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
+This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.
+After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p.
+591]{neym34}.
+Merely focusing on distance measures would open the door for invalid inferences.
+
+Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.
+For many outcome-driven imputation task a single valid imputation may suffice.
+Measures of predictive accuracy or the RMSE (root mean squared error) of predicted values could then serve as sensible evaluation criteria.
+%%%%%%%%%%%%%%%%%
+
+% The evaluation criteria may depend on the estimand.
+When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
+Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
+
+% For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.
+
+% [Add segway between evaluation of imputation and actual performance measures.]
+% [Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]
+% The choice of performance measures may inadvertently distort statistical properties of the imputed data.
+Developers often only inquire about the `accuracy' (i.e.
+how well can the method reproduce the original data).
+[Add RMSE at cell level here?]
+
+% [All imputation models should be congenial, but this is difficult to test and often ignored.
+Checks of imputation model fit are often omitted.
+Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others.
+Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies.
+Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]
 
@@
 
-[In geveral, adhere to ADEMP \citep{morr18}.]
+Standardizing the evaluation of imputation methodology requires simulators to consider a couple of key aspects in their simulation workflows.
+Table \ref{table:steps} outlines some suggested steps to adopt.
+Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach.
+Our suggestions are the accumulated result of scientific literature, research experience, and discussion.
+For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}.
+We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology.
+
+\begin{table}[tb]
+\begin{center}
+\caption{Steps to consider in imputation simulation studies.}
+\label{table:steps}
+\begin{tabular}{lll}
+\hline
+&Step                    & Simulation aspects under consideration \\
+\hline  
+1&Set scope            & Aim(s), missing data method(s) to evaluate, simulation design \\
+2&Obtain truth         & Data generating mechanism(s), estimand(s), sampling variance \\
+3&Induce missingness   & Missingness mechanism(s), missing data pattern(s) \\
+4&Apply methods        & Imputation model(s), analysis model(s) \\
+5&Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
+ &                      & model fit \\
+6&Evaluate performance & Performance measures, qualitative judgement \\
+7&Report               & Text, visualization, checklist, simulation script \\
+\hline
+\end{tabular}
+\end{center}
+% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data generating mechanisms and missingness proportions).}
+\end{table}
 
@@
 
-\subsection{Step 1. Obtain truth}
+\subsection{Set scope}
 
-The evaluator must to decide whether there is a need for sampling variance in the simulations scheme. If sampling variance cannot be omitted from the simulation scheme, two approaches are possible:
-\begin{enumerate}
-  \item Model-based simulation: draw samples from a known probability distribution, such as the multivariate normal distribution.
-The theoretical parameters under which the samples are obtained will serve as the comparative truth during the simulations.
-  \item Design-based simulation: sample data from a sufficiently large set.
-The parameters of the sufficiently large set will serve as the comparative truth during the simulations.
-This approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
-Applications of design-based simulation are often found in official statistics.
-A benefit of design-based simulation is the ability to use real-life observed data structures.
-\end{enumerate}
-For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches.
-The parameters of that single complete set will serve as comparative truth.
-This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations.
-One needs to realize, however, that the conventional pooling rules \citep[cf.][p.
-76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
+Before setting up a simulation study, the simulator should clearly define the scope of their evaluations.
+A simulation study aimed at imputation method comparison will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
+The choice of imputation method(s) under evaluation is naturally intertwined with the simulation aims and set-up.
+
+One specific simulation study aim (that we have thus far ignored) is to evaluate the predictive performance of pairs of imputation methods and estimation methods in empirical, incomplete data.
+These kinds of simulation studies do not start out from a complete data set in which missingness is induced by the simulator.
+Rather, the evaluations use the observed values of an outcome variable (or target) as the estimand, which is estimated from the incomplete predictor (or feature) space by first imputing the missingness and then applying a prediction method.
+Pairs of imputation and prediction methods with high predictive accuracy in one or more benchmark data sets are deemed as `good' \citep{liu21}.
+Note that such a simulation design does not have a `ground truth', so only the comparative performance of methods may be established.
+We do not recommend this approach if the inferential validity of the imputations is of interest.
+
+% RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
+Refer to missing data chapter.
+In this context, there is no talk of retrieving the `true' but unobserved values at all.
+The DGM is by definition design-based and there is no sampling variance.
+The missingness generation step is skipped and the only performance measure that can be used between methods is RMSE and maybe comparison with respect to CCA in terms of bias.]
+
+Some general choices in the simulation set-up to consider at this stage are the simulation design and the number of simulation repetitions.
+The simulation design should ideally be fully factorial, i.e.
+varying each simulation condition against all other conditions.
+There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions).
+If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
+Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation, which makes for fairer comparisons.
+The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
+as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
 
@@
 
-\subsection{Step 2. Induce missingness}
+\subsection{Obtain truth}
 
-First, one should always consider MCAR missingness, i.e.
-the scenario where the missing values are missing at random and the observed values are observed at random.
-Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
+% 1a. choose DGM (if model-based: choose n_obs, n_var, relations between var, var types, etc; if design-based: choose n_obs, missing data handling strategy)
+% 1b. decide if sampling variance is needed
+% 1c. draw complete sample(s)
 
-Next, missing data should be induced conform a model that is dependent on the observed data.
-A straightforward technique for inducing different forms of univariate MAR missingness is described in \citet[][p.
-63]{fimd} and a generalization to multivariate MAR missingness can be found in \citet[][Appendix B]{buur06} and \citet[][\S 5.2.3]{brand1999development}.
-If the missingness is to be induced in longitudinal data, autoregressive MAR models \citep[e.g.
-cf.][model 2 and model 3]{shara2015randomly} can be useful.
+The simulator should choose their data generating mechanism(s) in line with the study scope.
+Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived.
+A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
 
-Third, it is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
-Given the simulated data-distributions, one random missingness model may be far more disastrous to the observed information than another model.
-This may influence the performance of some (but not necessarily all) imputation routines For example, inference from  hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
-It would be a shame to overlook such results due to the focus on a single MAR mechanism.
+Estimand(s) or other simulation targets should be defined in the context of the study aim and data-generating mechanism.
+As a general rule, it would be wise to include one or more multivariate estimands.
+It is a realistic expectation for imputation methods to preserve both marginal and conditional distributions with respect to the comparative truth.
+Imputation methods that fail to do so, should not be considered general purpose methods (e.g., mean imputation).
 
-Fourth, the amount of missingness must be varied.
-Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
-Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
+The comparative truth of the estimand(s) is determined by the choice of data generating mechanism and whether sampling variance is included in the simulation set-up, see Table \ref{table:dgm_truth}.
+If sampling variance cannot be omitted from the simulation scheme, multiple samples should be drawn from the data generating mechanism (i.e., one sample per simulation repetition, a standard simulation set-up).
+If sampling variance is not of interest, a single complete data set can be obtained from the data generating mechanism (i.e., one sample for all simulation repetitions).
+This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
+We highly value the additional flexibility that such a simulation set-up offers.
+Especially in the case of data transformations it can be challenging to derive the true parametric references to evaluate against.
+Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
+% The simulator should first choose an appropriate data generating mechanism and then decide if sampling variance can be omitted from the simulation scheme to draw one or more samples from the data generating mechanism, respectively.
+When deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
+76-77]{rubi87} do not apply.
+Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
 
-The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
-If there is no missingness - or if there is no data, for that matter - the inference does not depend on the assumption.
-Alternatively, the validity of assumptions become increasingly important when the missingness increases.
-Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
-I therefor propose, for all mechanisms, to evaluate at least the following univariate missingness percentages when evaluating imputation routines:
-
-\begin{itemize}
-  \item 10\% missingness: Depending on the size of the data, this percentage can be considered as a lower bound of realistic evaluation.
-Anything less than 10\% may be of little influence on the true data inference.
-Performance of a missing data method should at least be acceptable for most missing data problems.
-  \item 25\% missingness: This is a fair amount of missingness and will, depending on the observed data information, have a noticeable influence on the completed data inference.
-When compared to the condition with 10\% missingness, the inference obtained under 25\% missingness should be less certain (i.e.
-confidence/credibility interval width should increase), but estimates should still be properly covered and the statistical properties of the missing data method should be sound.
-In practice, at least to my experience in social sciences and official statistics, 25\% univariate missingness can easily be considered as a realistic missingness percentage.
-  \item 50\% missingness: Performance under 50\% percent simulated missingness will most likely be impacted severely.
-Depending on how the missingness mechanism interacts with the simulated data, some imputation techniques may yield estimates that are under-covered such that the completed data inference should not be deemed valid anymore.
-If a method yields acceptable inference under 50\% MAR missingness, we can determine that the statistical properties of the imputation methodology are sound.
-\end{itemize}
-Although I limit the focus here to ignorable non-response, the above suggested proportions are equally applicable to simulations under non-ignorable non-response.
-
-Fifth, the used missingness mechanism should be detailed, either graphically or written as a function of the data.
-Often, when inducing missingness, authors remain vague about the actual missingness mechanism under investigation and, even worse, some authors only report something like
-\begin{quote}
-\emph{We generated missing data following a MAR missingness mechanism.}
-\end{quote}
-This should be considered unacceptable as claims about the validity of the multiple imputation inference depends heavily on the simulated missingness mechanism.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Step 3. Evaluate performance}
-
-It is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
-We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance.
-The evaluation criteria may depend on the estimand.
-When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
-Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
-In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
-\begin{itemize}
-  \item \emph{Bias:} Results should preferable be unbiased.
-However, the way bias is considered can greatly influence the interpretation of the results.
-For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.
-The way bias is computed should therefore be carefully chosen and described.
-  \item \emph{Interval width:} The way the confidence interval is calculated should be described.
-Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference.
-However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.
-  \item \emph{Coverage:} Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage of 95\% would be most efficient.
-Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
-  \item \emph{RMSE:} [RMSE at cell level and at prediction level (if applicable).]
-  \item \emph{Distributional characteristics:} In practice, the distribution of the incomplete data may differ greatly from the observed data.
-Under anything but the MCAR assumption, this can be expected.
-When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for an detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}.
-When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.
-  \item \emph{Plausibility of the imputed values:} Plausible imputations - imputations that could be real values if they had been observed - are not a necessary condition for obtaining valid inference.
-However, in practice, especially when the imputer and the analyst are different persons, plausibility of imputations may be a desired property.
-When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
-  \item \emph{Fit of the imputation model:} [PPC, ref Mingyang, \citep{nguy17, zhao22}.]
-  \item \emph{Convergence of the algorithm:} Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
-The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.
-  \end{itemize}
-For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems.
-This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
-Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
-However, the real-world missingness would then be omitted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Step 4. Report}
-
-Report the following.
-
-[Put this in a table, like this one]
-
-\begin{table}[htb]
+\begin{table}[tb]
 \begin{center}
-\caption{The caption of a table.}
+\caption{Source of the comparative truth in simulation studies.}
+\label{table:dgm_truth}
 \begin{tabular}{lll}
 \hline
-Description 1 & Description 2 & Description 3\\
+               & With sampling variance      & Without sampling variance \\
+               & (multiple samples drawn)    & (one sample drawn) \\
 \hline  
-Row 1, Col 1 & Row 1, Col 2 & Row 1, Col 3\\
-Row 2, Col 1 & Row 2, Col 2 & Row 2, Col 3\\
+Model-based simulation   & data-generating model         & single sample \\
+Design-based simulation  & sufficiently large data set   & single sample \\
 \hline
@@
 
-1. set-up
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-- aim
+\subsection{Induce missingness}
 
-- number of simulations
+% 2a. choose missingness mechanism(s): at least consider MCAR, vary MAR types
+% 2b. choose missing data pattern(s): introduce non-monotone multivariate missingness
+% 2c. choose missingness proportion(s): at least consider 10, 25 and 50\%
 
-2. data generation
+Missingness should be induced according to several sets of missing data conditions.
+We encourage simulators to consider different missingness mechanisms (including missingness types), and different missingness patterns (including missingness proportions).
+Although not every missingness mechanism is realistically assumed in practice, they can all offer valuable insights as simulation condition.
 
-- model vs design based
+First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions).
+This mechanism should be used as a reference condition.
+Every imputation method should yield valid imputations under MCAR.
+We also know that, under MCAR, the distribution of imputed values should be equivalent to the distribution of observed values for a given variable, which may offer an indication of imputation model (mis)fit.
+[TODO: leg uit!]
+% First, one should always consider MCAR missingness [as reference, because all imputation methods should yield valid inference under MCAR and we can use this mechanism to check imputation model fit] %, i.e.
+the scenario where the missing values are missing at random and the observed values are observed at random.
+% Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
 
-- data types
+Next, missing data should be induced conform a model that is dependent on the observed data (i.e.
+a MAR mechanism).
+Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully.
+A straightforward technique for inducing univariate MAR missingness is described in \citet[][p.
+63]{fimd}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
+If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly} can be useful.
+It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
+The effects of different types of MAR mechanisms are described in \citet{scho18}.% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
+This may influence the performance of some (but not necessarily all) imputation routines.
+For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+It would be a shame to overlook results due to the focus on a single MAR mechanism.
 
-- sample vs population (vs bootstrapping??)
+Third, a non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice.
+Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time.
+It would, therefore, be wise to include MNAR missingness in the simulation study just in case.
+Alternatively, sensitivity analyses may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
 
-- number of observations
+Apart from the missingness mechanisms, the missingness pattern must be varied.
+[TODO: beter bruggetje?] Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
+The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
+If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
+Alternatively, the validity of assumptions become increasingly important when the missingness increases.
+Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
+We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness (see Table \ref{table:prop}).
+% Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.
 
-- number of variables
+\begin{table}[htb]
+\begin{center}
+\caption{Suggested univariate missingness proportions.}
+\label{table:prop}
+\begin{tabular}{lll}
+\hline
+Prop. & Reasoning for suggested missingness proportion \\
+\hline  
+10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
+      & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
+      & the true data inference. Performance of a missing data method should at least be \\
+      & acceptable for most missing data problems. \\ % & \\
+25\%  & This is a fair amount of missingness and will, depending on the observed data \\
+      & information, have a noticeable influence on the completed data inference. When \\
+      & compared to the condition with 10\% missingness, the inference obtained under \\
+      & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
+      & should increase), but estimates should still be properly covered and the statistical \\
+      & properties of the missing data method should be sound. In practice, at least to our \\
+      & at least to my experience in social sciences and official statistics, 25\% \\
+      & univariate missingness can easily be considered as a realistic missingness percentage. \\ % & \\
+50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
+      & severely. Depending on how the missingness mechanism interacts with the simulated \\
+      & data, some imputation techniques may yield estimates that are under-covered such that \\
+      & the completed data inference should not be deemed valid anymore. If a method yields \\
+      & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
+      & properties of the imputation methodology are sound. \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
-- coherence between variables
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-3. missingness generation
+\subsection{Apply methods}
 
-- monotone/univariate vs multivariate
+% 3a. Define missing data methods: at least consider complete case analysis as benchmark
+% 3b. Choose imputation model parameters: appropriate n_imp and n_it (if applicable), maybe vary predictor matrices?
+% 3c. Impute the data
+% 3d. Obtain estimates
 
-- missingness mechanism
+The imputation methods under evaluation should be applied to the generated incomplete data, after which an analysis model may be fitted to estimate the estimand(s).
+% It is important to choose the imputation model parameters carefully (e.g., number of imputations, number of iterations, regularization, etc.).
+After imputation, an analysis model may be fitted to estimate the estimand(s).
+%[Use CCA as benchmark method: the imputation method should outperform CCA.
+Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matrix, regularization if applicable, etc.).
+Estimate estimand(s) using the analysis model, use correct pooling rules.]
+In addition, the simulator should always perform complete case analysis.
+We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance.
+Complete case analysis may therefore serve as a benchmark method against which performance should be evaluated.
+Moreover, pairing complete case analysis with MCAR and MAR mechanisms allows the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.% out-performing it should be a minimal requirement for the imputation method(s).
+%it is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
 
-- type of M(N)AR
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-- missingness proportion
+\subsection{Evaluate imputations}
 
-4. missing data methods
+% 4a. Evaluate imputations: check failed methods, algorithmic convergence (if applicable)
+% 4b. Check imputation model fit: anomalies in distributional characteristics, plausibility of imputed values (if requested), PPC
 
-- imputation method
+[Some suggested minimal requirements for imputation methods: the absence of algorithmic non-convergence, the imputation model fits the observed parts of the data (PPC), under a MCAR mechanism the imputed values are distributed equivalent to the observed values.]
 
-- imputation model
+% = evaluation of imputation generating process
+The absence of non-convergence in the imputation models is a minimum requirement for imputation methods.
+Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
+The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations might be invalid.
 
-- number of imputations
+[Distributional characteristics] In practice, the distribution of the incomplete data may differ greatly from the observed data.
+Under anything but the MCAR assumption, this can be expected.
+When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for a detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}.
+When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.
 
-- number of iterations
+The fit of the imputation model may be verified with the help of a posterior predictive check \citep[PPC][]{nguy17, zhao22}.
+A PPC in the context of imputation methodology could be the multiple over-imputation of observed data values.
+The reasoning is that if the over-imputed values have equivalent statistical properties to the true data values, the imputation model fits the observed part of the incomplete data well.
+We might then assume that the imputation model would be able to produce good imputations for missing values too.
 
-5. performance evaluation
+[Plausibility of the imputed values] Plausible imputations--imputations that could be real values if they had been observed--are not a necessary condition for obtaining valid inference.
+However, in practice, especially when the imputer and the analyst are different persons, plausible imputations may be a desired property.
+One would prefer an imputation technique to yield both valid inference and plausible imputations.
+It should be studied if an imputation method is prone to deliver such impractical results, and if so, under what conditions.
+When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
 
-- analysis model
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-- estimand
+\subsection{Evaluate performance}
 
-- diagnostic
+% 4c. Apply performance measures: at least consider bias, coverage, CI length, if applicable also RMSE (at prediction and/or cell level)
+% fraction of missing information (= model parameter), variance inflation
 
-- ppc
+[Some suggested minimum requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis.]
+
+% sim perf and model perf
+[The evaluation criteria may depend on the estimand.
+When descriptive statistics are the goal and when statistical inference would not be of interest, the bias of the estimates would still apply, but standard errors are generally ignored.
+Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.]
+
+In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
+ (i) Bias: Results should preferable be unbiased.
+However, the way bias is considered can greatly influence the interpretation of the results.
+For example, a negligible absolute bias for a parameter for which the true value is zero would yield infinite bias when relative bias is considered.
+The way bias is computed should therefore be carefully chosen and described.
+(ii) Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
+Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
+(iii) Interval width: The way the confidence interval is calculated should be described.
+Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference.
+However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.
+% [Add RMSE at cell level and at prediction level (if applicable).]
+
+Last, when the evaluator is on the verge of drawing conclusions about the performance of the imputation routine, the performance should be carefully qualified.
+Comparing the performance of an imputation routine given a population (or true) parameter allows for quantitative evaluation.
+However, in order to pose qualitative statements about the performance on simulated conditions, comparative methodology is required.
+For example, when claiming that imputation performance is unacceptable when deviations from normality become rather stringent, such performance is highly dependent on the simulation conditions that are used.
+For a well-balanced judgment about the severity of the performance drop, comparative simulations with e.g.
+nonparametric models should be executed.
+A method may perform badly, but if it still outperforms every other approach, it may yet be of great practical relevance.
+
+It could be convincing to demonstrate a method's applicability to real-world missing data problems.
+This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
+Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
+However, the real-world missingness would then be omitted.
+[TODO: check waar dit al eerder staat!]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Report}
+
+[Be explicit in choices.
+Describe the simulation textually, use pseudo-code, a flowchart, visualize missingness and simulation results.
+For example, the used missingness mechanism should be detailed, either graphically or written as a function of the data.
+And the procedure to generate missing data is often insufficiently described in resulting publications.
+Fill in the following checklist (in appendix).]
 
@@
 
+This manuscript aims to establish a common ground for the evaluation of imputation routines.
+Such a common ground would be the basis of a standardized evaluation.
+This would allow for fairer and more efficient comparisons between imputation techniques.
+Ultimately, it would be desirable to evaluate every imputation routine against the same standardized set in order to quantify the statistical properties across imputation routines.
+If properly executed, this would allow for careful matching of imputation methodologies to new missing data problems.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{acknowledgement}
-[Maybe add Gelman here?]
+We thank the Amices team for the fruitful discussions.
 \end{acknowledgement}
@@
 
-\noindent {\it{The authors have declared no conflict of interest. (or please state any conflicts of interest)}}
+\noindent {\it{The authors have declared no conflict of interest.}}
 
-% \section*{Appendix {\it(please insert here, if applicable)}}
-% 
-% \subsection*{A.1.\enspace Second level heading}
-% 
-% Please insert appendices before the references.
+\section*{Appendix}
+
+\subsection*{A.1.\enspace Checklist for the evaluation of imputation methodology}
+
+1. Set-up
+
+- aim
+
+- target/estimand + analysis model
+
+- precision level + number of simulations
+
+2. Data generation
+
+- model vs design based
+
+- sampling variance
+
+- number of observations
+
+- number of variables
+
+- data types + structures (e.g. clustering/time-series)
+
+- coherence between variables
+
+3. Missingness generation
+
+- missingness mechanism + type (if applicable)
+
+- missingness patterns (rows = univariate/multivariate; columns = proportion)
+
+4. Missing data methods
+
+- imputation methods + list-wise deletion
+
+- imputation method parameters (e.g. number of imputations)
+
+- imputation model parameters (e.g. predictor matrix)
+
+5. Imputation evaluation
+
+- imputation process (e.g. non-convergence)
+
+- imputation model fit (PPC)
+
+- distribution of imputations (e.g. plausibility)
+
+- imputation accuracy (e.g. RMSE at cell level)
+
+6. Performance evaluation
+
+- statistical properties (e.g. confidence validity)
+
+- comparative performance (e.g. RMSE of predictions)
+
+7. Report
 

### 2022-03-31 Formatted/Formatted.tex
Process last feedback Gerko 31-03

@@
 \noindent \hspace*{-4pc} Supporting Information for this article is available online from \\
-\hspace*{-4pc} \underline{github.com/gerkovink/StandardizedEvaluation}. 
+\hspace*{-4pc} \underline{gerkovink.com/evaluation}. 
 }  %%% semicolon and fullpoint added here for keyword style
 
-\title[Standardized evaluation]{Towards a standardized evaluation of imputation methodology}
+\title[Towards a standardized evaluation of imputation methodology]{Towards a standardized evaluation of imputation methodology}
 %% Information for the first author.
@@
 Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
-In this paper we propose a move towards a standardized evaluation of imputation methods.
+In this paper, we propose a move towards a standardized evaluation of imputation methods.
 To demonstrate the need for standardization, we highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
@@
 by the detailed manual created by the National Research Council \citep*{little2012prevention}.
-Although a top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers and data scientists are embracing imputation techniques.
+Although top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers and data scientists are embracing imputation techniques.
 After all, the principle of imputation is very intuitive.
@@
 In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods.
-However, no golden standard has been established to evaluate imputation routines and, as a result, the validity and aim of the simulation set-ups may differ tremendously from one developer to another.
+However, no golden standard has been established to evaluate imputation routines and, as a result, the validity and aim of the simulation setups may differ tremendously from one developer to another.
 Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced.
@@
 % [Nowadays there are a lot of imp alg being propagated, esp ai/ml, whose perf often times seems quite promising.
-But may not be a fair comparison because of the simulation set-up/dev in ai/ml so fast that steps may be skipped.
+But may not be a fair comparison because of the simulation setup/dev in ai/ml so fast that steps may be skipped.
 Especially in the field of ML/AI, there are promising new imputation methods, which may yield even sharper imputations than now-standard (semi-)parametric imputation methods]
@@
 To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established.
-Those who perform simulation studies are in the luxury position to establish the truth beforehand by choosing a data generating mechanism.
-Data generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
+Those who perform simulation studies are in the luxury position to establish the truth beforehand by choosing a data-generating mechanism.
+Data-generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
 There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as from official registers.
 
-The problem with model-based data generating mechanisms is that a method's performance on simulated data may not translate to empirical data.
+The problem with model-based data-generating mechanisms is that a method's performance on simulated data may not translate to empirical data.
 Real-life data hardly ever follow a given theoretical distribution, so there is no guarantee that simulation results are generalizable.
@@
 
-After deciding upon a data generating mechanism, another influential--but often overlooked--decision awaits.
-Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations, and only use the missing data generation procedure as source of Monte Carlo variation.
+After deciding upon a data-generating mechanism, another influential--but often overlooked--decision awaits.
+Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations and only use the missing data generation procedure as the source of Monte Carlo variation.
 After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
 Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}.
-The only required change to simulation set-ups is to remove sampling variance from performance evaluation.
+The only required change to simulation setups is to remove sampling variance from performance evaluation.
 A lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
@@
 Evaluating imputation methodology requires a missing data problem to be solved.
-After establishing a comparative truth from a data generating mechanism, some form of missingness therefore has to be induced.
+After establishing a comparative truth from a data-generating mechanism, some form of missingness therefore has to be induced.
 Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like:
@@
 
-Even the terminology on missingness generation can be confusing.
-Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
-In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation.
-The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).
-Moreover, the value of the actual missingness proportion may be diffusing too.
-Some studies use only 10\% missingness where other studies push the limits to additionally investigate the performance under missingness proportions of at least 50\%.
-The inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
-This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
-
-% Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
-Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
-The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
-If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
-Alternatively, the validity of assumptions become increasingly important when the missingness increases.
-Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
-
-Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables.
-It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways.
-Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to sub-optimal advice in practice.
-For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that list-wise deletion outperforms their imputation method(s).
-Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described in][\S 2.7]{fimd}, while biasing inferences in other cases.
-Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e.
-a pattern with uniformly increasing missingness proportions along variables;][]{litt20}.
-If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations.
-A final example is multivariate missingness that is generated using step-wise univariate missingness induction.
-The resulting incomplete data may then not have the desired statistical properties \citep[e.g.
-skewness;][]{ampute}.
-This complicates any definitive conclusions about imputation method performance in relation to the data generating mechanism.
-
-% is the   [Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.] For example, iterative imputation algorithms converge instantly if there is only one variable with missingness or when the pattern is monotone, but require iteration in all other situations.
-And very specific missing data patterns may misguide which imputation method prevails in performance: some ad hoc methods are known to yield valid inferences under certain restricted conditions \citep[e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively][]{fimd}, while biasing inferences in other cases.
-% [Missingness mechanisms paragraphs: 1) definitions of mechanisms (now in table), 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low,  4) that MNAR is often ignored, while it may be the most realistic in practice, and 5) finally add something about types of M(N)AR (ref Rianne's Dance paper).]
-% MCAR missingness: The scenario where the missing values are missing at random and the observed values are observed at random. 
-%  the validity of assumptions become increasingly important when the missingness increases.
-
 \begin{table}[tb]
@@
 \hline
-Mech. & Interpretation and consequences of the mechanism \\
+      & Interpretation and consequences of the mechanism \\
 \hline  
@@
 
+Even the terminology on missingness generation can be confusing.
+Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry?
+In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation.
+The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).
+Moreover, the value of the actual missingness proportion may be diffusing too.
+Some studies use only 10\% missingness whereas other studies push the limits to additionally investigate the performance under missingness proportions of at least 50\%.
+The inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
+This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
+
+% Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
+The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
+If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
+Alternatively, the validity of assumptions become increasingly important when the missingness increases.
+Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
+
+Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables.
+It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways.
+Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to sub-optimal advice in practice.
+For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that list-wise deletion outperforms their imputation method(s).
+Such a conclusion would, unfortunately, only translate to data that adhere to the same special case while biasing inferences in other cases \citep[][\S 2.7]{buur18}.
+Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e.
+a pattern with uniformly increasing missingness proportions along variables;][]{litt20}.
+If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations.
+A final example is multivariate missingness which is generated using step-wise univariate missingness induction.
+The resulting incomplete data may then not have the desired statistical properties \citep[e.g.
+skewness;][]{ampute}.
+This complicates any definitive conclusions about imputation method performance in relation to the data-generating mechanism.
+
+% is the   [Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.] For example, iterative imputation algorithms converge instantly if there is only one variable with missingness or when the pattern is monotone, but require iteration in all other situations.
+And very specific missing data patterns may misguide which imputation method prevails in performance: some ad hoc methods are known to yield valid inferences under certain restricted conditions \citep[e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively][]{buur18}, while biasing inferences in other cases.
+% [Missingness mechanisms paragraphs: 1) definitions of mechanisms (now in table), 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low,  4) that MNAR is often ignored, while it may be the most realistic in practice, and 5) finally add something about types of M(N)AR (ref Rianne's Dance paper).]
+% MCAR missingness: The scenario where the missing values are missing at random and the observed values are observed at random. 
+%  the validity of assumptions become increasingly important when the missingness increases.
+
 Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions).
@@
 On the one hand, there is MCAR missingness, which may be unlikely as the true missingness generating mechanism, but would yield insightful evaluations in a simulation setting.
-On the other hand, there is MNAR, which is arguably a quite reasonable source of many real-world missingness scenarios, but is generally ignored except for those evaluations that are specifically targeted at non-ignorable applications.
+On the other hand, there is MNAR, which is arguably a quite reasonable source of many real-world missingness scenarios but is generally ignored except for those evaluations that are specifically targeted at non-ignorable applications.
 A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.
@@
 
-Although MAR missingness is often considered as simulation condition, the problem of spurious MAR is generally overlooked.
+Although MAR missingness is often considered as a simulation condition, the problem of spurious MAR is generally overlooked.
 With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g.
@@
 The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random.
-In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p.
-48]{fimd}: the missingness does not depend on the incomplete variable.
+In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][\S 2.7]{buur18}: the missingness does not depend on the incomplete variable.
 This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is pointless.
@@
 This is not surprising as people from different fields could have a different focus on the problem at hand.
-There are, however, some over-arching issues with assessing imputation method performance.
-In the first place, evaluations of the imputation-generating process (e.g., an iterative imputation algorithm) are generally left out of simulation study results.
-In the second place, there are performance measures pitfalls to consider.
+There are, however, some overarching issues with assessing imputation method performance.
+In the first place, diagnostic evaluation of the generated imputations is often left out of simulation study results.
+Problems with the imputation-generating process (e.g., an iterative imputation algorithm) may then be overlooked.
+In the second place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model (i.e.
+the complete data model used to estimate the comparative truth).
+The choice of performance measures might [dictate (TODO: re-read and find better word!)] the results of the simulation study.
 
@@
 As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated.
-Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{fimd} is neither efficient nor failproof.
+Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor failproof.
 As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method.
-Moreover, failing to produce stable imputations is not an exclusive quality iterative aimputation algorithms.
-For example if a method does not yield any imputations at all in a certain simulation repetition (e.g.
-due to over-parameterization), this can be seen as failure in the imputation-generating process too.
+Problems with producing stable imputations is not an exclusive quality iterative imputation algorithms.
+Any method may run into failures of the imputation-generating process and subsequently lack results (e.g.
+due to over-parameterization errors).
 
-The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e.
-the complete data model used to estimate a target or estimand).
+Omitting further evaluation of the imputation-generating process may yield sub-optimal imputations.
+The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models.
 When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
-Congeniality is difficult to diagnose, but ignoring it may yield sub-optimal imputations.
-Methods to asses the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}, which is an infeasible endeavor in many simulation studies.
-Therefore, the performance of imputation procedures on distributional properties is often ignored in simulation studies.
+The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
+This makes its assessment an infeasible endeavor in many simulation studies.
+Moreover, the performance of imputation procedures on distributional properties is often ignored too.
 Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
@@
 
+The second source of problems with evaluating simulation outcomes is the choice of performance measures.
+Performance measures should express how well the obtained estimates approximate the comparative truth, but not every metric is suited for the evaluation of imputation methodology.
+For example, using measures of predictive accuracy or the RMSE (root mean squared error) as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{buur18}.
+And, under certain simulation conditions, an otherwise useful metric such as relative bias can distort the interpretation of results (i.e., a negligible absolute bias may yield an infinite relative bias if the true value of the estimand is zero).
+The usefulness of evaluation criteria may thus depend on the simulation goal and estimand.
+
 If the goal is inference--not prediction--then the uncertainty about estimates needs to be properly quantified.
-A single imputation would then not suffice.
+To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation.
 The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
@@
 591]{neym34}.
+
+The choice of performance measures may inadvertently distort statistical properties of the imputed data.
+Developers often only inquire about the `accuracy' (i.e.
+how well can the method reproduce the original data).
 Merely focusing on the discrepancy between the true and imputed data would open the door for invalid inferences.
 
-Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.
-For many outcome-driven imputation task a single valid imputation may perhaps suffice.
-We would like to note, however, that using measures of predictive accuracy or the RMSE (root mean squared error) of predicted values as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{fimd}.
+% The evaluation criteria may depend on the estimand.
+When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
+
+% Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.  
+% For many outcome-driven imputation tasks a single valid imputation may perhaps suffice.
 
@@
 \hline
-&Step                    & Simulation aspects under consideration \\
+                        & Simulation aspects under consideration \\
 \hline  
-1&Set scope            & Aim(s), missing data method(s) to evaluate, simulation design \\
-2&Obtain truth         & Data generating mechanism(s), estimand(s), sampling variance \\
-3&Induce missingness   & Missingness mechanism(s), missing data pattern(s) \\
-4&Apply methods        & Imputation model(s), analysis model(s) \\
-5&Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
- &                      & model fit \\
-6&Evaluate performance & Performance measures, qualitative judgement \\
-7&Report               & Text, visualization, checklist, simulation script \\
+1. Set scope            & Aim(s), missing data method(s) to evaluate, simulation design \\
+2. Obtain truth         & Data-generating mechanism(s), estimand(s), sampling variance \\
+3. Induce missingness   & Missingness mechanism(s), missing data pattern(s) \\
+4. Apply methods        & Imputation method(s), analytic method(s) \\
+5. Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
+                        & model fit \\
+6. Evaluate performance & Performance measures, qualitative judgment \\
+7. Report               & Text, visualization, checklist, simulation script \\
 \hline
@@
 \end{center}
-% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data generating mechanisms and missingness proportions).}
+% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data-generating mechanisms and missingness proportions).}
 \end{table}
@@
 A simulation study aimed at imputation method comparison will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
-The choice of imputation method(s) under evaluation is naturally intertwined with the simulation aims and set-up.
+The choice of imputation method(s) under evaluation is naturally intertwined with the simulation aims and setup.
 
@@
 
-Some general choices in the simulation set-up to consider at this stage are the simulation design and the number of simulation repetitions.
+Some general choices in the simulation setup to consider at this stage are the simulation design and the number of simulation repetitions.
 The simulation design should ideally be fully factorial, i.e.
@@
 
-The simulator should choose their data generating mechanism(s) in line with the study scope.
+The simulator should choose their data-generating mechanism(s) in line with the study scope.
 Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived.
@@
 It is a realistic expectation for imputation methods to preserve both marginal and conditional distributions with respect to the comparative truth.
-Imputation methods that fail to do so, should not be considered general purpose methods (e.g., mean imputation).
+Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).
 
-The comparative truth of the estimand(s) is determined by the choice of data generating mechanism and whether sampling variance is included in the simulation set-up, see Table \ref{table:dgm_truth}.
-If sampling variance cannot be omitted from the simulation scheme, multiple samples should be drawn from the data generating mechanism (i.e., one sample per simulation repetition, a standard simulation set-up).
-If sampling variance is not of interest, a single complete data set can be obtained from the data generating mechanism (i.e., one sample for all simulation repetitions).
+The comparative truth of the estimand(s) is determined by the choice of data-generating mechanism and whether sampling variance is included in the simulation setup, see Table \ref{table:dgm_truth}.
+If sampling variance cannot be omitted from the simulation scheme, multiple samples should be drawn from the data-generating mechanism (i.e., one sample per simulation repetition, a standard simulation setup).
+If sampling variance is not of interest, a single complete data set can be obtained from the data-generating mechanism (i.e., one sample for all simulation repetitions).
 This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
-We highly value the additional flexibility that such a simulation set-up offers.
-Especially in the case of data transformations it can be challenging to derive the true parametric references to evaluate against.
+We highly value the additional flexibility that such a simulation setup offers.
+Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against.
 Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
-% The simulator should first choose an appropriate data generating mechanism and then decide if sampling variance can be omitted from the simulation scheme to draw one or more samples from the data generating mechanism, respectively.
 When deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
@@
 Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully.
-A straightforward technique for inducing univariate MAR missingness is described in \citet[][p.
-63]{fimd}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
+A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
 If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly} can be useful.
@@
 If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
-Alternatively, the validity of assumptions become increasingly important when the missingness increases.
+Alternatively, the validity of assumptions becomes increasingly important when the missingness increases.
 Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
@@
 \hline
-Prop. & Reasoning for suggested missingness proportion \\
+      & Reasoning for suggested missingness proportion \\
 \hline  
@@
       & properties of the missing data method should be sound. In practice, at least to our \\
-      & at least to my experience in social sciences and official statistics, 25\% \\
+      & at least to our experience in social sciences and official statistics, 25\% \\
       & univariate missingness can easily be considered as a realistic missingness percentage. \\ % & \\
@@
 We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance.
-Complete case analysis may therefore serve as a benchmark method against which performance should be evaluated.
-Moreover, pairing complete case analysis with MCAR and MAR mechanisms allows the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.% out-performing it should be a minimal requirement for the imputation method(s).
+Complete case analysis may, therefore, serve as a benchmark method against which imputation performance should be evaluated.
+Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.% out-performing it should be a minimal requirement for the imputation method(s).
 %it is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
@@
 
-[Some suggested minimal requirements for imputation methods: the absence of algorithmic non-convergence, the imputation model fits the observed parts of the data (PPC), under a MCAR mechanism the imputed values are distributed equivalent to the observed values.]
+Every imputation workflow should contain an evaluation of the obtained imputations.
+Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend to consider the following aspects:
+% A minimum requirement for any imputation method is the absence of non-convergence in the imputation-generating process.
+If non-convergence is suspected, the inference resulting from the imputations might be invalid.
+% [Some suggested minimal requirements for imputation methods: the absence of algorithmic non-convergence, the imputation model fits the observed parts of the data (PPC), under a MCAR mechanism the imputed values are distributed equivalent to the observed values.]
 
-% = evaluation of imputation generating process
-The absence of non-convergence in the imputation models is a minimum requirement for imputation methods.
-Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}.
-The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations might be invalid.
+(i) The absence of non-convergence in the imputation-generating process is a minimum requirement for any imputation method.
+If non-convergence is suspected, the inference resulting from the imputations might be invalid.
+However, preliminary work suggests that iterative imputation algorithms could achieve inferential validity before reaching a stable state \citep{ober21}.
 
-[Distributional characteristics] In practice, the distribution of the incomplete data may differ greatly from the observed data.
+(ii) The fit of the imputation model may be verified with the help of a posterior predictive check \citep[][]{nguy17, zhao22}.
+A straightforward posterior predictive check for imputation methodology is the multiple over-imputation of observed data values.
+If the statistical properties of the over-imputed values are equivalent to those of the observed data values, one could infer that the imputation model fits the observed part of the incomplete data reasonably well.
+Then, by extension, one could assume that the imputation model might be able to produce good imputations for the missing part of the data too.
+
+(iii) The distributional characteristics of the imputations should be inspected for anomalies.
+The distribution of the incomplete data may differ greatly from the observed data.
 Under anything but the MCAR assumption, this can be expected.
@@
 
-The fit of the imputation model may be verified with the help of a posterior predictive check \citep[PPC][]{nguy17, zhao22}.
-A PPC in the context of imputation methodology could be the multiple over-imputation of observed data values.
-The reasoning is that if the over-imputed values have equivalent statistical properties to the true data values, the imputation model fits the observed part of the incomplete data well.
-We might then assume that the imputation model would be able to produce good imputations for missing values too.
-
-[Plausibility of the imputed values] Plausible imputations--imputations that could be real values if they had been observed--are not a necessary condition for obtaining valid inference.
+(iv) Finally, the plausibility of the imputed values may be evaluated.
+Plausible imputations--imputations that could be real values if they had been observed--are not a necessary condition for obtaining valid inference.
 However, in practice, especially when the imputer and the analyst are different persons, plausible imputations may be a desired property.
@@
 
-[Some suggested minimum requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis.]
+Imputation method performance should be quantified using appropriate measures.
+It depends on the specifics of each simulation study which performance measures would be most suitable.
+In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
 
-% sim perf and model perf
-[The evaluation criteria may depend on the estimand.
+(i) The estimates should preferably be unbiased.
+Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results.
+Unbiased estimates under MCAR are a minimum requirement for any imputation method.
+
+(ii) The estimates should have a proper coverage rate.
+Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
+Under-coverage (when estimation is too liberal and intervals are too narrow) indicates that the procedure is not confidence valid and may lead to invalid inference.
+Over-coverage (when estimation is too conservative and intervals are too wide) tells us that efficiency could still be gained.
+
+(iii) The width of the confidence or credible interval may convey statistical efficiency, which should be considered to compare imputation methods.
+Wider intervals are associated with more uncertainty whereas a more narrow interval that is still properly covered indicates a sharper inference.
+However, inference from a wider interval that is properly covered is to be considered more valid than a more narrow interval that is not properly covered anymore.
+
+We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values \citep[][\S 2.5]{buur18}.
+However, if a study is aimed at obtaining predictions or classification and inferential validity is not of interest, the RMSE of the predictions may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
+% [Some suggested minimum requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis.]
+% 
+% % sim perf and model perf
+% [The evaluation criteria may depend on the estimand.
 When descriptive statistics are the goal and when statistical inference would not be of interest, the bias of the estimates would still apply, but standard errors are generally ignored.
 Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.]
-
-In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
- (i) Bias: Results should preferable be unbiased.
+% 
+% 
+% (i) Bias: Results should preferably be unbiased.
 However, the way bias is considered can greatly influence the interpretation of the results.
@@
 The way bias is computed should therefore be carefully chosen and described.
-(ii) Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
+% (ii) Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
 Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
-(iii) Interval width: The way the confidence interval is calculated should be described.
+% (iii) Interval width: The way the confidence interval is calculated should be described.
 Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference.
 However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.
-% [Add RMSE at cell level and at prediction level (if applicable).]
+% % [Add RMSE at cell level and at prediction level (if applicable).]
 
-Last, when the evaluator is on the verge of drawing conclusions about the performance of the imputation routine, the performance should be carefully qualified.
+Last, when the simulator is on the verge of drawing conclusions about the performance of the imputation methodology, the performance should be carefully qualified.
 Comparing the performance of an imputation routine given a population (or true) parameter allows for quantitative evaluation.
-However, in order to pose qualitative statements about the performance on simulated conditions, comparative methodology is required.
+Yet, in order to pose qualitative statements about the performance on simulated conditions, comparative methodology is required.
 For example, when claiming that imputation performance is unacceptable when deviations from normality become rather stringent, such performance is highly dependent on the simulation conditions that are used.
@@
 
-It could be convincing to demonstrate a method's applicability to real-world missing data problems.
+% It could be convincing to demonstrate a method's applicability to real-world missing data problems.
 This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
@@
 
-[Be explicit in choices.
+After the careful consideration and execution of all simulation aspects, the evaluations should be properly reported.
+We encourage simulators to document their choices and to be explicit in their descriptions.
+For example, the simulation design may be presented textually, in a flow chart or as a block of pseudo-code, whereas missingness mechanisms may be written as a function of the data or displayed graphically.
+Ideally, this should be supplemented by an online repository with all of the data and code required to reproduce the simulation results.
+To aid simulators in reporting and move towards standardization in evaluation, we provide a draft version for reporting guidelines in Appendix A.1 (also available from \underline{gerkovink.com/evaluation}).
+We invite the readers of this paper to contribute to its development.
+
+% [Be explicit in choices.
 Describe the simulation textually, use pseudo-code, a flowchart, visualize missingness and simulation results.
-For example, the used missingness mechanism should be detailed, either graphically or written as a function of the data.
-And the procedure to generate missing data is often insufficiently described in resulting publications.
+For example, the procedure to generate missing data is often insufficiently described in publications.
+The used missingness mechanism should be detailed, either graphically or written as a function of the data.
 Fill in the following checklist (in appendix).]
@@
 
-This manuscript aims to establish a common ground for the evaluation of imputation routines.
+This manuscript aims to elicit critical thinking about incomplete data simulation and to establish a common ground for the evaluation of imputation routines.
 Such a common ground would be the basis of a standardized evaluation.
@@
 
-\subsection*{A.1.\enspace Checklist for the evaluation of imputation methodology}
+% \subsection*{A.1.\enspace Simulation setup}
 
-1. Set-up
+% Let $Y=(Y_{obs},Y_{mis})$ be an incomplete semicontinuous variable, where $Y_{obs}$ and $Y_{mis}$ denote the observed values and the missing values in $Y$, respectively.
+Further, $X=(X_1,\dots,X_k)$ is a set of $k$ fully observed covariates, where $X_{obs}$ and $X_{mis}$ correspond to the observed an missing parts in $Y$ and where $X_j$ would indicate the $j$th variable in $X$, with $j=1,\dots, k$.
+Finally, let $R$ be a response indicator that is 1 if $Y$ is observed and 0 if $Y$ is missing.
+% 
+% To prove the point, we generated a single data set from the multivariate normal distribution with means
+% \begin{equation*}\label{meanmatrix}
+% \mu= \bordermatrix{&	 \cr
+%  Y	&5	\cr
+% { X}_1	&10	\cr
+% { X}_2	&10	},
+% \end{equation*}
+% and covariance matrix
+% \begin{equation*}\label{covmatrix}
+% \Sigma = \bordermatrix{& Y &{ X}_1 &{ X}_2	 \cr
+%  Y	&1	&0.8	&0\cr
+% { X}_1	&0.8	&1	&0\cr
+% { X}_2	&0	&0	&1}.
+% \end{equation*}
+% 
+% Three incomplete sets were then generated, wherein random missingness was imposed in $Y$ in three ways.
+First, we left the observed data \emph{observed at random} to impose MCAR missingness following the mechanism
+% \begin{equation*}
+% P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0)=.50.
+% \end{equation*}
+% Next, I created random missingness in $Y$ according to the following MAR missingness mechanism
+% \begin{equation*}
+% P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0|Y_{obs}, X_j),
+% \end{equation*}
+% by using a random draw from a binomial distribution of the same length as $Y$ and of size 1 with missingness probability equal to the inverse logit
+% \begin{equation*}
+% P(R=0)=\frac{e^{a}}{(1+e^{a})}.
+% \end{equation*}
+% where $a=(\mu_{X_j}-X_{ij})/\sqrt{\mathbb{E}[X_{ij} - \mu_{X_j}]}$ in order to create approximately 50\% right-tailed MAR missingness. I followed this procedure twice; once for each $X_j$.
 
-- aim
+% The three resulting incomplete data sets were then imputed with package \texttt{mice} \citep[][v2.25]{mice} in \texttt{R} \citep[][v3.2.3]{R}, with Bayesian linear regression imputation (\texttt{mice.impute.norm}) as the imputation routine.
+The completed-data analyses were combined into a single inference following the rules described in \citet{vink2014pooling}.
 
-- target/estimand + analysis model
+\subsection*{A.1.\enspace Reporting guidelines}
 
-- precision level + number of simulations
-
-2. Data generation
-
-- model vs design based
-
-- sampling variance
-
-- number of observations
-
-- number of variables
-
-- data types + structures (e.g. clustering/time-series)
-
-- coherence between variables
-
-3. Missingness generation
-
-- missingness mechanism + type (if applicable)
-
-- missingness patterns (rows = univariate/multivariate; columns = proportion)
-
-4. Missing data methods
-
-- imputation methods + list-wise deletion
-
-- imputation method parameters (e.g. number of imputations)
-
-- imputation model parameters (e.g. predictor matrix)
-
-5. Imputation evaluation
-
-- imputation process (e.g. non-convergence)
-
-- imputation model fit (PPC)
-
-- distribution of imputations (e.g. plausibility)
-
-- imputation accuracy (e.g. RMSE at cell level)
-
-6. Performance evaluation
-
-- statistical properties (e.g. confidence validity)
-
-- comparative performance (e.g. RMSE of predictions)
-
-7. Report
+\begin{table}[h]
+\caption{Checklist for reporting on imputation methodology evaluations}
+\begin{tabular}{ll}
+% \hline
+%   & Suggested aspects to report                                                       \\
+\hline
+1 & Simulation scope                                                                   \\ \cline{2-2}
+  & Aim                                                                                \\
+  & Design (incl. pseudo-code or flow diagram)                                         \\
+  & Number of simulation repetitions                                                   \\ \hline
+2 & Comparative truth                                                                  \\ \cline{2-2}
+  & Data-generating mechanism (model-based or design-based)                            \\
+  & Sampling variance                                                                  \\
+%  & Data characteristics (number of observations, number of variables, variable types, \\
+%  & coherence between variables incl. data structures e.g. clustering)                 \\
+  & Data characteristics (incl. multivariate relations and structures e.g. clustering) \\
+  & Estimand                                                                           \\ \hline 
+3 & Induced missingness                                                                \\ \cline{2-2}
+  & Missingness mechanism (incl. type or functional form of the missing data   model)  \\
+  & Missingness pattern (incl. missingness proportion)                                 \\ \hline 
+4 & Applied methods                                                                    \\ \cline{2-2}
+  & Imputation methods (incl. parameters e.g. the number of imputations)               \\
+  & Analytic methods (incl. calculation of standard errors e.g. pooling   rules)       \\
+  & Reference method (e.g. complete case analysis)                                    \\ \hline 
+5 & Imputation evaluation                                                              \\ \cline{2-2}
+  & Imputation-generating process (e.g. algorithmic non-convergence)                   \\
+  & Imputation model fit (e.g. posterior predictive checks)                            \\
+  & Distributional characteristics (e.g. plausibility of imputed values)               \\ \hline 
+6 & Performance evaluation                                                             \\ \cline{2-2}
+  & Statistical properties (e.g. confidence validity)                                  \\
+  & Comparative performance (e.g. predictive accuracy)                                 \\ \hline
+\end{tabular}
+\end{table}
 

### 2022-03-31 Formatted/Formatted.tex
Before cleaning up comments in the manuscript

@@
 % } %%% semicolon and fullpoint added here for keyword style
-\noindent \hspace*{-4pc} Supporting Information for this article is available online from \\
-\hspace*{-4pc} \underline{gerkovink.com/evaluation}. 
+\noindent \hspace*{-4pc} Supporting Information for this article is available from \underline{gerkovink.com/evaluation}.\\
+% \hspace*{-4pc} \underline{gerkovink.com/evaluation}. 
 }  %%% semicolon and fullpoint added here for keyword style
@@
 
-\begin{table}[h]
+Table \ref{table:check} provides a checklist for reporting on imputation methodology evaluations.
+
+\begin{table}[ht]
 \caption{Checklist for reporting on imputation methodology evaluations}
+\label{table:check}
 \begin{tabular}{ll}

### 2022-03-31 Formatted/Formatted.tex
After removing comments, before fixing TODOs

@@
 
-% [Nowadays there are a lot of imp alg being propagated, esp ai/ml, whose perf often times seems quite promising.
-But may not be a fair comparison because of the simulation setup/dev in ai/ml so fast that steps may be skipped.
-Especially in the field of ML/AI, there are promising new imputation methods, which may yield even sharper imputations than now-standard (semi-)parametric imputation methods]
-
 The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology.
@@
 After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
-Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}.
+Therefore, taking sampling variation into account is neither necessary for obtaining information about a method's ability to handle the missing data problem, nor for objectively comparing methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}.
 The only required change to simulation setups is to remove sampling variance from performance evaluation.
-A lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
-
-% Table \ref{table:dgm_truth} displays the source of the comparative truth in simulation studies with and without sampling variance.When deviating conventional pooling rules \citep[cf.][p.
-76-77]{rubi87} do not apply.
-Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
-If the sampling variance is not handled correctly, subsequent inferences may be invalid and the simulation results cannot be trusted.
+Still, a lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
 
@@
 
-% Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
-Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
-The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
-If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
-Alternatively, the validity of assumptions become increasingly important when the missingness increases.
-Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
-
 Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables.
@@
 
-% is the   [Other problems with missingness patterns: monotone/univariate vs multivariate, sporadic vs systematic, and univariately generated missing data may not have the expected multivariate patterns.] For example, iterative imputation algorithms converge instantly if there is only one variable with missingness or when the pattern is monotone, but require iteration in all other situations.
-And very specific missing data patterns may misguide which imputation method prevails in performance: some ad hoc methods are known to yield valid inferences under certain restricted conditions \citep[e.g., list-wise deletion may outperform imputation methods if there is missingness in the outcome variable of the analysis model exclusively][]{buur18}, while biasing inferences in other cases.
-% [Missingness mechanisms paragraphs: 1) definitions of mechanisms (now in table), 2) that MCAR is often ignored, while it's the minimum requirement and reference, and sometimes realistic in practice, 3) that MAR can be generated \textit{spuriously} if the correlation between variables is low,  4) that MNAR is often ignored, while it may be the most realistic in practice, and 5) finally add something about types of M(N)AR (ref Rianne's Dance paper).]
-% MCAR missingness: The scenario where the missing values are missing at random and the observed values are observed at random. 
-%  the validity of assumptions become increasingly important when the missingness increases.
-
 Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions).
@@
 A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.
-
-% Out of the three mechanisms , random missingness (MAR) is usually generated for the evaluation of imputation methodology.
-Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored, except for those methods that are specifically targeted at non-ignorable applications.
-
-% MAR missingness is usually investigated in simulations and assumed in practice.
-Although MCAR might not reasonably be the true missingness generating mechanism and naturally not typically assumed, the mechanism  can be very insightful in the context of imputation methodology evaluation.
-By contrast, MNAR
-
-% usually assumed to be random (MAR) or completely random (MCAR).
 
@@
 
-% Figure \ref{fig:MAR} demonstrates two realizations of missingness mechanisms under MAR mechanisms derived from data with different multivariate relations.
-
-% \begin{figure}[htb]
-% \begin{center}
-% \resizebox{\textwidth}{!}{
-% \includegraphics[]{figures/spurious_MAR.pdf}}
-% \caption{Three missingness mechanisms that yield approximately 50 percent missingness ($N=500$).
-Displayed are a MCAR mechanism, a right-tailed MAR mechanism generated from bivariate normal data with correlation $\rho=.8$ and the same MAR mechanism, now generated from bivariate normal data with correlation $\rho= 0$.}
-% \end{center}
-% \label{fig:MAR}
-% \end{figure}
-% \vspace*{1pc}
-
 Finally, although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
 As stated, this mechanism is not usually included in evaluations.
-% The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).
-
-% Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
-Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
-The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
-If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
-Alternatively, the validity of assumptions become increasingly important when the missingness increases.
-Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
 
@@
 
-% The evaluation criteria may depend on the estimand.
-When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
-
-% Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification.  
-% For many outcome-driven imputation tasks a single valid imputation may perhaps suffice.
-
-% The evaluation criteria may depend on the estimand.
-When descriptive statistics are the goal and when statistical inference would not be of interest, bias of the estimates would still apply, but standard errors are generally ignored.
-Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.
-
-% For example, a negligible absolute bias for a parameter for which the true value is zero, would yield infinite bias when relative bias is considered.
-
-% [Add segway between evaluation of imputation and actual performance measures.]
-% [Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]
-% The choice of performance measures may inadvertently distort statistical properties of the imputed data.
-Developers often only inquire about the `accuracy' (i.e.
-how well can the method reproduce the original data).
-[Add RMSE at cell level here?]
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{center}
-% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data-generating mechanisms and missingness proportions).}
 \end{table}
@@
 We do not recommend this approach if the inferential validity of the imputations is of interest.
-
-% RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods.
-Refer to missing data chapter.
-In this context, there is no talk of retrieving the `true' but unobserved values at all.
-The DGM is by definition design-based and there is no sampling variance.
-The missingness generation step is skipped and the only performance measure that can be used between methods is RMSE and maybe comparison with respect to CCA in terms of bias.]
 
@@
 [TODO: leg uit!]
-% First, one should always consider MCAR missingness [as reference, because all imputation methods should yield valid inference under MCAR and we can use this mechanism to check imputation model fit] %, i.e.
-the scenario where the missing values are missing at random and the observed values are observed at random.
-% Under MCAR, the statistical properties of the observed data given the missing data are known and any imputation routine that cannot at least mimic the performance of the observed data inference, should be deemed inefficient in the scope of the simulation.
 
@@
 It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
-The effects of different types of MAR mechanisms are described in \citet{scho18}.% Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model.
-This may influence the performance of some (but not necessarily all) imputation routines.
-For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
-It would be a shame to overlook results due to the focus on a single MAR mechanism.
+The effects of different types of MAR mechanisms are described in \citet{scho18}.
 
@@
 We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness (see Table \ref{table:prop}).
-% Although we limit the focus here to ignorable non-response, the suggested proportions are equally applicable to simulations under non-ignorable non-response.
 
@@
 The imputation methods under evaluation should be applied to the generated incomplete data, after which an analysis model may be fitted to estimate the estimand(s).
-% It is important to choose the imputation model parameters carefully (e.g., number of imputations, number of iterations, regularization, etc.).
-After imputation, an analysis model may be fitted to estimate the estimand(s).
-%[Use CCA as benchmark method: the imputation method should outperform CCA.
-Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matrix, regularization if applicable, etc.).
-Estimate estimand(s) using the analysis model, use correct pooling rules.]
 In addition, the simulator should always perform complete case analysis.
@@
 Complete case analysis may, therefore, serve as a benchmark method against which imputation performance should be evaluated.
-Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.% out-performing it should be a minimal requirement for the imputation method(s).
-%it is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.
+Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.
 
@@
 Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend to consider the following aspects:
-% A minimum requirement for any imputation method is the absence of non-convergence in the imputation-generating process.
-If non-convergence is suspected, the inference resulting from the imputations might be invalid.
-% [Some suggested minimal requirements for imputation methods: the absence of algorithmic non-convergence, the imputation model fits the observed parts of the data (PPC), under a MCAR mechanism the imputed values are distributed equivalent to the observed values.]
 
@@
 However, if a study is aimed at obtaining predictions or classification and inferential validity is not of interest, the RMSE of the predictions may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
-% [Some suggested minimum requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis.]
-% 
-% % sim perf and model perf
-% [The evaluation criteria may depend on the estimand.
-When descriptive statistics are the goal and when statistical inference would not be of interest, the bias of the estimates would still apply, but standard errors are generally ignored.
-Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.]
-% 
-% 
-% (i) Bias: Results should preferably be unbiased.
-However, the way bias is considered can greatly influence the interpretation of the results.
-For example, a negligible absolute bias for a parameter for which the true value is zero would yield infinite bias when relative bias is considered.
-The way bias is computed should therefore be carefully chosen and described.
-% (ii) Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
-Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained.
-% (iii) Interval width: The way the confidence interval is calculated should be described.
-Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference.
-However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.
-% % [Add RMSE at cell level and at prediction level (if applicable).]
 
@@
 However, the real-world missingness would then be omitted.
-[TODO: check waar dit al eerder staat!]
 
@@
 We invite the readers of this paper to contribute to its development.
-
-% [Be explicit in choices.
-Describe the simulation textually, use pseudo-code, a flowchart, visualize missingness and simulation results.
-For example, the procedure to generate missing data is often insufficiently described in publications.
-The used missingness mechanism should be detailed, either graphically or written as a function of the data.
-Fill in the following checklist (in appendix).]
 
@@
 \section*{Appendix}
-
-% \subsection*{A.1.\enspace Simulation setup}
-
-% Let $Y=(Y_{obs},Y_{mis})$ be an incomplete semicontinuous variable, where $Y_{obs}$ and $Y_{mis}$ denote the observed values and the missing values in $Y$, respectively.
-Further, $X=(X_1,\dots,X_k)$ is a set of $k$ fully observed covariates, where $X_{obs}$ and $X_{mis}$ correspond to the observed an missing parts in $Y$ and where $X_j$ would indicate the $j$th variable in $X$, with $j=1,\dots, k$.
-Finally, let $R$ be a response indicator that is 1 if $Y$ is observed and 0 if $Y$ is missing.
-% 
-% To prove the point, we generated a single data set from the multivariate normal distribution with means
-% \begin{equation*}\label{meanmatrix}
-% \mu= \bordermatrix{&	 \cr
-%  Y	&5	\cr
-% { X}_1	&10	\cr
-% { X}_2	&10	},
-% \end{equation*}
-% and covariance matrix
-% \begin{equation*}\label{covmatrix}
-% \Sigma = \bordermatrix{& Y &{ X}_1 &{ X}_2	 \cr
-%  Y	&1	&0.8	&0\cr
-% { X}_1	&0.8	&1	&0\cr
-% { X}_2	&0	&0	&1}.
-% \end{equation*}
-% 
-% Three incomplete sets were then generated, wherein random missingness was imposed in $Y$ in three ways.
-First, we left the observed data \emph{observed at random} to impose MCAR missingness following the mechanism
-% \begin{equation*}
-% P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0)=.50.
-% \end{equation*}
-% Next, I created random missingness in $Y$ according to the following MAR missingness mechanism
-% \begin{equation*}
-% P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0|Y_{obs}, X_j),
-% \end{equation*}
-% by using a random draw from a binomial distribution of the same length as $Y$ and of size 1 with missingness probability equal to the inverse logit
-% \begin{equation*}
-% P(R=0)=\frac{e^{a}}{(1+e^{a})}.
-% \end{equation*}
-% where $a=(\mu_{X_j}-X_{ij})/\sqrt{\mathbb{E}[X_{ij} - \mu_{X_j}]}$ in order to create approximately 50\% right-tailed MAR missingness. I followed this procedure twice; once for each $X_j$.
-
-% The three resulting incomplete data sets were then imputed with package \texttt{mice} \citep[][v2.25]{mice} in \texttt{R} \citep[][v3.2.3]{R}, with Bayesian linear regression imputation (\texttt{mice.impute.norm}) as the imputation routine.
-The completed-data analyses were combined into a single inference following the rules described in \citet{vink2014pooling}.
 

### 2022-03-31 Formatted/Formatted.tex
Fix TODOs

@@
 Missingness mechanisms describe the relationship between missing entries and observed data values and are generally categorized into MCAR, MAR and MNAR mechanisms \citep[][see Table \ref{table:mech}]{rubi76}.
-[TODO: maybe remove mechanisms here?] Missingness patterns concern the location of missing entries across incomplete data \citep[][p.
+Missingness patterns concern the location of missing entries across incomplete data \citep[][p.
 8]{litt20}.
@@
 \begin{center}
-\caption{Missingness mechanisms. [TODO: add PMAR?]}
+\caption{Missingness mechanisms.}
 \label{table:mech}
@@
 the complete data model used to estimate the comparative truth).
-The choice of performance measures might [dictate (TODO: re-read and find better word!)] the results of the simulation study.
+The choice of performance measures might inadvertently influence simulation results.
 
@@
 This mechanism should be used as a reference condition.
-Every imputation method should yield valid imputations under MCAR.
-We also know that, under MCAR, the distribution of imputed values should be equivalent to the distribution of observed values for a given variable, which may offer an indication of imputation model (mis)fit.
-[TODO: leg uit!]
+Every imputation method must be able to yield valid imputations under MCAR, both in terms of distributional characteristics as well as statistical inference.
 
@@
 It would, therefore, be wise to include MNAR missingness in the simulation study just in case.
-Alternatively, sensitivity analyses may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
+Alternatively, a sensitivity analysis may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
 
-Apart from the missingness mechanisms, the missingness pattern must be varied.
-[TODO: beter bruggetje?] Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+In addition to varying the missingness mechanisms, each simulated mechanism should be combined with different missingness patterns.
+Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
 Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.

### 2022-03-31 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Submitted version

--- 2022-03-31 Formatted/Formatted.tex
@@
 % } %%% semicolon and fullpoint added here for keyword style
-\noindent \hspace*{-4pc} Supporting Information for this article is available from \underline{gerkovink.com/evaluation}.\\
+\noindent \hspace*{-4pc} Supporting Information for this article is available from \underline{www.gerkovink.com/evaluation}.\\
 % \hspace*{-4pc} \underline{gerkovink.com/evaluation}. 
@@
 Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
-In this paper, we propose a move towards a standardized evaluation of imputation methods.
+In this paper, we propose a move towards a standardized evaluation of imputation methodology.
 To demonstrate the need for standardization, we highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
-This may lead to sub-optimal use of multiple imputation in practice.
+This may lead to sub-optimal use of  imputation in practice.
 Additionally, we suggest a course of action for simulating and evaluating missing data problems.
@@
 Data-generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
-There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as from official registers.
+There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as official registers.
 
@@
 Real-life data hardly ever follow a given theoretical distribution, so there is no guarantee that simulation results are generalizable.
-Moreover, data are often generated such that the problem that is being studied is most pronounced, e.g.
+Moreover, data are often generated such that the problem being studied is most pronounced, e.g.
 with consistently high correlations between groups of variables.
@@
 Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.
-Just to be clear: leaving the missingness as-is and inducing additional missing values to impute could not work here, because we wouldn't have a real and unbiased comparative truth.
+Just to be clear: leaving the missingness as-is and inducing additional missing values to impute is no option here, because we would not have a real and unbiased comparative truth.
 
@@
       & cases. In other words, the missing values are missing at random and the observed \\
-      & values are observed at random. \\
+      & values are observed at random. \\ \noalign{\smallskip}
 MAR   & Missing at random: The probability to be missing is the same within groups of \\
@@
       & the data inform the missingness, such that MCAR may be assumed within the \\ 
-      & observed groups. \\
+      & observed groups. \\ \noalign{\smallskip}
 MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
@@
 However, the missingness mechanisms that are induced in simulation studies do not necessarily match any true missingness generating mechanisms in incomplete empirical data, nor the mechanisms that are typically assumed when imputing such data.
-On the one hand, there is MCAR missingness, which may be unlikely as the true missingness generating mechanism, but would yield insightful evaluations in a simulation setting.
+On the one hand, there is MCAR missingness, which is unlikely to be the true missingness generating mechanism, but would yield insightful evaluations in a simulation setting.
 On the other hand, there is MNAR, which is arguably a quite reasonable source of many real-world missingness scenarios but is generally ignored except for those evaluations that are specifically targeted at non-ignorable applications.
@@
 The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random.
-In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][\S 2.7]{buur18}: the missingness does not depend on the incomplete variable.
+In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than imputation \citep[see e.g.][\S 2.7]{buur18}: the missingness does not depend on the incomplete variable.
 This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is pointless.
@@
 4. Apply methods        & Imputation method(s), analytic method(s) \\
-5. Evaluate imputations & Algorithmic convergence, distributional characteristics, imputation \\
-                        & model fit \\
+5. Evaluate imputations & Imputation-generating process, imputation model fit \\
 6. Evaluate performance & Performance measures, qualitative judgment \\
@@
 A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
-If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly} can be useful.
+If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
 It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
@@
       & the true data inference. Performance of a missing data method should at least be \\
-      & acceptable for most missing data problems. \\ % & \\
+      & acceptable for most missing data problems. \\ \noalign{\smallskip} % & \\
 25\%  & This is a fair amount of missingness and will, depending on the observed data \\
@@
       & at least to our experience in social sciences and official statistics, 25\% \\
-      & univariate missingness can easily be considered as a realistic missingness percentage. \\ % & \\
+      & univariate missingness can easily be considered as a realistic missingness percentage. \\ \noalign{\smallskip} % & \\
 50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
@@
 It depends on the specifics of each simulation study which performance measures would be most suitable.
-In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
+In general, we recommend to evaluate at least the following points:
 
@@
 Ideally, this should be supplemented by an online repository with all of the data and code required to reproduce the simulation results.
-To aid simulators in reporting and move towards standardization in evaluation, we provide a draft version for reporting guidelines in Appendix A.1 (also available from \underline{gerkovink.com/evaluation}).
+To aid simulators in reporting and move towards standardization in evaluation, we provide a draft version for reporting guidelines in Appendix A.1 (also available from \underline{www.gerkovink.com/evaluation}).
 We invite the readers of this paper to contribute to its development.
@@
 \begin{table}[ht]
-\caption{Checklist for reporting on imputation methodology evaluations}
+\caption{Checklist for reporting on imputation methodology evaluations.}
 \label{table:check}

### 2022-03-31 index.Rmd

---
title: "Towards a standardized evaluation of imputation routines"
author: "Hanne Oberman and Gerko Vink"
date: "Department of Methodology & Statistics @ Utrecht University"
output: html_document
---

<style type="text/css">

body{ /* Normal  */
  font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 18px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 18px;
}
h2 { /* Header 2 */
  font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 18px;
}
code.r{ /* Code block */
  font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
  font-size: 14px;
}
</style>

---

This website accompanies the manuscript *Towards a standardized evaluation of imputation routines*. In this manuscript, we aim to elicit critical thinking about simulations with incomplete data and we propose the use of a set of evaluations and descriptions that would move towards a standardized evaluation of imputation routines. We summarize this set in a convenient checklist.

The [full publication archive can be found here](https://github.com/gerkovink/evaluation). Further links are:

- [The manuscript](LINK TO MANUSCRIPT)
- [The checklist](LINK TO CHECKLIST)

All the best,

[Hanne](https://hanneoberman.github.io) and [Gerko](https://www.gerkovink.com)

---

### 2022-03-31 index.Rmd
Add link to manuscript and checklist

@@
 
-- [The manuscript](LINK TO MANUSCRIPT)
-- [The checklist](LINK TO CHECKLIST)
+- [The manuscript](https://github.com/gerkovink/evaluation/blob/main/Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.pdf)
+- [The checklist](https://github.com/gerkovink/evaluation/blob/main/Checklist/Checklist.pdf)
 

### 2022-07-15 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Add reviewer comments

@@
 Additionally, we suggest a course of action for simulating and evaluating missing data problems.
+[TODO: explain how the standardisation leads to more neutrality, but that this is just a set of elements to consider/report, not a complete 'cookbook'.]
 \end{abstract}
@@
 
+[TODO: Narrowing the scope of the manuscript to the evaluation of methodology for statistical inference, not prediction or causal inference (although inferential validity encompasses predictive validity and may thus be generalized?).
+Emphasizing that the ideal evaluation of imputation methodology depends on the aim of the specific study (i.e.
+how the evaluated methods are used in practice).]
+
 Imputation is a state-of-the-art technique for drawing valid conclusions from incomplete data.
@@
 In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods.
-However, no golden standard has been established to evaluate imputation routines and, as a result, the validity and aim of the simulation setups may differ tremendously from one developer to another.
+However, no gold standard has been established to evaluate imputation routines and, as a result, the validity and aim of the simulation setups may differ tremendously from one developer to another.
 Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced.
@@
 But for sake of progress, it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties.
+[TODO: explain how the standardisation leads to more neutrality, but that this is just a set of elements to consider/report, not a complete 'cookbook'.]
 
@@
 \section{Why some evaluations should not be trusted}
+
+[TODO: Split up the problems with the simulation setup/evaluation and the problems with the interpretation/extrapolation from these simulation studies.
+See e.g.
+\citet{gree17} on cognitive problem of generelization versus methodological problems.]
 
@@
 Still, a lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
+[TODO: add nuance because this doesn't always hold, see Tims review!]
 
@@
 \begin{center}
-\caption{Missingness mechanisms.}
+\caption{Missingness mechanisms. [TODO: add refs \citet{seam13, meal15, dore18} for clarification and alternatives outlined in \citet{moha21, scho21, more18}]}
 \label{table:mech}
@@
 When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
-The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
+[TODO: change this to being able to embed the analysis procedure and the imputation model within a Bayesian model; the nonresponse mdel may be misspecified while the imputation procedure is congenial.]The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
 This makes its assessment an infeasible endeavor in many simulation studies.
@@
 591]{neym34}.
+[TODO: add nuance about coverage rate > 95\% irt CIW and efficiency.]
 
@@
 As a general rule, it would be wise to include one or more multivariate estimands.
-It is a realistic expectation for imputation methods to preserve both marginal and conditional distributions with respect to the comparative truth.
+[TODO: make this part about the study aim and refer to \citet{pete14} about clear and unambiguous estimand, and to focus on this estimand.] It is a realistic expectation for imputation methods to preserve both marginal and conditional distributions with respect to the comparative truth.
 Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).
@@
 \begin{center}
-\caption{Suggested univariate missingness proportions.}
+\caption{Suggested univariate missingness proportions.
+[TODO: change this to 'more than one missingness proportion, depending on the context' and/or add references for the given proportions (e.g.
+about  that 30\% is a good range for MI).]}
 \label{table:prop}
@@
 Then, by extension, one could assume that the imputation model might be able to produce good imputations for the missing part of the data too.
+[TODO: add that this is implemented in amelia.]
 
@@
 When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
+[TODO: add nuance here because valid inference, not plausible imputed values should be the aim.]
 
@@
 
+% "Note that Neyman's original description of confidence intervals defined the property of randomisation validity as exactly 100(1_alpha) of intervals containing theta.44-46 Confidence validity is the property that the true percentage is at least 100(1_alpha).
+This latter definition is less well known than the former, with the result that overcoverage and undercoverage are sometimes regarded as similarly bad.47 Of course, randomisation validity would usually be preferred over confidence validity because it implies shorter intervals, but this is not always the case.
+There are examples of procedures that return both shorter intervals and higher coverage.45, 46" ADEMP paper
+
 (iii) The width of the confidence or credible interval may convey statistical efficiency, which should be considered to compare imputation methods.
@@
 However, if a study is aimed at obtaining predictions or classification and inferential validity is not of interest, the RMSE of the predictions may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
+[TODO: make this an optional bullet point for prediction only: out of sample mean squared prediction error can be compared between methods.]
 

### 2022-07-21 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Copy paste sections

@@
 
+Standardizing the evaluation of imputation methodology requires simulators to consider a couple of key aspects in their simulation workflows.
+Table \ref{table:steps} outlines some suggested steps to adopt.
+Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach.
+Our suggestions are the accumulated result of scientific literature, research experience, and discussion.
+For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}.
+We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{Data generation problems}
+\subsection{Study scope}
+
+[Problem: parameters unclear, extrapolation beyond scope of simulations, misinterpretation \citep[see "investigator bias"][]{gree17}.]
+
+Before setting up a simulation study, the simulator should clearly define the scope of their evaluations.
+A simulation study aimed at imputation method comparison will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
+The choice of imputation method(s) under evaluation is naturally intertwined with the simulation aims and setup.
+
+One specific simulation study aim (that we have thus far ignored) is to evaluate the predictive performance of pairs of imputation methods and estimation methods in empirical, incomplete data.
+These kinds of simulation studies do not start out from a complete data set in which missingness is induced by the simulator.
+Rather, the evaluations use the observed values of an outcome variable (or target) as the estimand, which is estimated from the incomplete predictor (or feature) space by first imputing the missingness and then applying a prediction method.
+Pairs of imputation and prediction methods with high predictive accuracy in one or more benchmark data sets are deemed as `good' \citep{liu21}.
+Note that such a simulation design does not have a `ground truth', so only the comparative performance of methods may be established.
+We do not recommend this approach if the inferential validity of the imputations is of interest.
+
+Some general choices in the simulation setup to consider at this stage are the simulation design and the number of simulation repetitions.
+The simulation design should ideally be fully factorial, i.e.
+varying each simulation condition against all other conditions.
+There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions).
+If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
+Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation, which makes for fairer comparisons.
+The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
+as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Data generation}
 
@@
 
-\subsection{Missingness generation problems}
+% 1a. choose DGM (if model-based: choose n_obs, n_var, relations between var, var types, etc; if design-based: choose n_obs, missing data handling strategy)
+% 1b. decide if sampling variance is needed
+% 1c. draw complete sample(s)
+
+The simulator should choose their data-generating mechanism(s) in line with the study scope.
+Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived.
+A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
+
+Estimand(s) or other simulation targets should be defined in the context of the study aim and data-generating mechanism.
+As a general rule, it would be wise to include one or more multivariate estimands.
+[TODO: make this part about the study aim and refer to \citet{pete14} about clear and unambiguous estimand, and to focus on this estimand.] It is a realistic expectation for imputation methods to preserve both marginal and conditional distributions with respect to the comparative truth.
+Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).
+
+The comparative truth of the estimand(s) is determined by the choice of data-generating mechanism and whether sampling variance is included in the simulation setup, see Table \ref{table:dgm_truth}.
+If sampling variance cannot be omitted from the simulation scheme, multiple samples should be drawn from the data-generating mechanism (i.e., one sample per simulation repetition, a standard simulation setup).
+If sampling variance is not of interest, a single complete data set can be obtained from the data-generating mechanism (i.e., one sample for all simulation repetitions).
+This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
+We highly value the additional flexibility that such a simulation setup offers.
+Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against.
+Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
+When deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
+76-77]{rubi87} do not apply.
+Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
+
+\begin{table}[tb]
+\begin{center}
+\caption{Source of the comparative truth in simulation studies.}
+\label{table:dgm_truth}
+\begin{tabular}{lll}
+\hline
+               & With sampling variance      & Without sampling variance \\
+               & (multiple samples drawn)    & (one sample drawn) \\
+\hline  
+Model-based simulation   & data-generating model         & single sample \\
+Design-based simulation  & sufficiently large data set   & single sample \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Missingness generation}
+
+[TODO: change this to 'be clear which mechanisms are of interest, how they are generated, if they are suitably generated']
 
@@
       & observed groups. \\ \noalign{\smallskip}
+v-MAR & Conditional on the fully observed variables missingness occurs at random 
+      & \citep[][p. 1026]{moha21}.\\
 MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
@@
 
-\subsection{Evaluation problems}
+% 2a. choose missingness mechanism(s): at least consider MCAR, vary MAR types
+% 2b. choose missing data pattern(s): introduce non-monotone multivariate missingness
+% 2c. choose missingness proportion(s): at least consider 10, 25 and 50\%
+
+Missingness should be induced according to several sets of missing data conditions.
+We encourage simulators to consider different missingness mechanisms (including missingness types), and different missingness patterns (including missingness proportions).
+Although not every missingness mechanism is realistically assumed in practice, they can all offer valuable insights as simulation condition.
+
+First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions).
+This mechanism should be used as a reference condition.
+Every imputation method must be able to yield valid imputations under MCAR, both in terms of distributional characteristics as well as statistical inference.
+
+Next, missing data should be induced conform a model that is dependent on the observed data (i.e.
+a MAR mechanism).
+Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully.
+A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
+If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
+It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
+The effects of different types of MAR mechanisms are described in \citet{scho18}.
+
+Third, a non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice.
+Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time.
+It would, therefore, be wise to include MNAR missingness in the simulation study just in case.
+Alternatively, a sensitivity analysis may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
+
+In addition to varying the missingness mechanisms, each simulated mechanism should be combined with different missingness patterns.
+Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
+The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
+If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
+Alternatively, the validity of assumptions becomes increasingly important when the missingness increases.
+Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
+We, therefore, propose to evaluate imputation methodology under several missingness proportions to amulate a realistic range in severity of the missingness problem.
+Depending on how the missingness mechanism interacts with the simulated data, higher missingness proportions may yield biased or invalid completed data inferences.
+The missingness proportions should be considered carefully.
+
+% \begin{table}[htb]
+% \begin{center}
+% \caption{Suggested univariate missingness proportions.
+[TODO: change this to 'more than one missingness proportion, depending on the context' and/or add references for the given proportions (e.g.
+about  that 30\% is a good range for MI).]}
+% \label{table:prop}
+% \begin{tabular}{lll}
+% \hline
+%       & Reasoning for suggested missingness proportion \\
+% \hline  
+% 10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
+%       & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
+%       & the true data inference. Performance of a missing data method should at least be \\
+%       & acceptable for most missing data problems. \\ \noalign{\smallskip} % & \\
+% 25\%  & This is a fair amount of missingness and will, depending on the observed data \\
+%       & information, have a noticeable influence on the completed data inference. When \\
+%       & compared to the condition with 10\% missingness, the inference obtained under \\
+%       & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
+%       & should increase), but estimates should still be properly covered and the statistical \\
+%       & properties of the missing data method should be sound. In practice, at least to our \\
+%       & at least to our experience in social sciences and official statistics, 25\% \\
+%       & univariate missingness can easily be considered as a realistic missingness percentage. \\ \noalign{\smallskip} % & \\
+% 50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
+%       & severely. Depending on how the missingness mechanism interacts with the simulated \\
+%       & data, some imputation techniques may yield estimates that are under-covered such that \\
+%       & the completed data inference should not be deemed valid anymore. If a method yields \\
+%       & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
+%       & properties of the imputation methodology are sound. \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Imputation methods}
+
+% 3a. Define missing data methods: at least consider complete case analysis as benchmark
+% 3b. Choose imputation model parameters: appropriate n_imp and n_it (if applicable), maybe vary predictor matrices?
+% 3c. Impute the data
+% 3d. Obtain estimates
+
+The imputation methods under evaluation should be applied to the generated incomplete data, after which an analysis model may be fitted to estimate the estimand(s).
+In addition, the simulator should always perform complete case analysis.
+We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance.
+Complete case analysis may, therefore, serve as a benchmark method against which imputation performance should be evaluated.
+Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Evaluation}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SOLUTIONS
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Suggested course of action}
-
-Standardizing the evaluation of imputation methodology requires simulators to consider a couple of key aspects in their simulation workflows.
-Table \ref{table:steps} outlines some suggested steps to adopt.
-Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach.
-Our suggestions are the accumulated result of scientific literature, research experience, and discussion.
-For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}.
-We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology.
-
-\begin{table}[tb]
-\begin{center}
-\caption{Steps to consider in imputation simulation studies.}
-\label{table:steps}
-\begin{tabular}{lll}
-\hline
-                        & Simulation aspects under consideration \\
-\hline  
-1. Set scope            & Aim(s), missing data method(s) to evaluate, simulation design \\
-2. Obtain truth         & Data-generating mechanism(s), estimand(s), sampling variance \\
-3. Induce missingness   & Missingness mechanism(s), missing data pattern(s) \\
-4. Apply methods        & Imputation method(s), analytic method(s) \\
-5. Evaluate imputations & Imputation-generating process, imputation model fit \\
-6. Evaluate performance & Performance measures, qualitative judgment \\
-7. Report               & Text, visualization, checklist, simulation script \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Set scope}
-
-Before setting up a simulation study, the simulator should clearly define the scope of their evaluations.
-A simulation study aimed at imputation method comparison will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
-The choice of imputation method(s) under evaluation is naturally intertwined with the simulation aims and setup.
-
-One specific simulation study aim (that we have thus far ignored) is to evaluate the predictive performance of pairs of imputation methods and estimation methods in empirical, incomplete data.
-These kinds of simulation studies do not start out from a complete data set in which missingness is induced by the simulator.
-Rather, the evaluations use the observed values of an outcome variable (or target) as the estimand, which is estimated from the incomplete predictor (or feature) space by first imputing the missingness and then applying a prediction method.
-Pairs of imputation and prediction methods with high predictive accuracy in one or more benchmark data sets are deemed as `good' \citep{liu21}.
-Note that such a simulation design does not have a `ground truth', so only the comparative performance of methods may be established.
-We do not recommend this approach if the inferential validity of the imputations is of interest.
-
-Some general choices in the simulation setup to consider at this stage are the simulation design and the number of simulation repetitions.
-The simulation design should ideally be fully factorial, i.e.
-varying each simulation condition against all other conditions.
-There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions).
-If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
-Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation, which makes for fairer comparisons.
-The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
-as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Obtain truth}
-
-% 1a. choose DGM (if model-based: choose n_obs, n_var, relations between var, var types, etc; if design-based: choose n_obs, missing data handling strategy)
-% 1b. decide if sampling variance is needed
-% 1c. draw complete sample(s)
-
-The simulator should choose their data-generating mechanism(s) in line with the study scope.
-Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived.
-A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
-
-Estimand(s) or other simulation targets should be defined in the context of the study aim and data-generating mechanism.
-As a general rule, it would be wise to include one or more multivariate estimands.
-[TODO: make this part about the study aim and refer to \citet{pete14} about clear and unambiguous estimand, and to focus on this estimand.] It is a realistic expectation for imputation methods to preserve both marginal and conditional distributions with respect to the comparative truth.
-Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).
-
-The comparative truth of the estimand(s) is determined by the choice of data-generating mechanism and whether sampling variance is included in the simulation setup, see Table \ref{table:dgm_truth}.
-If sampling variance cannot be omitted from the simulation scheme, multiple samples should be drawn from the data-generating mechanism (i.e., one sample per simulation repetition, a standard simulation setup).
-If sampling variance is not of interest, a single complete data set can be obtained from the data-generating mechanism (i.e., one sample for all simulation repetitions).
-This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
-We highly value the additional flexibility that such a simulation setup offers.
-Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against.
-Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
-When deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
-76-77]{rubi87} do not apply.
-Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
-
-\begin{table}[tb]
-\begin{center}
-\caption{Source of the comparative truth in simulation studies.}
-\label{table:dgm_truth}
-\begin{tabular}{lll}
-\hline
-               & With sampling variance      & Without sampling variance \\
-               & (multiple samples drawn)    & (one sample drawn) \\
-\hline  
-Model-based simulation   & data-generating model         & single sample \\
-Design-based simulation  & sufficiently large data set   & single sample \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Induce missingness}
-
-% 2a. choose missingness mechanism(s): at least consider MCAR, vary MAR types
-% 2b. choose missing data pattern(s): introduce non-monotone multivariate missingness
-% 2c. choose missingness proportion(s): at least consider 10, 25 and 50\%
-
-Missingness should be induced according to several sets of missing data conditions.
-We encourage simulators to consider different missingness mechanisms (including missingness types), and different missingness patterns (including missingness proportions).
-Although not every missingness mechanism is realistically assumed in practice, they can all offer valuable insights as simulation condition.
-
-First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions).
-This mechanism should be used as a reference condition.
-Every imputation method must be able to yield valid imputations under MCAR, both in terms of distributional characteristics as well as statistical inference.
-
-Next, missing data should be induced conform a model that is dependent on the observed data (i.e.
-a MAR mechanism).
-Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully.
-A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
-If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
-It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
-The effects of different types of MAR mechanisms are described in \citet{scho18}.
-
-Third, a non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice.
-Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time.
-It would, therefore, be wise to include MNAR missingness in the simulation study just in case.
-Alternatively, a sensitivity analysis may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
-
-In addition to varying the missingness mechanisms, each simulated mechanism should be combined with different missingness patterns.
-Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
-Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
-The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
-If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
-Alternatively, the validity of assumptions becomes increasingly important when the missingness increases.
-Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
-We, therefore, propose to evaluate imputation methodology under 10\%, 25\% and 50\% univariate missingness (see Table \ref{table:prop}).
-
-\begin{table}[htb]
-\begin{center}
-\caption{Suggested univariate missingness proportions.
-[TODO: change this to 'more than one missingness proportion, depending on the context' and/or add references for the given proportions (e.g.
-about  that 30\% is a good range for MI).]}
-\label{table:prop}
-\begin{tabular}{lll}
-\hline
-      & Reasoning for suggested missingness proportion \\
-\hline  
-10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
-      & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
-      & the true data inference. Performance of a missing data method should at least be \\
-      & acceptable for most missing data problems. \\ \noalign{\smallskip} % & \\
-25\%  & This is a fair amount of missingness and will, depending on the observed data \\
-      & information, have a noticeable influence on the completed data inference. When \\
-      & compared to the condition with 10\% missingness, the inference obtained under \\
-      & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
-      & should increase), but estimates should still be properly covered and the statistical \\
-      & properties of the missing data method should be sound. In practice, at least to our \\
-      & at least to our experience in social sciences and official statistics, 25\% \\
-      & univariate missingness can easily be considered as a realistic missingness percentage. \\ \noalign{\smallskip} % & \\
-50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
-      & severely. Depending on how the missingness mechanism interacts with the simulated \\
-      & data, some imputation techniques may yield estimates that are under-covered such that \\
-      & the completed data inference should not be deemed valid anymore. If a method yields \\
-      & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
-      & properties of the imputation methodology are sound. \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Apply methods}
-
-% 3a. Define missing data methods: at least consider complete case analysis as benchmark
-% 3b. Choose imputation model parameters: appropriate n_imp and n_it (if applicable), maybe vary predictor matrices?
-% 3c. Impute the data
-% 3d. Obtain estimates
-
-The imputation methods under evaluation should be applied to the generated incomplete data, after which an analysis model may be fitted to estimate the estimand(s).
-In addition, the simulator should always perform complete case analysis.
-We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance.
-Complete case analysis may, therefore, serve as a benchmark method against which imputation performance should be evaluated.
-Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Evaluate imputations}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Evaluate performance}
 
@@
 
-\subsection{Report}
+\subsection{Reporting}
 
@@
 We invite the readers of this paper to contribute to its development.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SOLUTIONS
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[tb]
+\begin{center}
+\caption{Steps to consider in imputation simulation studies.}
+\label{table:steps}
+\begin{tabular}{lll}
+\hline
+                        & Simulation aspects under consideration \\
+\hline  
+1. Set scope            & Aim(s), missing data method(s) to evaluate, simulation design \\
+2. Obtain truth         & Data-generating mechanism(s), estimand(s), sampling variance \\
+3. Induce missingness   & Missingness mechanism(s), missing data pattern(s) \\
+4. Apply methods        & Imputation method(s), analytic method(s) \\
+5. Evaluate imputations & Imputation-generating process, imputation model fit \\
+6. Evaluate performance & Performance measures, qualitative judgment \\
+7. Report               & Text, visualization, checklist, simulation script \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 

### 2022-07-22 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
add refs review morris

@@
 
+[TODO: add that if the goal is statistical inference, the uncertainty due to missingness should be quantified/captured by the missing data method.
+Refer to \citet{sper20} for difference in context of prediction.
+And paraphrase "the quality of a technique can only be evaluated with respect to the aims of the problem/s it's intended to solve in practice".
+we focus on comparative sim studies]
+
 The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
@@
 Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced.
-Although these promising new methods seem to yield even sharper imputations than now-standard (semi-)parametric imputation methods, the comparison may not be fair due to the simulation setup.
+Although these promising new methods seem to yield even sharper imputations than now-standard imputation methods, the comparison may not be fair due to the simulation setup.
 
@@
 These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals.
-Second, to provide imputation methodologists with a suggested course of action when simulating missing data problems.
+Second, to provide imputation methodologists with a suggested course of action when using simulation studies to evaluate imputation techniques for missing data problems.
 This suggested approach should identify common ground but is in no way intended as an absolute solution.
@@
 We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology.
+
+[TODO: \citep{pawe22} pawel pre-reg simulation protocols? per-protocol results]
 
@@
 
+[TODO: split up into design, execution and reporting? \citep{pawe22}]
+
 \begin{table}[tb]

### 2022-07-26 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Add TODOs for revision

--- 2022-07-21 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
@@
 \usepackage[authoryear]{natbib}
+\usepackage{amsfonts}
 % \setlength{\bibsep}{2pt}
@@
 Additionally, we suggest a course of action for simulating and evaluating missing data problems.
-[TODO: explain how the standardisation leads to more neutrality, but that this is just a set of elements to consider/report, not a complete 'cookbook'.]
+[TODO: Narrowing the scope of the manuscript to the evaluation of methodology for statistical inference, not prediction or causal inference (although inferential validity encompasses predictive validity and may thus be generalized?).
+Emphasizing that the ideal evaluation of imputation methodology depends on the aim of the specific study (i.e.
+how the evaluated methods are used in practice).
+Split up the problems with the simulation setup/evaluation and the problems with the interpretation/extrapolation from these simulation studies.
+See e.g.
+\citet{gree17} on cognitive problem of generelization versus methodological problems.
+Explain how the standardisation leads to more neutrality, but that this is just a set of elements to consider/report, not a complete 'cookbook'.
+Add that we focus on  simulation studies evaluating imputation methodology for statistical inference.]
 \end{abstract}
@@
 
-[TODO: Narrowing the scope of the manuscript to the evaluation of methodology for statistical inference, not prediction or causal inference (although inferential validity encompasses predictive validity and may thus be generalized?).
-Emphasizing that the ideal evaluation of imputation methodology depends on the aim of the specific study (i.e.
-how the evaluated methods are used in practice).]
-
 Imputation is a state-of-the-art technique for drawing valid conclusions from incomplete data.
@@
 After all, the principle of imputation is very intuitive.
-
 The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.
-A popular variant is multiple imputation \citep{rubi76}, whereby each missing value is imputed multiple times.
-The completed data that are thus obtained can be analyzed by standard techniques and, in the case of multiple imputation, the analysis results can be combined into a single inference using Rubin's rules \citep[][p.
-76]{rubi87}.
-In contrast to ad hoc methods for dealing with missing values (e.g.
-list-wise deletion, mean imputation, regression imputation, last observation carried forward, indicator method), multiple imputation properly takes into account the sources of uncertainty that are related to the missingness problem.
-
+The completed data that are thus obtained can be analyzed by standard techniques, which conveniently splits up the process of solving the missing data problem and the complete-data problem.
 The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
@@
 
+In contrast to ad hoc methods for dealing with missing values (e.g.
+list-wise deletion), imputation allows analysts to account for the sources of uncertainty that are related to the missingness problem.
+With the popular variant multiple imputation \citep{rubi76}, each missing value is replaced by a plausible data value several times, resulting in multiple completed data sets.
+The estimated parameters of interest for each of these completed sets can be combined into a single inference using Rubin's rules \citep[][p.
+76]{rubi87}.
+[TODO: add that if the goal is statistical inference, the uncertainty due to missingness should be quantified/captured/propagated by the missing data method.
+Refer to \citet{sper20} for the context of prediction, where multiple imputation is not always necessary.] [TODO: paraphrase ``the quality of a technique can only be evaluated with respect to the aims of the problem/s it's intended to solve in practice".]
+
 When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies.
-In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods.
-However, no gold standard has been established to evaluate imputation routines and, as a result, the validity and aim of the simulation setups may differ tremendously from one developer to another.
-Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced.
-Although these promising new methods seem to yield even sharper imputations than now-standard (semi-)parametric imputation methods, the comparison may not be fair due to the simulation setup.
+Such studies are typically composed of an analysis pipeline in which data are generated, made incomplete, and imputed repeatedly under different simulation conditions, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods.
+However, there is no consensus on how to design, execute and report simulation studies aimed to evaluate imputation methodology.
+Without a `gold standard' for the evaluation of imputation routines, the validity and comparability of simulation setups may differ tremendously from one developer to another.
+This brings forth a chain of potential problems in the objective assessment of imputation method performance within and across studies, which may lead to sub-optimal use of imputation in practice.
+
+% Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced.
+Although these promising new methods seem to yield even sharper imputations than now-standard imputation methods, the comparison may not be fair due to the simulation setup.
 
@@
 We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology.
+[TODO: Add advice to pre-register simulation protocols, which allows for fair comparisons of per-protocol results \citep{pawe22}.]
 
@@
 
-[TODO: Split up the problems with the simulation setup/evaluation and the problems with the interpretation/extrapolation from these simulation studies.
-See e.g.
-\citet{gree17} on cognitive problem of generelization versus methodological problems.]
+% [TODO: split up into design, execution and reporting? \citep{pawe22}]
 
-As of today, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
-Typically, the developer of a new imputation routine does some tests by simulations, but these tests differ across developers.
-This brings forth a chain of potential problems in the objective assessment of imputation method performance, within and across studies, which may lead to sub-optimal use of imputation in practice.
-To demonstrate the broad impact of these problems, we subdivide the problems into the following three distinct categories: problems with data generation, problems with missingness generation, and problems with performance evaluation.
+The ideal evaluation of imputation methodology depends on the aim of the specific study (i.e.
+how the evaluated methods are used in practice).
+A simulation study developed for the comparison of imputation methods will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
+The suitability of the evaluations should be assessed with the imputation methods' purpose in mind, since the simulation aims and setup are naturally intertwined with the choice of imputation method(s) under evaluation.
+
+We limit the scope of this paper to comparative simulation studies in the context of statistical inference.
+We thereby exclude the evaluation of imputation methodology for prediction and causal inference.
+Simulation study designs aimed at comparing different methods' predictive performance in incomplete data typically do not start out from a complete data set in which missingness is induced by the simulator \citep{liu21}.
+Rather, the methods are evaluated against the observed values of a certain outcome variable (or target) as the estimand.
+After imputing the missingness in the incomplete predictor (or feature) space, a prediction method is applied to estimate the outcome.
+Pairs of imputation and prediction methods with high predictive accuracy in one or more benchmark data sets are deemed as `good'.
+Note that such a simulation design does not have a `ground truth', so only the comparative performance of methods may be established.
+We do not recommend this approach if the inferential validity of the imputations is of interest.
+For an overview of missingness in the prediction context see e.g.
+\citet{sper20}.
+Missing data within a causal inference framework has been described by e.g.
+\cite{more18} and \cite{moha21}.
+
+Within the scope of comparative simulation studies for the evaluation of imputation methodology for statistical inference, we observe several problems.
+To demonstrate the broad impact of these problems, we subdivide the problems into the following four distinct categories: problems with simulation design, problems with data generation, problems with missingness generation, and problems with performance evaluation.
 We further detail the impact each of these problems may have on the validity of evaluations.
@@
 
-\subsection{Study scope}
+\subsection{Simulation design}
 
-[Problem: parameters unclear, extrapolation beyond scope of simulations, misinterpretation \citep[see "investigator bias"][]{gree17}.]
+% Before setting up a simulation study, the simulator should clearly define the scope of their evaluations. 
+Estimand(s) or other simulation targets should be defined in the context of the study aim. [TODO: refer to \citet{pete14} about the need for a clear and unambiguous estimand.]
 
-Before setting up a simulation study, the simulator should clearly define the scope of their evaluations.
-A simulation study aimed at imputation method comparison will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
-The choice of imputation method(s) under evaluation is naturally intertwined with the simulation aims and setup.
+[TODO: describe problems with unclear simulation parameters, extrapolation beyond scope of simulations, QRPs \citep{pawe22} and misinterpretation \citep[see "investigator bias"][]{gree17}.]
 
-One specific simulation study aim (that we have thus far ignored) is to evaluate the predictive performance of pairs of imputation methods and estimation methods in empirical, incomplete data.
-These kinds of simulation studies do not start out from a complete data set in which missingness is induced by the simulator.
-Rather, the evaluations use the observed values of an outcome variable (or target) as the estimand, which is estimated from the incomplete predictor (or feature) space by first imputing the missingness and then applying a prediction method.
-Pairs of imputation and prediction methods with high predictive accuracy in one or more benchmark data sets are deemed as `good' \citep{liu21}.
-Note that such a simulation design does not have a `ground truth', so only the comparative performance of methods may be established.
-We do not recommend this approach if the inferential validity of the imputations is of interest.
+[TODO: rewrite sampling variance paragraphs and combine with ``If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
+Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation, which makes for fairer comparisons."]
 
-Some general choices in the simulation setup to consider at this stage are the simulation design and the number of simulation repetitions.
-The simulation design should ideally be fully factorial, i.e.
-varying each simulation condition against all other conditions.
-There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions).
-If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
-Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation, which makes for fairer comparisons.
-The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
-as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Data generation}
-
-To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established.
-Those who perform simulation studies are in the luxury position to establish the truth beforehand by choosing a data-generating mechanism.
-Data-generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
-There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as official registers.
-
-The problem with model-based data-generating mechanisms is that a method's performance on simulated data may not translate to empirical data.
-Real-life data hardly ever follow a given theoretical distribution, so there is no guarantee that simulation results are generalizable.
-Moreover, data are often generated such that the problem being studied is most pronounced, e.g.
-with consistently high correlations between groups of variables.
-This results in simulated data that contain such valuable information structures that, no matter what type of missingness would subsequently be induced, the observed parts of the data will still hold much (if not all) of the information about the missing part.
-Unsurprisingly, the performance of any imputation method will then be evaluated as good.
-Another threat to the generalizability of model-based simulations is the use of a single model for both data generation and imputation.
-If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied.
-Other (unfair) comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s), and the coherence between variables.
-In contrast to design-based studies, such characteristics are not always explicit simulation conditions, which may give a false sense of objectivity.
-
-An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be very challenging.
-Most real-world data contains at least some missing entries, for which the true underlying missing data model is--by definition--unknown.
-Therefore, the simulator needs to deal with missingness in \textit{some} way before incomplete empirical data can serve as comparative truth in the simulations.
-It may seem like an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples.
-However, there may be inherent differences between cases with and cases without any missing values, due to the unknown missing data model.
-Only sampling complete cases from the data set may thus result in samples that fail to capture all relevant real-world conditions, which in turn refutes the main reason for using design-based simulation.
-Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset to draw samples from.
-Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.
-Just to be clear: leaving the missingness as-is and inducing additional missing values to impute is no option here, because we would not have a real and unbiased comparative truth.
-
-After deciding upon a data-generating mechanism, another influential--but often overlooked--decision awaits.
-Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations and only use the missing data generation procedure as the source of Monte Carlo variation.
+[TODO: add nuance because this doesn't always hold, see Tims review!] Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations and only use the missing data generation procedure as the source of Monte Carlo variation.
 After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
@@
 Still, a lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
-[TODO: add nuance because this doesn't always hold, see Tims review!]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% 1a. choose DGM (if model-based: choose n_obs, n_var, relations between var, var types, etc; if design-based: choose n_obs, missing data handling strategy)
-% 1b. decide if sampling variance is needed
-% 1c. draw complete sample(s)
-
-The simulator should choose their data-generating mechanism(s) in line with the study scope.
-Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived.
-A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
-
-Estimand(s) or other simulation targets should be defined in the context of the study aim and data-generating mechanism.
-As a general rule, it would be wise to include one or more multivariate estimands.
-[TODO: make this part about the study aim and refer to \citet{pete14} about clear and unambiguous estimand, and to focus on this estimand.] It is a realistic expectation for imputation methods to preserve both marginal and conditional distributions with respect to the comparative truth.
-Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).
-
 The comparative truth of the estimand(s) is determined by the choice of data-generating mechanism and whether sampling variance is included in the simulation setup, see Table \ref{table:dgm_truth}.
@@
 
+The simulation design should ideally be fully factorial, i.e.
+varying each simulation condition against all other conditions.
+There may, for example, be interactions between the source and amount of missingness in the incomplete data and the efficacy of imputation methods (i.e., the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions).
+The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
+as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Data generation}
+
+% [TODO: split up into data and missingness part of the DGM, with choices between data sources, missingness mechanisms and proportions.
+And divide problems between simulation and interpretation, see Reviewer 2: "whether method A (e.g.
+imputation) outperforms method B (e.g.
+listwise deletion) depends on the estimand, purpose of the study, used distributions (in the DGP, imputation model, possibly also analysis model), and simulation setup (e.g.
+dependency structure) - if a human misinterprets findings in a comparison of method A and B, then this is a cognitive problem and not the problem on how missingness was generated in the DGP."]
+
+[TODO; describe problems with data generation and interpretation.]
+
+- If the source of the data is model-based, the correlations in the data may be so high that there is no effective missingness due to the predictive power in the other variables, and the data source may be unfair if the data generating model is equal to one of the imputation models under evaluation.
+
+- With a design-based simulation, the data source is unlikely to have no missingness and the choice of initial missing data method may propagate into the evaluations.
+
+To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established.
+Those who perform simulation studies are in the luxury position to establish the truth beforehand by choosing a data-generating mechanism.
+Data-generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
+There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as official registers.
+The simulator should choose their data-generating mechanism(s) in line with the study scope.
+Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived.
+A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
+
+The problem with model-based data-generating mechanisms is that a method's performance on simulated data may not translate to empirical data.
+Real-life data hardly ever follow a given theoretical distribution, so there is no guarantee that simulation results are generalizable.
+Moreover, data are often generated such that the problem being studied is most pronounced, e.g.
+with consistently high correlations between groups of variables.
+This results in simulated data that contain such valuable information structures that, no matter what type of missingness would subsequently be induced, the observed parts of the data will still hold much (if not all) of the information about the missing part.
+Unsurprisingly, the performance of any imputation method will then be evaluated as good.
+Another threat to the generalizability of model-based simulations is the use of a single model for both data generation and imputation.
+If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied.
+Other (unfair) comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s), and the coherence between variables.
+In contrast to design-based studies, such characteristics are not always explicit simulation conditions, which may give a false sense of objectivity.
+
+An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be very challenging.
+Most real-world data contains at least some missing entries, for which the true underlying missing data model is--by definition--unknown.
+Therefore, the simulator needs to deal with missingness in \textit{some} way before incomplete empirical data can serve as comparative truth in the simulations.
+It may seem like an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples.
+However, there may be inherent differences between cases with and cases without any missing values, due to the unknown missing data model.
+Only sampling complete cases from the data set may thus result in samples that fail to capture all relevant real-world conditions, which in turn refutes the main reason for using design-based simulation.
+Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset to draw samples from.
+Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.
+Just to be clear: leaving the missingness as-is and inducing additional missing values to impute is no option here, because we would not have a real and unbiased comparative truth.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% 1a. choose DGM (if model-based: choose n_obs, n_var, relations between var, var types, etc; if design-based: choose n_obs, missing data handling strategy)
+% 1b. decide if sampling variance is needed
+% 1c. draw complete sample(s)
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-[TODO: change this to 'be clear which mechanisms are of interest, how they are generated, if they are suitably generated']
+[TODO: paraphrase ``be clear which mechanisms are of interest, how they are generated, if they are suitably generated".]
 
-Evaluating imputation methodology requires a missing data problem to be solved.
+- Missingness mechanisms can be spurious (spurious MAR when correlations are too low, or when missingness proportion is too low).
+In reporting, the functional form or type of missingness is often omitted.
+As advice, MCAR can be a very informative mechnism for reference and MNAR may be useful for practice.
+
+- Missingness proportions can be misinterpreted.
+
+- Missingness pattern may have unexpected consequences (e.g., univariately generated multivariate missingness). In interpretation, watch out for generalizations from unintended special cases.
+
+% Evaluating imputation methodology requires a missing data problem to be solved.
 After establishing a comparative truth from a data-generating mechanism, some form of missingness therefore has to be induced.
-Often, however, reports of simulation studies remain vague about the actual missingness conditions under investigation and, even worse, some authors only report something like:
+
+Often reports of simulation studies remain vague about the missingness conditions under investigation and, even worse, some authors only report something like:
 \begin{quote}
@@
 \end{quote}
-This should be considered unacceptable as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness mechanisms and missingness patterns.
-Missingness mechanisms describe the relationship between missing entries and observed data values and are generally categorized into MCAR, MAR and MNAR mechanisms \citep[][see Table \ref{table:mech}]{rubi76}.
-Missingness patterns concern the location of missing entries across incomplete data \citep[][p.
+This should be considered unacceptable, as claims about the validity of the imputation inference heavily depend on the simulated missingness conditions, such as missingness mechanisms and missingness patterns.
+Missingness mechanisms describe the relationship between missing entries and observed data values whereas missingness patterns concern the location of missing entries across incomplete data \citep[][p.
 8]{litt20}.
@@
 
-\begin{table}[tb]
-\begin{center}
-\caption{Missingness mechanisms. [TODO: add refs \citet{seam13, meal15, dore18} for clarification and alternatives outlined in \citet{moha21, scho21, more18}]}
-\label{table:mech}
-\begin{tabular}{lll}
-\hline
-      & Interpretation and consequences of the mechanism \\
-\hline  
-MCAR  & Missing completely at random: The probability to be missing is the same for all \\
-      & cases. In other words, the missing values are missing at random and the observed \\
-      & values are observed at random. \\ \noalign{\smallskip}
-MAR   & Missing at random: The probability to be missing is the same within groups of \\
-      & cases defined by the observed data only. The essence is that observed relations in \\
-      & the data inform the missingness, such that MCAR may be assumed within the \\ 
-      & observed groups. \\ \noalign{\smallskip}
-v-MAR & Conditional on the fully observed variables missingness occurs at random 
-      & \citep[][p. 1026]{moha21}.\\
-MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
-      & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
-      & inferred from the observed data. This is considered non-ignorable missingness \\
-      & \citep[see e.g.][]{rubi76}. \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
+% \begin{table}[tb]
+% \begin{center}
+% \caption{Missingness mechanisms. [TODO: add refs \citet{seam13, meal15, dore18} for clarification and alternatives outlined in \citet{moha21, scho21, more18}]}
+% \label{table:mech}
+% \begin{tabular}{lll}
+% \hline
+%       & Interpretation and consequences of the mechanism \\
+% \hline  
+% MCAR  & Missing completely at random: The probability to be missing is the same for all \\
+%       & cases. In other words, the missing values are missing at random and the observed \\
+%       & values are observed at random. \\ \noalign{\smallskip}
+% MAR   & Missing at random: The probability to be missing is the same within groups of \\
+%       & cases defined by the observed data only. The essence is that observed relations in \\
+%       & the data inform the missingness, such that MCAR may be assumed within the \\ 
+%       & observed groups. \\ \noalign{\smallskip}
+% v-MAR & Conditional on the fully observed variables missingness occurs at random 
+%       & \citep[][p. 1026]{moha21}.\\
+% MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
+%       & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
+%       & inferred from the observed data. This is considered non-ignorable missingness \\
+%       & \citep[see e.g.][]{rubi76}. \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \end{table}
 
@@
 The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).
-Moreover, the value of the actual missingness proportion may be diffusing too.
+
+% Moreover, the value of the actual missingness proportion may be diffusing too.
 Some studies use only 10\% missingness whereas other studies push the limits to additionally investigate the performance under missingness proportions of at least 50\%.
@@
 
-\subsection{Imputation methods}
+% \subsection{Imputation methods}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Evaluation}
+
 The imputation methods under evaluation should be applied to the generated incomplete data, after which an analysis model may be fitted to estimate the estimand(s).
@@
 Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Evaluation}
 
@@
 While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.
+Under many circumstances, imputation methods may be realistically expected to preserve both marginal and conditional distributions with respect to the comparative truth.
+Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).
 
@@
 
-\subsection{Reporting}
+% \subsection{Reporting}
+
+% 
+% \begin{table}[tb]
+% \begin{center}
+% \caption{Steps to consider in imputation simulation studies.}
+% \label{table:steps}
+% \begin{tabular}{lll}
+% \hline
+%                         & Simulation aspects under consideration \\
+% \hline  
+% 1. Set scope            & Aim(s), missing data method(s) to evaluate, simulation design \\
+% 2. Obtain truth         & Data-generating mechanism(s), estimand(s), sampling variance \\
+% 3. Induce missingness   & Missingness mechanism(s), missing data pattern(s) \\
+% 4. Apply methods        & Imputation method(s), analytic method(s) \\
+% 5. Evaluate imputations & Imputation-generating process, imputation model fit \\
+% 6. Evaluate performance & Performance measures, qualitative judgment \\
+% 7. Report               & Text, visualization, checklist, simulation script \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% DISCUSSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Suggested course of action}
 
 After the careful consideration and execution of all simulation aspects, the evaluations should be properly reported.
+The inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
+This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
+
 We encourage simulators to document their choices and to be explicit in their descriptions.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SOLUTIONS
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}[tb]
-\begin{center}
-\caption{Steps to consider in imputation simulation studies.}
-\label{table:steps}
-\begin{tabular}{lll}
-\hline
-                        & Simulation aspects under consideration \\
-\hline  
-1. Set scope            & Aim(s), missing data method(s) to evaluate, simulation design \\
-2. Obtain truth         & Data-generating mechanism(s), estimand(s), sampling variance \\
-3. Induce missingness   & Missingness mechanism(s), missing data pattern(s) \\
-4. Apply methods        & Imputation method(s), analytic method(s) \\
-5. Evaluate imputations & Imputation-generating process, imputation model fit \\
-6. Evaluate performance & Performance measures, qualitative judgment \\
-7. Report               & Text, visualization, checklist, simulation script \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% DISCUSSION
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Discussion}
-
-This manuscript aims to elicit critical thinking about incomplete data simulation and to establish a common ground for the evaluation of imputation routines.
+Ultimately, this manuscript aims to elicit critical thinking about incomplete data simulation and to establish a common ground for the evaluation of imputation routines.
 Such a common ground would be the basis of a standardized evaluation.
@@
 2 & Comparative truth                                                                  \\ \cline{2-2}
-  & Data-generating mechanism (model-based or design-based)                            \\
+  & Data source (model-based or design-based)                                           \\
   & Sampling variance                                                                  \\

### 2022-07-26 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Revisions version shared with Gerko

@@
 
-[TODO; describe problems with data generation and interpretation.]
-
-- If the source of the data is model-based, the correlations in the data may be so high that there is no effective missingness due to the predictive power in the other variables, and the data source may be unfair if the data generating model is equal to one of the imputation models under evaluation.
-
-- With a design-based simulation, the data source is unlikely to have no missingness and the choice of initial missing data method may propagate into the evaluations.
+[TODO; describe problems with data generation and interpretation: If the source of the data is model-based, the correlations in the data may be so high that there is no effective missingness due to the predictive power in the other variables, and the data source may be unfair if the data generating model is equal to one of the imputation models under evaluation.
+With a design-based simulation, the data source is unlikely to have no missingness and the choice of initial missing data method may propagate into the evaluations.]
 
@@
 
-[TODO: paraphrase ``be clear which mechanisms are of interest, how they are generated, if they are suitably generated".]
-
-- Missingness mechanisms can be spurious (spurious MAR when correlations are too low, or when missingness proportion is too low).
+[TODO: Make clear in this section that missingness mechanisms can be spurious (spurious MAR when correlations are too low, or when missingness proportion is too low).
 In reporting, the functional form or type of missingness is often omitted.
 As advice, MCAR can be a very informative mechnism for reference and MNAR may be useful for practice.
-
-- Missingness proportions can be misinterpreted.
-
-- Missingness pattern may have unexpected consequences (e.g., univariately generated multivariate missingness). In interpretation, watch out for generalizations from unintended special cases.
+Missingness proportions can be misinterpreted.
+Missingness pattern may have unexpected consequences (e.g., univariately generated multivariate missingness).
+In interpretation, watch out for generalizations from unintended special cases.]
 
@@
 The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).
+Simulators should therefore be explicit in their description of the missing data generating model fitting the aim of their simulations.
 
@@
 
-Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables.
+% Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables.
 It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways.
-Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to sub-optimal advice in practice.
-For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that list-wise deletion outperforms their imputation method(s).
-Such a conclusion would, unfortunately, only translate to data that adhere to the same special case while biasing inferences in other cases \citep[][\S 2.7]{buur18}.
-Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e.
-a pattern with uniformly increasing missingness proportions along variables;][]{litt20}.
+
+Extrapolating simulation results from a specific missing data generating model to more intricate (empirical) missingness could lead to sub-optimal advice in practice.
+For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may inadvertedly induce missingness according to the special case descibed by \citet[][p.
+\S 2.7]{buur18}.
+Under this specific missingness model, list-wise deletion may outperform any imputation method.
+Such a conclusion would, unfortunately, only translate to data that adhere to the same special case, while biasing inferences in other cases.
+% Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e.
+a pattern with uniformly increasinmissingness proportions along variables;][]{litt20}.
 If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations.
-A final example is multivariate missingness which is generated using step-wise univariate missingness induction.
+Another example of potential side-effects of missingness generation may occur when multivariate missingness is generated using step-wise univariate missingness induction.
 The resulting incomplete data may then not have the desired statistical properties \citep[e.g.
 skewness;][]{ampute}.
-This complicates any definitive conclusions about imputation method performance in relation to the data-generating mechanism.
+This complicates any definitive conclusions about imputation method performance in relation to the data generation parameters.
 
-Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions).
+% Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions).
 However, the missingness mechanisms that are induced in simulation studies do not necessarily match any true missingness generating mechanisms in incomplete empirical data, nor the mechanisms that are typically assumed when imputing such data.
@@
 
-MCAR is a necessary simulation condition for evaluating the performance of imputation procedures, yet often omitted.
-Under MCAR, the statistical properties of the observed data given the missing data are known.
+Missingness should be induced according to several sets of missing data conditions.
+We encourage simulators to consider different missingness patterns and mechanisms.
+Missingness mechanisms were first defined in Rubin's seminal work \citep{rubi76}.
+For more recent considerations see e.g.
+\citet{seam13, meal15, dore18, more18, moha21, scho21}.
+Although not every missingness mechanism is realistically assumed in practice, they can all offer valuable insights as simulation condition.
+A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.
+
+[TODO: add that focusing on a single mechanism (even if the mechanism suits the study aim) is wasteful?] Truly random missingness across all data entries (`missing completely at random' or MCAR under Rubin's definition) may be considered a necessary simulation condition for the evaluation of imputation procedures, since the statistical properties of the observed data given the missing data are known.
 Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation.
 If an imputation method is not able to solve the problem (i.e.
-yield valid inference) under MCAR, the statistical properties of the procedure are not sound.
-Sadly, the straightforward case of MCAR is often neglected from simulation studies and focus is drawn to the evaluation of MAR mechanisms only.
-Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in our view may be far too simplistic.
+yield valid inference) under MCAR, the statistical properties of the procedure are not universally sound.
+Sadly, the straightforward case of MCAR is often neglected from simulation studies, while it offers an informative reference condition in many simulation setups.
+
+[TODO: maybe add definitions of missingness mechanisms back in???] A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
+If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
+It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
+The effects of different types of MAR mechanisms are described in \citet{scho18}.
+Inducing a MAR mechanism presents the simulator with a choice, namely the type or functional form of the missingness model.
+Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}.
+This may influence the performance of some (but not necessarily all) imputation routines.
+For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
+It would be a shame to overlook such results due to the focus on a single functional form of the missingness-generating model.
 
@@
 
-Inducing a MAR mechanism presents the simulator with another choice, namely the type or functional form of the missingness model.
-Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}.
-This may influence the performance of some (but not necessarily all) imputation routines.
-For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques.
-It would be a shame to overlook such results due to the focus on a single MAR mechanism.
-
-Finally, although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
-As stated, this mechanism is not usually included in evaluations.
+A non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice.
+Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time.
+After all, for every MNAR mechanism there is a MAR mechanism with equal fit and one cannot definitively verify that empirical missingness is random \citep{molenberghs2008every}.
+Therefore, it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
+It may be wise to include MNAR missingness in the simulation study just in case.
+Alternatively, a sensitivity analysis may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
 
@@
 
-Missingness should be induced according to several sets of missing data conditions.
-We encourage simulators to consider different missingness mechanisms (including missingness types), and different missingness patterns (including missingness proportions).
-Although not every missingness mechanism is realistically assumed in practice, they can all offer valuable insights as simulation condition.
-
-First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions).
+% First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions).
 This mechanism should be used as a reference condition.
 Every imputation method must be able to yield valid imputations under MCAR, both in terms of distributional characteristics as well as statistical inference.
-
-Next, missing data should be induced conform a model that is dependent on the observed data (i.e.
+% 
+% Next, missing data should be induced conform a model that is dependent on the observed data (i.e.
 a MAR mechanism).
 Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully.
-A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
-If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
-It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
-The effects of different types of MAR mechanisms are described in \citet{scho18}.
-
-Third, a non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice.
-Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time.
-It would, therefore, be wise to include MNAR missingness in the simulation study just in case.
-Alternatively, a sensitivity analysis may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
 
@@
 Depending on how the missingness mechanism interacts with the simulated data, higher missingness proportions may yield biased or invalid completed data inferences.
-The missingness proportions should be considered carefully.
+The missingness proportions should therefore be considered carefully.
 
+[TODO: paraphrase ``be clear which mechanisms are of interest, how they are generated, if they are suitably generated" as conclusion to this section.]
 % \begin{table}[htb]
@@
 
-\subsection{Evaluation}
+\subsection{Performance evaluation}
 
-The imputation methods under evaluation should be applied to the generated incomplete data, after which an analysis model may be fitted to estimate the estimand(s).
-In addition, the simulator should always perform complete case analysis.
-We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance.
-Complete case analysis may, therefore, serve as a benchmark method against which imputation performance should be evaluated.
-Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.
+[TODO: make this section less stringent and focus more on the suitability of the performancxe evaluation in the context of the study aim.]
 
@@
 There are, however, some overarching issues with assessing imputation method performance.
-In the first place, diagnostic evaluation of the generated imputations is often left out of simulation study results.
-Problems with the imputation-generating process (e.g., an iterative imputation algorithm) may then be overlooked.
-In the second place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model (i.e.
-the complete data model used to estimate the comparative truth).
-The choice of performance measures might inadvertently influence simulation results.
+In the first place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model after imputation (e.g.
+focusing on one performance measure over another, may yield different conclusions about imputation efficacy).
+In the second place, diagnostic evaluation of the generated imputations is often left out of simulation study results.
+Identifying problems with the imputation-generating process (e.g., an iterative imputation algorithm) may offer explanations for under-performance of imputation methods.
+Such evaluations, however, are typically omitted and valuable insights into the imputation method(s) may then be overlooked.
+%(i.e.
+the complete-data model used to estimate the comparative truth).
+% The choice of performance measures might inadvertently influence simulation results.
 
-Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations.
-As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated.
-Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor failproof.
-As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method.
-Problems with producing stable imputations is not an exclusive quality iterative imputation algorithms.
-Any method may run into failures of the imputation-generating process and subsequently lack results (e.g.
-due to over-parameterization errors).
+At minimum, imputation method performance should be quantified using appropriate measures.
+It depends on the specifics of each simulation study which performance measures would be most suitable.
+We recommend to evaluate at least the following points:
 
-Omitting further evaluation of the imputation-generating process may yield sub-optimal imputations.
-The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models.
-When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
-[TODO: change this to being able to embed the analysis procedure and the imputation model within a Bayesian model; the nonresponse mdel may be misspecified while the imputation procedure is congenial.]The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
-This makes its assessment an infeasible endeavor in many simulation studies.
-Moreover, the performance of imputation procedures on distributional properties is often ignored too.
-Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
-For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large.
-While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.
-Under many circumstances, imputation methods may be realistically expected to preserve both marginal and conditional distributions with respect to the comparative truth.
-Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).
+(i) The estimates should preferably be unbiased.
+Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results.
+In most cases, unbiased estimates may be expected under a MCAR missingness mechanism.
 
-The second source of problems with evaluating simulation outcomes is the choice of performance measures.
-Performance measures should express how well the obtained estimates approximate the comparative truth, but not every metric is suited for the evaluation of imputation methodology.
+(ii) The estimates should have a proper coverage rate.
+Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
+Under-coverage (when estimation is too liberal and intervals are too narrow) indicates that the procedure is not confidence valid and may lead to invalid inference.
+Over-coverage (when estimation is too conservative and intervals are too wide) tells us that efficiency could still be gained.
+
+% "Note that Neyman's original description of confidence intervals defined the property of randomisation validity as exactly 100(1_alpha) of intervals containing theta.44-46 Confidence validity is the property that the true percentage is at least 100(1_alpha).
+This latter definition is less well known than the former, with the result that overcoverage and undercoverage are sometimes regarded as similarly bad.47 Of course, randomisation validity would usually be preferred over confidence validity because it implies shorter intervals, but this is not always the case.
+There are examples of procedures that return both shorter intervals and higher coverage.45, 46" ADEMP paper
+
+(iii) The width of the confidence or credible interval may convey statistical efficiency, which should be considered to compare imputation methods.
+Wider intervals are associated with more uncertainty whereas a more narrow interval that is still properly covered indicates a sharper inference.
+However, inference from a wider interval that is properly covered is to be considered more valid than a more narrow interval that is not properly covered anymore.
+
+(iv) Predictive accuracy may be quantified using the root mean squared error (RMSE).
+We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values \citep[][\S 2.5]{buur18}.
+However, if a study is aimed at obtaining predictions or classification and inferential validity is not of interest, the RMSE of the predictions may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
+[TODO: make this an optional bullet point for prediction only: out of sample mean squared prediction error can be compared between methods.]
+
+[TODO: intrgrate this paragraph into the above!] Performance measures should express how well the obtained estimates approximate the comparative truth, but not every metric is suited for the evaluation of imputation methodology.
 For example, using measures of predictive accuracy or the RMSE (root mean squared error) as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{buur18}.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%
+Next to imputation method performance, the imputation-generating process may be evaluated.
+Omitting further evaluation of the imputation-generating process may yield sub-optimal imputations.
+The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models.
+When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
+[TODO: change this to being able to embed the analysis procedure and the imputation model within a Bayesian model; the nonresponse mdel may be misspecified while the imputation procedure is congenial.] The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
+This makes its assessment an infeasible endeavor in many simulation studies.
+Moreover, the performance of imputation procedures on distributional properties is often ignored too.
+Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
+For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large.
+While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.
+Under many circumstances, imputation methods may be realistically expected to preserve both marginal and conditional distributions with respect to the comparative truth.
+Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).
 
-% 4a. Evaluate imputations: check failed methods, algorithmic convergence (if applicable)
-% 4b. Check imputation model fit: anomalies in distributional characteristics, plausibility of imputed values (if requested), PPC
+[TODO: add segway to this paragraph.] Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations.
+As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated.
+Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor failproof.
+As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method.
+Problems with producing stable imputations is not an exclusive quality iterative imputation algorithms.
+Any method may run into failures of the imputation-generating process and subsequently lack results (e.g.
+due to over-parameterization errors).
 
-Every imputation workflow should contain an evaluation of the obtained imputations.
+[TODO: add segway to this paragraph and/or intrgrate into text above!] Every imputation workflow should contain an evaluation of the obtained imputations.
 Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend to consider the following aspects:
@@
 
+[TODO: add segway to reference method paragraph and add that performance on the complete data ofefrs an upper limit for performance.] In addition, the simulator should always perform complete case analysis.
+We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance.
+Complete case analysis may, therefore, serve as a benchmark method against which imputation performance should be evaluated.
+Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% 4a. Evaluate imputations: check failed methods, algorithmic convergence (if applicable)
+% 4b. Check imputation model fit: anomalies in distributional characteristics, plausibility of imputed values (if requested), PPC
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % fraction of missing information (= model parameter), variance inflation
-
-Imputation method performance should be quantified using appropriate measures.
-It depends on the specifics of each simulation study which performance measures would be most suitable.
-In general, we recommend to evaluate at least the following points:
-
-(i) The estimates should preferably be unbiased.
-Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results.
-Unbiased estimates under MCAR are a minimum requirement for any imputation method.
-
-(ii) The estimates should have a proper coverage rate.
-Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
-Under-coverage (when estimation is too liberal and intervals are too narrow) indicates that the procedure is not confidence valid and may lead to invalid inference.
-Over-coverage (when estimation is too conservative and intervals are too wide) tells us that efficiency could still be gained.
-
-% "Note that Neyman's original description of confidence intervals defined the property of randomisation validity as exactly 100(1_alpha) of intervals containing theta.44-46 Confidence validity is the property that the true percentage is at least 100(1_alpha).
-This latter definition is less well known than the former, with the result that overcoverage and undercoverage are sometimes regarded as similarly bad.47 Of course, randomisation validity would usually be preferred over confidence validity because it implies shorter intervals, but this is not always the case.
-There are examples of procedures that return both shorter intervals and higher coverage.45, 46" ADEMP paper
-
-(iii) The width of the confidence or credible interval may convey statistical efficiency, which should be considered to compare imputation methods.
-Wider intervals are associated with more uncertainty whereas a more narrow interval that is still properly covered indicates a sharper inference.
-However, inference from a wider interval that is properly covered is to be considered more valid than a more narrow interval that is not properly covered anymore.
-
-We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values \citep[][\S 2.5]{buur18}.
-However, if a study is aimed at obtaining predictions or classification and inferential validity is not of interest, the RMSE of the predictions may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
-[TODO: make this an optional bullet point for prediction only: out of sample mean squared prediction error can be compared between methods.]
 

### 2022-07-27 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
After feedack Gerko on revisions

@@
 The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.
-The completed data that are thus obtained can be analyzed by standard techniques, which conveniently splits up the process of solving the missing data problem and the complete-data problem.
+The completed data that are thus obtained can be analyzed by standard techniques, which conveniently separates the missing data problem from the analysis.
+When prediction is the goal, a single imputation may be sufficient \citet{sper20}.
+For inferential analyses, multiple imputation \citep{rubi76} has proven to be a valuable technique for obtaining valid inferences on incomplete datasets.
+In the case of multiple imputation each missing datum is imputed more than once, leading to multiple completed data sets.
+The resulting multiple inferences can be combined into a single inference using Rubin's rules \citep[][p.
+76]{rubi87}.
 The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
@@
 
-In contrast to ad hoc methods for dealing with missing values (e.g.
+% In contrast to ad hoc methods for dealing with missing values (e.g.
 list-wise deletion), imputation allows analysts to account for the sources of uncertainty that are related to the missingness problem.
@@
 When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies.
-Such studies are typically composed of an analysis pipeline in which data are generated, made incomplete, and imputed repeatedly under different simulation conditions, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods.
+Such studies are typically composed of an analysis pipeline in which data are generated, made incomplete, and imputed repeatedly under different simulation conditions.
+A set of evaluation criteria is then postulated to evaluate the performance of one or more missing data methods.
 However, there is no consensus on how to design, execute and report simulation studies aimed to evaluate imputation methodology.
@@
 But for sake of progress, it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties.
-[TODO: explain how the standardisation leads to more neutrality, but that this is just a set of elements to consider/report, not a complete 'cookbook'.]
 
-Standardizing the evaluation of imputation methodology requires simulators to consider a couple of key aspects in their simulation workflows.
+Standardizing the evaluation of imputation methodology requires simulators to consider common aspects in their simulation workflows.
 Table \ref{table:steps} outlines some suggested steps to adopt.
@@
 
-Within the scope of comparative simulation studies for the evaluation of imputation methodology for statistical inference, we observe several problems.
-To demonstrate the broad impact of these problems, we subdivide the problems into the following four distinct categories: problems with simulation design, problems with data generation, problems with missingness generation, and problems with performance evaluation.
+Within the scope of comparative simulation studies aimed at statistical inference, we observe several problems.
+To demonstrate the broad impact of these problems, we recognize the following four distinct categories: problems with simulation design, problems with data generation, problems with missingness generation, and problems with performance evaluation.
 We further detail the impact each of these problems may have on the validity of evaluations.
@@
 Another example of potential side-effects of missingness generation may occur when multivariate missingness is generated using step-wise univariate missingness induction.
-The resulting incomplete data may then not have the desired statistical properties \citep[e.g.
-skewness;][]{ampute}.
+The resulting incomplete data may then not have the desired statistical properties \citep{ampute}.
 This complicates any definitive conclusions about imputation method performance in relation to the data generation parameters.
@@
 For more recent considerations see e.g.
-\citet{seam13, meal15, dore18, more18, moha21, scho21}.
+\citet{seam13, meal15, dore18, more18, scho18, litt20, moha21, scho21}.
 Although not every missingness mechanism is realistically assumed in practice, they can all offer valuable insights as simulation condition.
@@
 
-[TODO: add segway to this paragraph and/or intrgrate into text above!] Every imputation workflow should contain an evaluation of the obtained imputations.
+[TODO: integrate this into text above!] Every imputation workflow should contain an evaluation of the obtained imputations.
 Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend to consider the following aspects:
@@
 When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
-[TODO: add nuance here because valid inference, not plausible imputed values should be the aim.]
+[TODO: add nuance here because valid inference, not plausible imputed values, should be the aim (although plausibility is nice).]
 
@@
 
-Ultimately, this manuscript aims to elicit critical thinking about incomplete data simulation and to establish a common ground for the evaluation of imputation routines.
+We aim to elicit critical thinking about incomplete data simulation and to establish a common ground for the evaluation of imputation routines.
 Such a common ground would be the basis of a standardized evaluation.
@@
 Ultimately, it would be desirable to evaluate every imputation routine against the same standardized set in order to quantify the statistical properties across imputation routines.
-If properly executed, this would allow for careful matching of imputation methodologies to new missing data problems.
+If properly executed, such evaluations would allow for careful matching of imputation methodologies to new missing data problems.
 

### 2022-07-27 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Fix referencing issues, add check boxes to check list

@@
 Standardizing the evaluation of imputation methodology requires simulators to consider common aspects in their simulation workflows.
-Table \ref{table:steps} outlines some suggested steps to adopt.
+Table \ref{table:check} outlines some suggested steps to adopt.
 Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach.
@@
 1 & Simulation scope                                                                   \\ \cline{2-2}
-  & Aim                                                                                \\
-  & Design (incl. pseudo-code or flow diagram)                                         \\
-  & Number of simulation repetitions                                                   \\ \hline
+$\square$  & Aim                                                                                \\
+$\square$  & Design (incl. pseudo-code or flow diagram)                                         \\
+$\square$  & Number of simulation repetitions                                                   \\ \hline
 2 & Comparative truth                                                                  \\ \cline{2-2}
-  & Data source (model-based or design-based)                                           \\
-  & Sampling variance                                                                  \\
+$\square$  & Data source (model-based or design-based)                                           \\
+$\square$  & Sampling variance                                                                  \\
 %  & Data characteristics (number of observations, number of variables, variable types, \\
 %  & coherence between variables incl. data structures e.g. clustering)                 \\
-  & Data characteristics (incl. multivariate relations and structures e.g. clustering) \\
-  & Estimand                                                                           \\ \hline 
+$\square$  & Data characteristics (incl. multivariate relations and structures e.g. clustering) \\
+$\square$  & Estimand                                                                           \\ \hline 
 3 & Induced missingness                                                                \\ \cline{2-2}
-  & Missingness mechanism (incl. type or functional form of the missing data   model)  \\
-  & Missingness pattern (incl. missingness proportion)                                 \\ \hline 
+$\square$  & Missingness mechanism (incl. type or functional form of the missing data   model)  \\
+$\square$  & Missingness pattern (incl. missingness proportion)                                 \\ \hline 
 4 & Applied methods                                                                    \\ \cline{2-2}
-  & Imputation methods (incl. parameters e.g. the number of imputations)               \\
-  & Analytic methods (incl. calculation of standard errors e.g. pooling   rules)       \\
-  & Reference method (e.g. complete case analysis)                                    \\ \hline 
+$\square$  & Imputation methods (incl. parameters e.g. the number of imputations)               \\
+$\square$  & Analytic methods (incl. calculation of standard errors e.g. pooling   rules)       \\
+$\square$  & Reference method (e.g. complete case analysis)                                    \\ \hline 
 5 & Imputation evaluation                                                              \\ \cline{2-2}
-  & Imputation-generating process (e.g. algorithmic non-convergence)                   \\
-  & Imputation model fit (e.g. posterior predictive checks)                            \\
-  & Distributional characteristics (e.g. plausibility of imputed values)               \\ \hline 
+$\square$  & Imputation-generating process (e.g. algorithmic non-convergence)                   \\
+$\square$  & Imputation model fit (e.g. posterior predictive checks)                            \\
+$\square$  & Distributional characteristics (e.g. plausibility of imputed values)               \\ \hline 
 6 & Performance evaluation                                                             \\ \cline{2-2}
-  & Statistical properties (e.g. confidence validity)                                  \\
-  & Comparative performance (e.g. predictive accuracy)                                 \\ \hline
+$\square$  & Statistical properties (e.g. confidence validity)                                  \\
+$\square$  & Comparative performance (e.g. predictive accuracy)                                 \\ \hline
 \end{tabular}

### 2022-08-01 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Revisions (cont.)

@@
 The completed data that are thus obtained can be analyzed by standard techniques, which conveniently separates the missing data problem from the analysis.
-When prediction is the goal, a single imputation may be sufficient \citet{sper20}.
+
 For inferential analyses, multiple imputation \citep{rubi76} has proven to be a valuable technique for obtaining valid inferences on incomplete datasets.
@@
 76]{rubi87}.
+When prediction is the goal of the analysis effort, a single imputation may be sufficient \citep{nijm20}.
 The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
@@
 For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}.
-We recommend adhering to their proposed ADEMP structure (aims, data-generating mechanisms, estimands, methods, performance measures) for planning and reporting simulation studies, but highlight and add some specific recommendations for the evaluation of imputation methodology.
-[TODO: Add advice to pre-register simulation protocols, which allows for fair comparisons of per-protocol results \citep{pawe22}.]
+We recommend adhering to their proposed structure for planning and reporting simulation studies and \citet{pawe22}'s advice to pre-register simulation protocols.
+We henceforth highlight and add some recommendations for the evaluation of imputation methodology.
+% [TODO: Add advice to pre-register simulation protocols, which allows for fair comparisons of per-protocol results \citep{pawe22}.]
 
@@
 
-The ideal evaluation of imputation methodology depends on the aim of the specific study (i.e.
-how the evaluated methods are used in practice).
-A simulation study developed for the comparison of imputation methods will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
-The suitability of the evaluations should be assessed with the imputation methods' purpose in mind, since the simulation aims and setup are naturally intertwined with the choice of imputation method(s) under evaluation.
+The ideal evaluation of imputation methodology is study specific.
+A simulation study developed for the comparison of several imputation methods will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
+How the evaluated methods are used in practice thereby governs the simulation procedure.
+Since the simulation aims and setup are intertwined with the choice of imputation methods under evaluation, the suitability of the evaluations should be assessed with the imputation methods' purpose in mind.
 
 We limit the scope of this paper to comparative simulation studies in the context of statistical inference.
-We thereby exclude the evaluation of imputation methodology for prediction and causal inference.
-Simulation study designs aimed at comparing different methods' predictive performance in incomplete data typically do not start out from a complete data set in which missingness is induced by the simulator \citep{liu21}.
-Rather, the methods are evaluated against the observed values of a certain outcome variable (or target) as the estimand.
-After imputing the missingness in the incomplete predictor (or feature) space, a prediction method is applied to estimate the outcome.
-Pairs of imputation and prediction methods with high predictive accuracy in one or more benchmark data sets are deemed as `good'.
-Note that such a simulation design does not have a `ground truth', so only the comparative performance of methods may be established.
+With that, we exclude the evaluation of imputation methodology for prediction or causal inference.
+The type of simulation study under consideration in this paper requires a form of `comparative truth' to assess the inferential validity of imputation methods.
+This is in contrast to simulation studies aimed at comparing the predictive performance of imputation and prediction method pairs.
+Such a simulation design typically does not start out from a `ground truth' (e.g., a complete data set in which missingness is subsequently induced by the simulator).
+Rather, the imputation and prediction method pairs are evaluated on their ability to yield high predictive accuracy in one or more incomplete benchmark data sets \citep{liu21}.
+In these designs, only the comparative performance of methods may be established.
 We do not recommend this approach if the inferential validity of the imputations is of interest.
 For an overview of missingness in the prediction context see e.g.
-\citet{sper20}.
+\citet{wood15} and \citet{sper20}.
 Missing data within a causal inference framework has been described by e.g.
@@
 
-% Before setting up a simulation study, the simulator should clearly define the scope of their evaluations. 
-Estimand(s) or other simulation targets should be defined in the context of the study aim. [TODO: refer to \citet{pete14} about the need for a clear and unambiguous estimand.]
+Before setting up a simulation study, the simulator should clearly define and describe the scope of their evaluations.
+If the simulation parameters are unclear, methodological problems with simulation study design may be obfuscated by the cognitive problem of generalization \citep{gree17}.
+For example, misinterpretation and (unintended) questionable research practices can prompt extrapolation beyond the scope of the simulations \citep{gree17, pawe22}.
+Imputation methods that seem to work well in simulations may then not be suitable in practice, when applied to incomplete empirical data.
 
-[TODO: describe problems with unclear simulation parameters, extrapolation beyond scope of simulations, QRPs \citep{pawe22} and misinterpretation \citep[see "investigator bias"][]{gree17}.]
+One potential reason for discrepancies between imputation method performance in simulations versus applications is the use of sampling variance in the simulation design.
+Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations and only use the missing data generation procedure as the source of Monte Carlo variation \citep{vink14}.
+If sampling variance is not of interest, a single complete data set can serve as `comparative truth' in all simulation repetitions (as opposed to a `standard' simulation setup with a different complete data set per repetition).
+After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
+Removing sampling variance from performance evaluation is computationally convenient and may potentially sharpen comparisons between imputation methods, clarifying inconclusive results on their ability to correct for missingness.
+However, when deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
+76-77]{rubi87} do not apply and alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
+This opens the door towards human error in the implementation and interpretation of such a simulation study.
+On the one hand, we highly value the additional flexibility that this design offers.
+Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against.
+Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
+On the other hand, we want to warn simulators against potential mishaps, e.g.
+when calculating and comparing standard errors.
 
-[TODO: rewrite sampling variance paragraphs and combine with ``If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
-Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation, which makes for fairer comparisons."]
+% [TODO: describe problems with unclear simulation parameters, extrapolation beyond scope of simulations, QRPs \citep{pawe22} and misinterpretation \citep[see "investigator bias"][]{gree17}.]
 
-[TODO: add nuance because this doesn't always hold, see Tims review!] Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations and only use the missing data generation procedure as the source of Monte Carlo variation.
-After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
-Therefore, taking sampling variation into account is neither necessary for obtaining information about a method's ability to handle the missing data problem, nor for objectively comparing methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}.
+% Therefore, taking sampling variation into account is neither necessary for obtaining information about a method's ability to handle the missing data problem, nor for objectively comparing methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}.
 The only required change to simulation setups is to remove sampling variance from performance evaluation.
 Still, a lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
-The comparative truth of the estimand(s) is determined by the choice of data-generating mechanism and whether sampling variance is included in the simulation setup, see Table \ref{table:dgm_truth}.
-If sampling variance cannot be omitted from the simulation scheme, multiple samples should be drawn from the data-generating mechanism (i.e., one sample per simulation repetition, a standard simulation setup).
-If sampling variance is not of interest, a single complete data set can be obtained from the data-generating mechanism (i.e., one sample for all simulation repetitions).
-This process is computationally convenient because only a single complete data set has to be considered during all of the simulations.
-We highly value the additional flexibility that such a simulation setup offers.
-Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against.
-Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
-When deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
-76-77]{rubi87} do not apply.
-Instead, alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
 
-\begin{table}[tb]
-\begin{center}
-\caption{Source of the comparative truth in simulation studies.}
-\label{table:dgm_truth}
-\begin{tabular}{lll}
-\hline
-               & With sampling variance      & Without sampling variance \\
-               & (multiple samples drawn)    & (one sample drawn) \\
-\hline  
-Model-based simulation   & data-generating model         & single sample \\
-Design-based simulation  & sufficiently large data set   & single sample \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
+% The comparative truth of the estimand(s) is determined by the choice of data-generating mechanism and whether sampling variance is included in the simulation setup, see Table \ref{table:dgm_truth}.
 
+% If sampling variance cannot be omitted from the simulation scheme, multiple samples should be drawn from the data-generating mechanism (i.e., one sample per simulation repetition, a standard simulation setup).
+
+% If sampling variance is not of interest, a single complete data set can be obtained from the data-generating mechanism (i.e., one sample for all simulation repetitions).
+
+% \begin{table}[tb]
+% \begin{center}
+% \caption{Source of the comparative truth in simulation studies.}
+% \label{table:dgm_truth}
+% \begin{tabular}{lll}
+% \hline
+%                & With sampling variance      & Without sampling variance \\
+%                & (multiple samples drawn)    & (one sample drawn) \\
+% \hline  
+% Model-based simulation   & data-generating model         & single sample \\
+% Design-based simulation  & sufficiently large data set   & single sample \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \end{table}
+%
+
+[TODO: fix segway!] For fair evaluations, the following simulation design choices may be considered.
+Estimands and other simulation targets should be clearly and unambiguously defined in the context of the study aim \citet{pete14}.
 The simulation design should ideally be fully factorial, i.e.
@@
 as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
+If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
+Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation, which makes for fairer comparisons.
 
@@
 
-[TODO: add segway to reference method paragraph and add that performance on the complete data ofefrs an upper limit for performance.] In addition, the simulator should always perform complete case analysis.
+[TODO: add segway to reference method paragraph and add that performance on the complete data offers an upper limit for performance.] In addition, the simulator should always perform complete case analysis.
 We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance.

### 2022-09-07 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
sync before meeting

@@
 Before setting up a simulation study, the simulator should clearly define and describe the scope of their evaluations.
-If the simulation parameters are unclear, methodological problems with simulation study design may be obfuscated by the cognitive problem of generalization \citep{gree17}.
+If the simulation parameters are unclear, any methodological problems with simulation study design may be obfuscated by the cognitive problem of generalization \citep{gree17}.
 For example, misinterpretation and (unintended) questionable research practices can prompt extrapolation beyond the scope of the simulations \citep{gree17, pawe22}.
-Imputation methods that seem to work well in simulations may then not be suitable in practice, when applied to incomplete empirical data.
+Imputation methods that seem to work well in simulations may then not be suitable in practice, when applied on incomplete empirical data.
 
@@
 Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations and only use the missing data generation procedure as the source of Monte Carlo variation \citep{vink14}.
-If sampling variance is not of interest, a single complete data set can serve as `comparative truth' in all simulation repetitions (as opposed to a `standard' simulation setup with a different complete data set per repetition).
+If sampling variance is not of interest, a single complete data set can serve as comparative truth in all simulation repetitions (as opposed to a `standard' simulation setup, in which data is drawn every repetition).
 After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
@@
 
-[TODO: fix segway!] For fair evaluations, the following simulation design choices may be considered.
+A well-designed simulation does not guarantee fair evaluations.
+Still, we propose to consider the following simulation design choices.
 Estimands and other simulation targets should be clearly and unambiguously defined in the context of the study aim \citet{pete14}.
+The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
+as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
 The simulation design should ideally be fully factorial, i.e.
 varying each simulation condition against all other conditions.
-There may, for example, be interactions between the source and amount of missingness in the incomplete data and the efficacy of imputation methods (i.e., the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions).
-The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
-as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
+There may, for example, be interactions between the source and amount of missingness in the incomplete data and the efficacy of imputation methods (i.e., the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions) [TODO: add ref dance of the mechanisms!].
 If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.

### 2022-09-15 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Resubmit changes

@@
 by the detailed manual created by the National Research Council \citep*{little2012prevention}.
-Although top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers and data scientists are embracing imputation techniques.
+Although top-down enforcement of valid ways to handle missing data is not yet very pronounced, an increasing amount of researchers and data scientists are embracing and developing imputation methodology.
 After all, the principle of imputation is very intuitive.
-The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.
+The idea behind imputation is to impute (fill in) missing values to obtain a valid estimate of what could have been.
 The completed data that are thus obtained can be analyzed by standard techniques, which conveniently separates the missing data problem from the analysis.
 
-For inferential analyses, multiple imputation \citep{rubi76} has proven to be a valuable technique for obtaining valid inferences on incomplete datasets.
-In the case of multiple imputation each missing datum is imputed more than once, leading to multiple completed data sets.
-The resulting multiple inferences can be combined into a single inference using Rubin's rules \citep[][p.
+For some data analyses a single imputation may suffice, while for other data analyses multiple imputations are needed.
+For example, in the case of inferential analyses, such multiple imputations \citep{rubi76} have proven to be a valuable technique for obtaining valid inferences on incomplete datasets.
+The resulting multiple inferences on multiple completed data sets can be combined into a single inference using Rubin's rules \citep[][p.
 76]{rubi87}.
-When prediction is the goal of the analysis effort, a single imputation may be sufficient \citep{nijm20}.
-The quality of a solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which an imputation procedure is able to capture these properties when modeling missing values.
+When generating predicted values is the goal of the data analysis, a single imputation may already generate sufficient \citep{nijm20} predicted values.
+
+In all scenarios it can be said that the quality of the solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which the imputation procedure is able to capture these properties when modeling missing values.
 In general, it holds that modeling missing data becomes more challenging when the amount of missingness increases.
@@
 
-When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies.
-Such studies are typically composed of an analysis pipeline in which data are generated, made incomplete, and imputed repeatedly under different simulation conditions.
+When evaluating the statistical properties - and thereby the practical applicability - of imputation methodology, researchers most often make use of simulation studies.
+Such studies are typically composed of an analysis pipeline in which data are subsequently generated, made incomplete, and imputed repeatedly under different simulation conditions.
 A set of evaluation criteria is then postulated to evaluate the performance of one or more missing data methods.
+Recent attention on best practices for method evaluation by means of simulation \citep{morr18}.
+and **REF MISSING** \citet{pawe22}'s advice to pre-register simulation protocols will guide methodologist in conceptualizing their simulation workflow.
 However, there is no consensus on how to design, execute and report simulation studies aimed to evaluate imputation methodology.
-Without a `gold standard' for the evaluation of imputation routines, the validity and comparability of simulation setups may differ tremendously from one developer to another.
-This brings forth a chain of potential problems in the objective assessment of imputation method performance within and across studies, which may lead to sub-optimal use of imputation in practice.
+Without a `gold standard' for evaluation, the validity and comparability of simulation setups may differ tremendously from one developer to another.
+This brings forth a chain of potential problems in the objective assessment of imputation method performance within and across studies, which could lead to sub-optimal use of imputation in practice.
 
@@
 
-The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology.
-These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals.
-Second, to provide imputation methodologists with a suggested course of action when using simulation studies to evaluate imputation techniques for missing data problems.
-This suggested approach should identify common ground but is in no way intended as an absolute solution.
-This identifies the third purpose of this paper: discussion.
+The purpose of this paper is threefold: First, we raise some concerns with respect to evaluating imputation methodology.
+These concerns stem from careful consideration with fellow `imputers' and from encounters as reviewers for statistical journals.
+Second, we provide imputation methodologists with a suggested course of action when using simulation studies to evaluate imputation techniques for missing data problems.
+This suggested approach should identify common ground, but is in no way intended as an absolute solution.
+This identifies the third purpose of our paper: discussion.
 We hope to elicit critical thinking regarding the problems at hand.
@@
 
-Standardizing the evaluation of imputation methodology requires simulators to consider common aspects in their simulation workflows.
+IK HEB EEN BEETJE MOEITE MET ONDERSTAANDE PARAGRAAF: HET GAAT OM DE REF VAN TIM EN DE PREREG. DIE KUNNEN WE OOK 2 ALINEAS EERDER INFIETSEN. IK HEB EEN POGING GEDAAN.
+Standardizing the evaluation of imputation methodology requires simulators adopt common aspects in their simulation and evaluation workflows.
 Table \ref{table:check} outlines some suggested steps to adopt.
@@
 
-The ideal evaluation of imputation methodology is study specific.
+The ideal evaluation of imputation methodology may differ between studies.
 A simulation study developed for the comparison of several imputation methods will typically have a different design than one aimed at establishing the inferential validity of a single (novel) method.
 How the evaluated methods are used in practice thereby governs the simulation procedure.
-Since the simulation aims and setup are intertwined with the choice of imputation methods under evaluation, the suitability of the evaluations should be assessed with the imputation methods' purpose in mind.
+Since the simulation aims and setup are intertwined with the choice of the evaluated imputation methods, one should always asses the suitablity of the chosen evaluation metrics with the imputation methods' purpose in mind.
 
@@
 If sampling variance is not of interest, a single complete data set can serve as comparative truth in all simulation repetitions (as opposed to a `standard' simulation setup, in which data is drawn every repetition).
-After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies.
-Removing sampling variance from performance evaluation is computationally convenient and may potentially sharpen comparisons between imputation methods, clarifying inconclusive results on their ability to correct for missingness.
+After all, we are interested in solving for the missingness and have no need for the noise induced by the sampling mechanism for evaluation in such studies.
+
+Removing sampling variance from performance evaluation is computationally convenient and may potentially sharpen comparisons between imputation methods, clarifying inconclusive results on their ability to solve for missingness.
 However, when deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
@@
 A well-designed simulation does not guarantee fair evaluations.
-Still, we propose to consider the following simulation design choices.
+Still, we propose to consider the following simulation design choices --GV WHICH CHOICES?
+Ik snap deze eerste zin niet helemaal--.
 Estimands and other simulation targets should be clearly and unambiguously defined in the context of the study aim \citet{pete14}.
@@
 
-To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established.
+To evaluate the ability of an imputation routine to handle missingness, a form of ground  truth has to be established.
 Those who perform simulation studies are in the luxury position to establish the truth beforehand by choosing a data-generating mechanism.
 Data-generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
-There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as official registers.
-The simulator should choose their data-generating mechanism(s) in line with the study scope.
+There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) GV IS DIT NODIG?
+from a sufficiently large observed set, such as official registers.
+The simulator should choose their data-generating mechanism in line with the study scope.
 Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived.
@@
 
-The problem with model-based data-generating mechanisms is that a method's performance on simulated data may not translate to empirical data.
-Real-life data hardly ever follow a given theoretical distribution, so there is no guarantee that simulation results are generalizable.
+The challenge with model-based data-generating mechanisms is that a method's performance on simulated data may not translate to empirical data.
+Real-life data hardly ever follow a known theoretical distribution, so there is no guarantee that simulation results are generalizable.
 Moreover, data are often generated such that the problem being studied is most pronounced, e.g.
@@
 Unsurprisingly, the performance of any imputation method will then be evaluated as good.
+
 Another threat to the generalizability of model-based simulations is the use of a single model for both data generation and imputation.
 If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied.
-Other (unfair) comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s), and the coherence between variables.
+Other (unfair) GV TEVEEL HAAKJES comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s), and the coherence between variables.
 In contrast to design-based studies, such characteristics are not always explicit simulation conditions, which may give a false sense of objectivity.

### 2022-09-15 Manuscript/plots/functions/evaluate.R

@@ -1,0 +1,36 @@
+evaluate <- function(object, population){
+  OUT <- list(NA)
+  N <- nrow(object[[1]]$imp$data)
+  pb <- txtProgressBar(min = 0, max = nsim, style = 3)  
+  for(i in 1:length(object)){
+    # extract mean of Y - name it Q cf. Rubin (1987, pp 76)
+    Q1           <- unlist(with(object[[i]]$imp, mean(Y))$analyses) 
+    # extract variance of Q - name it U cf. Rubin (1987, pp 76)
+    U1           <- unlist(with(object[[i]]$imp, var(Y))$analyses) / N
+    pool1         <- pool.finite(Q1, U1)
+    pool1$lower   <- pool1$qbar - qt(.975, pool1$df) * sqrt(pool1$t)
+    pool1$upper   <- pool1$qbar + qt(.975, pool1$df) * sqrt(pool1$t)
+    pool1$lambda  <- (pool1$b + (pool1$b / pool1$m)) / pool1$t
+    # extract CCA mean of Y - name it Q cf. Rubin (1987, pp 76)
+    CCA          <- na.omit(object[[i]]$incompl$data)
+    Q2           <- rep(mean(CCA$Y), object[[i]]$imp$m)
+    # extract CCA variance of Q - name it U cf. Rubin (1987, pp 76)
+    U2           <- rep(var(CCA$Y), object[[i]]$imp$m)  / 500
+    pool2         <- pool.scalar(Q2, U2)
+    pool2$lower   <- pool2$qbar - qt(.975, nrow(CCA) - 1) * sqrt(pool2$t)
+    pool2$upper   <- pool2$qbar + qt(.975, nrow(CCA) - 1) * sqrt(pool2$t)
+    pool2$lambda  <- (pool2$b + (pool2$b / pool2$m)) / pool2$t
+    #save evaluations to list
+    res      <- as.data.frame(rbind(unlist(pool1), unlist(pool2)))
+    rownames(res) <- c("imp", "CCA")
+    res$pop.mean <- mean(population)
+    res$bias <- res$pop.mean - res$qbar
+    res$cov <- res$lower < res$pop.mean & res$pop.mean < res$upper
+    res$ciw <- res$upper - res$lower
+    OUT[[i]] <- res
+    setTxtProgressBar(pb, i)
+  }
+  close(pb)
+  POOL <- Reduce("+", OUT) / length(OUT)
+  return(list(sim = OUT, pool = POOL))
+}

### 2022-09-15 Manuscript/plots/functions/pool.finite.R

@@ -1,0 +1,15 @@
+#adjusted pool.scalar function
+pool.finite <- function (Q, U, finite = TRUE) {
+  m <- length(Q)
+  qbar <- mean(Q)
+  ubar = 0  
+  b <- var(Q)
+  t <- ubar + (m + 1) * b/m
+  r <- (1 + 1/m) * b/ubar
+  lambda <- (1 + 1/m) * b/t
+  df <- (m - 1) * (1 + 1/r)^2
+  fmi <- (r + 2/(df + 3))/(r + 1)
+  fit <- list(m = m, qhat = Q, u = U, qbar = qbar, ubar = ubar, 
+              b = b, t = t, r = r, df = df, fmi = fmi, lambda = lambda)
+  return(fit)
+}

### 2022-09-16 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
changes

@@
 76]{rubi87}.
-When generating predicted values is the goal of the data analysis, a single imputation may already generate sufficient \citep{nijm20} predicted values.
+When generating predicted values is the goal of the data analysis, a single imputation may already generate sufficient predicted values \citep{nijm20}.
 
@@
 But for sake of progress, it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties.
-
-IK HEB EEN BEETJE MOEITE MET ONDERSTAANDE PARAGRAAF: HET GAAT OM DE REF VAN TIM EN DE PREREG. DIE KUNNEN WE OOK 2 ALINEAS EERDER INFIETSEN. IK HEB EEN POGING GEDAAN.
-Standardizing the evaluation of imputation methodology requires simulators adopt common aspects in their simulation and evaluation workflows.
-Table \ref{table:check} outlines some suggested steps to adopt.
-Please note that there is--of course--no `silver bullet' and that we do not claim to present a universally applicable approach.
-Our suggestions are the accumulated result of scientific literature, research experience, and discussion.
-For an excellent overview of general best practices for method evaluation by means of simulation, see \citet{morr18}.
-We recommend adhering to their proposed structure for planning and reporting simulation studies and \citet{pawe22}'s advice to pre-register simulation protocols.
-We henceforth highlight and add some recommendations for the evaluation of imputation methodology.
 % [TODO: Add advice to pre-register simulation protocols, which allows for fair comparisons of per-protocol results \citep{pawe22}.]
@@
 A well-designed simulation does not guarantee fair evaluations.
-Still, we propose to consider the following simulation design choices --GV WHICH CHOICES?
-Ik snap deze eerste zin niet helemaal--.
 Estimands and other simulation targets should be clearly and unambiguously defined in the context of the study aim \citet{pete14}.
@@
 Data-generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
-There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) GV IS DIT NODIG?
-from a sufficiently large observed set, such as official registers.
+There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled from a sufficiently large observed set, such as official registers.
 The simulator should choose their data-generating mechanism in line with the study scope.
@@
 If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied.
-Other (unfair) GV TEVEEL HAAKJES comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s), and the coherence between variables.
+Other (unfair) GV TEVEEL HAAKJES comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the measurement level, and the coherence between variables.
 In contrast to design-based studies, such characteristics are not always explicit simulation conditions, which may give a false sense of objectivity.
 
-An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be very challenging.
+An obvious problem with design-based simulation is that obtaining a large dataset without missing entries can be challenging.
 Most real-world data contains at least some missing entries, for which the true underlying missing data model is--by definition--unknown.
@@
 The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).
-Simulators should therefore be explicit in their description of the missing data generating model fitting the aim of their simulations.
+Simulators should therefore be explicit in their description of the missing data generating model and should make sure that the model fits the aim of their simulations.
 
@@
 
-[TODO: add that focusing on a single mechanism (even if the mechanism suits the study aim) is wasteful?] Truly random missingness across all data entries (`missing completely at random' or MCAR under Rubin's definition) may be considered a necessary simulation condition for the evaluation of imputation procedures, since the statistical properties of the observed data given the missing data are known.
+Truly random missingness across all data entries (`missing completely at random' or MCAR under Rubin's definition) may be considered a necessary simulation condition for the evaluation of imputation procedures, since the statistical properties of the observed data given the missing data are known.
 Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation.
@@
 
-[TODO: maybe add definitions of missingness mechanisms back in???] A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
+GV STREAMLINEN [TODO: maybe add definitions of missingness mechanisms back in???] A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
 If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
@@
 These relations in the observed data may, however, be weak or non-existent.
-If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid.
+If MAR is induced based on weaker relations in the true data, claims for a method's applicability to situations where the missingness is random become less valid.
 The most extreme example would be when MAR is induced from data without multivariate relations.
-The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random.
+The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is generated random GV PLAATJE WEER ER IN BRENGEN!.
 In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than imputation \citep[see e.g.][\S 2.7]{buur18}: the missingness does not depend on the incomplete variable.

### 2022-09-16 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
update plot

@@
 \usepackage{amsfonts}
+\usepackage{graphicx, latexsym}
 % \setlength{\bibsep}{2pt}
@@
 The most extreme example would be when MAR is induced from data without multivariate relations.
-The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is generated random GV PLAATJE WEER ER IN BRENGEN!.
+The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is generated randomly (see Figure 1).
 In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than imputation \citep[see e.g.][\S 2.7]{buur18}: the missingness does not depend on the incomplete variable.
 This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is pointless.
+
+\begin{figure}[h]
+  \resizebox{\textwidth}{!}{ %notice the \resizebox{} command
+        \includegraphics{plots/plot_mmech.pdf}
+  }      
+  \caption{Three different realizations of missingness generation for MCAR and right-tailed MAR missingness mechanisms for different levels of correlation $\rho$ between the variable to ampute and the covariates that guide the probability to be missing.
+Note that the conditional probabilities of MCAR and right-tailed MAR ($rho=0$) are identical.}
+    \label{Fig4}
+\end{figure}
 

### 2022-09-16 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex

@@
 
-[TODO: make this section less stringent and focus more on the suitability of the performancxe evaluation in the context of the study aim.]
+[TODO: make this section less stringent and focus more on the suitability of the performance evaluation in the context of the study aim.
+The take-away should be that simulation results cannot be easily compared if the performance measures differ.
+So we suggest to use CCA as benchmark method, and if possible also complete-data inference, to provide a lower and upper limit for expected imputation method performance.
+And even if the goal of the simulation effort is imputation method comparison, one should not just evaluate relative performance between imputation methods, but also report statistical properties of estimates such as bias, confidence intervals, and the coverage rate of the confidence intervals.]
 
@@
 (i) The estimates should preferably be unbiased.
-Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results.
+Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results (i.e., a negligible absolute bias may yield an infinite relative bias if the true value of the estimand is zero).
 In most cases, unbiased estimates may be expected under a MCAR missingness mechanism.
@@
 (iv) Predictive accuracy may be quantified using the root mean squared error (RMSE).
-We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values \citep[][\S 2.5]{buur18}.
-However, if a study is aimed at obtaining predictions or classification and inferential validity is not of interest, the RMSE of the predictions may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
-[TODO: make this an optional bullet point for prediction only: out of sample mean squared prediction error can be compared between methods.]
-
-[TODO: intrgrate this paragraph into the above!] Performance measures should express how well the obtained estimates approximate the comparative truth, but not every metric is suited for the evaluation of imputation methodology.
-For example, using measures of predictive accuracy or the RMSE (root mean squared error) as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{buur18}.
-And, under certain simulation conditions, an otherwise useful metric such as relative bias can distort the interpretation of results (i.e., a negligible absolute bias may yield an infinite relative bias if the true value of the estimand is zero).
-The usefulness of evaluation criteria may thus depend on the simulation goal and estimand.
+We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may increase the rate of false positives \citep[][]{buur18}.
+However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between method, which may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
 
@@
 This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.
-After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p.
-591]{neym34}.
-[TODO: add nuance about coverage rate > 95\% irt CIW and efficiency.]
-
-The choice of performance measures may inadvertently distort statistical properties of the imputed data.
-Developers often only inquire about the `accuracy' (i.e.
-how well can the method reproduce the original data).
-Merely focusing on the discrepancy between the true and imputed data would open the door for invalid inferences.
 

### 2022-09-19 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
update 2.3 and refs

@@
 If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied.
-Other (unfair) GV TEVEEL HAAKJES comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the measurement level, and the coherence between variables.
+Other potentially unfair comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the measurement level, and the coherence between variables.
 In contrast to design-based studies, such characteristics are not always explicit simulation conditions, which may give a false sense of objectivity.
@@
 
-GV STREAMLINEN [TODO: maybe add definitions of missingness mechanisms back in???] A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
-If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
+Besides MCAR missingness, data dependent missingness mechanisms, such as missing at random (MAR) missingness are paramount in statistical inference investigations.
+A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
+If the missingness is to be induced in longitudinal data, generating the missingness through autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
 It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness.
 The effects of different types of MAR mechanisms are described in \citet{scho18}.
+
 Inducing a MAR mechanism presents the simulator with a choice, namely the type or functional form of the missingness model.
@@
 With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g.
-weight is made incomplete based on observed gender to emulate a situation wherein one gender is less likely to disclose their weight).
+weight is made incomplete based on observed gender to mimic a situation wherein one gender is less likely to disclose their weight).
 These relations in the observed data may, however, be weak or non-existent.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht!]
   \resizebox{\textwidth}{!}{ %notice the \resizebox{} command
@@
 In addition to varying the missingness mechanisms, each simulated mechanism should be combined with different missingness patterns.
-Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process.
+Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process \citep[Corollary 6.1A]{litt20}.
 Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data.
 The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
-If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption.
+If there is no missingness --or if there is no data, for that matter-- the inference does not depend on the assumption.
+
 Alternatively, the validity of assumptions becomes increasingly important when the missingness increases.
@@
 
-[TODO: paraphrase ``be clear which mechanisms are of interest, how they are generated, if they are suitably generated" as conclusion to this section.]
+The above emphasizes the need for both thorough evaluation and thorough documentation of the intended and used missingness mechanisms.
+Simulators must be make clear which mechanisms are of interest.
+To be able to verify these mechanisms, simulators myst also be transparant about the process that generated the missingness.
+For example by sharing the code or detailing the process that generated the missing values.
+Alternatively, missingness generation devices, such as the \texttt{mice::ampute()} function can be used to generate valid, pattern-based missingness in any structured data set \citep{mice, ampute}.
+Finally, simulators must include simulation conditions and evaluation parameters that make it possible to identify the proper generation of missing values.
+This would prevent that errors in the missingness generation would go unnoticed.
+
 % \begin{table}[htb]

### 2022-09-19 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Restyled some paragraphs

@@
 There are, however, some overarching issues with assessing imputation method performance.
-In the first place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model after imputation (e.g.
-focusing on one performance measure over another, may yield different conclusions about imputation efficacy).
+In the first place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model after imputation.
+For example, focusing on one performance measure over another, may yield different conclusions about imputation efficacy.
 In the second place, diagnostic evaluation of the generated imputations is often left out of simulation study results.
@@
 (i) The estimates should preferably be unbiased.
-Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results (i.e., a negligible absolute bias may yield an infinite relative bias if the true value of the estimand is zero).
-In most cases, unbiased estimates may be expected under a MCAR missingness mechanism.
+Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results.
+A negligible absolute bias, for example, may already yield an almost infinite relative bias if the true value of the estimand is zero.
+In most cases, unbiased estimates may be expected under a MCAR missingness mechanism which can be easily verified by including MCAR missingness to the scope of missingness investigations.
 
 (ii) The estimates should have a proper coverage rate.
-Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient.
-Under-coverage (when estimation is too liberal and intervals are too narrow) indicates that the procedure is not confidence valid and may lead to invalid inference.
-Over-coverage (when estimation is too conservative and intervals are too wide) tells us that efficiency could still be gained.
+Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient \citep{neym34}.
+Under-coverage indicates that the procedure is not confidence valid and may lead to invalid inference.
+Undercoverage may occur when the estimation procedure is either biased, too liberal leading to intervals that are too narrow, or both.
+Over-coverage, though technically not confidence invalid, would indicate that efficiency could still be gained.
+In such cases, a narrower interval exists for which the method would still yield confidence valid result.
 
@@
 However, inference from a wider interval that is properly covered is to be considered more valid than a more narrow interval that is not properly covered anymore.
+That said, with valid nominal coverage, the method that yields the narrowest interval would be most efficient.
 
 (iv) Predictive accuracy may be quantified using the root mean squared error (RMSE).
-We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may increase the rate of false positives \citep[][]{buur18}.
+We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may increase the rate of false positives \citep[Chapter 2.6][]{buur18}.
 However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between method, which may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
 
-If the goal is inference--not prediction--then the uncertainty about estimates needs to be properly quantified.
+If the goal is inference --not prediction-- then the uncertainty about estimates needs to be properly quantified.
 To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation.
 The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied.
+This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
+
+GV CONTINUE HERE
 

### 2022-09-19 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
latest

@@
 This is not surprising as people from different fields could have a different focus on the problem at hand.
-There are, however, some overarching issues with assessing imputation method performance.
+There are, however, some overarching issues when assessing imputation method performance.
 In the first place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model after imputation.
@@
 
-(ii) The estimates should have a proper coverage rate.
+(ii) The intervals for estimates should have a valid coverage of the population (i.e.
+true) value.
 Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient \citep{neym34}.

### 2022-09-20 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
latest changes

@@
 The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
-as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.
+as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error; see][]{morr18}.
 The simulation design should ideally be fully factorial, i.e.
 varying each simulation condition against all other conditions.
-There may, for example, be interactions between the source and amount of missingness in the incomplete data and the efficacy of imputation methods (i.e., the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions) [TODO: add ref dance of the mechanisms!].
+There may, for example, be interactions between the source and amount of missingness in the incomplete data and the efficacy of imputation methods (i.e., the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions) \citep{scho18}.
 If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
@@
 
-[TODO: Make clear in this section that missingness mechanisms can be spurious (spurious MAR when correlations are too low, or when missingness proportion is too low).
+% [TODO: Make clear in this section that missingness mechanisms can be spurious (spurious MAR when correlations are too low, or when missingness proportion is too low).
 In reporting, the functional form or type of missingness is often omitted.
@@
   \caption{Three different realizations of missingness generation for MCAR and right-tailed MAR missingness mechanisms for different levels of correlation $\rho$ between the variable to ampute and the covariates that guide the probability to be missing.
-Note that the conditional probabilities of MCAR and right-tailed MAR ($rho=0$) are identical.}
+Note that the conditional probabilities of MCAR and right-tailed MAR ($\rho=0$) are identical.}
     \label{Fig4}
@@
 The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness.
-If there is no missingness --or if there is no data, for that matter-- the inference does not depend on the assumption.
+If there is no missingness--or if there is no data, for that matter-- the inference does not depend on the assumption.
 
@@
 Simulators must be make clear which mechanisms are of interest.
-To be able to verify these mechanisms, simulators myst also be transparant about the process that generated the missingness.
+To be able to verify these mechanisms, simulators must also be transparant about the process that generated the missingness.
 For example by sharing the code or detailing the process that generated the missing values.
@@
 \subsection{Performance evaluation}
-
-[TODO: make this section less stringent and focus more on the suitability of the performance evaluation in the context of the study aim.
-The take-away should be that simulation results cannot be easily compared if the performance measures differ.
-So we suggest to use CCA as benchmark method, and if possible also complete-data inference, to provide a lower and upper limit for expected imputation method performance.
-And even if the goal of the simulation effort is imputation method comparison, one should not just evaluate relative performance between imputation methods, but also report statistical properties of estimates such as bias, confidence intervals, and the coverage rate of the confidence intervals.]
-
 The evaluation criteria used to assess imputation performance vary from one simulator to another.
@@
 true) value.
-Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient \citep{neym34}.
+Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient \citep[p.
+589]{neym34}.
 Under-coverage indicates that the procedure is not confidence valid and may lead to invalid inference.
-Undercoverage may occur when the estimation procedure is either biased, too liberal leading to intervals that are too narrow, or both.
+Under-coverage may occur when the estimation procedure is either biased, too liberal leading to intervals that are too narrow, or both.
 Over-coverage, though technically not confidence invalid, would indicate that efficiency could still be gained.
@@
 
-If the goal is inference --not prediction-- then the uncertainty about estimates needs to be properly quantified.
+If the goal is inference--not prediction-- then the uncertainty about estimates needs to be properly quantified.
 To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation.
@@
 
-GV CONTINUE HERE
-
-Next to imputation method performance, the imputation-generating process may be evaluated.
-Omitting further evaluation of the imputation-generating process may yield sub-optimal imputations.
-The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models.
-When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
-[TODO: change this to being able to embed the analysis procedure and the imputation model within a Bayesian model; the nonresponse mdel may be misspecified while the imputation procedure is congenial.] The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
-This makes its assessment an infeasible endeavor in many simulation studies.
-Moreover, the performance of imputation procedures on distributional properties is often ignored too.
+After establishing the statistical validity of an imputation routine on the level of the analysis estimates, it remains paramount to also study the imputation-generating process.
+Omitting such investigations may yield sub-optimal results.
+One simple example arises often in practice: generating implausible imputed values.
 Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
@@
 Under many circumstances, imputation methods may be realistically expected to preserve both marginal and conditional distributions with respect to the comparative truth.
-Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).
+Imputation methods that fail to do so, should not be considered general-purpose methods.
+Many techniques can yield valid inferences, but the techniques that samples realistic or plausible values may be preferable in practice.
 
-[TODO: add segway to this paragraph.] Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations.
+The evaluation of simulated results goes far beyond plausibility and distributional characteristics.
+The quality and validity of an imputation task will also rely on the relation between the missing data models, imputation models, and analysis models.
+When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
+[TODO: change this to being able to embed the analysis procedure and the imputation model within a Bayesian model; the nonresponse model may be misspecified while the imputation procedure is congenial.] The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
+This makes its assessment a challenging and timeconsuming endeavor in many simulation studies.
+
+When algorithms are used the simulator should study the validity of the algorithmic process.
+Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations.
 As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated.
+\citet{ober21} demonstrates that statistical validity can happen before algorithmic convergence is reached, but this is not a guarantee.
 Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor failproof.
 As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method.
-Problems with producing stable imputations is not an exclusive quality iterative imputation algorithms.
-Any method may run into failures of the imputation-generating process and subsequently lack results (e.g.
-due to over-parameterization errors).
+Problems with producing stable imputations are not exclusive to iterative imputation algorithms.
+Any method may run into failures of the imputation-generating process and subsequently lack results, for example due to over-parameterization errors.
 
-[TODO: integrate this into text above!] Every imputation workflow should contain an evaluation of the obtained imputations.
+Every imputation workflow should therefore contain an evaluation of the obtained imputations.
 Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend to consider the following aspects:
@@
 (ii) The fit of the imputation model may be verified with the help of a posterior predictive check \citep[][]{nguy17, zhao22}.
-A straightforward posterior predictive check for imputation methodology is the multiple over-imputation of observed data values.
+A straightforward posterior predictive check for imputation methodology is the multiple over-imputation of observed data values \citep{caiPPC}.
 If the statistical properties of the over-imputed values are equivalent to those of the observed data values, one could infer that the imputation model fits the observed part of the incomplete data reasonably well.
 Then, by extension, one could assume that the imputation model might be able to produce good imputations for the missing part of the data too.
-[TODO: add that this is implemented in amelia.]
+This overimputation procedure is straigtforward with the \texttt{overimpute()} function in \texttt{Amelia II} \citep{amelia} and the \texttt{where} argument in \texttt{mice} \citep{mice, caiPPC}.
 
@@
 When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.
-[TODO: add nuance here because valid inference, not plausible imputed values, should be the aim (although plausibility is nice).]
+We must note, however, that the aim of imputation should be valid inference in the first place, and plausibility in the second place.
 
-[TODO: add segway to reference method paragraph and add that performance on the complete data offers an upper limit for performance.] In addition, the simulator should always perform complete case analysis.
-We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance.
+There are parallels between evaluating the performance and evaluating the missingness.
+Just like using MCAR as a condition to benchmark the validity of the missing data solution, we can use the simulation's complete data (i.e.
+without missing values) as an upper limit for our performance evaluation.
+In addition, the simulator should also perform complete case analysis, i.e.
+analysis on the incomplete data without solving for the missingness.
+We know the theoretical properties of complete case analysis, which makes the ad hoc technique useful as a lower limit for evaluating imputation performance.
 Complete case analysis may, therefore, serve as a benchmark method against which imputation performance should be evaluated.
-Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.
+Realistically, the simulator would expect the imputation solution to perform not worse than the complete case analysis and, preferably, mimic the complete data results.
+Moreover, pairing complete case analysis with MCAR and MAR mechanisms allows the simulator to evaluate the impact of the missing data generation process under the chosen analysis model.
 

### 2022-09-20 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
final text/bib/img update

@@
 \usepackage{amsfonts}
+\usepackage{epstopdf}
 \usepackage{graphicx, latexsym}
@@
 Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
+In part this may be due to different aims between fields and studies.
+For example, when evaluating imputation techniques aimed at prediction, different aims may be formulated than when statistical inference is of interest.
+The lack of consensus may also stem from different personal preference or scientific background.
+All in all, the lack of a common ground in evaluating imputation methodology may lead to sub-optimal use in practice.
 In this paper, we propose a move towards a standardized evaluation of imputation methodology.
 To demonstrate the need for standardization, we highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
-This may lead to sub-optimal use of  imputation in practice.
 Additionally, we suggest a course of action for simulating and evaluating missing data problems.
-[TODO: Narrowing the scope of the manuscript to the evaluation of methodology for statistical inference, not prediction or causal inference (although inferential validity encompasses predictive validity and may thus be generalized?).
-Emphasizing that the ideal evaluation of imputation methodology depends on the aim of the specific study (i.e.
-how the evaluated methods are used in practice).
-Split up the problems with the simulation setup/evaluation and the problems with the interpretation/extrapolation from these simulation studies.
-See e.g.
-\citet{gree17} on cognitive problem of generelization versus methodological problems.
-Explain how the standardisation leads to more neutrality, but that this is just a set of elements to consider/report, not a complete 'cookbook'.
-Add that we focus on  simulation studies evaluating imputation methodology for statistical inference.]
+Our suggested course of action is by no means meant to serve as a complete cookbook, but rather meant to incite critical thinking and a move to objective and fair evaluations of imputation methodology.
+We invite the readers of this paper to contribute to the suggested course of action.
 \end{abstract}
@@
 A set of evaluation criteria is then postulated to evaluate the performance of one or more missing data methods.
-Recent attention on best practices for method evaluation by means of simulation \citep{morr18}.
-and **REF MISSING** \citet{pawe22}'s advice to pre-register simulation protocols will guide methodologist in conceptualizing their simulation workflow.
+Recent attention on best practices for method evaluation by means of simulation \citep{morr18} and \citet{pawe22}'s advice to pre-register simulation protocols will guide methodologist in conceptualizing their simulation workflow.
 However, there is no consensus on how to design, execute and report simulation studies aimed to evaluate imputation methodology.
@@
 For example, misinterpretation and (unintended) questionable research practices can prompt extrapolation beyond the scope of the simulations \citep{gree17, pawe22}.
-Imputation methods that seem to work well in simulations may then not be suitable in practice, when applied on incomplete empirical data.
+Imputation methods that seem to work well in simulations may then not be suitable in practice, when applied to incomplete empirical data.
 
@@
 Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations and only use the missing data generation procedure as the source of Monte Carlo variation \citep{vink14}.
-If sampling variance is not of interest, a single complete data set can serve as comparative truth in all simulation repetitions (as opposed to a `standard' simulation setup, in which data is drawn every repetition).
+If sampling variance is not of interest, a single complete data set can serve as comparative truth in all simulation repetitions--as opposed to a `standard' simulation setup, in which data is drawn every repetition.
 After all, we are interested in solving for the missingness and have no need for the noise induced by the sampling mechanism for evaluation in such studies.
@@
 varying each simulation condition against all other conditions.
-There may, for example, be interactions between the source and amount of missingness in the incomplete data and the efficacy of imputation methods (i.e., the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions) \citep{scho18}.
+There may, for example, be interactions between the source and amount of missingness in the incomplete data and the efficacy of imputation methods--i.e., the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions \citep{scho18}.
 If there is more than one imputation method under evaluation, the simulation design should apply each method to every incomplete data set.
@@
 dependency structure) - if a human misinterprets findings in a comparison of method A and B, then this is a cognitive problem and not the problem on how missingness was generated in the DGP."]
-
-[TODO; describe problems with data generation and interpretation: If the source of the data is model-based, the correlations in the data may be so high that there is no effective missingness due to the predictive power in the other variables, and the data source may be unfair if the data generating model is equal to one of the imputation models under evaluation.
+%
+%[TODO; describe problems with data generation and interpretation: If the source of the data is model-based, the correlations in the data may be so high that there is no effective missingness due to the predictive power in the other variables, and the data source may be unfair if the data generating model is equal to one of the imputation models under evaluation.
 With a design-based simulation, the data source is unlikely to have no missingness and the choice of initial missing data method may propagate into the evaluations.]
@@
 
-\begin{figure}[ht!]
+\begin{figure}[t!]
   \resizebox{\textwidth}{!}{ %notice the \resizebox{} command
@@
   \caption{Three different realizations of missingness generation for MCAR and right-tailed MAR missingness mechanisms for different levels of correlation $\rho$ between the variable to ampute and the covariates that guide the probability to be missing.
-Note that the conditional probabilities of MCAR and right-tailed MAR ($\rho=0$) are identical.}
+Note that the conditional probability of right-tailed MAR ($\rho=0$) mimics that of MCAR.}
     \label{Fig4}
@@
 The quality and validity of an imputation task will also rely on the relation between the missing data models, imputation models, and analysis models.
-When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.
-[TODO: change this to being able to embed the analysis procedure and the imputation model within a Bayesian model; the nonresponse model may be misspecified while the imputation procedure is congenial.] The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
+When a joint model exists whose conditionals would include the imputation and analysis model, then the imputation model and analysis are said to be congenial \citep{meng94, bart2015}.
+The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
 This makes its assessment a challenging and timeconsuming endeavor in many simulation studies.
@@
 
+\newpage
+
 \section*{Appendix}
@@
 
-\begin{table}[ht]
+\begin{table}[ht!]
 \caption{Checklist for reporting on imputation methodology evaluations.}

### 2022-09-20 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
reviewer 2 response

@@
 
-Besides MCAR missingness, data dependent missingness mechanisms, such as missing at random (MAR) missingness are paramount in statistical inference investigations.
+Besides MCAR missingness, observed-data dependent missingness mechanisms--such as missing at random (MAR) missingness--are paramount in statistical inference investigations.
 A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
@@
 
-A non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice.
+A non-ignorable or MNAR mechanism--where the unobserved values themselves may also play a role in the probability to be missing--might fall outside many studies' scope, yet could yield informative insights for daily practice.
 Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time.

### 2022-09-20 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
update response reviewer 1

@@
 Unfortunately, there is no consensus on how to perform simulation studies to evaluate the properties of imputation methods.
-In part this may be due to different aims between fields and studies.
+In part, this may be due to different aims between fields and studies.
 For example, when evaluating imputation techniques aimed at prediction, different aims may be formulated than when statistical inference is of interest.
@@
 In this paper, we propose a move towards a standardized evaluation of imputation methodology.
-To demonstrate the need for standardization, we highlight a set of potential pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
+To demonstrate the need for standardization, we highlight a set of possible pitfalls that bring forth a chain of potential problems in the objective assessment of the performance of imputation routines.
 Additionally, we suggest a course of action for simulating and evaluating missing data problems.

### 2022-09-20 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Finish all but one TODO in reviewer responses

@@
 
-% In contrast to ad hoc methods for dealing with missing values (e.g.
-list-wise deletion), imputation allows analysts to account for the sources of uncertainty that are related to the missingness problem.
-With the popular variant multiple imputation \citep{rubi76}, each missing value is replaced by a plausible data value several times, resulting in multiple completed data sets.
-The estimated parameters of interest for each of these completed sets can be combined into a single inference using Rubin's rules \citep[][p.
-76]{rubi87}.
-[TODO: add that if the goal is statistical inference, the uncertainty due to missingness should be quantified/captured/propagated by the missing data method.
-Refer to \citet{sper20} for the context of prediction, where multiple imputation is not always necessary.] [TODO: paraphrase ``the quality of a technique can only be evaluated with respect to the aims of the problem/s it's intended to solve in practice".]
-
 When evaluating the statistical properties - and thereby the practical applicability - of imputation methodology, researchers most often make use of simulation studies.
@@
 This brings forth a chain of potential problems in the objective assessment of imputation method performance within and across studies, which could lead to sub-optimal use of imputation in practice.
-
-% Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced.
-Although these promising new methods seem to yield even sharper imputations than now-standard imputation methods, the comparison may not be fair due to the simulation setup.
 
@@
 But for sake of progress, it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties.
-% [TODO: Add advice to pre-register simulation protocols, which allows for fair comparisons of per-protocol results \citep{pawe22}.]
 
@@
 \section{Why some evaluations should not be trusted}
-
-% [TODO: split up into design, execution and reporting? \citep{pawe22}]
 
@@
 
-% [TODO: describe problems with unclear simulation parameters, extrapolation beyond scope of simulations, QRPs \citep{pawe22} and misinterpretation \citep[see "investigator bias"][]{gree17}.]
-
-% Therefore, taking sampling variation into account is neither necessary for obtaining information about a method's ability to handle the missing data problem, nor for objectively comparing methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}.
-The only required change to simulation setups is to remove sampling variance from performance evaluation.
-Still, a lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.
-
-% The comparative truth of the estimand(s) is determined by the choice of data-generating mechanism and whether sampling variance is included in the simulation setup, see Table \ref{table:dgm_truth}.
-
-% If sampling variance cannot be omitted from the simulation scheme, multiple samples should be drawn from the data-generating mechanism (i.e., one sample per simulation repetition, a standard simulation setup).
-
-% If sampling variance is not of interest, a single complete data set can be obtained from the data-generating mechanism (i.e., one sample for all simulation repetitions).
-
-% \begin{table}[tb]
-% \begin{center}
-% \caption{Source of the comparative truth in simulation studies.}
-% \label{table:dgm_truth}
-% \begin{tabular}{lll}
-% \hline
-%                & With sampling variance      & Without sampling variance \\
-%                & (multiple samples drawn)    & (one sample drawn) \\
-% \hline  
-% Model-based simulation   & data-generating model         & single sample \\
-% Design-based simulation  & sufficiently large data set   & single sample \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \end{table}
-%
-
 A well-designed simulation does not guarantee fair evaluations.
@@
 \subsection{Data generation}
-
-% [TODO: split up into data and missingness part of the DGM, with choices between data sources, missingness mechanisms and proportions.
-And divide problems between simulation and interpretation, see Reviewer 2: "whether method A (e.g.
-imputation) outperforms method B (e.g.
-listwise deletion) depends on the estimand, purpose of the study, used distributions (in the DGP, imputation model, possibly also analysis model), and simulation setup (e.g.
-dependency structure) - if a human misinterprets findings in a comparison of method A and B, then this is a cognitive problem and not the problem on how missingness was generated in the DGP."]
-%
-%[TODO; describe problems with data generation and interpretation: If the source of the data is model-based, the correlations in the data may be so high that there is no effective missingness due to the predictive power in the other variables, and the data source may be unfair if the data generating model is equal to one of the imputation models under evaluation.
-With a design-based simulation, the data source is unlikely to have no missingness and the choice of initial missing data method may propagate into the evaluations.]
 
@@
 
-% 1a. choose DGM (if model-based: choose n_obs, n_var, relations between var, var types, etc; if design-based: choose n_obs, missing data handling strategy)
-% 1b. decide if sampling variance is needed
-% 1c. draw complete sample(s)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \subsection{Missingness generation}
-
-% [TODO: Make clear in this section that missingness mechanisms can be spurious (spurious MAR when correlations are too low, or when missingness proportion is too low).
-In reporting, the functional form or type of missingness is often omitted.
-As advice, MCAR can be a very informative mechnism for reference and MNAR may be useful for practice.
-Missingness proportions can be misinterpreted.
-Missingness pattern may have unexpected consequences (e.g., univariately generated multivariate missingness).
-In interpretation, watch out for generalizations from unintended special cases.]
-
-% Evaluating imputation methodology requires a missing data problem to be solved.
-After establishing a comparative truth from a data-generating mechanism, some form of missingness therefore has to be induced.
 
@@
 
-% \begin{table}[tb]
-% \begin{center}
-% \caption{Missingness mechanisms. [TODO: add refs \citet{seam13, meal15, dore18} for clarification and alternatives outlined in \citet{moha21, scho21, more18}]}
-% \label{table:mech}
-% \begin{tabular}{lll}
-% \hline
-%       & Interpretation and consequences of the mechanism \\
-% \hline  
-% MCAR  & Missing completely at random: The probability to be missing is the same for all \\
-%       & cases. In other words, the missing values are missing at random and the observed \\
-%       & values are observed at random. \\ \noalign{\smallskip}
-% MAR   & Missing at random: The probability to be missing is the same within groups of \\
-%       & cases defined by the observed data only. The essence is that observed relations in \\
-%       & the data inform the missingness, such that MCAR may be assumed within the \\ 
-%       & observed groups. \\ \noalign{\smallskip}
-% v-MAR & Conditional on the fully observed variables missingness occurs at random 
-%       & \citep[][p. 1026]{moha21}.\\
-% MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
-%       & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
-%       & inferred from the observed data. This is considered non-ignorable missingness \\
-%       & \citep[see e.g.][]{rubi76}. \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \end{table}
-
 Even the terminology on missingness generation can be confusing.
@@
 Simulators should therefore be explicit in their description of the missing data generating model and should make sure that the model fits the aim of their simulations.
-
-% Moreover, the value of the actual missingness proportion may be diffusing too.
-Some studies use only 10\% missingness whereas other studies push the limits to additionally investigate the performance under missingness proportions of at least 50\%.
-The inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
-This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
-
-% Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables.
-It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways.
 
@@
 This complicates any definitive conclusions about imputation method performance in relation to the data generation parameters.
-
-% Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions).
-However, the missingness mechanisms that are induced in simulation studies do not necessarily match any true missingness generating mechanisms in incomplete empirical data, nor the mechanisms that are typically assumed when imputing such data.
-On the one hand, there is MCAR missingness, which is unlikely to be the true missingness generating mechanism, but would yield insightful evaluations in a simulation setting.
-On the other hand, there is MNAR, which is arguably a quite reasonable source of many real-world missingness scenarios but is generally ignored except for those evaluations that are specifically targeted at non-ignorable applications.
-A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.
 
@@
 The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is generated randomly (see Figure 1).
-In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than imputation \citep[see e.g.][\S 2.7]{buur18}: the missingness does not depend on the incomplete variable.
+In fact, this would amount to one of the special cases described by \citet{buur18} under which complete case analysis would be more efficient than imputation: the missingness does not depend on the incomplete variable.
 This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is pointless.
@@
 Alternatively, a sensitivity analysis may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% 2a. choose missingness mechanism(s): at least consider MCAR, vary MAR types
-% 2b. choose missing data pattern(s): introduce non-monotone multivariate missingness
-% 2c. choose missingness proportion(s): at least consider 10, 25 and 50\%
-
-% First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions).
-This mechanism should be used as a reference condition.
-Every imputation method must be able to yield valid imputations under MCAR, both in terms of distributional characteristics as well as statistical inference.
-% 
-% Next, missing data should be induced conform a model that is dependent on the observed data (i.e.
-a MAR mechanism).
-Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully.
 
@@
 
-% \begin{table}[htb]
-% \begin{center}
-% \caption{Suggested univariate missingness proportions.
-[TODO: change this to 'more than one missingness proportion, depending on the context' and/or add references for the given proportions (e.g.
-about  that 30\% is a good range for MI).]}
-% \label{table:prop}
-% \begin{tabular}{lll}
-% \hline
-%       & Reasoning for suggested missingness proportion \\
-% \hline  
-% 10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
-%       & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
-%       & the true data inference. Performance of a missing data method should at least be \\
-%       & acceptable for most missing data problems. \\ \noalign{\smallskip} % & \\
-% 25\%  & This is a fair amount of missingness and will, depending on the observed data \\
-%       & information, have a noticeable influence on the completed data inference. When \\
-%       & compared to the condition with 10\% missingness, the inference obtained under \\
-%       & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
-%       & should increase), but estimates should still be properly covered and the statistical \\
-%       & properties of the missing data method should be sound. In practice, at least to our \\
-%       & at least to our experience in social sciences and official statistics, 25\% \\
-%       & univariate missingness can easily be considered as a realistic missingness percentage. \\ \noalign{\smallskip} % & \\
-% 50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
-%       & severely. Depending on how the missingness mechanism interacts with the simulated \\
-%       & data, some imputation techniques may yield estimates that are under-covered such that \\
-%       & the completed data inference should not be deemed valid anymore. If a method yields \\
-%       & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
-%       & properties of the imputation methodology are sound. \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \subsection{Imputation methods}
-
-% 3a. Define missing data methods: at least consider complete case analysis as benchmark
-% 3b. Choose imputation model parameters: appropriate n_imp and n_it (if applicable), maybe vary predictor matrices?
-% 3c. Impute the data
-% 3d. Obtain estimates
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Such evaluations, however, are typically omitted and valuable insights into the imputation method(s) may then be overlooked.
-%(i.e.
-the complete-data model used to estimate the comparative truth).
-% The choice of performance measures might inadvertently influence simulation results.
 
@@
 
-(i) The estimates should preferably be unbiased.
+(i) The methods should preferably be unbiased.
 Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results.
@@
 
-% "Note that Neyman's original description of confidence intervals defined the property of randomisation validity as exactly 100(1_alpha) of intervals containing theta.44-46 Confidence validity is the property that the true percentage is at least 100(1_alpha).
-This latter definition is less well known than the former, with the result that overcoverage and undercoverage are sometimes regarded as similarly bad.47 Of course, randomisation validity would usually be preferred over confidence validity because it implies shorter intervals, but this is not always the case.
-There are examples of procedures that return both shorter intervals and higher coverage.45, 46" ADEMP paper
-
 (iii) The width of the confidence or credible interval may convey statistical efficiency, which should be considered to compare imputation methods.
@@
 (iv) Predictive accuracy may be quantified using the root mean squared error (RMSE).
-We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may increase the rate of false positives \citep[Chapter 2.6][]{buur18}.
+We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may increase the type I error rate of the statistical inferences \citep[Chapter 2.6][]{buur18}.
 However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between method, which may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
@@
 
-% 4a. Evaluate imputations: check failed methods, algorithmic convergence (if applicable)
-% 4b. Check imputation model fit: anomalies in distributional characteristics, plausibility of imputed values (if requested), PPC
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% 4c. Apply performance measures: at least consider bias, coverage, CI length, if applicable also RMSE (at prediction and/or cell level)
-% fraction of missing information (= model parameter), variance inflation
-
 Last, when the simulator is on the verge of drawing conclusions about the performance of the imputation methodology, the performance should be carefully qualified.
@@
 A method may perform badly, but if it still outperforms every other approach, it may yet be of great practical relevance.
-
-% It could be convincing to demonstrate a method's applicability to real-world missing data problems.
-This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set.
-Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data.
-However, the real-world missingness would then be omitted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \subsection{Reporting}
-
-% 
-% \begin{table}[tb]
-% \begin{center}
-% \caption{Steps to consider in imputation simulation studies.}
-% \label{table:steps}
-% \begin{tabular}{lll}
-% \hline
-%                         & Simulation aspects under consideration \\
-% \hline  
-% 1. Set scope            & Aim(s), missing data method(s) to evaluate, simulation design \\
-% 2. Obtain truth         & Data-generating mechanism(s), estimand(s), sampling variance \\
-% 3. Induce missingness   & Missingness mechanism(s), missing data pattern(s) \\
-% 4. Apply methods        & Imputation method(s), analytic method(s) \\
-% 5. Evaluate imputations & Imputation-generating process, imputation model fit \\
-% 6. Evaluate performance & Performance measures, qualitative judgment \\
-% 7. Report               & Text, visualization, checklist, simulation script \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \end{table}
 

### 2022-09-21 Manuscript/plots/2. Createplot.R
recompile revised manuscript
change theme of figure to remove grey background, put review files in separate folder

@@ -22,6 +22,7 @@ levels(data$mech) <- c("MCAR",
 p <- ggplot(data, aes(Y, p.mis, colour = factor(mech)) ) + 
   geom_point() + 
   geom_smooth(method = "lm", colour = "black") + 
+  theme_classic() +
   scale_colour_discrete(guide="none") + 
   facet_grid(. ~ mech, labeller = label_parsed) + 
   ylab("Probability to be missing") + 

### 2022-09-21 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
update
update rev1

@@
 Another example of potential side-effects of missingness generation may occur when multivariate missingness is generated using step-wise univariate missingness induction.
-The resulting incomplete data may then not have the desired statistical properties \citep{ampute}.
+The resulting incomplete data may then not have the desired statistical properties or unintentional non-ignorable missingness may be generated \citep{ampute, robins1997}.
 This complicates any definitive conclusions about imputation method performance in relation to the data generation parameters.
@@
 
-Truly random missingness across all data entries (`missing completely at random' or MCAR under Rubin's definition) may be considered a necessary simulation condition for the evaluation of imputation procedures, since the statistical properties of the observed data given the missing data are known.
+Truly random missingness across all data entries (`missing completely at random' or MCAR under Rubin's (1976) \nocite{rubi76} definition) may be considered a necessary simulation condition for the evaluation of imputation procedures, since the statistical properties of the observed data given the missing data are known.
 Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation.
@@
 Realistically, the simulator would expect the imputation solution to perform not worse than the complete case analysis and, preferably, mimic the complete data results.
-Moreover, pairing complete case analysis with MCAR and MAR mechanisms allows the simulator to evaluate the impact of the missing data generation process under the chosen analysis model.
+Moreover, pairing complete case analysis with MCAR and MAR mechanisms allows the simulator to evaluate the impact of the missing data generation process under the chosen analysis model and to identify the unintentional generation of accidental MNAR.
 

### 2022-09-21 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Fix typos/referencing errors, combine paragraphs 1 and 2 from section 3

--- 2022-09-20 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
@@
 A well-designed simulation does not guarantee fair evaluations.
-Estimands and other simulation targets should be clearly and unambiguously defined in the context of the study aim \citet{pete14}.
-The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g.
+Estimands and other simulation targets should be clearly and unambiguously defined in the context of the study aim \citep{pete14}.
+The study aim and the required level of precision may also inform the number of simulation repetitions \citep[e.g.
 as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error; see][]{morr18}.
@@
 Therefore, the simulator needs to deal with missingness in \textit{some} way before incomplete empirical data can serve as comparative truth in the simulations.
-It may seem like an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples.
+It might seem like an intuitive solution to only draw complete cases from the large dataset, which would indeed yield complete samples.
 However, there may be inherent differences between cases with and cases without any missing values, due to the unknown missing data model.
@@
 In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation.
-The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).
+The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations, because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).
 Simulators should therefore be explicit in their description of the missing data generating model and should make sure that the model fits the aim of their simulations.
@@
 Extrapolating simulation results from a specific missing data generating model to more intricate (empirical) missingness could lead to sub-optimal advice in practice.
-For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may inadvertedly induce missingness according to the special case descibed by \citet[][p.
-\S 2.7]{buur18}.
+For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may inadvertedly induce missingness according to the special case descibed by \citet[][\S 2.7]{buur18}.
 Under this specific missingness model, list-wise deletion may outperform any imputation method.
@@
 Besides MCAR missingness, observed-data dependent missingness mechanisms--such as missing at random (MAR) missingness--are paramount in statistical inference investigations.
-A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}.
+A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}; generalizations to multivariate MAR missingness can be found in \citet{ampute}.
 If the missingness is to be induced in longitudinal data, generating the missingness through autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
@@
 If there is no missingness--or if there is no data, for that matter-- the inference does not depend on the assumption.
-
 Alternatively, the validity of assumptions becomes increasingly important when the missingness increases.
@@
 Depending on how the missingness mechanism interacts with the simulated data, higher missingness proportions may yield biased or invalid completed data inferences.
-The missingness proportions should therefore be considered carefully.
+The missingness proportions should thus be considered carefully.
 
@@
 It depends on the specifics of each simulation study which performance measures would be most suitable.
-We recommend to evaluate at least the following points:
+If the goal is inference--not prediction--the uncertainty about estimates needs to be properly quantified.
+To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation.
+The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
+This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
+We therefore recommend to evaluate the following points:
 
@@
 A negligible absolute bias, for example, may already yield an almost infinite relative bias if the true value of the estimand is zero.
-In most cases, unbiased estimates may be expected under a MCAR missingness mechanism which can be easily verified by including MCAR missingness to the scope of missingness investigations.
+In most cases, unbiased estimates may be expected under a MCAR missingness mechanism, which can be easily verified by including MCAR missingness to the scope of missingness investigations.
 
@@
 true) value.
-Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient \citep[p.
+Coverage of a 95\% interval should in theory be $\geq 95$\%, where a coverage rate of 95\% would be most efficient \citep[p.
 589]{neym34}.
@@
 (iv) Predictive accuracy may be quantified using the root mean squared error (RMSE).
-We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may increase the type I error rate of the statistical inferences \citep[Chapter 2.6][]{buur18}.
-However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between method, which may yield valuable information about the methods' efficiency in terms of both accuracy and precision.
-
-If the goal is inference--not prediction-- then the uncertainty about estimates needs to be properly quantified.
-To capture this uncertainty, the standard errors of the estimates should be correctly calculated, which requires multiple imputation.
-The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
+We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may inflate the type I error rate of statistical inferences \citep[Chapter 2.6][]{buur18}.
+However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between methods.
+Then, the RMSE may yield valuable information about the methods' comparative efficiency in terms of accuracy and precision.
 
@@
 Imputation methods that fail to do so, should not be considered general-purpose methods.
-Many techniques can yield valid inferences, but the techniques that samples realistic or plausible values may be preferable in practice.
+Many techniques can yield valid inferences, but techniques that sample realistic or plausible values may be preferable in practice.
 
@@
 
-After the careful consideration and execution of all simulation aspects, the evaluations should be properly reported.
-The inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
-This ultimately may lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
-
-We encourage simulators to document their choices and to be explicit in their descriptions.
-For example, the simulation design may be presented textually, in a flow chart or as a block of pseudo-code, whereas missingness mechanisms may be written as a function of the data or displayed graphically.
-Ideally, this should be supplemented by an online repository with all of the data and code required to reproduce the simulation results.
+We encourage simulators to carefully consider and document their choices in the evaluation of imputation methodology.
+Simulation studies should not only be well-executed, but also properly reported.
+For example, the inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
+This may ultimately lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
+Simulators should therefore to be explicit in the descriptions of their evaluations, but deciding what and how to report may be challenging.
+The simulation design could be presented textually, in a flow chart or as a block of pseudo-code, whereas missingness mechanisms could be written as a function of the data or displayed graphically.
+Ideally, the evaluations should be supplemented by an online repository with all of the data and code required to reproduce the simulation results.
 To aid simulators in reporting and move towards standardization in evaluation, we provide a draft version for reporting guidelines in Appendix A.1 (also available from \underline{www.gerkovink.com/evaluation}).

### 2022-09-22 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
edit checklist to match ADEMP format

@@
 $\square$  & Number of simulation repetitions                                                   \\ \hline
-2 & Comparative truth                                                                  \\ \cline{2-2}
-$\square$  & Data source (model-based or design-based)                                           \\
-$\square$  & Sampling variance                                                                  \\
+2 & Data generation                                                                  \\ \cline{2-2}
+$\square$  & Data source (e.g. model-based or design-based, sampling variance)                                           \\
 %  & Data characteristics (number of observations, number of variables, variable types, \\
@@
 $\square$  & Estimand                                                                           \\ \hline 
-3 & Induced missingness                                                                \\ \cline{2-2}
+3 & Missingness generation                                                                \\ \cline{2-2}
 $\square$  & Missingness mechanism (incl. type or functional form of the missing data model)  \\
-$\square$  & Missingness pattern (incl. missingness proportion)                                 \\ \hline 
-4 & Applied methods                                                                    \\ \cline{2-2}
+$\square$  & Missingness pattern (incl. missingness proportions)                                 \\ \hline 
+4 & Methods                                                                    \\ \cline{2-2}
 $\square$  & Imputation methods (incl. parameters e.g. the number of imputations)               \\
 $\square$  & Analytic methods (incl. calculation of standard errors e.g. pooling rules)       \\
-$\square$  & Reference method (e.g. complete case analysis)                                    \\ \hline 
-5 & Imputation evaluation                                                              \\ \cline{2-2}
+$\square$  & Reference methods (e.g. complete case analysis)                                    \\ \hline 
+6 & Performance evaluation                                                             \\ \cline{2-2}
+$\square$  & Statistical properties (e.g. confidence validity)                                  \\
+$\square$  & Comparative performance (e.g. predictive accuracy)                                 \\ \hline
 $\square$  & Imputation-generating process (e.g. algorithmic non-convergence)                   \\
@@
 $\square$  & Distributional characteristics (e.g. plausibility of imputed values)               \\ \hline 
-6 & Performance evaluation                                                             \\ \cline{2-2}
-$\square$  & Statistical properties (e.g. confidence validity)                                  \\
-$\square$  & Comparative performance (e.g. predictive accuracy)                                 \\ \hline
 \end{tabular}

### 2022-09-22 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Update checklist in manuscript

@@
 Table \ref{table:check} provides a checklist for reporting on imputation methodology evaluations.
+We have structured the checklist according to simulation study reporting format proposed by \citet{morr18}.
 
@@
 \begin{tabular}{ll}
-% \hline
-%   & Suggested aspects to report                                                       \\
 \hline
-1 & Simulation scope                                                                   \\ \cline{2-2}
-$\square$  & Aim                                                                                \\
-$\square$  & Design (incl. pseudo-code or flow diagram)                                         \\
-$\square$  & Number of simulation repetitions                                                   \\ \hline
-2 & Data generation                                                                  \\ \cline{2-2}
-$\square$  & Data source (e.g. model-based or design-based, sampling variance)                                           \\
-%  & Data characteristics (number of observations, number of variables, variable types, \\
-%  & coherence between variables incl. data structures e.g. clustering)                 \\
-$\square$  & Data characteristics (incl. multivariate relations and structures e.g. clustering) \\
-$\square$  & Estimand                                                                           \\ \hline 
-3 & Missingness generation                                                                \\ \cline{2-2}
-$\square$  & Missingness mechanism (incl. type or functional form of the missing data model)  \\
-$\square$  & Missingness pattern (incl. missingness proportions)                                 \\ \hline 
-4 & Methods                                                                    \\ \cline{2-2}
-$\square$  & Imputation methods (incl. parameters e.g. the number of imputations)               \\
-$\square$  & Analytic methods (incl. calculation of standard errors e.g. pooling rules)       \\
-$\square$  & Reference methods (e.g. complete case analysis)                                    \\ \hline 
-6 & Performance evaluation                                                             \\ \cline{2-2}
-$\square$  & Statistical properties (e.g. confidence validity)                                  \\
-$\square$  & Comparative performance (e.g. predictive accuracy)                                 \\ \hline
-$\square$  & Imputation-generating process (e.g. algorithmic non-convergence)                   \\
-$\square$  & Imputation model fit (e.g. posterior predictive checks)                            \\
-$\square$  & Distributional characteristics (e.g. plausibility of imputed values)               \\ \hline 
+$\square$ & \textbf{Aims}                                                                                  \\
+  & e.g. simulation scope                                                                          \\ \cline{2-2} 
+  & $\square$ simulation design (incl. pseudo-code or flow diagram)                                        \\
+  & $\square$ required level of precision (incl. number of simulation   repetitions)                       \\ \hline
+$\square$ & \textbf{Data-generating mechanisms}                                                            \\
+  & e.g. data generation and missingness generation                                                \\ \cline{2-2} 
+  & $\square$ data source (incl. model-based or design-based, sampling   variance)                         \\
+  & $\square$ data characteristics (incl. multivariate relations and data   structures such as clustering) \\
+  & $\square$ missingness mechanisms (incl. type or functional form of the   missing data model)           \\
+  & $\square$ missingness patterns (incl. missingness proportion)                                          \\ \hline
+$\square$ & \textbf{Estimands}                                                                             \\
+  & e.g. complete data target                                                                      \\ \hline
+$\square$ & \textbf{Methods}                                                                               \\
+  & e.g. missing data methods and analytic method                                                  \\ \cline{2-2} 
+  & $\square$ imputation methods (incl. parameters such as the number of   imputations)                    \\
+  & $\square$ estimation method (incl. reference methods such as complete   case analysis)                 \\ \hline
+$\square$ & \textbf{Performance measures}                                                                  \\
+  & e.g. evaluation of estimates and imputed values                                                \\ \cline{2-2} 
+  & $\square$ statistical properties (incl. comparative performance, if   applicable)                      \\
+  & $\square$ validity of imputations (incl. imputation model fit and   distributional characteristics)    \\ \hline
 \end{tabular}

### 2022-09-22 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Fix typos/spelling errors

@@
 For example, when evaluating imputation techniques aimed at prediction, different aims may be formulated than when statistical inference is of interest.
-The lack of consensus may also stem from different personal preference or scientific background.
-All in all, the lack of a common ground in evaluating imputation methodology may lead to sub-optimal use in practice.
+The lack of consensus may also stem from different personal preferences or scientific backgrounds.
+All in all, the lack of common ground in evaluating imputation methodology may lead to sub-optimal use in practice.
 In this paper, we propose a move towards a standardized evaluation of imputation methodology.
@@
 Imputation is a state-of-the-art technique for drawing valid conclusions from incomplete data.
-The technique has earned a permanent spot in research and policymaking, demonstrated e.g.
+The technique has earned a permanent spot in research and policy-making, demonstrated e.g.
 by the detailed manual created by the National Research Council \citep*{little2012prevention}.
@@
 
-In all scenarios it can be said that the quality of the solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which the imputation procedure is able to capture these properties when modeling missing values.
-In general, it holds that modeling missing data becomes more challenging when the amount of missingness increases.
+In all scenarios, it can be said that the quality of the solution obtained by imputation depends on the statistical properties of the incomplete data and the degree to which the imputation procedure is able to capture these properties when modelling missing values.
+In general, it holds that modelling missing data becomes more challenging when the amount of missingness increases.
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
@@
 
-When evaluating the statistical properties - and thereby the practical applicability - of imputation methodology, researchers most often make use of simulation studies.
+When evaluating the statistical properties--and thereby the practical applicability--of imputation methodology, researchers most often make use of simulation studies.
 Such studies are typically composed of an analysis pipeline in which data are subsequently generated, made incomplete, and imputed repeatedly under different simulation conditions.
 A set of evaluation criteria is then postulated to evaluate the performance of one or more missing data methods.
-Recent attention on best practices for method evaluation by means of simulation \citep{morr18} and \citet{pawe22}'s advice to pre-register simulation protocols will guide methodologist in conceptualizing their simulation workflow.
+Recent attention on best practices for method evaluation by means of simulation \citep{morr18} and \citet{pawe22}'s advice to pre-register simulation protocols will guide methodologists in conceptualizing their simulation workflow.
 However, there is no consensus on how to design, execute and report simulation studies aimed to evaluate imputation methodology.
@@
 How the evaluated methods are used in practice thereby governs the simulation procedure.
-Since the simulation aims and setup are intertwined with the choice of the evaluated imputation methods, one should always asses the suitablity of the chosen evaluation metrics with the imputation methods' purpose in mind.
+Since the simulation aims and setup are intertwined with the choice of the evaluated imputation methods, one should always assess the suitability of the chosen evaluation metrics with the imputation methods' purpose in mind.
 
@@
 Another threat to the generalizability of model-based simulations is the use of a single model for both data generation and imputation.
-If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied.
-Other potentially unfair comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the measurement level, and the coherence between variables.
+If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favour of the problem that is studied.
+Other potentially unfair comparative advantages in favour of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the measurement level, and the coherence between variables.
 In contrast to design-based studies, such characteristics are not always explicit simulation conditions, which may give a false sense of objectivity.
@@
 Another way to deal with missingness in a design-based simulation is to impute the incomplete dataset once, to obtain a single completed dataset to draw samples from.
-Unfortunately, this practice may favor the imputation method that was used in this initial imputation step throughout any further evaluations.
+Unfortunately, this practice may favour the imputation method that was used in this initial imputation step throughout any further evaluations.
 Just to be clear: leaving the missingness as-is and inducing additional missing values to impute is no option here, because we would not have a real and unbiased comparative truth.
@@
 Extrapolating simulation results from a specific missing data generating model to more intricate (empirical) missingness could lead to sub-optimal advice in practice.
-For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may inadvertedly induce missingness according to the special case descibed by \citet[][\S 2.7]{buur18}.
+For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may inadvertently induce missingness according to the special case described by \citet[][\S 2.7]{buur18}.
 Under this specific missingness model, list-wise deletion may outperform any imputation method.
@@
 Besides MCAR missingness, observed-data dependent missingness mechanisms--such as missing at random (MAR) missingness--are paramount in statistical inference investigations.
-A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}; generalizations to multivariate MAR missingness can be found in \citet{ampute}.
+A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, whereas generalizations to multivariate MAR missingness can be found in \citet{ampute}.
 If the missingness is to be induced in longitudinal data, generating the missingness through autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}.
@@
 Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions.
-We, therefore, propose to evaluate imputation methodology under several missingness proportions to amulate a realistic range in severity of the missingness problem.
+We, therefore, propose to evaluate imputation methodology under several missingness proportions to emulate a realistic range in severity of the missingness problem.
 Depending on how the missingness mechanism interacts with the simulated data, higher missingness proportions may yield biased or invalid completed data inferences.
@@
 The above emphasizes the need for both thorough evaluation and thorough documentation of the intended and used missingness mechanisms.
-Simulators must be make clear which mechanisms are of interest.
-To be able to verify these mechanisms, simulators must also be transparant about the process that generated the missingness.
+Simulators must make clear which mechanisms are of interest.
+To be able to verify these mechanisms, simulators must also be transparent about the process that generated the missingness.
 For example by sharing the code or detailing the process that generated the missing values.
@@
 This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
-We therefore recommend to evaluate the following points:
+We therefore recommend evaluating the following points:
 
@@
 A negligible absolute bias, for example, may already yield an almost infinite relative bias if the true value of the estimand is zero.
-In most cases, unbiased estimates may be expected under a MCAR missingness mechanism, which can be easily verified by including MCAR missingness to the scope of missingness investigations.
+In most cases, unbiased estimates may be expected under a MCAR missingness mechanism, which can be easily verified by including MCAR missingness in the scope of missingness investigations.
 
@@
 The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}.
-This makes its assessment a challenging and timeconsuming endeavor in many simulation studies.
+This makes its assessment a challenging and time-consuming endeavour in many simulation studies.
 
@@
 As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated.
-\citet{ober21} demonstrates that statistical validity can happen before algorithmic convergence is reached, but this is not a guarantee.
-Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor failproof.
+\citet{ober21} demonstrate that statistical validity can happen before algorithmic convergence is reached, but this is not a guarantee.
+Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor fail-proof.
 As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method.
@@
 Every imputation workflow should therefore contain an evaluation of the obtained imputations.
-Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend to consider the following aspects:
+Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend simulators to consider the following aspects:
 
@@
 Then, by extension, one could assume that the imputation model might be able to produce good imputations for the missing part of the data too.
-This overimputation procedure is straigtforward with the \texttt{overimpute()} function in \texttt{Amelia II} \citep{amelia} and the \texttt{where} argument in \texttt{mice} \citep{mice, caiPPC}.
+This over-imputation procedure is straightforward with the \texttt{overimpute()} function in \texttt{Amelia II} \citep{amelia} and the \texttt{where} argument in \texttt{mice} \citep{mice, caiPPC}.
 
@@
 Simulation studies should not only be well-executed, but also properly reported.
-For example, the inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favorable because of less stringent simulation conditions.
+For example, the inconsistent display of simulation conditions may impact the objectivity of meta-evaluations over imputation methods, as one method's performance may appear to be favourable because of less stringent simulation conditions.
 This may ultimately lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.

### 2022-09-22 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
More typos

@@
 Problems with producing stable imputations are not exclusive to iterative imputation algorithms.
-Any method may run into failures of the imputation-generating process and subsequently lack results, for example due to over-parameterization errors.
+Any method may run into failures of the imputation-generating process and subsequently lack results, for example due to over-parameterisation errors.
 
 Every imputation workflow should therefore contain an evaluation of the obtained imputations.
-Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend simulators to consider the following aspects:
+Even though inspecting each imputation may be labour-intensive due to the number of imputations generated in a simulation study, we highly recommend simulators consider the following aspects:
 

### 2022-09-22 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Add eps version of image

@@
 For a well-balanced judgment about the severity of the performance drop, comparative simulations with e.g.
-nonparametric models should be executed.
+non-parametric models should be executed.
 A method may perform badly, but if it still outperforms every other approach, it may yet be of great practical relevance.
@@
 This may ultimately lead to statisticians recommending a less efficient method to applied researchers, thereby limiting the efficiency of the imputation approach and unnecessarily lowering the statistical power.
-Simulators should therefore to be explicit in the descriptions of their evaluations, but deciding what and how to report may be challenging.
+Simulators should therefore be explicit in the descriptions of their evaluations.
+However, deciding what and how to report may be challenging.
 The simulation design could be presented textually, in a flow chart or as a block of pseudo-code, whereas missingness mechanisms could be written as a function of the data or displayed graphically.

### 2022-11-02 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
reformat figure code

@@
 After all, the principle of imputation is very intuitive.
-The idea behind imputation is to impute (fill in) missing values to obtain a valid estimate of what could have been.
+The idea behind imputation is to impute (fill in) missing values, which may yield a valid estimate of what could have been.
 The completed data that are thus obtained can be analyzed by standard techniques, which conveniently separates the missing data problem from the analysis.
@@
 However, when (strong) relations in the data are present, the observed parts can hold great predictive power for the models that estimate the missingness.
-In that case, imputation would be substantially more efficient than the ubiquitous complete case analysis.
+In that case, imputation may be substantially more efficient than the ubiquitous complete case analysis.
 
@@
 We limit the scope of this paper to comparative simulation studies in the context of statistical inference.
-With that, we exclude the evaluation of imputation methodology for prediction or causal inference.
+With that, we exclude the evaluation of imputation methodology for detailed aspects of causal inference and prediction.
 The type of simulation study under consideration in this paper requires a form of `comparative truth' to assess the inferential validity of imputation methods.
@@
 Data-generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition.
-There are two general approaches to generating complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled from a sufficiently large observed set, such as official registers.
-The simulator should choose their data-generating mechanism in line with the study scope.
+We highlight two general data generation approaches: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled from a sufficiently large observed set, such as official registers.
 Model-based data-generating mechanisms have advantages in flexibility and precision, since data are generated from a known statistical model and the true theoretical parameters can be derived.
 A design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest.
+Of course, the simulator should choose their data-generating mechanism in line with the study scope.
+Other methods to simulate data may be more suitable, e.g.
+synthetization of the actual data-generating process \citep{li22; volk21}.
+% "A nonparametric regression method, undersmoothed highly adaptive lasso, is used to generate the simulated distribution which preserves important features from the observed data and reproduces a set of true target parameters."
 
@@
 Real-life data hardly ever follow a known theoretical distribution, so there is no guarantee that simulation results are generalizable.
-Moreover, data are often generated such that the problem being studied is most pronounced, e.g.
+[**TODO: change this to 'We cannot know so safer not to pretend we do'**] Moreover, data are often generated such that the problem being studied is most pronounced, e.g.
 with consistently high correlations between groups of variables.

### 2022-11-17 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex

@@
 After all, the principle of imputation is very intuitive.
-The idea behind imputation is to impute (fill in) missing values, which may yield a valid estimate of what could have been.
+The idea behind imputation is to impute (fill in) missing values to obtain a valid estimate of what could have been.
 The completed data that are thus obtained can be analyzed by standard techniques, which conveniently separates the missing data problem from the analysis.
@@
 Real-life data hardly ever follow a known theoretical distribution, so there is no guarantee that simulation results are generalizable.
-[**TODO: change this to 'We cannot know so safer not to pretend we do'**] Moreover, data are often generated such that the problem being studied is most pronounced, e.g.
+Moreover, data are often generated such that the problem being studied is most pronounced, e.g.
 with consistently high correlations between groups of variables.

### 2022-11-17 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
updates Gerko so far

@@
 \usepackage{graphicx, latexsym}
+\usepackage{color}
+
 % \setlength{\bibsep}{2pt}
@@
 One potential reason for discrepancies between imputation method performance in simulations versus applications is the use of sampling variance in the simulation design.
-Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations and only use the missing data generation procedure as the source of Monte Carlo variation \citep{vink14}.
-If sampling variance is not of interest, a single complete data set can serve as comparative truth in all simulation repetitions--as opposed to a `standard' simulation setup, in which data is drawn every repetition.
+Simulation studies on missing data methodology have the unique option to exclude conventional sampling schemes and focus on a single finite population.
+When a finite population is made incomplete, sampling variance must be excluded from inference evaluations and the missing data generation procedure serves as the necessary source of Monte Carlo variation \citep{vink14, raghunathan2003multiple}.
+\citet{vink14} propose pooling rules for multiple imputation on finite populations and demonstrate that valid inferences are then obtained by excluding the sampling variance components from the calculation of total variance about $(Q - \bar{Q})$, where $Q$ would be the estimand and $\bar{Q} = \sum_{l=1}^m \hat{Q}/m$ is the average over the $m$ imputed estimates \citep[cf.][p.
+76-77]{rubi87}.
+The total variance $T=B+B/m$ would then solely rely on the between imputation variance $B = \sum_{l=1}^m (\hat{Q}_l - \bar{Q})^\prime(\hat{Q}_l - \bar{Q}) / m-1$.
+
+When sampling variation is excluded from the simulation design, a single complete data set can serve as comparative truth in all simulation repetitions--as opposed to a `standard' simulation setup, in which data is drawn every repetition.
+This is computationally convenient and may potentially sharpen comparisons between imputation methods, clarifying inconclusive results on their ability to solve for missingness.
 After all, we are interested in solving for the missingness and have no need for the noise induced by the sampling mechanism for evaluation in such studies.
 
-Removing sampling variance from performance evaluation is computationally convenient and may potentially sharpen comparisons between imputation methods, clarifying inconclusive results on their ability to solve for missingness.
-However, when deviating from the standard simulation workflow, conventional pooling rules for multiple imputation \citep[cf.][p.
-76-77]{rubi87} do not apply and alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.
-This opens the door towards human error in the implementation and interpretation of such a simulation study.
-On the one hand, we highly value the additional flexibility that this design offers.
+We highly value the additional flexibility that the above outlined design offers.
 Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against.
@@
 Other methods to simulate data may be more suitable, e.g.
-synthetization of the actual data-generating process \citep{li22; volk21}.
+synthetization of the actual data-generating process \citep{li22,volk21}.
 % "A nonparametric regression method, undersmoothed highly adaptive lasso, is used to generate the simulated distribution which preserves important features from the observed data and reproduces a set of true target parameters."
@@
 The most extreme example would be when MAR is induced from data without multivariate relations.
-The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is generated randomly (see Figure 1).
+The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is generated randomly.
+Figure \ref{Fig} demonstrates this: Both univariate MAR mechanisms in Figure \ref{Fig} are induced based on the right tail, but only the mechanisms for which the \emph{variable to ampute} and the \emph{missingness covariate} have a positive correlation will yield valid MAR mechanisms.
+In the case of zero correlation between the data columns, iterative univariate non-MCAR missingness schemes may induce invalid missingness mechanisms that mimic MCAR.
+Table \ref{Tab} demonstrates this and highlights the importance of including complete case analysis as an evaluation analysis to allow for identifying such invalidly generated missingness mechanisms in simulation.
 In fact, this would amount to one of the special cases described by \citet{buur18} under which complete case analysis would be more efficient than imputation: the missingness does not depend on the incomplete variable.
@@
   }      
-  \caption{Three different realizations of missingness generation for MCAR and right-tailed MAR missingness mechanisms for different levels of correlation $\rho$ between the variable to ampute and the covariates that guide the probability to be missing.
-Note that the conditional probability of right-tailed MAR ($\rho=0$) mimics that of MCAR.}
-    \label{Fig4}
+  \caption{Three different realizations of missingness generation for MCAR and right-tailed MAR missingness mechanisms for different levels of correlation $\rho$ between the \emph{variable to ampute} and the \emph{missingness covariate} that guide the probability to be missing.
+Note that the conditional probability of right-tailed MAR ($\rho=0$) mimics that of MCAR because the variable to ampute and the missingness covariate have $\rho=0$.}
+    \label{Fig}
 \end{figure}
+
+% latex table generated in R 3.2.2 by xtable 1.8-0 package
+% Thu Feb  4 16:27:54 2016
+\begin{table}[t!]
+\caption{Inferences obtained by complete case analysis (CCA) and multiple imputation over 1000 simulations.
+Displayed are the bias of the mean, coverage rate (cov) of the corresponding 95\% confidence interval and the confidence interval width (ciw).
+Imputations are generated by Bayesian linear regression imputation.}
+\centering
+\begin{tabular}{rlccccccc}
+  \hline
+ 	& &\multicolumn{3}{c}{CCA} &&\multicolumn{3}{c}{imputation}	\\\cline{3-5}\cline{7-9}
+ \multicolumn{2}{c}{mechanism}	& bias & cov & ciw && bias & cov & ciw \\ 
+  \hline
+  $[\rho = .8]$ &mcar & 0.001 & 0.959 & 0.179 && -0.001 & 0.946 & 0.156 \\ 
+  $[\rho = .8]$ &mar & {\color{red} 0.334} & {\color{red} 0.000} & 0.169 && 0.004 & 0.955 & 0.195 \\ 
+  $[\rho = .0]$ &mar & 0.003 & 0.973 &{\color{red} 0.180} && 0.013 & 0.947 & 0.311 \\    \hline
+\end{tabular}
+\label{Tab}
+\end{table}
 
@@
 Therefore, it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
-It may be wise to include MNAR missingness in the simulation study just in case.
+It may be wise to include MNAR missingness in the simulation study just in case: A method that performs well under MAR and some cases of MNAR may be preferable over a method that only works under MAR, and not under MAR.
+Such `late-phase` methodological investigations \citep[see][]{hei22} can provide an understanding of an imputation method's robustness.
 Alternatively, a sensitivity analysis may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.

### 2022-11-17 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
added to text and table

@@
 When a finite population is made incomplete, sampling variance must be excluded from inference evaluations and the missing data generation procedure serves as the necessary source of Monte Carlo variation \citep{vink14, raghunathan2003multiple}.
-\citet{vink14} propose pooling rules for multiple imputation on finite populations and demonstrate that valid inferences are then obtained by excluding the sampling variance components from the calculation of total variance about $(Q - \bar{Q})$, where $Q$ would be the estimand and $\bar{Q} = \sum_{l=1}^m \hat{Q}/m$ is the average over the $m$ imputed estimates \citep[cf.][p.
+\citet{vink14} propose pooling rules for multiple imputation on finite populations and demonstrate that valid inferences are then obtained by excluding the sampling variance components from the calculation of the total variance about $(Q - \bar{Q})$, where $Q$ would be the estimand and $\bar{Q} = \sum_{l=1}^m \hat{Q}/m$ is the average over the $m$ imputed estimates \citep[cf.][p.
 76-77]{rubi87}.
@@
 The challenge with model-based data-generating mechanisms is that a method's performance on simulated data may not translate to empirical data.
-Real-life data hardly ever follow a known theoretical distribution, so there is no guarantee that simulation results are generalizable.
+Real-life data may not follow a known theoretical distribution, so there is no guarantee that simulation results are generalizable.
 Moreover, data are often generated such that the problem being studied is most pronounced, e.g.
@@
 Table \ref{Tab} demonstrates this and highlights the importance of including complete case analysis as an evaluation analysis to allow for identifying such invalidly generated missingness mechanisms in simulation.
-In fact, this would amount to one of the special cases described by \citet{buur18} under which complete case analysis would be more efficient than imputation: the missingness does not depend on the incomplete variable.
+GV: REF NAAR SIMULATIE In fact, this would amount to one of the special cases described by \citet{buur18} under which complete case analysis would be more efficient than imputation: the missingness does not depend on the incomplete variable.
 This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is pointless.
@@
 
-% latex table generated in R 3.2.2 by xtable 1.8-0 package
-% Thu Feb  4 16:27:54 2016
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu Nov 17 17:52:47 2022
 \begin{table}[t!]
@@
   \hline
-  $[\rho = .8]$ &mcar & 0.001 & 0.959 & 0.179 && -0.001 & 0.946 & 0.156 \\ 
-  $[\rho = .8]$ &mar & {\color{red} 0.334} & {\color{red} 0.000} & 0.169 && 0.004 & 0.955 & 0.195 \\ 
-  $[\rho = .0]$ &mar & 0.003 & 0.973 &{\color{red} 0.180} && 0.013 & 0.947 & 0.311 \\    \hline
+$[\rho = .8]$ &mcar & 0.000 & 0.940 & 0.178 & & -0.000 & 0.941 & 0.160 \\ 
+$[\rho = .8]$ &mar   & 0.335 & 0.000 & 0.169 & & 0.004 & 0.954 & 0.190 \\ 
+$[\rho = .0]$ &mar   & 0.001 & 0.973 & 0.180 & & 0.011 & 0.939 & 0.311 \\ 
+   \hline
 \end{tabular}
@@
 Therefore, it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios.
-It may be wise to include MNAR missingness in the simulation study just in case: A method that performs well under MAR and some cases of MNAR may be preferable over a method that only works under MAR, and not under MAR.
+It may be wise to include MNAR missingness in the simulation study just in case: A method that performs well under MAR and some cases of MNAR may be preferable over a method that only works under MAR, and not under MNAR.
 Such `late-phase` methodological investigations \citep[see][]{hei22} can provide an understanding of an imputation method's robustness.
@@
 The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
+This means that, given the framework provided by \citet{rubi87}, and for any $(Q-\bar(Q)$, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
 We therefore recommend evaluating the following points:
@@
 A negligible absolute bias, for example, may already yield an almost infinite relative bias if the true value of the estimand is zero.
-In most cases, unbiased estimates may be expected under a MCAR missingness mechanism, which can be easily verified by including MCAR missingness in the scope of missingness investigations.
+In most cases, unbiased estimation may be expected under a MCAR missingness mechanism, which can be easily verified by including MCAR missingness in the scope of missingness investigations.
 
-(ii) The intervals for estimates should have a valid coverage of the population (i.e.
+(ii) The intervals around estimates should have a valid coverage of the population (i.e.
 true) value.
@@
 
-(iv) Predictive accuracy may be quantified using the root mean squared error (RMSE).
+(iv) Resemblence to the true data values may be quantified using the root mean squared error (RMSE).
 We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may inflate the type I error rate of statistical inferences \citep[Chapter 2.6][]{buur18}.
@@
 
-After establishing the statistical validity of an imputation routine on the level of the analysis estimates, it remains paramount to also study the imputation-generating process.
+After establishing the statistical validity of an imputation routine on the level of the estimand, it remains paramount to also study the imputation-generating process.
 Omitting such investigations may yield sub-optimal results.
 One simple example arises often in practice: generating implausible imputed values.
-Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
+Even though the estimation on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers.
 For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large.
@@
 Moreover, pairing complete case analysis with MCAR and MAR mechanisms allows the simulator to evaluate the impact of the missing data generation process under the chosen analysis model and to identify the unintentional generation of accidental MNAR.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 $\square$ & \textbf{Methods}                                                                               \\
-  & e.g. missing data methods and analytic method                                                  \\ \cline{2-2} 
+  & e.g. missing data methods, analytic method and inference pooling rules                                    \\ \cline{2-2} 
   & $\square$ imputation methods (incl. parameters such as the number of   imputations)                    \\
-  & $\square$ estimation method (incl. reference methods such as complete   case analysis)                 \\ \hline
+  & $\square$ estimation method (incl. reference methods such as complete   case analysis)                 \\
+  & $\square$ methods used to construct standard errors and confidence intervals \\ 
+  & $\;\;$ $-$ Such as rules by \citet{rubi87}, \citet{bar99}, \citet{rei03}, etc.    \\ \hline
 $\square$ & \textbf{Performance measures}                                                           \\      
@@
   & $\square$ statistical properties (incl. comparative performance, if   applicable)                      \\
-  & $\square$ validity of imputations (incl. imputation model fit and   distributional characteristics)    \\ \hline
+  & $\square$ validity of imputations (incl. imputation model fit and distributional characteristics)    \\ 
+  & $\;\;$ $-$ Does the imputation model fit well to the observed values? (e.g. posterior predictive check)   \\ 
+  & $\;\;$ $-$ Are there unrealistic values in the data? (e.g. negative where positive expected)   \\ 
+  & $\;\;$ $-$ Are there implausible combinations in the data? (e.g. pregnant grandfathers)   \\ \hline
+
 \end{tabular}

### 2022-11-21 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
submitted revisions round 2

@@
 % } %%% semicolon and fullpoint added here for keyword style
-\noindent \hspace*{-4pc} Supporting Information for this article is available from \underline{www.gerkovink.com/evaluation}.\\
+\noindent \hspace*{-4pc} Supporting information and a simulation archive are available from \underline{www.gerkovink.com/evaluation}.\\
 % \hspace*{-4pc} \underline{gerkovink.com/evaluation}. 
@@
 
-We highly value the additional flexibility that the above outlined design offers.
+We highly value the additional flexibility that the above-outlined design offers.
 Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against.
@@
 yield valid inference) under MCAR, the statistical properties of the procedure are not universally sound.
-Sadly, the straightforward case of MCAR is often neglected from simulation studies, while it offers an informative reference condition in many simulation setups.
+Sadly, the straightforward case of MCAR is often neglected in simulation studies, while it offers an informative reference condition in many simulation setups.
 
@@
 The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-This means that, given the framework provided by \citet{rubi87}, and for any $(Q-\bar(Q)$, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
+This means that, given the framework provided by \citet{rubi87}, and for any $(Q-\bar{Q}$, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
 We therefore recommend evaluating the following points:
@@
 
-(iv) Resemblence to the true data values may be quantified using the root mean squared error (RMSE).
+(iv) Resemblance to the true data values may be quantified using the root mean squared error (RMSE).
 We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values and it may inflate the type I error rate of statistical inferences \citep[Chapter 2.6][]{buur18}.
-However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the predicted values can be compared between methods.
+However, if a study is aimed at obtaining predictions, and inferential validity is not of interest, the RMSE of the prediction errors (residuals) can be compared between methods.
 Then, the RMSE may yield valuable information about the methods' comparative efficiency in terms of accuracy and precision.

### 2022-11-23 Manuscript/response_reviewer_2.md

Reviewer 2

Thanks to the authors for addressing my comments on the first submission.

The authors choose to focus on the first option that I listed at the previous review: Incomplete covariate values in observational studies with focus on unbiased, efficient parameter estimation. This seems a sensible decision. This authors be made explicit in the title and abstract, both of which promise to be far more general.

- I noted a problem with a specific suggestion in section 2.1, which is to remove sampling variation by fixing the complete data and only simulating missing values. The authors have partially addressed this but I don???t think been clear enough. There are two important points that I think readers would miss:

  1. The suitability of this approach depends on the performance measures to be used. Specifically, any that involve sampling variation are of no interest (i.e. it???s only ok if you are only evaluating bias). The problem can be
2. This approach is restricted to using a selection-model factorisation when generating data. If using a pattern-mixture approach, you cannot do this. Perhaps this point is obvious but it???s not explicit.

**We have rewritten the paragraph to avoid confusion and have outlined that this scenario only applies to simulation studies where the simulator has the flexibility to induce missing data uncertainty in a finite population. In that case, the Rubin's rules about the total variance (Q-\bar{Q}) have to be adapted to T = B + B/m. We have also detailed this in the revised manuscript. If a finite population is repeatedly made incomplete in simulation, such a finite population may serve as comparative truth and valid inference are obtained by the above mentioned pooling rule. This scenario applies to all performance measures and data generating schemes and has been a part of our simulation toolbox for years. The publication by Vink & Van Buuren (2014) demonstrates application and validity.**

- In section 2.2, I like the recommendation to explore MNAR even if something is primarily intended to work with MAR. However, I think there may be two points to add. First, explicitly say that this is a way of understanding robustness: method A may perform ok under MAR and some MNAR mechanisms, while method B performs well under MAR but very poorly under any MNAR mechanism. In this case we may favour A over B. Second, this issue is IMO applicable to ???late-phase??? methodological research (see Heinze et al. https://arxiv.org/abs/2209.13358); in early-phase work, I think it's perfectly acceptable to say 'this method works well in the setting it's intended to work well in'.

**Thanks. This is a useful addition to this paragraph. We have taken and implemented your suggestions.**

- It took a while to get my head around figure 1, and I???d suggest a little more explanation. The ???MCAR??? and ???Right-tailed MAR (??=0.8)??? panels are clear, but the ???Right-tailed MAR (??=0)??? is not. First, I cannot see what is ???right-tailed??? about it. Second, while I agree that MCAR requires an identical probability of missingness for each individual, how is this realised in a simulation study involving repetitions? The right panel of figure 1 seems to be ???MAR given individual???. Are notional ???individuals??? retained for each repetition? I don???t know the answer here, but two points:
1. fixing something across repetitions vs. drawing it each time can make a difference (e.g. comment above about only simulating missing data).
2. If you choose not to fix this ???MAR??? mechanism across reps, it becomes operationally indistinguishable from MCAR.

**We understand the confusion as the right-tailed nature of the missingness is lost through the zero correlation between the variable to ampute and the variable that serves as missingness covariate. This is exactly the point of the figure, but we apparently have not made this sufficiently clear. Ww have adjusted the text and have added a table that includes a small set of evaluation metrics that demonstrate both the importance of valid missingness and the ability to infer invalid missingness mechanisms through the comparison between results obtained on MAR/MNAR with CCA and multiple imputation.**

- I like the properties suggested in section 2.4. Two comments:
1. It would be good to note that these properties are for an estimand of interest rather than just any parameter in the analysis model
2. (Root) mean squared error is arguably not for predictive accuracy, which you have said is not the target here, but a trade-off between bias and variance when a biased procedure has lower variance than an unbiased procedure.

**We've noted (1) and changed predictive accuracy to 'resemblence to the true values' for (2)**

  - In table 1, I wondered about adding a point under methods for ???how SEs and confidence intervals are to be constructed???. Rubin???s rules are likely the default but people might use Robins & Wang???s rules, Reiter???s rules, Efron???s full-mechanism bootstrap, von Hippel and Bartlett???s bootstrap???
**added**

  - The point about validity of imputations is interesting. Rubin makes the point that the aim of MI is validity of inference, not of imputations,  Do the authors have examples of how to assess the validity of imputations? I broadly agree with this as a useful diagnostic but ideas of how to operationalise it in simulation studies would be useful.
**We've added some suggestions to text and table**

MISC./MINOR COMMENTS
- The introduction says ???The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.??? This is behind some imputation but certainly not multiple imputation. Rubin explicitly talks about the idea as being valid inference, not to recreate the missing values.
**Although we fully agree with you, we believe this wording is correct as we do not yet speak of multiple imputation in particular and we talk about "valid estimate of what could have been" and not values that could have been. We also very clearly separate single and multiple imputation in the following paragraphs and do not wish to confuse the reader with semantics in the first paragraph.**

- The statement ???Real-life data hardly ever follow a known theoretical distribution??? seems impossible to falsify. Perhaps safer to describe this as ???We cannot know so safer not to pretend we do???.
**The revised manuscript relaxes this particular phrasing**

- Since reviewing the first submission, I've seen some simulation studies where people have MAR and positivity vs. MAR and non-positivity (or near non-positivity). No pressure but I think it might be worth mentioning this in section 2.3.
**It is not clear to us to what you refer with positivity. Do you refer by positivity to any unique combination of confounding variables having a non-zero probability of missingness? Or is your comment referring to (non)positive covariance matrices? That said, we realize that given the timeframe of the review process, many potentially interesting notes could be made about newly published or pre-printed research. References that redefined the state-of-the-art in imputation methodology evaluation or were vital to the validity of our manuscript, were added between review rounds. For reasons of brevity we cannot include all. If you feel that we have missed such a must-include reference, we would be delighted to hear from you. After all, your comments have been very valuable to us.**

### 2022-12-21 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
revisions 3

@@
 One potential reason for discrepancies between imputation method performance in simulations versus applications is the use of sampling variance in the simulation design.
-Simulation studies on missing data methodology have the unique option to exclude conventional sampling schemes and focus on a single finite population.
-When a finite population is made incomplete, sampling variance must be excluded from inference evaluations and the missing data generation procedure serves as the necessary source of Monte Carlo variation \citep{vink14, raghunathan2003multiple}.
-\citet{vink14} propose pooling rules for multiple imputation on finite populations and demonstrate that valid inferences are then obtained by excluding the sampling variance components from the calculation of the total variance about $(Q - \bar{Q})$, where $Q$ would be the estimand and $\bar{Q} = \sum_{l=1}^m \hat{Q}/m$ is the average over the $m$ imputed estimates \citep[cf.][p.
+For example, when dealing with infinite populations, the conventional pooling rules proposed by \citep[cf.][p.
+76-77]{rubi87} apply.
+When dealing with a finite incomplete population, some adjustment is needed for pooling rules to yield correct inferences.
+In such cases, valid inferences can be obtained by excluding the sampling variance components from the calculation of the total variance about $(Q - \bar{Q})$ \citep{vink14, raghunathan2003multiple}, where $Q$ would be the estimand and $\bar{Q} = \sum_{l=1}^m \hat{Q}/m$ is the average over the $m$ imputed estimates \citep[cf.][p.
 76-77]{rubi87}.
-The total variance $T=B+B/m$ would then solely rely on the between imputation variance $B = \sum_{l=1}^m (\hat{Q}_l - \bar{Q})^\prime(\hat{Q}_l - \bar{Q}) / m-1$.
+\citet{vink14} propose that the total variance $T=B+B/m$ would then solely rely on the between imputation variance $B = \sum_{l=1}^m (\hat{Q}_l - \bar{Q})^\prime(\hat{Q}_l - \bar{Q}) / m-1$ and demonstrate that the degrees of freedom would then equal $m-1$.
 
-When sampling variation is excluded from the simulation design, a single complete data set can serve as comparative truth in all simulation repetitions--as opposed to a `standard' simulation setup, in which data is drawn every repetition.
-This is computationally convenient and may potentially sharpen comparisons between imputation methods, clarifying inconclusive results on their ability to solve for missingness.
-After all, we are interested in solving for the missingness and have no need for the noise induced by the sampling mechanism for evaluation in such studies.
+We can make use of this unique real-world data property when simulating the performance of incomplete data methodology.
+When we take a single drawn complete data set as our comparitive truth, just inducing simulated missingness would suffice, and the necessary Monte Carlo variance for our evaluations would stem from the sampled missingness.
+The above detailed pooling rules can be used to obtain valid inferences in those simulations.
+A detailed overview of simulation strategies that adopt this approach, as well as conventional simulation strategies can be found in \citet{simmiss}.
 
-We highly value the additional flexibility that the above-outlined design offers.
-Especially in the case of data transformations, it can be challenging to derive the true parametric references to evaluate against.
-Simply using a single generated complete set in which missingness is induced, and evaluating against that generated data set avoids a plethora of procedural problems and computational challenges.
-On the other hand, we want to warn simulators against potential mishaps, e.g.
-when calculating and comparing standard errors.
+We would like to note that it is not necessary to exclude sampling variance from missing data simulations, as the additional variance induced by the sampling step would be ignorable.
+In practice, however, we have found it to be convenient to allow for a single data set to serve as comparative truth in all simulation repetitions.
+Especially for data-driven simulation strategies with complex data transformations, it can be challenging to derive the true parametric references to evaluate against.
+Evaluating against a single generated set would then avoid many procedural problems and computational challenges.
+In practice, there are also other benefits to this simulation approach as it is computationally convenient, requires less data archiving and may potentially sharpen comparisons between imputation methods, thereby clarifying inconclusive results on method imputation performance.
+After all, we are interested in solving for the missingness and can do without the ignorable noise induced by the sampling mechanism for evaluation in such studies.
 
@@
 The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete.
-This means that, given the framework provided by \citet{rubi87}, and for any $(Q-\bar{Q}$, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
+This means that, given the framework provided by \citet{rubi87}, and for any $(Q-\bar{Q})$, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should always be studied.
 We therefore recommend evaluating the following points:
@@
 \begin{acknowledgement}
-We thank the Amices team for the fruitful discussions.
+We thank the Amices team for the fruitful discussions and highly value the comments and suggestions from the reviewers and the editors. 
 \end{acknowledgement}

### 2022-12-22 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Fix spelling error

@@
 We can make use of this unique real-world data property when simulating the performance of incomplete data methodology.
-When we take a single drawn complete data set as our comparitive truth, just inducing simulated missingness would suffice, and the necessary Monte Carlo variance for our evaluations would stem from the sampled missingness.
-The above detailed pooling rules can be used to obtain valid inferences in those simulations.
+When we take a single drawn complete data set as our comparative truth, just inducing simulated missingness would suffice, and the necessary Monte Carlo variance for our evaluations would stem from the sampled missingness.
+The above-detailed pooling rules can be used to obtain valid inferences in those simulations.
 A detailed overview of simulation strategies that adopt this approach, as well as conventional simulation strategies can be found in \citet{simmiss}.
